### 2016-07-22 Berkeley_PD_Stop.Rmd
updated 072216

---
title: "Test"
author: "Amy Zhu"
date: "July 19, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)

stop_df <- read.csv("~/Desktop/Berkeley_PD_-_Stop_Data.csv")

#disps <- as.character(stop_df$Dispositions)
#max(nchar(disps))
#disps[which(nchar(disps) == 42)]

stop_df <- mutate(stop_df, Dispositions = as.character(stop_df$Dispositions))

#entries with "zz do not use Assist" 
stop_df <- mutate(stop_df, Dispositions = str_replace_all(stop_df$Dispositions, "zz do not use Assist", ""))
#entries with "FA - TIGHT"
stop_df <- mutate(stop_df, Dispositions = str_replace_all(stop_df$Dispositions, "FA - TIGHT", ""))
#fixing entries that were input incorrectly
stop_df <- mutate(stop_df, Dispositions = str_replace_all(stop_df$Dispositions, "[0]{6}[1]?", "00000"))

#deal with special cases first 
stop_df <- mutate(stop_df, Other = str_replace_all(stop_df$Dispositions, "[ ,]{0,2}[a-zA-Z0-9]{6}", ""))
stop_df <- mutate(stop_df, Other = str_replace_all(stop_df$Other, "^[ ,]{0,4}", ""))

stop_df <- mutate(stop_df, Dispositions = str_extract_all(stop_df$Dispositions, "[a-zA-Z0-9]{6}"))

View(stop_df)

dispositions <- separate(stop_df, Dispositions, into = c("Individual_1", "Individual_2", "Individual_3", "Individual_4", "Individual_5", "Individual_6"), extra = "merge", fill = "right")

#deconstructing_6_characters <- function (x) {
 # dispositions %>%
  #  filter(nchar(x) == 6) %>%
   #   mutate(Race_x = substr(x,1,1)) %>%
    #    mutate(Gender_x = substr (x,2,2)) %>%
     #     mutate(AgeRange_x = substr (x,3,3)) %>%
      #      mutate(Reason_x =substr(x,4,4)) %>%
       #       mutate(Enforcement_x = substr(x,5,5)) %>%
        #         mutate(CarSearch_x = substr(x,6,6))
#}
#lapply

dispositions_6_characters_1 <- dispositions %>%
    filter(nchar(Individual_1) == 6) %>%
      mutate(Race_1 = substr(Individual_1,1,1)) %>%
        mutate(Gender_1 = substr (Individual_1,2,2)) %>%
          mutate(AgeRange_1 = substr (Individual_1,3,3)) %>%
            mutate(Reason_1 =substr(Individual_1,4,4)) %>%
              mutate(Enforcement_1 = substr(Individual_1,5,5)) %>%
                 mutate(CarSearch_1 = substr(Individual_1,6,6))

dispositions_6_characters_2 <- dispositions %>%
    filter(nchar(Individual_2) == 6) %>%
      mutate(Race_2 = substr(Individual_2,1,1)) %>%
        mutate(Gender_2 = substr (Individual_2,2,2)) %>%
          mutate(AgeRange_2 = substr (Individual_2,3,3)) %>%
            mutate(Reason_2 =substr(Individual_2,4,4)) %>%
              mutate(Enforcement_2 = substr(Individual_2,5,5)) %>%
                 mutate(CarSearch_2 = substr(Individual_2,6,6))

dispositions_6_characters_3 <- dispositions %>%
    filter(nchar(Individual_3) == 6) %>%
      mutate(Race_3 = substr(Individual_3,1,1)) %>%
        mutate(Gender_3 = substr (Individual_3,2,2)) %>%
          mutate(AgeRange_3 = substr (Individual_3,3,3)) %>%
            mutate(Reason_3 =substr(Individual_3,4,4)) %>%
              mutate(Enforcement_3 = substr(Individual_3,5,5)) %>%
                 mutate(CarSearch_3 = substr(Individual_3,6,6))

dispositions_6_characters_4 <- dispositions %>%
    filter(nchar(Individual_4) == 6) %>%
      mutate(Race_4 = substr(Individual_4,1,1)) %>%
        mutate(Gender_4 = substr (Individual_4,2,2)) %>%
          mutate(AgeRange_4 = substr (Individual_4,3,3)) %>%
            mutate(Reason_4 =substr(Individual_4,4,4)) %>%
              mutate(Enforcement_4 = substr(Individual_4,5,5)) %>%
                 mutate(CarSearch_4 = substr(Individual_4,6,6))

dispositions_6_characters_5 <- dispositions %>%
    filter(nchar(Individual_5) == 6) %>%
      mutate(Race_5 = substr(Individual_5,1,1)) %>%
        mutate(Gender_5 = substr (Individual_5,2,2)) %>%
          mutate(AgeRange_5 = substr (Individual_5,3,3)) %>%
            mutate(Reason_5 =substr(Individual_5,4,4)) %>%
              mutate(Enforcement_5 = substr(Individual_5,5,5)) %>%
                 mutate(CarSearch_5 = substr(Individual_5,6,6))

dispositions_6_characters_6 <- dispositions %>%
    filter(nchar(Individual_6) == 6) %>%
      mutate(Race_6 = substr(Individual_6,1,1)) %>%
        mutate(Gender_6 = substr (Individual_6,2,2)) %>%
          mutate(AgeRange_6 = substr (Individual_6,3,3)) %>%
            mutate(Reason_6 =substr(Individual_6,4,4)) %>%
              mutate(Enforcement_6 = substr(Individual_6,5,5)) %>%
                 mutate(CarSearch_6 = substr(Individual_6,6,6))

#View(dispositions_1)
#View(dispositions_2)
#View(dispositions_3)
#View(dispositions_4)
#View(dispositions_5)
#View(dispositions_6)

#deconstructing_other_characters <- function(x) {
#}
#lapply

dispositions_other_characters_1 <- dispositions  %>%
  filter(nchar(Individual_1) == 1 | nchar(Individual_1) == 2 | nchar(Individual_1) == 3 | nchar(Individual_1) == 5) %>%
    mutate(Other_1 = Individual_1)

dispositions_other_characters_2 <- dispositions  %>%
  filter(nchar(Individual_2) == 1 | nchar(Individual_2) == 2 | nchar(Individual_2) == 3 | nchar(Individual_2) == 5) %>%
    mutate(Other_2 = Individual_2)

dispositions_other_characters_3 <- dispositions  %>%
  filter(nchar(Individual_3) == 1 | nchar(Individual_3) == 2 | nchar(Individual_3) == 3 | nchar(Individual_3) == 5) %>%
    mutate(Other_3 = Individual_3)

dispositions_other_characters_4 <- dispositions  %>%
  filter(nchar(Individual_4) == 1 | nchar(Individual_4) == 2 | nchar(Individual_4) == 3 | nchar(Individual_4) == 5) %>%
    mutate(Other_4 = Individual_4)

dispositions_other_characters_5 <- dispositions  %>%
  filter(nchar(Individual_5) == 1 | nchar(Individual_5) == 2 | nchar(Individual_5) == 3 | nchar(Individual_5) == 5) %>%
    mutate(Other_5 = Individual_5)

dispositions_other_characters_6 <- dispositions  %>%
  filter(nchar(Individual_6) == 1 | nchar(Individual_6) == 2 | nchar(Individual_6) == 3 | nchar(Individual_6) == 5) %>%
    mutate(Other_6 = Individual_6)

#View(dispositions_other_characters_1)
#View(dispositions_other_characters_2)
#View(dispositions_other_characters_3)
#View(dispositions_other_characters_4)
#View(dispositions_other_characters_5)
#View(dispositions_other_characters_6)

dispositions <- left_join(dispositions, dispositions_1)
dispositions <- left_join(dispositions, dispositions_other_characters_1)
dispositions <- left_join(dispositions, dispositions_2)
dispositions <- left_join(dispositions, dispositions_other_characters_2)
dispositions <- left_join(dispositions, dispositions_3)
dispositions <- left_join(dispositions, dispositions_other_characters_3)
dispositions <- left_join(dispositions, dispositions_4)
dispositions <- left_join(dispositions, dispositions_other_characters_4)
dispositions <- left_join(dispositions, dispositions_5)
dispositions <- left_join(dispositions, dispositions_other_characters_5)
dispositions <- left_join(dispositions, dispositions_6)
dispositions <- left_join(dispositions, dispositions_other_characters_6)

dispositions <- dispositions[,c(1:4, 11:17, 6, 18:24, 7, 25:31, 8, 32:38, 9, 39:45, 10, 46:52)]
View(dispositions)

```

### 2016-07-22 Berkeley_PD_Stop.Rmd
updated 072216

@@
 
+#mutate dispositions column to only contain six character cases
 stop_df <- mutate(stop_df, Dispositions = str_extract_all(stop_df$Dispositions, "[a-zA-Z0-9]{6}"))
+#removes empty cases
+stop_df <- mutate(stop_df, Dispositions = str_replace_all(stop_df$Dispositions, "character(0)", ""))
+
+dispositions <- separate(stop_df, Dispositions, into = c("Individual_1", "Individual_2", "Individual_3", "Individual_4", "Individual_5", "Individual_6"), extra = "merge", fill = "right")
 
 View(stop_df)
+View(dispositions)
+###
 

### 2016-07-22 Rebeccas_location_Code/B_stop_location_script.R
updated dataset

@@ -2,6 +2,8 @@
 # Shane Lynn 10/10/2013 at http://www.r-bloggers.com/batch-geocoding-with-r-and-google-maps/
 # Modified by Rebecca Reus for use with Berkeley policing data for year 2016
 
+setwd("~/Desktop/Stats R/PROJECT/finalproject_angry_ladies/Rebeccas_location_Code")
+
 #load up the ggmap library and data wrangling / string analysis libraries:
 library(sp)
 library(ggmap)
@@ -13,7 +15,7 @@ library(lubridate)
 library(ggplot2)
 
 # row number to finish at (will start at the last spot):
-number_of_rows_to_do<-2496
+number_of_rows_to_do<-5000
 geocodeQueryCheck()
 
 # get the input data for police stops in Berkeley:

### 2016-07-22 Berkeley_PD_Stop.Rmd
clean Disposition

--- 2016-07-22 Berkeley_PD_Stop.Rmd
@@
 
+#read data
 stop_df <- read.csv("~/Desktop/Berkeley_PD_-_Stop_Data.csv")
 
-#disps <- as.character(stop_df$Dispositions)
-#max(nchar(disps))
-#disps[which(nchar(disps) == 42)]
+# Delete meaningless information and fix typos
+stop_df <- stop_df %>%
+  mutate(Dispositions = as.character(stop_df$Dispositions)) %>%
+  mutate(Dispositions = str_replace(Dispositions, "zz do not use Assist", "")) %>%
+  mutate(Dispositions = str_replace(Dispositions, "FA - TIGHT", "")) %>%
+  mutate(Dispositions = str_replace(Dispositions, "[0]{6}[1]?", "00000"))
 
-stop_df <- mutate(stop_df, Dispositions = as.character(stop_df$Dispositions))
+View(stop_df)
+# Split people information and other information
+stop_df <- stop_df %>%
+  mutate(Other = str_replace_all(Dispositions, "[ ,]{0,2}[a-zA-Z0-9]{6}", "")) %>%
+  mutate(Other = str_replace_all(Other, "^[ ,]{0,4}", "")) %>%
+  mutate(Other = str_replace(Other, ",$", ""))
 
-#entries with "zz do not use Assist" 
-stop_df <- mutate(stop_df, Dispositions = str_replace_all(stop_df$Dispositions, "zz do not use Assist", ""))
-#entries with "FA - TIGHT"
-stop_df <- mutate(stop_df, Dispositions = str_replace_all(stop_df$Dispositions, "FA - TIGHT", ""))
-#fixing entries that were input incorrectly
-stop_df <- mutate(stop_df, Dispositions = str_replace_all(stop_df$Dispositions, "[0]{6}[1]?", "00000"))
-
-#deal with special cases first 
-stop_df <- mutate(stop_df, Other = str_replace_all(stop_df$Dispositions, "[ ,]{0,2}[a-zA-Z0-9]{6}", ""))
-stop_df <- mutate(stop_df, Other = str_replace_all(stop_df$Other, "^[ ,]{0,4}", ""))
-
+OtherInformation <- c("P", "^M$", ", M$", "^M, P$")
+stop_df <- stop_df %>%
+  mutate(Dispositions = str_replace_all(Dispositions, ))
 stop_df <- mutate(stop_df, Dispositions = str_extract_all(stop_df$Dispositions, "[a-zA-Z0-9]{6}"))

### 2016-07-22 Berkeley_PD_Stop.Rmd

@@
 
-View(stop_df)
 # Split people information and other information
@@
 
-OtherInformation <- c("P", "^M$", ", M$", "^M, P$")
+OtherInformation <- c( "^M$" = "", ", M$" = "", "M, P$" = "", "P[, ]{0,2}" = "", "AR[, ]{0,2}" = "", "IN[, ]{0,2}" = "", "FC[, ]{0,2}" = "", "CO[, ]{0,2}" = "", "MH[, ]{0,2}" = "", "TOW[, ]{0,2}" = "", "0[, ]{0,2}" = "", "00000[, ]{0,2}" = "")
 stop_df <- stop_df %>%
-  mutate(Dispositions = str_replace_all(Dispositions, ))
-stop_df <- mutate(stop_df, Dispositions = str_extract_all(stop_df$Dispositions, "[a-zA-Z0-9]{6}"))
+  mutate(Dispositions = str_replace_all(Dispositions, OtherInformation))
 
 View(stop_df)
+
+#mutate dispositions column to only contain six character cases
+
+stop_df <- mutate(stop_df, Dispositions = str_extract_all(stop_df$Dispositions, "[a-zA-Z0-9]{6}"))
+#removes empty cases
+stop_df <- mutate(stop_df, Dispositions = str_replace_all(stop_df$Dispositions, "character(0)", ""))
+
+dispositions <- separate(stop_df, Dispositions, into = c("Individual_1", "Individual_2", "Individual_3", "Individual_4", "Individual_5", "Individual_6"), extra = "merge", fill = "right")
+
+View(stop_df)
+View(dispositions)
+###
 

### 2016-07-22 Berkeley_PD_Stop.Rmd
clean disposition

@@
 stop_df <- stop_df %>%
-  mutate(Dispositions = str_replace_all(Dispositions, OtherInformation))
+  mutate(Dispositions = str_replace_all(Dispositions, OtherInformation)) %>%
+  mutate(Dispositions = str_replace_all(Dispositions, "[, ]{0,2}$", ""))
 

### 2016-07-22 Berkeley_PD_Stop.Rmd
clean disposition

--- 2016-07-22 Berkeley_PD_Stop.Rmd
@@
   mutate(Dispositions = str_replace(Dispositions, "zz do not use Assist", "")) %>%
-  mutate(Dispositions = str_replace(Dispositions, "FA - TIGHT", "")) %>%
+  mutate(Dispositions = str_replace(Dispositions, "FA - TIGHT, ", "")) %>%
   mutate(Dispositions = str_replace(Dispositions, "[0]{6}[1]?", "00000"))
@@
 
+# Delete Other information from Disposition
 OtherInformation <- c( "^M$" = "", ", M$" = "", "M, P$" = "", "P[, ]{0,2}" = "", "AR[, ]{0,2}" = "", "IN[, ]{0,2}" = "", "FC[, ]{0,2}" = "", "CO[, ]{0,2}" = "", "MH[, ]{0,2}" = "", "TOW[, ]{0,2}" = "", "0[, ]{0,2}" = "", "00000[, ]{0,2}" = "")
 stop_df <- stop_df %>%
-  mutate(Dispositions = str_replace_all(Dispositions, OtherInformation))
+  mutate(Dispositions = str_replace_all(Dispositions, OtherInformation)) %>%
+  mutate(Dispositions = str_replace_all(Dispositions, "[, ]{0,2}$", ""))
 
@@
 #mutate dispositions column to only contain six character cases
-
-stop_df <- mutate(stop_df, Dispositions = str_extract_all(stop_df$Dispositions, "[a-zA-Z0-9]{6}"))
-#removes empty cases
-stop_df <- mutate(stop_df, Dispositions = str_replace_all(stop_df$Dispositions, "character(0)", ""))
 

### 2016-07-22 Berkeley_PD_Stop.Rmd
072216

@@
 
-#read data
 stop_df <- read.csv("~/Desktop/Berkeley_PD_-_Stop_Data.csv")
 
-# Delete meaningless information and fix typos
-stop_df <- stop_df %>%
-  mutate(Dispositions = as.character(stop_df$Dispositions)) %>%
-  mutate(Dispositions = str_replace(Dispositions, "zz do not use Assist", "")) %>%
-  mutate(Dispositions = str_replace(Dispositions, "FA - TIGHT, ", "")) %>%
-  mutate(Dispositions = str_replace(Dispositions, "[0]{6}[1]?", "00000"))
+#disps <- as.character(stop_df$Dispositions)
+#max(nchar(disps))
+#disps[which(nchar(disps) == 42)]
 
-# Split people information and other information
-stop_df <- stop_df %>%
-  mutate(Other = str_replace_all(Dispositions, "[ ,]{0,2}[a-zA-Z0-9]{6}", "")) %>%
-  mutate(Other = str_replace_all(Other, "^[ ,]{0,4}", "")) %>%
-  mutate(Other = str_replace(Other, ",$", ""))
+stop_df <- mutate(stop_df, Dispositions = as.character(stop_df$Dispositions))
 
-# Delete Other information from Disposition
-OtherInformation <- c( "^M$" = "", ", M$" = "", "M, P$" = "", "P[, ]{0,2}" = "", "AR[, ]{0,2}" = "", "IN[, ]{0,2}" = "", "FC[, ]{0,2}" = "", "CO[, ]{0,2}" = "", "MH[, ]{0,2}" = "", "TOW[, ]{0,2}" = "", "0[, ]{0,2}" = "", "00000[, ]{0,2}" = "")
-stop_df <- stop_df %>%
-  mutate(Dispositions = str_replace_all(Dispositions, OtherInformation)) %>%
-  mutate(Dispositions = str_replace_all(Dispositions, "[, ]{0,2}$", ""))
+#entries with "zz do not use Assist" 
+stop_df <- mutate(stop_df, Dispositions = str_replace_all(stop_df$Dispositions, "zz do not use Assist", ""))
+#entries with "FA - TIGHT"
+stop_df <- mutate(stop_df, Dispositions = str_replace_all(stop_df$Dispositions, "FA - TIGHT", ""))
+#fixing entries that were input incorrectly
+stop_df <- mutate(stop_df, Dispositions = str_replace_all(stop_df$Dispositions, "[0]{6}[1]?", "00000"))
 
-View(stop_df)
+#deal with special cases first 
+stop_df <- mutate(stop_df, Other = str_replace_all(stop_df$Dispositions, "[ ,]{0,2}[a-zA-Z0-9]{6}", ""))
+stop_df <- mutate(stop_df, Other = str_replace_all(stop_df$Other, "^[ ,]{0,4}", ""))
 
 #mutate dispositions column to only contain six character cases
+stop_df <- mutate(stop_df, Dispositions = str_match_all(stop_df$Dispositions, "[a-zA-Z0-9]{6}"))
 
-dispositions <- separate(stop_df, Dispositions, into = c("Individual_1", "Individual_2", "Individual_3", "Individual_4", "Individual_5", "Individual_6"), extra = "merge", fill = "right")
+dispositions <- separate(stop_df, Dispositions, into = c("Empty", "Individual_1", "Individual_2", "Individual_3", "Individual_4", "Individual_5"), extra = "merge", fill = "right")
 
@@
 View(dispositions)
-###
+
+###IGNORE EVERYTHING AFTER THIS
 

### 2016-07-22 Berkeley_PD_Stop.Rmd
update

--- 2016-07-22 Berkeley_PD_Stop.Rmd
@@
 
-# Delete Other information from Disposition
+# Remove Other information from Disposition
 OtherInformation <- c( "^M$" = "", ", M$" = "", "M, P$" = "", "P[, ]{0,2}" = "", "AR[, ]{0,2}" = "", "IN[, ]{0,2}" = "", "FC[, ]{0,2}" = "", "CO[, ]{0,2}" = "", "MH[, ]{0,2}" = "", "TOW[, ]{0,2}" = "", "0[, ]{0,2}" = "", "00000[, ]{0,2}" = "")
@@
   mutate(Dispositions = str_replace_all(Dispositions, "[, ]{0,2}$", ""))
+
+# Creating new entries for multiple people in one disposition
 

### 2016-07-22 Berkeley_PD_Stop.Rmd
072216

--- 2016-07-22 Berkeley_PD_Stop.Rmd
@@
 library(ggplot2)
+library(reshape)
 
@@
 
-# Split people information and other information
+# Split information about the individual from other information
 stop_df <- stop_df %>%
@@
 
-# Delete Other information from Disposition
+# Delete other information from Disposition
 OtherInformation <- c( "^M$" = "", ", M$" = "", "M, P$" = "", "P[, ]{0,2}" = "", "AR[, ]{0,2}" = "", "IN[, ]{0,2}" = "", "FC[, ]{0,2}" = "", "CO[, ]{0,2}" = "", "MH[, ]{0,2}" = "", "TOW[, ]{0,2}" = "", "0[, ]{0,2}" = "", "00000[, ]{0,2}" = "")
@@
 
+stop_df <- separate(stop_df, Dispositions, into = c("Individual_1", "Individual_2", "Individual_3", "Individual_4", "Individual_5"), extra = "merge", fill = "right")
+
+stop_df <- melt(stop_df, id = c("Incident.Number", "Call.Date.Time", "Location", "Incident.Type", "Other")) %>%
+  arrange(Incident.Number)
+
 View(stop_df)
 
-#mutate dispositions column to only contain six character cases
+=======
 
-dispositions <- separate(stop_df, Dispositions, into = c("Individual_1", "Individual_2", "Individual_3", "Individual_4", "Individual_5", "Individual_6"), extra = "merge", fill = "right")
-
-View(stop_df)
-View(dispositions)
-###
+###IGNORE EVERYTHING AFTER THIS
 

### 2016-07-22 Berkeley_PD_Stop.Rmd
update

--- 2016-07-22 Berkeley_PD_Stop.Rmd
@@
 #mutate dispositions column to only contain six character cases
+stop_df <- mutate(stop_df, Dispositions = str_match_all(stop_df$Dispositions, "[a-zA-Z0-9]{6}"))
 
-dispositions <- separate(stop_df, Dispositions, into = c("Individual_1", "Individual_2", "Individual_3", "Individual_4", "Individual_5", "Individual_6"), extra = "merge", fill = "right")
+dispositions <- separate(stop_df, Dispositions, into = c("Empty", "Individual_1", "Individual_2", "Individual_3", "Individual_4", "Individual_5"), extra = "merge", fill = "right")
 

### 2016-07-22 Berkeley_PD_Stop.Rmd

--- 2016-07-22 Berkeley_PD_Stop.Rmd
@@
 
-stop_df <- separate(stop_df, Dispositions, into = c("Individual_1", "Individual_2", "Individual_3", "Individual_4", "Individual_5"), extra = "merge", fill = "right")
-
-stop_df <- melt(stop_df, id = c("Incident.Number", "Call.Date.Time", "Location", "Incident.Type", "Other")) %>%
+# Creating new entries for multiple people in one disposition
+stop_df <- stop_df %>%
+  separate(Dispositions, into = c("1", "2", "3", "4", "5"), extra = "merge", fill = 
+             "right") %>%
+  gather(key = Individual, value = Dispositions, -Incident.Type, -Incident.Number, -Call.Date.Time, -Location, -Other) %>%
+  na.omit() %>%
   arrange(Incident.Number)
@@
 View(stop_df)
-
-=======
 

### 2016-07-22 Berkeley_PD_Stop.Rmd
072216

--- 2016-07-22 Berkeley_PD_Stop.Rmd
@@
 
-# Split information about the individual from other information
+# Split other information from information about the individual(s)
 stop_df <- stop_df %>%
@@
 stop_df <- melt(stop_df, id = c("Incident.Number", "Call.Date.Time", "Location", "Incident.Type", "Other")) %>%
+  na.omit() %>%
   arrange(Incident.Number)
@@
 
-dispositions <- separate(stop_df, Dispositions, into = c("Individual_1", "Individual_2", "Individual_3", "Individual_4", "Individual_5", "Individual_6"), extra = "merge", fill = "right")
+dispositions <- separate(stop_df, Dispositions, into = c("Individual_1", "Individual_2", "Individual_3", "Individual_4", "Individual_5"), extra = "merge", fill = "right")
 

### 2016-07-22 Berkeley_PD_Stop.Rmd
Separate multiple people
Separate information

--- 2016-07-22 Berkeley_PD_Stop.Rmd
@@
 stop_df <- stop_df %>%
-  separate(Dispositions, into = c("1", "2", "3", "4", "5"), extra = "merge", fill = 
-             "right") %>%
+  separate(Dispositions, into = c("1", "2", "3", "4", "5"), extra = "merge", fill = "right") %>%
   gather(key = Individual, value = Dispositions, -Incident.Type, -Incident.Number, -Call.Date.Time, -Location, -Other) %>%
@@
   arrange(Incident.Number)
+
+# Separate information about people
+stop_df<- stop_df %>%
+    filter(nchar(Dispositions) == 6) %>%
+      mutate(Race = substr(Dispositions,1,1)) %>%
+        mutate(Gender = substr (Dispositions,2,2)) %>%
+          mutate(AgeRange = substr (Dispositions,3,3)) %>%
+            mutate(Reason =substr(Dispositions,4,4)) %>%
+              mutate(Enforcement = substr(Dispositions,5,5)) %>%
+                 mutate(CarSearch = substr(Dispositions,6,6))
 

### 2016-07-22 Berkeley_PD_Stop.Rmd
Separate information

@@
 
-###IGNORE EVERYTHING AFTER THIS
-
-dispositions <- separate(stop_df, Dispositions, into = c("Individual_1", "Individual_2", "Individual_3", "Individual_4", "Individual_5", "Individual_6"), extra = "merge", fill = "right")
-
-#deconstructing_6_characters <- function (x) {
- # dispositions %>%
-  #  filter(nchar(x) == 6) %>%
-   #   mutate(Race_x = substr(x,1,1)) %>%
-    #    mutate(Gender_x = substr (x,2,2)) %>%
-     #     mutate(AgeRange_x = substr (x,3,3)) %>%
-      #      mutate(Reason_x =substr(x,4,4)) %>%
-       #       mutate(Enforcement_x = substr(x,5,5)) %>%
-        #         mutate(CarSearch_x = substr(x,6,6))
-#}
-#lapply
-
-dispositions_6_characters_1 <- dispositions %>%
-    filter(nchar(Individual_1) == 6) %>%
-      mutate(Race_1 = substr(Individual_1,1,1)) %>%
-        mutate(Gender_1 = substr (Individual_1,2,2)) %>%
-          mutate(AgeRange_1 = substr (Individual_1,3,3)) %>%
-            mutate(Reason_1 =substr(Individual_1,4,4)) %>%
-              mutate(Enforcement_1 = substr(Individual_1,5,5)) %>%
-                 mutate(CarSearch_1 = substr(Individual_1,6,6))
-
-dispositions_6_characters_2 <- dispositions %>%
-    filter(nchar(Individual_2) == 6) %>%
-      mutate(Race_2 = substr(Individual_2,1,1)) %>%
-        mutate(Gender_2 = substr (Individual_2,2,2)) %>%
-          mutate(AgeRange_2 = substr (Individual_2,3,3)) %>%
-            mutate(Reason_2 =substr(Individual_2,4,4)) %>%
-              mutate(Enforcement_2 = substr(Individual_2,5,5)) %>%
-                 mutate(CarSearch_2 = substr(Individual_2,6,6))
-
-dispositions_6_characters_3 <- dispositions %>%
-    filter(nchar(Individual_3) == 6) %>%
-      mutate(Race_3 = substr(Individual_3,1,1)) %>%
-        mutate(Gender_3 = substr (Individual_3,2,2)) %>%
-          mutate(AgeRange_3 = substr (Individual_3,3,3)) %>%
-            mutate(Reason_3 =substr(Individual_3,4,4)) %>%
-              mutate(Enforcement_3 = substr(Individual_3,5,5)) %>%
-                 mutate(CarSearch_3 = substr(Individual_3,6,6))
-
-dispositions_6_characters_4 <- dispositions %>%
-    filter(nchar(Individual_4) == 6) %>%
-      mutate(Race_4 = substr(Individual_4,1,1)) %>%
-        mutate(Gender_4 = substr (Individual_4,2,2)) %>%
-          mutate(AgeRange_4 = substr (Individual_4,3,3)) %>%
-            mutate(Reason_4 =substr(Individual_4,4,4)) %>%
-              mutate(Enforcement_4 = substr(Individual_4,5,5)) %>%
-                 mutate(CarSearch_4 = substr(Individual_4,6,6))
-
-dispositions_6_characters_5 <- dispositions %>%
-    filter(nchar(Individual_5) == 6) %>%
-      mutate(Race_5 = substr(Individual_5,1,1)) %>%
-        mutate(Gender_5 = substr (Individual_5,2,2)) %>%
-          mutate(AgeRange_5 = substr (Individual_5,3,3)) %>%
-            mutate(Reason_5 =substr(Individual_5,4,4)) %>%
-              mutate(Enforcement_5 = substr(Individual_5,5,5)) %>%
-                 mutate(CarSearch_5 = substr(Individual_5,6,6))
-
-dispositions_6_characters_6 <- dispositions %>%
-    filter(nchar(Individual_6) == 6) %>%
-      mutate(Race_6 = substr(Individual_6,1,1)) %>%
-        mutate(Gender_6 = substr (Individual_6,2,2)) %>%
-          mutate(AgeRange_6 = substr (Individual_6,3,3)) %>%
-            mutate(Reason_6 =substr(Individual_6,4,4)) %>%
-              mutate(Enforcement_6 = substr(Individual_6,5,5)) %>%
-                 mutate(CarSearch_6 = substr(Individual_6,6,6))
-
-#View(dispositions_1)
-#View(dispositions_2)
-#View(dispositions_3)
-#View(dispositions_4)
-#View(dispositions_5)
-#View(dispositions_6)
-
-#deconstructing_other_characters <- function(x) {
-#}
-#lapply
-
-dispositions_other_characters_1 <- dispositions  %>%
-  filter(nchar(Individual_1) == 1 | nchar(Individual_1) == 2 | nchar(Individual_1) == 3 | nchar(Individual_1) == 5) %>%
-    mutate(Other_1 = Individual_1)
-
-dispositions_other_characters_2 <- dispositions  %>%
-  filter(nchar(Individual_2) == 1 | nchar(Individual_2) == 2 | nchar(Individual_2) == 3 | nchar(Individual_2) == 5) %>%
-    mutate(Other_2 = Individual_2)
-
-dispositions_other_characters_3 <- dispositions  %>%
-  filter(nchar(Individual_3) == 1 | nchar(Individual_3) == 2 | nchar(Individual_3) == 3 | nchar(Individual_3) == 5) %>%
-    mutate(Other_3 = Individual_3)
-
-dispositions_other_characters_4 <- dispositions  %>%
-  filter(nchar(Individual_4) == 1 | nchar(Individual_4) == 2 | nchar(Individual_4) == 3 | nchar(Individual_4) == 5) %>%
-    mutate(Other_4 = Individual_4)
-
-dispositions_other_characters_5 <- dispositions  %>%
-  filter(nchar(Individual_5) == 1 | nchar(Individual_5) == 2 | nchar(Individual_5) == 3 | nchar(Individual_5) == 5) %>%
-    mutate(Other_5 = Individual_5)
-
-dispositions_other_characters_6 <- dispositions  %>%
-  filter(nchar(Individual_6) == 1 | nchar(Individual_6) == 2 | nchar(Individual_6) == 3 | nchar(Individual_6) == 5) %>%
-    mutate(Other_6 = Individual_6)
-
-#View(dispositions_other_characters_1)
-#View(dispositions_other_characters_2)
-#View(dispositions_other_characters_3)
-#View(dispositions_other_characters_4)
-#View(dispositions_other_characters_5)
-#View(dispositions_other_characters_6)
-
-dispositions <- left_join(dispositions, dispositions_1)
-dispositions <- left_join(dispositions, dispositions_other_characters_1)
-dispositions <- left_join(dispositions, dispositions_2)
-dispositions <- left_join(dispositions, dispositions_other_characters_2)
-dispositions <- left_join(dispositions, dispositions_3)
-dispositions <- left_join(dispositions, dispositions_other_characters_3)
-dispositions <- left_join(dispositions, dispositions_4)
-dispositions <- left_join(dispositions, dispositions_other_characters_4)
-dispositions <- left_join(dispositions, dispositions_5)
-dispositions <- left_join(dispositions, dispositions_other_characters_5)
-dispositions <- left_join(dispositions, dispositions_6)
-dispositions <- left_join(dispositions, dispositions_other_characters_6)
-
-dispositions <- dispositions[,c(1:4, 11:17, 6, 18:24, 7, 25:31, 8, 32:38, 9, 39:45, 10, 46:52)]
-View(dispositions)
-
 ```

### 2016-07-22 Berkeley_PD_Stop.Rmd
completed

@@
   separate(Dispositions, into = c("1", "2", "3", "4", "5"), extra = "merge", fill = "right") %>%
-  gather(key = Individual, value = Dispositions, -Incident.Type, -Incident.Number, -Call.Date.Time, -Location, -Other) %>%
+  gather(key = Individual, value = Dispositions, -Incident.Number, -Incident.Type, -Call.Date.Time, -Other, -Location) %>%
   na.omit() %>%
@@
 View(stop_df)
-
-```

### 2016-07-22 Berkeley_PD_Stop.Rmd
update

@@
 ---
-title: "Test"
-author: "Amy Zhu"
+title: "Cleaned Stop Data"
+author: "Amy Zhu, Mengyu Li"
 date: "July 19, 2016"
@@
 library(stringr)
-library(lubridate)
-library(ggplot2)
-library(reshape)
 
-#read data
+#Read data
 stop_df <- read.csv("~/Desktop/Berkeley_PD_-_Stop_Data.csv")
 
-# Delete meaningless information and fix typos
+# Delete meaningless information and fix data entry errors
 stop_df <- stop_df %>%
@@
 
-# Split other information from information about the individual(s)
+# Isolates information about the case into the Other column
 stop_df <- stop_df %>%
@@
 
-# Delete other information from Disposition
+# Isolates information about the individual(s) into the Dispositions column
 OtherInformation <- c( "^M$" = "", ", M$" = "", "M, P$" = "", "P[, ]{0,2}" = "", "AR[, ]{0,2}" = "", "IN[, ]{0,2}" = "", "FC[, ]{0,2}" = "", "CO[, ]{0,2}" = "", "MH[, ]{0,2}" = "", "TOW[, ]{0,2}" = "", "0[, ]{0,2}" = "", "00000[, ]{0,2}" = "")
@@
 
-# Creating new entries for multiple people in one disposition
+# Arranges information in the Dispositions column into separate objects for each individual involved 
 stop_df <- stop_df %>%
@@
 
-# Separate information about people
-stop_df<- stop_df %>%
+# Separates information about the individual into 6 columns describing race, gender, age, reason, enforcement, and car search
+stop_df <- stop_df %>%
     filter(nchar(Dispositions) == 6) %>%
@@
           mutate(AgeRange = substr (Dispositions,3,3)) %>%
-            mutate(Reason =substr(Dispositions,4,4)) %>%
+            mutate(Reason = substr(Dispositions,4,4)) %>%
               mutate(Enforcement = substr(Dispositions,5,5)) %>%

### 2016-07-22 Berkeley_PD_Stop.Rmd
complete cleaning

--- 2016-07-22 Berkeley_PD_Stop.Rmd
@@
 library(ggplot2)
-library(reshape)
 

### 2016-07-23 AnalysizeStop.Rmd
Analyze stop data

---
title: "Analyze Stop Data"
author: "Mengyu Li"
date: "July 22, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(lubridate)

# Read data
stop_by <- read_csv("cleaned_stop_data.csv")

stop_by$Call.Date.Time <- mdy_hm(stop_by$Call.Date.Time)
stop_by$AgeRange <- as.factor(stop_by$AgeRange)

stop_by <- stop_by %>%
  mutate(hour = hour(Call.Date.Time))

stop_race <- ggplot(stop_by, aes(x = Race, fill = Gender)) + geom_bar()
stop_age <- ggplot(stop_by, aes(x = AgeRange, fill = Gender)) + geom_bar()
stop_reason <- ggplot(stop_by, aes(x = Reason, fill = Gender)) + geom_bar()
stop_time <- ggplot(stop_by, aes(x = hour, fill = AgeRange)) + geom_bar()

View(stop_by)
```

### 2016-07-29 AnalysizeStop.Rmd
graph updates

@@
 
-stop_race <- ggplot(stop_by, aes(x = Race, fill = Gender)) + geom_bar()
+#Sample graph
+stop_race <- ggplot(stop_by, aes(x = Race, fill = Gender)) + 
+  geom_bar(aes(y = ..count../sum(..count..))) + 
+  labs(y = "Density") +
+  facet_wrap(~Reason)
+
 stop_age <- ggplot(stop_by, aes(x = AgeRange, fill = Gender)) + geom_bar() 
@@
 
+stop_race
+
 View(stop_by)

### 2016-07-29 Rebeccas_location_Code2/StopDataGetLocation.R
new location code

@@ -1,0 +1,312 @@
+###################################################################################################################################
+## Title: STOP DATA LOCATION SCRIPT FOR BERKELEY, CA
+## Author: Rebecca Reus, STAT 133 Summer 2016, UC Berkeley
+## Information:
+## Uses part of Geocoding script for large list of addresses created by Shane Lynn, 10/10/2013 at http://www.r-bloggers.com/batch-geocoding-with-r-and-google-maps/
+## Police Stop Data source: https://data.cityofberkeley.info/
+###################################################################################################################################
+
+
+###################################################################################################################################
+## LIBRARIES:
+## Before running, please make sure you have installed ALL of these packages:
+###################################################################################################################################
+library(sp)
+library(ggmap)
+library(tidyr)
+library(plyr)
+library(dplyr)
+library(stringr)
+library(lubridate)
+library(ggplot2)
+library(readr)
+library(data.table)
+library(rworldmap)
+library(maps)
+library(mapdata)
+library(maptools)
+library(scales)
+library(RgoogleMaps)
+library(tmap)
+library(rgdal)
+library(rgeos)
+
+###################################################################################################################################
+## DATA TIDYING:
+###################################################################################################################################
+
+# Get the data from the file in your working directory:
+infile <- "StopData"
+data <- read.csv(paste0('./', infile, '.csv'))
+
+# Filter out the data that is not already in coordinate form (doesn't have longitude and latitude values):
+stop_df <- data %>%
+  subset( !str_detect( Location, "~-"))
+
+# Since we removed some rows, we need to rename the row names to match the row number (used for indexing):
+rownames(stop_df) <- 1:nrow(stop_df)
+
+# We need to adjust the Location data to account for popular abbreviations and mispellings, in addition to some other data tidying:
+stop_df <- stop_df %>%
+  mutate( Location = as.character( Location ) ) %>%                  # location = character information
+  mutate( Location = paste0(Location, " , Berkeley, CA, USA") ) %>%
[…]

### 2016-07-29 AnalysizeStop.Rmd
07292016

@@
   geom_bar(aes(y = ..count../sum(..count..))) + 
+  scale_fill_manual(values = c("skyblue1", "pink")) +
   labs(y = "Density") +
   facet_wrap(~Reason)
+
+#Gender, race, and enforcement
 

### 2016-07-29 Rebeccas_location_Code2/StopDataGetLocation.R

@@ -155,7 +155,7 @@ stop_df <- stop_df %>%
 ####################################################################################################################################
 
 # Choose a row number to finish at (will start at the last spot using the temp file):
-number_of_rows_to_do <- 4300
+number_of_rows_to_do <- 9263
 
 # Select the addresses from the Location column of stop_df to run through Google:
 #address<-head(stop_df$Location,number_of_rows_to_do)

### 2016-07-29 AnalysizeStop.Rmd
072916

--- 2016-07-29 AnalysizeStop.Rmd
@@
 
-#Sample graph
+#Race, Gender, and Reason
 stop_race <- ggplot(stop_by, aes(x = Race, fill = Gender)) + 
   geom_bar(aes(y = ..count../sum(..count..))) + 
+  scale_fill_manual(values = c("skyblue1", "pink")) +
   labs(y = "Density") +
   facet_wrap(~Reason)
+#Remove input "E" in the R column
+
+#Reason, Gender, and Enforcement
+stop_reason <- ggplot(stop_by, aes(x = Reason, fill = Gender)) + 
+  geom_bar(aes(y = ..count../sum(..count..))) + 
+  scale_fill_manual(values = c("skyblue1", "pink")) +
+  labs(y = "Density") +
+  facet_wrap(~Enforcement)
 
@@
 stop_race
+stop_reason
 

### 2016-07-29 AnalysizeStop.Rmd
072916

--- 2016-07-29 AnalysizeStop.Rmd
@@
 
-#Sample graph
-stop_race <- ggplot(stop_by, aes(x = Race, fill = Gender)) + 
+#Remove input "E" from the R column
+
+stop_race <- ggplot(stop_by, aes(x = Reason, fill = Enforcement)) + 
   geom_bar(aes(y = ..count../sum(..count..))) + 
   labs(y = "Density") +
-  facet_wrap(~Reason)
+  facet_wrap(~Race)
 
-stop_age <- ggplot(stop_by, aes(x = AgeRange, fill = Gender)) + geom_bar() 
-stop_reason <- ggplot(stop_by, aes(x = Reason, fill = Gender)) + geom_bar()
-stop_time <- ggplot(stop_by, aes(x = hour, fill = AgeRange)) + geom_bar()
+stop_race <- ggplot(stop_by, aes(x = Reason, fill = Enforcement)) + 
+  geom_bar(aes(y = ..count../sum(..count..))) + 
+  labs(y = "Density") +
+  scale_fill_brewer(palette = "Set1") +
+  facet_grid(Race~Gender)
 

### 2016-07-29 Rebeccas_location_Code2/StopDataGetLocation.R

@@ -155,7 +155,7 @@ stop_df <- stop_df %>%
 ####################################################################################################################################
 
 # Choose a row number to finish at (will start at the last spot using the temp file):
-number_of_rows_to_do <- 9263
+number_of_rows_to_do <- 10500
 
 # Select the addresses from the Location column of stop_df to run through Google:
 #address<-head(stop_df$Location,number_of_rows_to_do)

### 2016-07-30 Rebeccas_location_Code2/StopDataGetLocation.R
updated to 12960

@@ -7,6 +7,9 @@
 ###################################################################################################################################
 
 
+# Jenny WD:
+# setwd("~/Desktop/Stats R/PROJECT/finalproject_angry_ladies/Rebeccas_location_Code2")
+
 ###################################################################################################################################
 ## LIBRARIES:
 ## Before running, please make sure you have installed ALL of these packages:
@@ -28,6 +31,7 @@ library(maptools)
 library(scales)
 library(RgoogleMaps)
 library(tmap)
+library(sp)
 library(rgdal)
 library(rgeos)
 
@@ -155,7 +159,7 @@ stop_df <- stop_df %>%
 ####################################################################################################################################
 
 # Choose a row number to finish at (will start at the last spot using the temp file):
-number_of_rows_to_do <- 10500
+number_of_rows_to_do <- 12960
 
 # Select the addresses from the Location column of stop_df to run through Google:
 #address<-head(stop_df$Location,number_of_rows_to_do)

### 2016-07-30 Rebeccas_location_Code2/StopDataGetLocation.R
completed 15514

@@ -159,7 +159,7 @@ stop_df <- stop_df %>%
 ####################################################################################################################################
 
 # Choose a row number to finish at (will start at the last spot using the temp file):
-number_of_rows_to_do <- 12960
+number_of_rows_to_do <- 15514
 
 # Select the addresses from the Location column of stop_df to run through Google:
 #address<-head(stop_df$Location,number_of_rows_to_do)

### 2016-07-30 AnalysizeStop.Rmd
before pull

@@
 
-#Remove input "E" from the R column
-
-stop_race <- ggplot(stop_by, aes(x = Reason, fill = Enforcement)) + 
-  geom_bar(aes(y = ..count../sum(..count..))) + 
-  labs(y = "Density") +
-  facet_wrap(~Race)
-
-stop_race <- ggplot(stop_by, aes(x = Reason, fill = Enforcement)) + 
-  geom_bar(aes(y = ..count../sum(..count..))) + 
-  labs(y = "Density") +
-  scale_fill_brewer(palette = "Set1") +
-  facet_grid(Race~Gender)
-
-stop_race
+stop_race <- ggplot(stop_by, aes(x = Race, fill = AgeRange)) + geom_bar()
+stop_age <- ggplot(stop_by, aes(x = AgeRange, fill = Gender)) + geom_bar()
+stop_reason <- ggplot(stop_by, aes(x = Reason, fill = Gender)) + geom_bar()
+stop_time <- ggplot(stop_by, aes(x = hour, fill = AgeRange)) + geom_bar()
 

### 2016-07-31 AnalysizeStop.Rmd

@@
 library(lubridate)
+library(stringr)
+library(ggvis)
 
+```
+
+```{r}
 # Read data
@@
 stop_by <- stop_by %>%
-  mutate(hour = hour(Call.Date.Time))
+  mutate(Hour = as.integer(hour(Call.Date.Time))) %>%
+  mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))
 
-stop_race <- ggplot(stop_by, aes(x = Race, fill = AgeRange)) + geom_bar()
-stop_age <- ggplot(stop_by, aes(x = AgeRange, fill = Gender)) + geom_bar()
-stop_reason <- ggplot(stop_by, aes(x = Reason, fill = Gender)) + geom_bar()
-stop_time <- ggplot(stop_by, aes(x = hour, fill = AgeRange)) + geom_bar()
+```
 
-View(stop_by)
+```{r}
+#Remove input "E" from the R column
+stop_race <- stop_by %>%
+  ggvis(x = ~Race , fill = ~AgeRange) %>%
+  layer_bars() %>%
+  add_tooltip(function(stop_by) stop_by$AgeRange) 
+stop_race
+
 ```
+
+```{r}
+stop_day <- stop_by %>%
+  ggvis(x = ~Day, fill = ~Race) %>%
+  layer_bars() %>%
+  add_tooltip(function(stop_by) stop_by$Race)
+stop_day
+```
+
+```{r}
+stop_time <- stop_by %>%
+  ggvis(x = ~Hour, fill := "lightpink") %>%
+  layer_histograms(width = input_slider(1, 12, value = 1, step = 2, label = "width"))
+stop_time
+#width = input_slider(1, 12, step = 2, label = "width")
+```
+
+```{r}
+
+stop_time_percentage <- ggplot(stop_by, aes(x = Hour, fill = AgeRange)) + 
+  geom_bar(aes(y = ..count../sum(..count..))) +
+  labs(y = "Density") +
+  facet_wrap(~AgeRange)
+stop_time_percentage
+
+```
+
+```{r}
+stop_by_hourCount <- stop_by %>%
+  group_by(Hour) %>%
+  summarise(hourcount = n())
+stop_by_ageRangeCount <- stop_by %>%
+  group_by(Hour, AgeRange) %>%
+  summarise(ageCount = n())
+stop_by_percentage <- left_join(stop_by_ageRangeCount, stop_by_hourCount)
+stop_by_percentage <- stop_by_percentage %>%
+  mutate(percentage = ageCount/hourcount)
+
+stop_time_percentage2 <- ggplot(stop_by_percentage, aes(x = Hour, fill = AgeRange)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  labs(y = "Density") +
+  facet_wrap(~AgeRange)
+stop_time_percentage2
+```
+
+```{r}
+stop_by_hourCount <- stop_by %>%
+  group_by(Hour) %>%
+  summarise(hourcount = n())
+stop_by_raceRangeCount <- stop_by %>%
+  group_by(Hour, Race) %>%
+  summarise(raceCount = n())
+stop_by_percentage1 <- left_join(stop_by_raceRangeCount, stop_by_hourCount)
+stop_by_percentage1 <- stop_by_percentage1 %>%
+  mutate(percentage = raceCount/hourcount)
+
+stop_time_percentage3 <- ggplot(stop_by_percentage1, aes(x = Hour, fill = Race)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  labs(y = "Density") +
+  facet_wrap(~Race)
+stop_time_percentage3
+```
+
+```{r}
+stop_by_reason_race <- stop_by %>%
+  group_by(Race, Reason) %>%
+  summarise(totalCount = n())
+stop_by_arrest <- stop_by %>%
+  filter(Enforcement == "A") %>%
+  group_by(Race, Reason) %>%
+  summarise(arrestCount = n())
+stop_by_percentage_race <- left_join(stop_by_reason_race, stop_by_arrest)
+stop_by_percentage_race <- stop_by_percentage_race %>%
+  mutate(percentage = arrestCount / totalCount)
+
+stop_arrest <- ggplot(stop_by_percentage_race, aes(x = Race, fill = Race)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  labs(y = "Density") +
+  facet_wrap(~Reason)
+stop_arrest
+```
+
+```{r}
+stop_by_reason_age <- stop_by %>%
+  group_by(AgeRange, Reason) %>%
+  summarise(totalCount = n())
+stop_by_arrest_age <- stop_by %>%
+  filter(Enforcement == "A") %>%
+  group_by(AgeRange, Reason) %>%
+  summarise(arrestCount = n())
+stop_by_percentage_age <- left_join(stop_by_reason_age, stop_by_arrest_age)
+stop_by_percentage_age <- stop_by_percentage_age %>%
+  mutate(percentage = arrestCount / totalCount)
+
+stop_arrest_age <- ggplot(stop_by_percentage_age, aes(x = AgeRange, fill = AgeRange)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  labs(y = "Density") +
+  facet_wrap(~Reason)
+stop_arrest_age
+```
+
+```{r}
+stop_by_reason_gender <- stop_by %>%
+  group_by(Gender, Reason) %>%
+  summarise(totalCount = n())
+stop_by_arrest_gender <- stop_by %>%
+  filter(Enforcement == "A") %>%
+  group_by(Gender, Reason) %>%
+  summarise(arrestCount = n())
+stop_by_percentage_gender <- left_join(stop_by_reason_gender, stop_by_arrest_gender)
+stop_by_percentage_gender <- stop_by_percentage_gender %>%
+  mutate(percentage = arrestCount / totalCount)
+
+stop_arrest_gender <- ggplot(stop_by_percentage_gender, aes(x = Gender, fill = Gender)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  labs(y = "Density") +
+  facet_wrap(~Reason)
+stop_arrest_gender
+```
+
+```{r}
+stop_race <- ggplot(stop_by, aes(x = Reason, fill = Enforcement)) + 
+  geom_bar(aes(y = ..count../sum(..count..))) + 
+  labs(y = "Density") +
+  facet_wrap(~Race)
+stop_race
+
+```
+
+```{r}
+
+stop_race2 <- ggplot(stop_by, aes(x = Reason, fill = Enforcement)) + 
+  geom_bar(aes(y = ..count../sum(..count..))) + 
+  labs(y = "Density") +
+  scale_fill_brewer(palette = "Set1") +
+  facet_grid(Race~Gender)
+
+stop_race2
+
+```

### 2016-07-31 Berkeley_PD_Stop.Rmd
complete cleaning
write cleaned csv files

--- 2016-07-22 Berkeley_PD_Stop.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE)
+```
+
+```{r}
 library(tidyr)
@@
 
+ageVector <- c("1" = "0-18", "2" = "18-29", "3" = "30-39", "4" = "40+")
+stop_df$AgeRange = str_replace_all(stop_df$AgeRange, ageVector)
+
+stop_df$Reason <- str_replace(stop_df$Reason, "E", "W")
+
+write.csv(stop_df, "cleaned_stop_data.csv", row.names = FALSE)
 View(stop_df)
+
+```

### 2016-07-31 Stop_Data_Analysis.Rmd

---
title: "Stop Data Analysis"
author: "Mengyu Li"
date: "July 29, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### About the Berkeley PD Stop Data

The Berkeley PD stop data reflects accidents that have been visited by the Berkeley Police Department.

### 2016-07-31 Jenny_Analysis/Final Project Writeup.Rmd
jennyedit

---
title: "Final Project Arrest & Jail Dataset"
author: "Shangjun Jiang"
date: "29 July 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Outline

## 1. Introduction: Amy

## 2. Graph + Findings 
      - Stop Data: Mengyu
      - Jail & Arrest: Jenny

## 3. Maps
      - Rebecca

```{r}
setwd("~/Desktop/Stats R/PROJECT/finalproject_angry_ladies/Jenny_Analysis")
```

## About the Berkeley Police Arrest Log

The 'Berkeley Police Arrest Log' dataset provides information for adults arrested by the Berkeley Police Department over the past 30 days. Arrests include both felonies and misdemeanors. Arrests include persons booked into a jail facility and those cited and released with a future court date. The data are updated each business day (Monday through Friday, excluding holidays) at 9:00 am. At this time, the arrest location, arrest city, arrest latitude and longitude are not available.

Links:

Person                   | Time
------------------------ | --------------------
Arrest Number            | A sequential index number
Date and Time            | Date and Time of the arrest person
Arrest Type              | The type of arrest that was made
Subject                  | The name of the arrest person
Race                     | The race of the arrest person
Sex                      | The gender of the arrest person
Age                      | The age of the arrest person
Status Type              | The statutory code the statute number is found in. 
Case Number              | The associated case number for the police report documenting the arrest.

![](Arrest_AgeRaceSex.png)

## About the Berkeley Police Jail Booking Log

The 'Berkeley Police Jail Booking Log' dataset provides information  for those adults booked into the Berkeley Police Department Jail facility over the past 30 days. Bookings may be for any law enforcement agency utilizing the Berkeley Jail (not exclusively the Berkeley Police Department).The data are updated each business day (Monday through Friday, excluding holidays) at 9:00 am. At this time, the arrest location, arrest city, arrest latitude and longitude are not available.

## arrest time and booking time

### 2016-08-02 AnalysizeStop.Rmd
analysiss of stop data

@@
   labs(y = "Density") +
-  facet_wrap(~AgeRange)
+  facet_wrap(~Race)
 stop_time_percentage
@@
 ```
-
-```{r}
-stop_race <- ggplot(stop_by, aes(x = Reason, fill = Enforcement)) + 
-  geom_bar(aes(y = ..count../sum(..count..))) + 
-  labs(y = "Density") +
-  facet_wrap(~Race)
-stop_race
-
-```
-
-```{r}
-
-stop_race2 <- ggplot(stop_by, aes(x = Reason, fill = Enforcement)) + 
-  geom_bar(aes(y = ..count../sum(..count..))) + 
-  labs(y = "Density") +
-  scale_fill_brewer(palette = "Set1") +
-  facet_grid(Race~Gender)
-
-stop_race2
-
-```

### 2016-08-02 Stop_Data_Analysis.Rmd
analysiss of stop data

@@
 
+```{r, echo=FALSE}
+library(readr)
+library(ggplot2)
+library(dplyr)
+library(lubridate)
+library(stringr)
+library(ggvis)
+library(tidyr)
+```
+
+```{r, echo=FALSE}
+# Read data
+stop_by <- read_csv("cleaned_stop_data.csv")
+
+stop_by$Call.Date.Time <- mdy_hm(stop_by$Call.Date.Time)
+stop_by$AgeRange <- as.factor(stop_by$AgeRange)
+
+stop_by <- stop_by %>%
+  mutate(Hour = as.integer(hour(Call.Date.Time))) %>%
+  mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))
+```
 ### About the Berkeley PD Stop Data
@@
 The Berkeley PD stop data reflects accidents that have been visited by the Berkeley Police Department.
+Raw data set is downloaded from City of Berkeley Open Data website and provided by Berkeley Police Department.
+
+Analysis in this part focuses on the information provided by the dispositions variable.
+The dispositions variable can be divided into three types.
+The first type only includes more than one six-character person message.
+Each character in each person message represents the race, gender and age range of the person, reason and enforcement of the stop, and the car search information during the stop.
+The second type of the dispositions includes only additional dispositions, such as P for primary cases.
+The third type of dispositions includes both person information and additional dispositions.
+
+For the analysis in this part, only person information for all the people in each incident and the Call Time/Date of each incident are analyzed.
+
+### Introduction
+
+items                                | count    | percentage
+---------------------------------    | -------- |-------------------
+total incidents reported             | 16255    |     
+dispositions of type one             | 13016    | 80.07%
+dispositions of type two             | 2346     | 14.43%
+dispositions of type three           | 845      | 5.20%
+blank dispositions                   | 48       | 0.30%
+incidents with person information    | 13861    | 85.27%
+  incidents with 1 person            | 13494    | 97.352%
+  incidents with 2 people            | 322      | 2.323%              
+  incidents with 3 people            | 33       | 0.238%
+  incidents with 4 people            | 6        | 0.043%
+  incidents with 5 people            | 6        | 0.043%
+total personnel reported             | 14291    |
+
+(1) From 01/26/2015 to 04/30/2016, the Berkeley Police Department(or BPD) stopped for 16225 incidents.
+80.07% of all stops report only personnel information, 14.43% report only additional information, 5.20% report both and 0.30% report none.
+
+(2) Among all the stops, 13851 incidents recorded 14291 personnel dispositions in total.
+Of the incidents with the personal dispositions, 97.352% involves only one person, 2.648% involves more than two people.
+There are at most 5 people involved in one incident.
+
+### Analysis of Race
+#### Picture: count of person recorded in each race
+```{r, echo=FALSE}
+stop_race <- stop_by %>%
+  ggvis(x = ~Race , fill = ~AgeRange) %>%
+  layer_bars() %>%
+  add_tooltip(function(stop_by) stop_by$AgeRange)
+
+stop_race
+```
+
+```{r, echo=FALSE, eval=FALSE}
+stop_race_total_table <- stop_by %>%
+    group_by(Race) %>%
+    summarise(totaLCount = n())
+stop_race_agerange_table <- stop_by %>%
+    group_by(Race, AgeRange) %>%
+    summarise(count = n())
+stop_race_table <- left_join(stop_race_agerange_table, stop_race_total_table, by = "Race")
+stop_race_table <- stop_race_table %>%
+    mutate(percentage = count/totaLCount) %>%
+    select(-count)
+stop_race_table <- stop_race_table %>%
+    spread(key = AgeRange, value = percentage)
+stop_race_table
+```
+
+####Table: count of person recorded in each race
+
+Race  | Count | 0-18 | 18-29 | 30-39 | 40+
+----  |------ |------|-------|-------|-------
+A     | 1141  | 8.50%| 43.47%| 21.03%| 26.99%
+B     | 4636  | 2.29%| 34.97%| 25.09%| 37.66%
+H     | 1676  | 3.22%| 46.12%| 29.18%| 21.47%
+O     | 1384  | 2.67%| 43.28%| 29.17%| 26.87%
+W     | 5454  | 1.72%| 29.45%| 25.39%| 43.44%
+
+#### Picture: count of person recorded of each race in each hour
+```{r, echo=FALSE}
+stop_time_percentage <- ggplot(stop_by, aes(x = Hour, fill = AgeRange)) + 
+  geom_bar(aes(y = ..count..)) +
+  labs(y = "Count") +
+  facet_wrap(~Race)
+stop_time_percentage
+```
+
+(1) Among the 14291 personnel information recorded, white people contributes the largest percentage.
+38.16% of all records, namely 5454 people are white.
+black people contributes the second largest percentage.
+32.44% of all records, namely 4636 people are black.
+
+(2) The percentage of people aged from 0 to 18 of each race is less than 9%. People aged from 0 to 18 contributes the smallest percentage in each race, which makes sense in daily life.
+
+(3) Among white people and black people recorded by the BPD, people aged greater than 40  contribute the largest percentage, which are 43.44% and 37.66% respectively.
+
+(4) Among Asian, Hispanic and other people recorded by the BPD, people aged from 18 to 29 contribute the largest percentage, which are 43.37%, 46.12% and 43.28%, respectively.
+
+(5) Black people contribute a especially high percentage of the incidents at night, while the white people contribute a especially high percentage of the incidents at the noon.
+
+### Analysis of Day
+####Picture: count of person recorded in each day in a week
+```{r, echo=FALSE}
+stop_day <- stop_by %>%
+  ggvis(x = ~Day, fill = ~Race) %>%
+  layer_bars() %>%
+  add_tooltip(function(stop_by) stop_by$Race)
+stop_day
+```
+
+```{r, eval=FALSE, echo=FALSE}
+stop_day_total_table <- stop_by %>%
+    group_by(Day) %>%
+    summarise(totaLCount = n())
+stop_day_race_table <- stop_by %>%
+    group_by(Day, Race) %>%
+    summarise(count = n())
+stop_day_table <- left_join(stop_day_race_table, stop_day_total_table, by = "Day")
+stop_day_table <- stop_day_table %>%
+    spread(key = Race, value = count)
+stop_day_table
+```
+####Picture: count of person recorded in each day in a week
+Day   | Count | A    | B    | H    | O    | W
+----  |------ |------|------|------|------|------
+1     | 1692  | 116  | 653  | 221  | 159  | 543
+2     | 1665  | 149  | 543  | 182  | 166  | 625
+3     | 2344  | 189  | 645  | 231  | 236  | 1043
+4     | 2287  | 206  | 616  | 251  | 244  | 970
+5     | 1941  | 152  | 592  | 223  | 168  | 806
+6     | 2231  | 186  | 764  | 268  | 203  | 810
+7     | 2131  | 143  | 823  | 300  | 208  | 657
+
+(1) Among all the day in a week, the number of incidents happened on Sunday and Monday are much less than those happened through Tuesday and Saturday.
+
+(2) The count of records of Asian people is much less than that of any other race of people in each day and the whole week.
+
+(3) Asian people are less liable to commit an incident on Sunday, and more liable to commit an incident on Wednesday.
+
+(4) Black and Hispanic people are less liable to commit an incident on Monday, and more liable to commit an incident on Saturday.
+
+(5) White people are less liable to commit an incident on Sunday, and more liable to commit an incident on Tuesday and Wednesday, which is similar to Asian people.
+
+### Analysis of incidents of each hour in a day
+#### Picture: count of person recorded in each hour in a day
+```{r, echo=FALSE}
+stop_time <- stop_by %>%
+  ggvis(x = ~Hour, fill := "lightpink") %>%
+  layer_histograms(width = input_slider(1, 12, value = 1, step = 2, label = "width"))
+stop_time
+```
+Note: you can change the value of the input slider to see the number of incidents happened in specific time span of the day.
+```{r, eval = FALSE, echo=FALSE}
+stop_hour <- stop_by %>%
+  group_by(Hour) %>%
+  summarise(count = n())
+stop_hour
+```
+
+#### Table: count of person recorded in each hour in a day
+Hour  | Count | Hour | Count| Hour | Count| Hour | Count
+----  |------ |------|------|------|------|------|------
+0     | 823   | 6    | 157  | 12   | 802  | 18   | 544
+1     | 618   | 7    | 341  | 13   | 695  | 19   | 685
+2     | 489   | 8    | 489  | 14   | 515  | 20   | 834
+3     | 289   | 9    | 582  | 15   | 542  | 21   | 971
+4     | 135   | 10   | 682  | 16   | 612  | 22   | 1098
+5     | 61    | 11   | 678  | 17   | 578  | 23   | 1071
+
+(1) time span of 22:00 to 22:59 in a day has the highest incidents count, and 23:00 to 23:59 has the second highest, which means that Berkeley area is most dangerous from 10pm to 12am in a day, which corresponds to daily life experience.
+(2) From 12am to 5am, incidents count decreases gradually.
+A reasonable conjecture of this fact is that more and more people choose to sleep as time goes by in this time span.
+For the similar reason, incidents count increases gradually from 5am to 11am.
+(3) After the evening until the midnight, the incidents number increases gradually again in a day. The tally of the statistics and observation in daily life is in a good agreement.
+
+#### Picture: Probability of BPD stop for a specific age range in a specific hour
+
+```{r, echo=FALSE}
+stop_by_hourCount <- stop_by %>%
+  group_by(Hour) %>%
+  summarise(hourcount = n())
+stop_by_ageRangeCount <- stop_by %>%
+  group_by(Hour, AgeRange) %>%
+  summarise(ageCount = n())
+stop_by_percentage <- left_join(stop_by_ageRangeCount, stop_by_hourCount, by = "Hour")
+stop_by_percentage <- stop_by_percentage %>%
+  mutate(percentage = ageCount/hourcount)
+```
+
+```{r, echo=FALSE}
+stop_time_percentage2 <- ggplot(stop_by_percentage, aes(x = Hour, fill = AgeRange)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  labs(y = "Density") +
+  facet_wrap(~AgeRange)
+stop_time_percentage2
+```
+
+(1) People aged from 18 to 29 are obviously liable to commit incidents at night.
+The average ratio of people aged from 18 to 29 to all people stoped by the Berkeley Police Department at night is greater than 40%.
+
+(2) People aged greater than 40 are obviously liable to commit incidents in the daytime.
+The average ratio of people aged greater than 40 to all people stoped by the Berkeley Police Deparment during the day is greater than 40%.
+
+(3) Ratio of incidents commited by people aged between 0 and 18 and people aged from 30 to 39 fluctuates during the daytime and the night, with an average ratio of 2.5% and 25% respectively.
+
+#### Probability of BPD stop for a specific race in a specific hour
+
+```{r, echo=FALSE}
+stop_by_hourCount <- stop_by %>%
+  group_by(Hour) %>%
+  summarise(hourcount = n())
+stop_by_raceRangeCount <- stop_by %>%
+  group_by(Hour, Race) %>%
+  summarise(raceCount = n())
+stop_by_percentage1 <- left_join(stop_by_raceRangeCount, stop_by_hourCount, by = "Hour")
+stop_by_percentage1 <- stop_by_percentage1 %>%
+  mutate(percentage = raceCount/hourcount)
+```
+```{r, echo=FALSE}
+stop_time_percentage3 <- ggplot(stop_by_percentage1, aes(x = Hour, fill = Race)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  labs(y = "Density") +
+  facet_wrap(~Race)
+stop_time_percentage3
+```
+
+(1) Black people are obviously liable to commit incidents at night. The average ratio of black people to all people stoped by the Berkeley Police Department at night is about 40%.
+
+(2) White people are obviously liable to commit incidents in the daytime. The average ratio of white people to all people stoped by the Berkeley Police Deparment during the day is about 45%.
+
+(3) Ratio of incidents commited by Asian people, Hispanic people and other people fluctuate during the daytime and the night, with an average ratio of 8%, 11% and 9% respectively.
+
+### Analysis of prefence of Berkeley Police Department arresting people
+#### Picture: Probability of arrested by BPD of a specific race for a specific reason
+```{r,echo=FALSE}
+stop_by_reason_race <- stop_by %>%
+  group_by(Race, Reason) %>%
+  summarise(totalCount = n())
+stop_by_arrest <- stop_by %>%
+  filter(Enforcement == "A") %>%
+  group_by(Race, Reason) %>%
+  summarise(arrestCount = n())
+stop_by_percentage_race <- left_join(stop_by_reason_race, stop_by_arrest, by = c("Race", "Reason"))
+stop_by_percentage_race <- stop_by_percentage_race %>%
+  mutate(percentage = arrestCount / totalCount)
+stop_arrest <- ggplot(stop_by_percentage_race, aes(x = Race, fill = Race)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  labs(y = "Density") +
+  facet_wrap(~Reason)
+stop_arrest
+```
+
+#### Table: Probability of arrested by BPD of a specific race for a specific reason
+P   | I      | K     | R     | T     | W     |
+----|------- |-------|-------|-------|-------|
+A   | 10.34% | 75%   | 10.34%| 1.24% | 17.60%|
+B   | 8.31%  | 15.79%| 6.44% | 2.28% | 50.00%|  
+H   | 6.98%  | 16.67%| 9.09% | 2.90% | 25.00%| 
+O   | 10.14% | 18.18%| 5.80% | 0.41% | 41.67%| 
+W   | 8.14%  | 26.79%| 3.40% | 1.04% | 32.56%|
+
+Note: in the picture and table above, I for Investigation, T for Traffic, R for Reasonable Suspicion, K for Probation/Parole, W for wanted.
+
+(1) The probability of arrested in a stop with the reason Traffic is much lower than other reasons. The average conditional probability of arrested given reason is traffic is 1.58%.
+
+(2) The probability of arrested in a stop with the reason Wanted is much higher than other reason. The average conditional probability of arrested given reason is wanted is 33.33%.
+
+(3) An interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and race Asian is 75%, which is much higher than that of any other race.
+Asian people is much more liable to be arrested by Berkeley Police Department during Probation or Parole if stopped by the BPD.
+
+#### Picture: Probability of arrested by BPD of a specific age range for a specific reason
+```{r, echo=FALSE}
+stop_by_reason_age <- stop_by %>%
+  group_by(AgeRange, Reason) %>%
+  summarise(totalCount = n())
+stop_by_arrest_age <- stop_by %>%
+  filter(Enforcement == "A") %>%
+  group_by(AgeRange, Reason) %>%
+  summarise(arrestCount = n())
+stop_by_percentage_age <- left_join(stop_by_reason_age, stop_by_arrest_age, by = c("AgeRange", "Reason"))
+stop_by_percentage_age <- stop_by_percentage_age %>%
+  mutate(percentage = arrestCount / totalCount)
+
+stop_arrest_age <- ggplot(stop_by_percentage_age, aes(x = AgeRange, fill = AgeRange)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  labs(y = "Density") +
+  facet_wrap(~Reason)
+stop_arrest_age
+```
+
+#### Table: Probability of arrested by BPD of a specific age range for a specific reason
+P     | I     | K     | R     | T     | W     
+------|-------|-------|-------|-------|-------
+0-18  | 17.95%| 67.57%| 13.95%| 3.24% | 50.00%  
+18-29 | 8.45% | 12.50%| 3.87% | 1.82% | 32.35%  
+30-39 | 6.32% | 25.00%| 5.04% | 1.49% | 26.92%  
+40+   | 7.87% | 23.86%| 6.13% | 1.41% | 31.82%
+
+(1) Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.
+
+(2) Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and age range 0-18 is 67.56%, which is much higher than that of any other age range.
+
+(3) The conditional probability of arrested given person involved aged from 0 to 18 in a stop given any reason is more than that of given person with any other age range.
+Therefore, in a stop, teenagers is more liable to be arrested by Berkeley Police Department.
+
+#### Picture: Probability of arrested by BPD of a specific gender for a specific reason
+```{r, echo=FALSE}
+stop_by_reason_gender <- stop_by %>%
+  group_by(Gender, Reason) %>%
+  summarise(totalCount = n())
+stop_by_arrest_gender <- stop_by %>%
+  filter(Enforcement == "A") %>%
+  group_by(Gender, Reason) %>%
+  summarise(arrestCount = n())
+stop_by_percentage_gender <- left_join(stop_by_reason_gender, stop_by_arrest_gender, by = c("Gender", "Reason"))
+stop_by_percentage_gender <- stop_by_percentage_gender %>%
+  mutate(percentage = arrestCount / totalCount)
+
+stop_arrest_gender <- ggplot(stop_by_percentage_gender, aes(x = Gender, fill = Gender)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  labs(y = "Density") +
+  facet_wrap(~Reason)
+stop_arrest_gender
+```
+
+#### Table: Probability of arrested by BPD of a specific gender range for a specific reason
+P     | I      | K     | R     | T     | W    
+----  |------- |-------|-------|-------|-------
+F     | 9.05%  | 50.00%| 5.00% | 1.10% | 28.57%  
+M     | 8.26%  | 20.11%| 5.42% | 1.85% | 37.04%
+
+(1) Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.
+
+(2) Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole of female is 50.00%, which is much higher than that of male, which is 20.11%.

### 2016-08-02 Jenny_Analysis/Jail.Rmd

---
title: "Jail"
author: "Shangjun Jiang"
date: "30 July 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("~/Desktop/Stats R/PROJECT/finalproject_angry_ladies/Jenny_Analysis")
```

```{r}
library(rvest)
library(xml2)
library(ggplot2)
library(lubridate)
library(stringr)
library(dplyr)
library(tidyr)
library(readr)
library(scales)
library(ggvis)
```

```{r}
jail <- read.csv("Berkeley_PD_Log_-_Jail_Bookings.csv")
jail$Age <- as.integer(jail$Age)
jail$Age[jail$Age<20] <- "0-19"
jail$Age[jail$Age>=20 & jail$Age<30] <- "20-29"
jail$Age[jail$Age>=30 & jail$Age<40] <- "30-39"
jail$Age[jail$Age>=40 & jail$Age<50] <- "40-49"
jail$Age[jail$Age>=50 & jail$Age<60] <- "50-59"
jail$Age[jail$Age>=60 & jail$Age<70] <- "60-69"
jail$Age[jail$Age>=70 & jail$Age<80] <- "70-79"
jail$Age <- as.factor(jail$Age)

#jail$Arrest.Date.and.Time<- as.POSIXct(as.character(jail$Arrest.Date.and.Time))
#jail$Booking.Date.and.Time <- as.POSIXct(as.character(jail$Booking.Date.and.Time))

#jail$Arrest.Date.and.Time <- mdy_hms(jail$Arrest.Date.and.Time)
#jail$Booking.Date.and.Time <- mdy_hms(jail$Booking.Date.and.Time)
#jail$Date.Of.Birth <- mdy(as.character(jail$Date.Of.Birth))

names(jail)
#[1] "Booking.Number"        "Booking.Date.and.Time" "Subject"              
#[4] "Race"                  "Sex"                   "Date.Of.Birth"        
#[7] "Age"                   "Height"                "Weight"               
#[10] "Hair"                  "Eyes"                  "Occupation"           
#[13] "Statute"               "Statute.Type"          "Statute.Description"  
#[16] "Arrest.Date.and.Time"  "Case.Number"           "Booking.Agency"       
#[19] "Disposition"  
jail[jail == ""] <- NA
najail <- na.omit(jail)

```

## race and disposition
```{r}
ggplot(najail)+
  geom_bar(aes(x=Disposition,
               fill = Sex))+
  facet_wrap(~Race)

ggplot(najail)+
  geom_bar(aes(x=Disposition,
               fill = Race))+
  facet_wrap(~Sex)

```

## demographics data
```{r}
ggplot(najail)+
  geom_bar(aes(x=Age, 
               fill = Race))+
  facet_wrap(~Sex)+
  labs(x= "Age",
       y = "Number of people",
       title = "Jail Data")

```

## occupation data
```{r, eval=F}
najail$Race <- as.factor(najail$Race)
ggplot(najail) +
  geom_bar(aes(x=Occupation,
           fill = Race))
hour(jail$Booking.Date.and.Time)

ggvis(jail,
      x=~hour(Booking.Date.and.Time),
      fill := "pink") %>% 
  layer_bars(width = input_slider(1,20,step=2,label="width"))

ggvis(jail,
      x=~hour(Arrest.Date.and.Time),
      fill := "pink") %>% 
  layer_bars(width = input_slider(1,20,step=2,label="width"))

  layer

```

```{r, eval=F}
#group_by hour, count

names(jail)

jail %>% 
  filter(Race == "Black")
?filter

jail<- jail %>% 
  mutate(arresthour =hour(Arrest.Date.and.Time))
#group by arrest hour
newjail<- jail%>% 
  group_by(arresthour) %>% 
  summarise(hourcount=n())
#group by race
new2jail <- jail %>% 
  group_by(Race, arresthour) %>% 
  summarise(racecount=n())
#group by age
new3jail <- jail %>% 
  group_by(Age, arresthour) %>% 
  summarise(agecount=n())

racearrestjail<-left_join(new2jail,newjail, by = "arresthour")

agearrestjail <-left_join(new3jail,newjail,by = "arresthour")

#race arrest plot
ggplot(na.omit(racearrestjail))+
  geom_bar(aes(x=arresthour, 
                     y=(racecount/hourcount),
                     fill = Race),
           stat="identity") +
  facet_wrap(~Race)

```

## Arrest date and booking date 
jail$Arrest.Date.and.Time <- mdy_hms(jail$Arrest.Date.and.Time)
jail$Booking.Date.and.Time <- mdy_hms(jail$Booking.Date.and.Time)

```{r, eval=F}
ggplot(jail)+
  geom_segment(aes(x= hour(Arrest.Date.and.Time), 
               xend = hour(Booking.Date.and.Time),
               y = day(Arrest.Date.and.Time),
               yend = day(Booking.Date.and.Time)))

ggplot(jail)+
  geom_segment(aes(x= hour(Arrest.Date.and.Time), 
                  xend = hour(Booking.Date.and.Time),
                  y= Race,
                  yend=Race))

ggplot(jail)+
  geom_segment(aes(x= hour(Arrest.Date.and.Time), 
                  xend = hour(Booking.Date.and.Time),
                  y= minute(Arrest.Date.and.Time),
                  yend=minute(Booking.Date.and.Time)))

ggplot(na.omit(jail))+
  geom_point(aes(x= hour(Arrest.Date.and.Time), 
                 y= hour(Booking.Date.and.Time),
                 color = Race,
                 shape = Sex))

najail <- na.omit(jail)
najail <- najail %>% 
  mutate(arresthour = hour(Arrest.Date.and.Time)) %>% 
  mutate(bookhour = hour(Booking.Date.and.Time))

nai
najail %>% 
  ggvis(~arresthour,
        ~bookhour,
         fill= ~Race,
        opacity := 0.7) %>%
  layer_points() %>%
  add_tooltip((function(naarrest) najail$arresthour))

```

### 2016-08-02 Rebeccas_location_Code2/StopData_Maps.Rmd
location maps

---
title: "Berkeley Police Stop Data Maps"
author: "Rebecca Reus"
date: "July 31, 2016"
output: 
  html_document:
    fig_width: 10
    fig_height: 10
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sp)
library(ggmap)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(data.table)
library(rworldmap)
library(maps)
library(mapdata)
library(maptools)
library(scales)
library(RgoogleMaps)
library(tmap)
library(sp)
library(rgdal)
library(rgeos)
library(RColorBrewer)
#library(raster)
#library(GISTools)
 #library(mgcv)
```

## Read in the Data:
```{r}
#finaldf <- readRDS( 'StopData_finaldf.rds' )
mergedf <- readRDS( 'StopData_merged.rds' )
```

## Location info:
```{r}
latmax <- max( mergedf$lat, na.rm = TRUE ) 
latmin <- min( mergedf$lat, na.rm = TRUE )
lonmax <- max( mergedf$long, na.rm = TRUE ) 
lonmin <- min( mergedf$long, na.rm = TRUE )
latvals <- c( latmin, latmax )
lonvals <- c( lonmin, lonmax )
```

## Make the variables factors:
```{r}
mergedf$CarSearch <- factor(mergedf$CarSearch)
levels(mergedf$CarSearch) <- c("No Search", "Search")

mergedf$Race <- factor(mergedf$Race)
levels(mergedf$Race) <- c("Asian", "Black", "Hispanic", "Other", "White")

mergedf$Enforcement <- factor(mergedf$Enforcement)
levels(mergedf$Enforcement) <- c("Arrest", "Citation", "Other", "Warning")

mergedf$Reason <- factor(mergedf$Reason)
levels(mergedf$Reason) <- c("Other", "Investigation", "Probation/Parole", "Reasonable Suscipcion", "Traffic", "Wanted")

mergedf <- mergedf %>%
  mutate(Arrest = ifelse( 
    as.character(Enforcement) == "Arrest", 
    "Arrested",
    "Not Arrested") ) %>%
  mutate(Arrest = factor(Arrest))

v <- factor(mergedf$Other)
levels(v)

mergedf <- mergedf %>%
  mutate(Emergency.Psych.Eval = ifelse( str_detect(as.character(Other), "MH"), 
                                        yes = "Yes",
                                        no = "No") )

mergedf$Emergency.Psych.Eval <- factor(mergedf$Emergency.Psych.Eval)

```

## Map 1
```{r}
newmap <- getMap( resolution="low" )
plot( newmap, xlim = lonvals, ylim = latvals, asp=1 )
points( mergedf$long, mergedf$lat, col="red", cex=.6 )
```

## Map 2
```{r}

berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14)
ggmap(berkMap) +
  geom_point(aes(x = long, y = lat, color = Incident.Type), data = mergedf, alpha = 0.2, size = 3) + 
  scale_colour_discrete( "Incident Type", labels = c("Pedestrian Stop", "Bicycle Stop", "Suspicious Vehicle Stop",  "Traffic Stop" ) ) +
  theme ( 
		#legend.position = c(0.05, 0.05), # put the legend INSIDE the plot area
		#legend.justification = c(0, 0),
		#legend.background = element_rect(colour = F, fill = "white"),
		#legend.key = element_rect (fill = F, colour = F),
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + 
  ggtitle("BPD Stops by Incident Type, 2015-2016") 
```

## MAP 2.1
```{r}
ggmap(berkMap) +
  geom_point(aes(x = long, y = lat, size = CarSearch, color = Race), data = mergedf, alpha = 0.2) + 
  scale_size_discrete( "Car Search", labels = c("No Search", "Search") ) +
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + 
  ggtitle("BPD Stop Car Searches by Race 2015-2016") +
  facet_wrap(~ Race, ncol = 3 ) 
```

## MAP 2.2
```{r}
df <- subset(mergedf, as.character(Enforcement) == "Arrest")

ggmap(berkMap) +
  geom_point(aes(x = long, y = lat, colour = Race), data = df, alpha = 0.5, size = 5) + 
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + 
  ggtitle("BPD Stops to Arrests, by Race 2015-2016") +
  facet_wrap(~ Race) 
```

## MAP 2.3
```{r}
ggmap(berkMap) +
  geom_point(aes(x = long, y = lat, colour = Race, size = Arrest), data = mergedf, alpha = 0.5)  + 
  scale_size_discrete( range = c(7, 3) ) +
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + 
  ggtitle("BPD Stops to Arrests, by Race 2015-2016") +
  facet_wrap(~ Race)

# p <- sum(as.character(mergedf$Arrest) == "Arrested")/nrow(Arrest)
# 
# print( str_c( Percent of People Arrested Out of Total People Stopped: ), )
```

## MAP 2.4
```{r}

df <- subset(mergedf, as.character(Emergency.Psych.Eval) == "Yes")

ggmap(berkMap) +
  geom_point(aes(x = long, y = lat, colour = Race), data = df, alpha = 0.5, size = 6) +
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + 
  ggtitle("BPD Stops: Emergency Psch Evaluation, by Race 2015-2016")

```

## MAP 3:
```{r}
ggmap(berkMap) +
  geom_point(aes(x = long, y = lat, colour = factor(Gender)), data = mergedf, alpha = 0.2, size = 3) + 
  scale_colour_discrete( "Gender", labels = c("Female", "Male") ) +
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + 
  ggtitle("BPD Stops by Gender, 2015-2016") 
```

## MAP 4:
```{r}
ggmap(berkMap) +
  geom_point(aes(x = long, y = lat, colour = Reason), data = mergedf, alpha = 0.7, size = 3) +
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + ggtitle("All BPD Stops, by Reason 2015-2016") +
    facet_wrap( ~ Reason , ncol = 2)
```

## MAP 4.1
```{r}
df <- subset(mergedf, as.character(Reason) == "Traffic")

ggmap(berkMap) +
  geom_point(aes(x = long, y = lat, colour = Race, size= Arrest), data = df, alpha = 0.5) + 
  scale_size_discrete( range = c(7, 3) ) +
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + 
  ggtitle("BPD Traffic Stops Only, by Race 2015-2016") +
  facet_grid(~ Race)
```

## MAP 5:
```{r}
# doing it with qmplot is even easier
qmplot(long, lat, data = mergedf, maptype = "toner-lite",
  color = factor(Race), size = factor(CarSearch), legend = "topleft"
) +
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + ggtitle("BPD Stops 2015-2016")  + 
  scale_colour_discrete( "Race", labels = c("Asian", "Black", "Hispanic", "Other", "White")  ) +
  scale_size_discrete("Car Search", labels = c("No Search", "Search"))
```

## MAP 6:
```{r}
ggmap(berkMap) +
  geom_point( data = mergedf, aes(x = long, y = lat, colour = factor(Race), size = factor(CarSearch) ), alpha = 0.5 ) + 
  scale_colour_discrete( "Race", labels = c("Asian", "Black", "Hispanic", "Other", "White")  ) +
  scale_size_discrete("CarSearch", labels = c("No Search", "Search"),
     range = c(1.75,6)) +
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + ggtitle("BPD Stops by Race and Car Search 2015-2016") +
  labs(size = "CarSearch")
```

## MAP 6.1:
```{r}
# a contour plot
ggmap(berkMap) +
  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
    size = 2, bins = 5, data = mergedf, geom = "polygon") +
  scale_fill_gradient(low = "black", high = "red") +
    theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + ggtitle("All BPD Stops Density, 2015-2016") +
  labs(alpha = element_blank())
```

## MAP 6.2
```{r}
qmplot(long, lat, data = mergedf) +
  ggtitle("All BPD Stops")
qmplot(long, lat, data = mergedf, facets = ~ Race) +
  ggtitle("All BPD Stops by Race")
```

## MAP 7
```{r, fig.height = 25, fig.width = 5}
# crime example by month
mergedf <- mergedf %>%
 # mutate( Call.Date.Time = mdy_hm( as.character(Call.Date.Time) ) ) %>%
 # mutate( Call.Hour = hour( as.character(Call.Date.Time) ) ) %>%
  #mutate( Call.Date = mdy( as.character(Call.Date.Time) ) ) %>%
  mutate( Call.Month = month( mdy_hm(Call.Date.Time) ) ) #%>%
 # mutate( Call.Weekday = wday( as.character(Call.Date.Time) ) )
mergedf$Call.Month <- factor(mergedf$Call.Month)

map7 <- get_map(location = "berkeley", zoom = 14, source = "osm", color = "bw")
BerkMap7 <- ggmap( map7,
  base_layer = ggplot(aes(x = long, y = lat), data = mergedf))
```

```{r}
BerkMap7 +
  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
    bins = I(5), geom = "polygon", data = mergedf ) +
  scale_fill_gradient2( "StopDensity",
    low = "white", mid = "orange", high = "red", midpoint = 500) +
  labs(x = "Longitude", y = "Latitude") + facet_wrap(~ Call.Month, ncol = 6) +
  scale_alpha(range = c(.2, .55), guide = FALSE) +
  ggtitle("BPD Stop Contour Map of Berkeley by Month") +
  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
```

## MAP 8
# ```{r}
# # crime density by race:
# mergedf <- mergedf %>%
#  # mutate( Call.Date.Time = mdy_hm( as.character(Call.Date.Time) ) ) %>%
#  # mutate( Call.Hour = hour( as.character(Call.Date.Time) ) ) %>%
#   #mutate( Call.Date = mdy( as.character(Call.Date.Time) ) ) %>%
#   mutate( Call.Month = month( mdy_hm(Call.Date.Time) ) ) #%>%
#  # mutate( Call.Weekday = wday( as.character(Call.Date.Time) ) )
# mergedf$Call.Month <- factor(mergedf$Call.Month)
# 
# map7 <- get_map(location = "berkeley", zoom = 6, source = "osm", color = "bw")
# BerkMap7 <- ggmap( map7,
#   base_layer = ggplot(aes(x = long, y = lat), data = mergedf))
# 
# BerkMap7 +
#   stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
#     bins = I(5), geom = "polygon", data = mergedf ) +
#   scale_fill_gradient2( "StopDensity",
#     low = "white", mid = "orange", high = "red", midpoint = 500) +
#   labs(x = "Longitude", y = "Latitude") + facet_wrap(~ Race ) +
#   scale_alpha(range = c(.2, .55), guide = FALSE) +
#   ggtitle("BPD Stop Contour Map of Berkeley by Race") +
#   guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
# ```

## MAP 9
```{r, fig.height = 25, fig.width = 5}
ggmap(berkMap) +
  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
    bins = I(5), geom = "polygon", data = mergedf ) +
  scale_fill_gradient2( "StopDensity",
    low = "white", mid = "orange", high = "red", midpoint = 100) +
  labs(x = "Longitude", y = "Latitude") + facet_wrap( ~ Race ) +
  scale_alpha(range = c(.2, .55), guide = FALSE) +
  ggtitle("BPD Stop Contour Map of Berkeley by Race") +
  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
```

## MAP 10 : AgeRange Density Facet
```{r, fig.height = 25, fig.width = 5}
ggmap(berkMap) +
  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
    bins = I(5), geom = "polygon", data = mergedf ) +
  scale_fill_gradient2( "Stop Density",
    low = "white", mid = "orange", high = "red", midpoint = 25) +
  labs(x = "Longitude", y = "Latitude") + facet_wrap(~ AgeRange) +
  scale_alpha(range = c(.2, .55), guide = FALSE) +
  ggtitle("BPD Stop Contour Map of Berkeley by Age Range") +
  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
```

## MAP 11: By Race Again
```{r, fig.height = 25, fig.width = 5}
# ggmap(berkMap) +
# geom_point(aes(x = long, y = lat, colour = Race, size = Race),
#   data = mergedf)

ggmap(berkMap) +
stat_bin2d(
aes(x = long, y = lat, colour = Race, fill = Race),
size = .25, bins = 100, alpha = .25,
data = mergedf
)

```

```{r}
ggmap(berkMap) +
stat_density2d(aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
bins = 5, geom = "polygon",
data = mergedf ) +
scale_fill_gradient(low = "black", high = "red")
```

# Get Shapefiles
```{r}
locationCensusFiles <- "C:/Users/Rebecca/Dropbox/School/2016_su_School/stat133/finalproject_angry_ladies/Rebeccas_location_Code2/Census_Block_Group_Polygons2010"
blocks <- readOGR(locationCensusFiles,"Census_blockgroups_2010", verbose = TRUE)
summary(blocks)
print(proj4string(blocks))
blocks.polygons <- blocks@polygons
blocks.data <- blocks@data
blocks.plot.order <- blocks@plotOrder
b2 <- spTransform(blocks, CRS("+proj=longlat +datum=WGS84"))
b3 <- fortify(b2)
```

## Map 12
```{r}
bMap12 <- ggmap(berkMap) + 
  geom_polygon(aes(x=long, y=lat, group=group), fill='grey', size=1,color='red', data=b3, alpha=0) +
  geom_point( data = mergedf, aes(x = long, y = lat), alpha = 0.2, size = 3 )
bMap12
```

## MAP 13
```{r}
bMap12 <- ggmap(berkMap) + 
geom_polygon(aes(x = long, y = lat, group = group), data = b3,
               colour = 'white', fill = 'black', alpha = .4, size = .3) +
  geom_point( data = mergedf, aes(x = long, y = lat), alpha = 0.2, size = 3, color = 'red' )
bMap12

```

# Get Shapefiles Again
```{r}
locationCensusFiles <- "C:/Users/Rebecca/Dropbox/School/2016_su_School/stat133/finalproject_angry_ladies/Rebeccas_location_Code2/Census_Tract_Polygons2010"
blocks <- readOGR(locationCensusFiles,"Census_tracts_2010", verbose = TRUE)
summary(blocks)
print(proj4string(blocks))
# blocks.polygons <- blocks@polygons
# blocks.data <- blocks@data
# blocks.plot.order <- blocks@plotOrder
b2 <- spTransform(blocks, CRS("+proj=longlat +datum=WGS84"))
b3 <- fortify(b2)
```

## Map 14
```{r}
bMap12 <- ggmap(berkMap) + 
geom_polygon(aes(x = long, y = lat, group = group), data = b3,
               colour = 'black', fill = 'white', alpha = .4, size = .3) +
  geom_point( data = mergedf, aes(x = long, y = lat), alpha = 0.2, size = 3, color = 'red' )
bMap12
```

## Population by census block group:
```{r}
shpfile <- "Census_tracts_2010.shp"
sh <- readShapePoly("C:/Users/Rebecca/Dropbox/School/2016_su_School/stat133/finalproject_angry_ladies/Rebeccas_location_Code2/Census_Tract_Polygons2010/Census_tracts_2010.shp")
plot(sh)
sh@data$ID <- as.numeric(sh@data$ID)

# Read in the demographic data and merge on Neighbourhood Id
demo <- read.csv(file="Census_tracts_2010.csv", header=T)
sh2 <- merge(sh, demo, by='ID')

# Set the palette
p <- colorRampPalette(c("white", "red"))(128)
palette(p)

# Scale the total population to the palette
pop <- sh2@data$TotalPop.x
cols <- (pop - min(pop))/diff(range(pop))*127+1
plot(sh, col=cols)
#points( mergedf$long, mergedf$lat, col="black", cex=.6 )

```

# Map 15
```{r}
j <- merge( b3, demo, by = "group"  )

# Plot
bMap12 <- ggmap(berkMap) +
  geom_polygon(aes(x=long,y=lat, group=group, fill=TotalPop), data=j, color='white') +
  scale_fill_gradient(low='white', high='orange') +
  geom_point( data = mergedf, aes(x = long, y = lat), alpha = 0.2, size = 3, color = 'red' )
bMap12
```

### 2016-08-02 Rebeccas_location_Code2/StopDataGetLocation.R

@@ -154,6 +154,8 @@ stop_df <- stop_df %>%
     yes = str_replace( Location, " BLOCK", "" ), 
     no = Location))                                                  # "BLOCK" Interferes with Google server info
 
+saveRDS( stop_df, 'StopData_stop_df.rds' ) # save this form of stopdf for later.
+
 ####################################################################################################################################
 ## GETTING LOCATION COORDINATES WITH GOOGLE:
 ####################################################################################################################################
@@ -310,6 +312,9 @@ points( finaldf$long, finaldf$lat, col="red", cex=.6 )
 
 # WRITE finaldf to a file, if you would like to use the Location info at another time:
 write.csv( finaldf, file = "StopData_w_Locations.csv")
+tempfilename_finaldf <- paste0( infile, '_temp_finaldf')
+saveRDS( finaldf, tempfilename_finaldf )
+
 
 # Use this code to see how many rows left Google will allow you for today:
 geocodeQueryCheck()

### 2016-08-02 Rebeccas_location_Code2/StopData_ManualFix.R

@@ -1,52 +1,191 @@
-#############################################################################################################################
-## FIXING OTHER PROBLEMS:
-#############################################################################################################################
+# StopData_ManualFix.R
+
+# Purpose: Use this file to manually change latitude and longitude errors by hand.
+#          Do NOT run this file unless you know EXACTLY which row you want to change, 
+#          since you must manually enter the row you want to change, and this file 
+#          updates that row in our temp file (the .rds file in this folder).
+# Author: Rebecca Reus, STAT 133 Group: The Angry Ladies, Summer 2016, UC Berkeley
+
+####################################################################################################
 
 #### IMPORTANT: ONLY RUN THIS CODE IF YOU ARE REBECCA: THIS IS FOR MANUAL FIXES! LET REBECCA FIX AT THE END.
 
+###################################################################################################################################
+## LIBRARIES:
+## Before running, please make sure you have installed ALL of these packages:
+###################################################################################################################################
+library(sp)
+library(ggmap)
+library(tidyr)
+library(plyr)
+library(dplyr)
+library(stringr)
+library(lubridate)
+library(ggplot2)
+library(readr)
+library(data.table)
+library(rworldmap)
+library(maps)
+library(mapdata)
+library(maptools)
+library(scales)
+library(RgoogleMaps)
+library(tmap)
+library(sp)
+library(rgdal)
+library(rgeos)
 
-## Only run this if you know what exactly to change by hand:
-finaldf[ which( finaldf$lat == latmin ),]
+####################################################################################################################################
+## CREATE SOME FUNCTIONS:
+####################################################################################################################################
 
+# Define a function that will process Google's server responses:
+getGeoDetails <- function( address ) {   
+  
+  # Use the gecode() function from the ggmap library to query Google's servers:
[…]

### 2016-08-02 Rebeccas_location_Code2/StopData_Maps.Rmd
map changes

@@
 		) + 
-  ggtitle("BPD Stops by Incident Type, 2015-2016") 
+  ggtitle("All BPD Stops by Incident Type, 2015-2016") 
 ```
@@
 		) + 
-  ggtitle("BPD Stop Car Searches by Race 2015-2016") +
+  ggtitle("BPD All Stops w/ Car Searches by Race 2015-2016") +
   facet_wrap(~ Race, ncol = 3 ) 
@@
 		) + 
-  ggtitle("BPD Stops: Emergency Psch Evaluation, by Race 2015-2016")
+  ggtitle("BPD Stops with Emergency Psch Evaluation, by Race 2015-2016")
 
@@
 		) + 
-  ggtitle("BPD Stops by Gender, 2015-2016") 
+  ggtitle("All BPD Stops by Gender, 2015-2016") 
 ```
@@
   geom_point(aes(x = long, y = lat, colour = Race, size= Arrest), data = df, alpha = 0.5) + 
-  scale_size_discrete( range = c(7, 3) ) +
+  scale_size_discrete( range = c(5, 2) ) +
   theme ( 
@@
 ## MAP 9
-```{r, fig.height = 25, fig.width = 5}
+```{r}
 ggmap(berkMap) +
@@
 ## MAP 10 : AgeRange Density Facet
-```{r, fig.height = 25, fig.width = 5}
+```{r}
 ggmap(berkMap) +
@@
 ## MAP 11: By Race Again
-```{r, fig.height = 25, fig.width = 5}
+```{r}
 # ggmap(berkMap) +
@@
 shpfile <- "Census_tracts_2010.shp"
-sh <- readShapePoly("C:/Users/Rebecca/Dropbox/School/2016_su_School/stat133/finalproject_angry_ladies/Rebeccas_location_Code2/Census_Tract_Polygons2010/Census_tracts_2010.shp")
+sh <- readShapePoly("Census_Tract_Polygons2010/Census_tracts_2010.shp")
 plot(sh)
@@
 # Read in the demographic data and merge on Neighbourhood Id
-demo <- read.csv(file="Census_tracts_2010.csv", header=T)
+demo <- read.csv(file="Census_tracts_2010.csv", header = TRUE)
 sh2 <- merge(sh, demo, by='ID')
@@
 
-# Map 15
-```{r}
-j <- merge( b3, demo, by = "group"  )
+<!-- # Map 15 -->
+<!-- ```{r} -->
+<!-- sh2data <- sh2@data -->
+<!-- j <- merge( sh2data, demo, by.x= "GEOID10.x", by.y="GEOID10"  ) -->
 
-# Plot
-bMap12 <- ggmap(berkMap) +
-  geom_polygon(aes(x=long,y=lat, group=group, fill=TotalPop), data=j, color='white') +
-  scale_fill_gradient(low='white', high='orange') +
-  geom_point( data = mergedf, aes(x = long, y = lat), alpha = 0.2, size = 3, color = 'red' )
-bMap12
-```
+<!-- # Plot -->
+<!-- bMap12 <- ggmap(berkMap) + -->
+<!--   geom_polygon(aes(x=long,y=lat, group=group, fill=TotalPop), data=sh2, color='white') + -->
+<!--   scale_fill_gradient(low='white', high='orange') + -->
+<!--   geom_point( data = mergedf, aes(x = long, y = lat), alpha = 0.2, size = 3, color = 'red' ) -->
+<!-- bMap12 -->
+<!-- ``` -->

### 2016-08-02 final_paper/BDP_Paperdraft.Rmd
finalpaper

--- 2016-08-02 Stop_Data_Analysis.Rmd
@@
 ---
-title: "Stop Data Analysis"
-author: "Mengyu Li"
-date: "July 29, 2016"
+title: "BPD_draftpaper"
+author: "Amy Zhu; Mengyu Li; Rebecca Reus; Shangjun Jiang"
+date: "2 August 2016"
 output: html_document
+runtime: shiny
 ---
@@
 knitr::opts_chunk$set(echo = TRUE)
+library(sp)
+library(ggmap)
+library(tidyr)
+library(readr)
+library(plyr)
+library(dplyr)
+library(ggvis)
+library(tidyr)
+library(stringr)
+library(lubridate)
+library(ggplot2)
+library(readr)
+library(data.table)
+library(rworldmap)
+library(maps)
+library(mapdata)
+library(maptools)
+library(scales)
+library(RgoogleMaps)
+library(tmap)
+library(sp)
+library(rgdal)
+library(rgeos)
+library(RColorBrewer)
 ```
 
-```{r, echo=FALSE}
-library(readr)
-library(ggplot2)
-library(dplyr)
-library(lubridate)
-library(stringr)
-library(ggvis)
-library(tidyr)
-```
+## Introduction
+
+ For the final project, our group cleaned, explored, and analyzed four different data sets from the City of Berkeley that contained information on stops, calls for service, arrests, and jail bookings made by the the police department in 2016 (and 2015 as well for the stop data).
+Rather than limiting the project to one set of data, four sets were chosen in order to gain a more holistic and comprehensive understanding of the data.
+The additional variables included expanded the project’s capacity for manipulating data, examining relationships, and improving result reliability.
+
+#### Data Collection
+
+* [Stop Data (16,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Stop-Data/6e9j-pj9p)
+* [Arrest (200)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Arrests/xi7q-nji6)
+* [Jail Bookings (250)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Jail-Bookings/7ykt-c32j)
+* [Calls for Service (4,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Calls-for-Service/k2nh-s5h5)
+
+ With the variety of resources and information made available by collating multiple data sets, the objective for the project was to study differences in police activity (in terms of call requests and patrols) and intensity of assessed offenses based on time, race, gender, age, and mental health.
+As a challenge, another project target was to create a map applet depicting the density of police activity in Berkeley with an interactive component allowing the visitor to input an address and observe their proximity and observe the types of incidences that occurred most commonly in the area.
+
+	For the stop data, there were 16,255 incidents assessed by the Berkeley police.
+In each case, the call date and time, location, incident type, and disposition(s) were recorded.
+The cases typically had a six character disposition, with each character conveying race, gender, age, reason, enforcement, and car search,  respectively, for each subject involved in the incident, although there were additional dispositions, ranging from one to three characters,  that could be input and conveyed other information.
+In order to prepare the data for exploration, the cleaning process included changing the call date and time to the lubridate format, transforming the addresses into longitude and latitude referencing Google Maps, and separating the information on dispositions into separate row entries for each individual assessed in a case and splitting it further by isolating “other” and six character dispositions into different columns.
+To improve readability, the acronyms and abbreviations used in the data set were mutated to convey their meanings  (will do soon).
+
+	The calls for service data (4913 offenses) contained information on the case number, the offense type with a description of said offense, the date and start and end time for each case and the date of case entry into the database, and the location.
+Because the data set contained little to no information on the subjects apprehended in each case, it was largely used for mapping purposes.
+
+	The data for arrests (205), and jail bookings (223) contained similar information on case/arrest/booking number, date and time, type, and subject information (name, race, sex, D.O.B., age, height, weight, hair, eyes, and occupation) and statute information (type, description, agency, and disposition).
+Cleaning required the dates and times to be put into lubridate format, for the two data sets to be compiled in a reasonable, and for other needed adjustments.
+The same code created for the stop data would be used again to convert address information to longitudinal and latitudinal coordinates.
+
+## Data Analysis
 
@@
 ```
+
 ### About the Berkeley PD Stop Data
@@
 For the analysis in this part, only person information for all the people in each incident and the Call Time/Date of each incident are analyzed.
-
-### Introduction
-
-items                                | count    | percentage
----------------------------------    | -------- |-------------------
-total incidents reported             | 16255    |     
-dispositions of type one             | 13016    | 80.07%
-dispositions of type two             | 2346     | 14.43%
-dispositions of type three           | 845      | 5.20%
-blank dispositions                   | 48       | 0.30%
-incidents with person information    | 13861    | 85.27%
-  incidents with 1 person            | 13494    | 97.352%
-  incidents with 2 people            | 322      | 2.323%              
-  incidents with 3 people            | 33       | 0.238%
-  incidents with 4 people            | 6        | 0.043%
-  incidents with 5 people            | 6        | 0.043%
-total personnel reported             | 14291    |
-
-(1) From 01/26/2015 to 04/30/2016, the Berkeley Police Department(or BPD) stopped for 16225 incidents.
-80.07% of all stops report only personnel information, 14.43% report only additional information, 5.20% report both and 0.30% report none.
-
-(2) Among all the stops, 13851 incidents recorded 14291 personnel dispositions in total.
-Of the incidents with the personal dispositions, 97.352% involves only one person, 2.648% involves more than two people.
-There are at most 5 people involved in one incident.
 
@@
 
-### Analysis of Day
-####Picture: count of person recorded in each day in a week
-```{r, echo=FALSE}
-stop_day <- stop_by %>%
-  ggvis(x = ~Day, fill = ~Race) %>%
-  layer_bars() %>%
-  add_tooltip(function(stop_by) stop_by$Race)
-stop_day
-```
-
-```{r, eval=FALSE, echo=FALSE}
-stop_day_total_table <- stop_by %>%
-    group_by(Day) %>%
-    summarise(totaLCount = n())
-stop_day_race_table <- stop_by %>%
-    group_by(Day, Race) %>%
-    summarise(count = n())
-stop_day_table <- left_join(stop_day_race_table, stop_day_total_table, by = "Day")
-stop_day_table <- stop_day_table %>%
-    spread(key = Race, value = count)
-stop_day_table
-```
-####Picture: count of person recorded in each day in a week
-Day   | Count | A    | B    | H    | O    | W
-----  |------ |------|------|------|------|------
-1     | 1692  | 116  | 653  | 221  | 159  | 543
-2     | 1665  | 149  | 543  | 182  | 166  | 625
-3     | 2344  | 189  | 645  | 231  | 236  | 1043
-4     | 2287  | 206  | 616  | 251  | 244  | 970
-5     | 1941  | 152  | 592  | 223  | 168  | 806
-6     | 2231  | 186  | 764  | 268  | 203  | 810
-7     | 2131  | 143  | 823  | 300  | 208  | 657
-
-(1) Among all the day in a week, the number of incidents happened on Sunday and Monday are much less than those happened through Tuesday and Saturday.
-
-(2) The count of records of Asian people is much less than that of any other race of people in each day and the whole week.
-
-(3) Asian people are less liable to commit an incident on Sunday, and more liable to commit an incident on Wednesday.
-
-(4) Black and Hispanic people are less liable to commit an incident on Monday, and more liable to commit an incident on Saturday.
-
-(5) White people are less liable to commit an incident on Sunday, and more liable to commit an incident on Tuesday and Wednesday, which is similar to Asian people.
-
-### Analysis of incidents of each hour in a day
+## Analysis of incidents of each hour in a day
 #### Picture: count of person recorded in each hour in a day
@@
 ```
+
 Note: you can change the value of the input slider to see the number of incidents happened in specific time span of the day.
+
 ```{r, eval = FALSE, echo=FALSE}
@@
 
-### Analysis of prefence of Berkeley Police Department arresting people
+### Analysis of preference of Berkeley Police Department arresting people
 #### Picture: Probability of arrested by BPD of a specific race for a specific reason
@@
 #### Table: Probability of arrested by BPD of a specific gender range for a specific reason
-P     | I      | K     | R     | T     | W    
-----  |------- |-------|-------|-------|-------
-F     | 9.05%  | 50.00%| 5.00% | 1.10% | 28.57%  
-M     | 8.26%  | 20.11%| 5.42% | 1.85% | 37.04%
+P     | I (Investigation) |K (Probation/ Parole)| R (Reasonable Suspicion)|T (Traffic)| W (Wanted) 
+----  |------------------ |---------------------|-------------------------|-----------|-------
+F     | 9.05%             | 50.00%              | 5.00%                   | 1.10%     | 28.57%  
+M     | 8.26%             | 20.11%              | 5.42%                   | 1.85%     | 37.04%
 
@@
 (2) Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole of female is 50.00%, which is much higher than that of male, which is 20.11%.
+
+```{r, echo=FALSE}
+## Read in the Data:
+#finaldf <- readRDS( 'StopData_finaldf.rds' )
+mergedf <- readRDS( 'StopData_merged.rds' )
+```
+
+```{r, echo=FALSE}
+## Location info:
+latmax <- max( mergedf$lat, na.rm = TRUE ) 
+latmin <- min( mergedf$lat, na.rm = TRUE )
+lonmax <- max( mergedf$long, na.rm = TRUE ) 
+lonmin <- min( mergedf$long, na.rm = TRUE )
+latvals <- c( latmin, latmax )
+lonvals <- c( lonmin, lonmax )
+```
+
+```{r, echo=FALSE}
+## Make the variables factors:
+mergedf$CarSearch <- factor(mergedf$CarSearch)
+levels(mergedf$CarSearch) <- c("No Search", "Search")
+
+mergedf$Race <- factor(mergedf$Race)
+levels(mergedf$Race) <- c("Asian", "Black", "Hispanic", "Other", "White")
+
+mergedf$Enforcement <- factor(mergedf$Enforcement)
+levels(mergedf$Enforcement) <- c("Arrest", "Citation", "Other", "Warning")
+
+mergedf$Reason <- factor(mergedf$Reason)
+levels(mergedf$Reason) <- c("Other", "Investigation", "Probation/Parole", "Reasonable Suscipcion", "Traffic", "Wanted")
+
+mergedf <- mergedf %>%
+  mutate(Arrest = ifelse( 
+    as.character(Enforcement) == "Arrest", 
+    "Arrested",
+    "Not Arrested") ) %>%
+  mutate(Arrest = factor(Arrest))
+
+v <- factor(mergedf$Other)
+levels(v)
+
+mergedf <- mergedf %>%
+  mutate(Emergency.Psych.Eval = ifelse( str_detect(as.character(Other), "MH"), 
+                                        yes = "Yes",
+                                        no = "No") )
+
+mergedf$Emergency.Psych.Eval <- factor(mergedf$Emergency.Psych.Eval)
+
+berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14)
+
+```
+
+## MAP 1:All BPD Stops Density, 2015-2016
+```{r, fig.width=10, fig.asp=0.8, echo=FALSE}
+# a contour plot
+ggmap(berkMap) +
+  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
+    size = 2, bins = 5, data = mergedf, geom = "polygon") +
+  scale_fill_gradient(low = "black", high = "red") +
+    theme (panel.grid.major = element_blank (), # remove major grid
+		       panel.grid.minor = element_blank (),  # remove minor grid
+		       axis.text = element_blank (), 
+		       axis.title = element_blank (),
+	         axis.ticks = element_blank ()) + 
+  ggtitle("All BPD Stops Density, 2015-2016") +
+  labs(alpha = element_blank())+
+  guides(alpha = FALSE)
+
+```
+
+## MAP 2: BPD Stop Contour Map of Berkeley by Race
+```{r, fig.height = 5, fig.width = 10, echo=FALSE}
+ggmap(berkMap) +
+  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
+    bins = I(5), geom = "polygon", data = mergedf ) +
+  scale_fill_gradient2( "StopDensity",
+    low = "white", mid = "orange", high = "red", midpoint = 100) +
+  labs(x = "Longitude", y = "Latitude") + facet_wrap( ~ Race ) +
+  scale_alpha(range = c(.2, .55), guide = FALSE) +
+  ggtitle("BPD Stop Contour Map of Berkeley by Race") +
+  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
+```
+
+## MAP 3: BPD Stop Contour Map of Berkeley by Age Range
+```{r, fig.height = 8, fig.width = 10, echo=FALSE}
+ggmap(berkMap) +
+  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
+    bins = I(5), geom = "polygon", data = mergedf ) +
+  scale_fill_gradient2( "Stop Density",
+    low = "white", mid = "orange", high = "red", midpoint = 25) +
+  labs(x = "Longitude", y = "Latitude") + facet_wrap(~ AgeRange) +
+  scale_alpha(range = c(.2, .55), guide = FALSE) +
+  ggtitle("BPD Stop Contour Map of Berkeley by Age Range") +
+  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
+```
+
+We found that the

### 2016-08-02 final_paper/Shinyapp.Rmd
finalpaper

---
title: "SHINY APP"
author: "Shangjun Jiang"
date: "2 August 2016"
output: html_document
runtime: shiny
---

```{r}
setwd("~/Desktop/Stats R/PROJECT/finalproject_angry_ladies/final_paper")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sp)
library(ggmap)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(data.table)
library(rworldmap)
library(maps)
library(mapdata)
library(maptools)
library(scales)
library(RgoogleMaps)
library(tmap)
library(sp)
library(rgdal)
library(rgeos)
library(RColorBrewer)

```

```{r}
## Read in the Data:
#finaldf <- readRDS( 'StopData_finaldf.rds' )
mergedf <- readRDS( 'StopData_merged.rds' )
```

```{r}
## Location info:
latmax <- max( mergedf$lat, na.rm = TRUE ) 
latmin <- min( mergedf$lat, na.rm = TRUE )
lonmax <- max( mergedf$long, na.rm = TRUE ) 
lonmin <- min( mergedf$long, na.rm = TRUE )
latvals <- c( latmin, latmax )
lonvals <- c( lonmin, lonmax )
```

```{r}
## Make the variables factors:
mergedf$CarSearch <- factor(mergedf$CarSearch)
levels(mergedf$CarSearch) <- c("No Search", "Search")

mergedf$Race <- factor(mergedf$Race)
levels(mergedf$Race) <- c("Asian", "Black", "Hispanic", "Other", "White")

mergedf$Enforcement <- factor(mergedf$Enforcement)
levels(mergedf$Enforcement) <- c("Arrest", "Citation", "Other", "Warning")

mergedf$Reason <- factor(mergedf$Reason)
levels(mergedf$Reason) <- c("Other", "Investigation", "Probation/Parole", "Reasonable Suscipcion", "Traffic", "Wanted")

mergedf <- mergedf %>%
  mutate(Arrest = ifelse( 
    as.character(Enforcement) == "Arrest", 
    "Arrested",
    "Not Arrested") ) %>%
  mutate(Arrest = factor(Arrest))

v <- factor(mergedf$Other)
levels(v)

mergedf <- mergedf %>%
  mutate(Emergency.Psych.Eval = ifelse( str_detect(as.character(Other), "MH"), 
                                        yes = "Yes",
                                        no = "No") )

mergedf$Emergency.Psych.Eval <- factor(mergedf$Emergency.Psych.Eval)

?write.csv

write.csv(mergedf, "mergedf.csv", row.names = F)

berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14)
```

## MAP 4:
```{r}

ggmap(berkMap) +
  geom_point(aes(x = long, y = lat, colour = Reason), data = mergedf, alpha = 0.7, size = 3) +
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + ggtitle("All BPD Stops, by Reason 2015-2016") +
    facet_wrap( ~ Reason , ncol = 2)
```

### 2016-08-03 Rebeccas_location_Code2/StopData_Maps.Rmd
draft paper

@@
 ```{r}
-locationCensusFiles <- "C:/Users/Rebecca/Dropbox/School/2016_su_School/stat133/finalproject_angry_ladies/Rebeccas_location_Code2/Census_Block_Group_Polygons2010"
-blocks <- readOGR(locationCensusFiles,"Census_blockgroups_2010", verbose = TRUE)
-summary(blocks)
-print(proj4string(blocks))
+#locationCensusFiles <- "C:/Users/Rebecca/Dropbox/School/2016_su_School/stat133/finalproject_angry_ladies/Rebeccas_location_Code2/Census_Block_Group_Polygons2010"
+blocks <- readOGR("Census_Block_Group_Polygons2010","Census_blockgroups_2010", verbose = TRUE)
+
 blocks.polygons <- blocks@polygons
@@
 locationCensusFiles <- "C:/Users/Rebecca/Dropbox/School/2016_su_School/stat133/finalproject_angry_ladies/Rebeccas_location_Code2/Census_Tract_Polygons2010"
+
+locationCensusFiles <- "Census_Tract_Polygons2010"
+blocks <- readOGR("Census_Tract_Polygons2010","Census_tracts_2010", verbose = TRUE)
+
+locationCensusFiles <- "Census_Tract_Polygons2010"
 blocks <- readOGR(locationCensusFiles,"Census_tracts_2010", verbose = TRUE)
@@
 sh <- readShapePoly("Census_Tract_Polygons2010/Census_tracts_2010.shp")
-plot(sh)
+#plot(sh)
 sh@data$ID <- as.numeric(sh@data$ID)
@@
 pop <- sh2@data$TotalPop.x
+#str(sh2)
+#sh2[1]
+#sh2@data$TotalPop.x
+#pop <- c(2925,3144,4196,4658,3558,3387,2007,3198,3563,1992,1756,3632,3118,3844,3716,3647,3640,8368,3633,5659,2642,2794,1593,2685,1552,2020,2172,3964,4396,4885,1215,4336,4685)
+
 cols <- (pop - min(pop))/diff(range(pop))*127+1

### 2016-08-03 final_paper/draft.Rmd
draft paper

--- 2016-08-02 final_paper/BDP_Paperdraft.Rmd
@@
 ---
-title: "BPD_draftpaper"
+title: "BDP draft paper"
 author: "Amy Zhu; Mengyu Li; Rebecca Reus; Shangjun Jiang"
@@
 output: html_document
-runtime: shiny
 ---
@@
 library(RColorBrewer)
+library(grid)
+library(gridExtra)
 ```
@@
 * [Calls for Service (4,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Calls-for-Service/k2nh-s5h5)
+* [Berkeley Census Data](https://data.cityofberkeley.info/)
 
@@
 In order to prepare the data for exploration, the cleaning process included changing the call date and time to the lubridate format, transforming the addresses into longitude and latitude referencing Google Maps, and separating the information on dispositions into separate row entries for each individual assessed in a case and splitting it further by isolating “other” and six character dispositions into different columns.
-To improve readability, the acronyms and abbreviations used in the data set were mutated to convey their meanings  (will do soon).
 
@@
 # Read data
+
+############################## stop data ############################## 
 stop_by <- read_csv("cleaned_stop_data.csv")
@@
   mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))
+
+############################## location data ############################## 
+mergedf <- readRDS( 'StopData_merged.rds' )
+
+## Location info:
+latmax <- max(mergedf$lat, na.rm = TRUE) 
+latmin <- min(mergedf$lat, na.rm = TRUE)
+lonmax <- max(mergedf$long, na.rm = TRUE) 
+lonmin <- min(mergedf$long, na.rm = TRUE)
+latvals <- c(latmin, latmax)
+lonvals <- c(lonmin, lonmax)
+
+## Make the variables factors:
+mergedf$CarSearch <- factor(mergedf$CarSearch)
+levels(mergedf$CarSearch) <- c("No Search", "Search")
+
+mergedf$Race <- factor(mergedf$Race)
+levels(mergedf$Race) <- c("Asian", "Black", "Hispanic", "Other", "White")
+
+mergedf$Enforcement <- factor(mergedf$Enforcement)
+levels(mergedf$Enforcement) <- c("Arrest", "Citation", "Other", "Warning")
+
+mergedf$Reason <- factor(mergedf$Reason)
+levels(mergedf$Reason) <- c("Other", "Investigation", "Probation/Parole", "Reasonable Suscipcion", "Traffic", "Wanted")
+
+mergedf <- mergedf %>%
+  mutate(Arrest = ifelse( 
+    as.character(Enforcement) == "Arrest", 
+    "Arrested",
+    "Not Arrested") ) %>%
+  mutate(Arrest = factor(Arrest))
+
+v <- factor(mergedf$Other)
+levels(v)
+
+mergedf <- mergedf %>%
+  mutate(Emergency.Psych.Eval = ifelse( str_detect(as.character(Other), "MH"), 
+                                        yes = "Yes",
+                                        no = "No") )
+
+mergedf$Emergency.Psych.Eval <- factor(mergedf$Emergency.Psych.Eval)
+
+### BerkeleyMap 
+berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14)
+
+############################## census data ##############################
+
+blocks <- readOGR("Census_Tract_Polygons2010","Census_tracts_2010", verbose = TRUE)
+
+locationCensusFiles <- "Census_Tract_Polygons2010"
+blocks <- readOGR(locationCensusFiles,"Census_tracts_2010", verbose = TRUE)
+b2 <- spTransform(blocks, CRS("+proj=longlat +datum=WGS84"))
+b3 <- fortify(b2)
+
+############################## arrest data ############################## 
+arrest <- read.csv("Berkeley_PD_Log_-_Arrests.csv")
+
+############################## jail data ############################## 
+jail <- read.csv("Berkeley_PD_Log_-_Jail_Bookings.csv")
+
+############################## multiplot function ############################## 
+multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
+  library(grid)
+
+  # Make a list from the ... arguments and plotlist
+  plots <- c(list(...), plotlist)
+
+  numPlots = length(plots)
+
+  # If layout is NULL, then use 'cols' to determine layout
+  if (is.null(layout)) {
+    # Make the panel
+    # ncol: Number of columns of plots
+    # nrow: Number of rows needed, calculated from # of cols
+    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
+                    ncol = cols, nrow = ceiling(numPlots/cols))
+  }
+
+ if (numPlots==1) {
+    print(plots[[1]])
+
+  } else {
+    # Set up the page
+    grid.newpage()
+    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
+
+    # Make each plot, in the correct location
+    for (i in 1:numPlots) {
+      # Get the i,j matrix positions of the regions that contain this subplot
+      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
+
+      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
+                                      layout.pos.col = matchidx$col))
+    }
+  }
+}
 ```
 
-### About the Berkeley PD Stop Data
+We found the census data from the Berkeley open data website and created a headmap of the Berkeley population.
+Similarly, we created a heat map based on the stop data.
+Interestingly, we found that the most dense place (around north Berkeley) is a relatively safe place.
+The place where people are more likely to be stopped (downtown berkeley area) is less dense.
+Since Downtown Berkeley area is a transportation hub, numerous people come and go around this area.
+Though there are more people living north Berkeley, mainly residents will visit the area.
+Therefore, the most likely place to get stopped is not the most dense living place.
+
+### Heat Map 1: Berkeley Census data.
+```{r, fig.width=10, fig.asp=0.8, echo=FALSE}
+ggmap(berkMap) +
+  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
+    size = 2, bins = 5, data = b3, geom = "polygon") +
+  scale_fill_gradient(low = "black", high = "red") +
+    theme (panel.grid.major = element_blank (), # remove major grid
+		       panel.grid.minor = element_blank ()  # remove minor grid
+		       ) + 
+  ggtitle("Population Density") +
+  labs(alpha = element_blank())+
+  guides(alpha = FALSE)
+```
+
+### Heat Map 2: All BPD Stops Density, 2015-2016
+```{r, fig.width=10, fig.asp=0.8, echo=FALSE}
+# a contour plot
+ggmap(berkMap) +
+  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
+    size = 2, bins = 5, data = mergedf, geom = "polygon") +
+  scale_fill_gradient(low = "black", high = "red") +
+    theme (panel.grid.major = element_blank (), # remove major grid
+		       panel.grid.minor = element_blank ()  # remove minor grid
+		       )+ 
+  ggtitle("All BPD Stops Density, 2015-2016") +
+  labs(alpha = element_blank())+
+  guides(alpha = FALSE)
+
+```
+
+Then we explored the stop data by age range (0-18; 18-29; 30-39; and 40+) and data.
+Similar to the all stop data density, the area that is most likely to be stopped is the same: Downtown Berkeley.
+The differences in age range and race don't play a crucial role in the possibility of being stopped in Berkeley.
+These two analysis confirmed our explanations for the all stop data heat map.
+
+### Heat Map 3: BPD Stop Contour Map of Berkeley by Age Range
+```{r, fig.height = 8, fig.width = 10, echo=FALSE}
+ggmap(berkMap) +
+  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
+    bins = I(5), geom = "polygon", data = mergedf ) +
+  scale_fill_gradient2( "Stop Density",
+    low = "white", mid = "orange", high = "red", midpoint = 25) +
+  labs(x = "Longitude", y = "Latitude") + facet_wrap(~ AgeRange) +
+  scale_alpha(range = c(.2, .55), guide = FALSE) +
+  ggtitle("BPD Stop Contour Map of Berkeley by Age Range") +
+  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
+```
+
+Note: 1 refers to age 0-18; 2 refers to age 18-29; 3 refers to age 30-39; and 4 refers to age 40+.
+
+## Heat Map 4: BPD Stop Contour Map of Berkeley by Race
+```{r, fig.height = 5, fig.width = 10, echo=FALSE}
+ggmap(berkMap) +
+  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
+    bins = I(5), geom = "polygon", data = mergedf ) +
+  scale_fill_gradient2( "StopDensity",
+    low = "white", mid = "orange", high = "red", midpoint = 100) +
+  labs(x = "Longitude", y = "Latitude") + facet_wrap( ~ Race ) +
+  scale_alpha(range = c(.2, .55), guide = FALSE) +
+  ggtitle("BPD Stop Contour Map of Berkeley by Race") +
+  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
+```
 
 The Berkeley PD stop data reflects accidents that have been visited by the Berkeley Police Department.
-Raw data set is downloaded from City of Berkeley Open Data website and provided by Berkeley Police Department.
-
-Analysis in this part focuses on the information provided by the dispositions variable.
-The dispositions variable can be divided into three types.
-The first type only includes more than one six-character person message.
-Each character in each person message represents the race, gender and age range of the person, reason and enforcement of the stop, and the car search information during the stop.
-The second type of the dispositions includes only additional dispositions, such as P for primary cases.
-The third type of dispositions includes both person information and additional dispositions.
-
-For the analysis in this part, only person information for all the people in each incident and the Call Time/Date of each incident are analyzed.
+After the initial exploration of the Berkeley stop data, we focus on on the information provided by the dispositions variable.
+The dispositions variable provides information on race, gender and age range, stop reason and enforcement of the stop, the car search information during the stop, and additional dispositions.
 
@@
 
-```{r, echo=FALSE, eval=FALSE}
-stop_race_total_table <- stop_by %>%
-    group_by(Race) %>%
-    summarise(totaLCount = n())
-stop_race_agerange_table <- stop_by %>%
-    group_by(Race, AgeRange) %>%
-    summarise(count = n())
-stop_race_table <- left_join(stop_race_agerange_table, stop_race_total_table, by = "Race")
-stop_race_table <- stop_race_table %>%
-    mutate(percentage = count/totaLCount) %>%
-    select(-count)
-stop_race_table <- stop_race_table %>%
-    spread(key = AgeRange, value = percentage)
-stop_race_table
-```
-
 ####Table: count of person recorded in each race
@@
    Race     | Count | 0-18 | 18-29 | 30-39 | 40+
-----  |------ |------|-------|-------|-------
-A     | 1141  | 8.50%| 43.47%| 21.03%| 26.99%
-B     | 4636  | 2.29%| 34.97%| 25.09%| 37.66%
-H     | 1676  | 3.22%| 46.12%| 29.18%| 21.47%
-O     | 1384  | 2.67%| 43.28%| 29.17%| 26.87%
-W     | 5454  | 1.72%| 29.45%| 25.39%| 43.44%
+----------  |------ |------|-------|-------|-------
+A (Asian)   | 1141  | 8.50%| 43.47%| 21.03%| 26.99%
+B (Black)   | 4636  | 2.29%| 34.97%| 25.09%| 37.66%
+H (Hispanic)| 1676  | 3.22%| 46.12%| 29.18%| 21.47%
+O (Other)   | 1384  | 2.67%| 43.28%| 29.17%| 26.87%
+W (White)   | 5454  | 1.72%| 29.45%| 25.39%| 43.44%
 
-#### Picture: count of person recorded of each race in each hour
+(1) Among the 14291 personnel information recorded, white people is the number one population group of being stopped by Berkeley Police.
+38.16% of all records, namely 5454 people are white.
+black people contributes the second largest percentage.
+32.44% of all records, namely 4636 people are black.
+
+(2) The percentage of people aged from 0 to 18 of each race is less than 9%.
+People aged from 0 to 18 contributes the smallest percentage in each race.
+These teengers are looked after by their parents most of the time.
+
+(3) Among white people and black people recorded by the BPD, people who are over 40 are the number one group that are being stopped, which are 43.44% and 37.66% respectively.
+
+(4) Among Asian, Hispanic and other people recorded by the BPD, people aged from 18 to 29 contribute the largest percentage, which are 43.37%, 46.12% and 43.28%, respectively.
+
+Then we further analyzed the stop data with and without car searches by race.
+Since white people are the largest group to be stopped, we suspect that their possibility of car search will also be the highest.
+However, black people seem to be more likely to be car searched.
+
+### MAP: BPD All Stops w/ Car Searches by Race 2015-2016
+```{r, echo=FALSE}
+ggmap(berkMap) +
+  geom_point(aes(x = long, y = lat, size = CarSearch, color = Race), data = mergedf, alpha = 0.2) + 
+  scale_size_discrete( "Car Search", labels = c("No Search", "Search") ) +
+  theme ( 
+		panel.grid.major = element_blank (), # remove major grid
+		panel.grid.minor = element_blank (),  # remove minor grid
+		axis.text = element_blank (), 
+		axis.title = element_blank (),
+		axis.ticks = element_blank ()
+		) + 
+  ggtitle("BPD All Stops w/ Car Searches by Race 2015-2016") +
+  facet_wrap(~ Race, ncol = 3) 
+```
+
+Even when we look at the arrest data, it is black people who are more likely to get arrest after being stopped.
+
+```{r, echo=FALSE}
+df <- subset(mergedf, as.character(Enforcement) == "Arrest")
+
+ggmap(berkMap) +
+  geom_point(aes(x = long, y = lat, colour = Race), data = df, alpha = 0.5, size = 5) + 
+  theme ( 
+		panel.grid.major = element_blank (), # remove major grid
+		panel.grid.minor = element_blank (),  # remove minor grid
+		axis.text = element_blank (), 
+		axis.title = element_blank (),
+		axis.ticks = element_blank ()
+		) + 
+  ggtitle("BPD Stops to Arrests, by Race 2015-2016") +
+  facet_wrap(~ Race) 
+```
+
+### Analysis of stop reasons 
+#### Map: All stop data by reasons.
+```{r, fig.height = 5, fig.width = 10, echo=FALSE}
+ggmap(berkMap) +
+  geom_point(aes(x = long, y = lat, colour = Reason), data = mergedf, alpha = 0.7, size = 3) +
+  theme ( 
+		panel.grid.major = element_blank (), # remove major grid
+		panel.grid.minor = element_blank (),  # remove minor grid
+		axis.text = element_blank (), 
+		axis.title = element_blank (),
+		axis.ticks = element_blank ()
+		) + ggtitle("All BPD Stops, by Reason 2015-2016") +
+    facet_wrap( ~ Reason , ncol = 2)
+```
+
+From the analysis of the stop reasons, we found that the number one reason that people are stopped is traffic reason.
+Then we explored traffic reasons within race.
+As shown, black people are the highest to be stopped and arrested.
+
+#### Map: stop by only traffic reasons
+```{r, fig.height = 5, fig.width = 10, echo=FALSE}
+df <- subset(mergedf, as.character(Reason) == "Traffic")
+
+ggmap(berkMap) +
+  geom_point(aes(x = long, y = lat, colour = Race, size= Arrest), data = df, alpha = 0.5) + 
+  scale_size_discrete( range = c(5, 2) ) +
+  theme ( 
+		panel.grid.major = element_blank (), # remove major grid
+		panel.grid.minor = element_blank (),  # remove minor grid
+		axis.text = element_blank (), 
+		axis.title = element_blank (),
+		axis.ticks = element_blank ()
+		) + 
+  ggtitle("BPD Traffic Stops Only, by Race 2015-2016") +
+  facet_grid(~ Race)
+```
+
+Such contrast of being stopped, being car searched, and being arrested is further demonstrated by the arrest and jail data.
+
+```{r, echo=FALSE}
+p1 <- ggplot(na.omit(arrest))+
+  geom_bar(aes(x=Age, 
+               fill = Race))+
+  facet_wrap(~Race)+
+ labs(x= "Age",
+       y = "Number of people",
+       title = "Arrest Data")
+
+p2 <- ggplot(na.omit(jail))+
+  geom_bar(aes(x=Age,
+               fill = Race))+
+  facet_wrap(~Race)
+
+multiplot(p1, p2,cols=2)
+
+```
+
+As we can clearly see from both graphs, black people are the largest group to be arrested and put into jail.
+
+Then we looked at the counts of people recorded of different races in every hour.Black people contribute a especially high percentage of the incidents at night, while the white people contribute a especially high percentage of the incidents at the noon.
+
 ```{r, echo=FALSE}
@@
   labs(y = "Count") +
-  facet_wrap(~Race)
+  facet_wrap(~Race)+
+  labs(title = "Counts of people recorded of each race in every hour")
 stop_time_percentage
@@
 
-(1) Among the 14291 personnel information recorded, white people contributes the largest percentage.
-38.16% of all records, namely 5454 people are white.
-black people contributes the second largest percentage.
-32.44% of all records, namely 4636 people are black.
+Such analysis made us further digged into the date information.
 
-(2) The percentage of people aged from 0 to 18 of each race is less than 9%. People aged from 0 to 18 contributes the smallest percentage in each race, which makes sense in daily life.
+### Analysis of Day
+```{r, echo=FALSE, eval = FALSE}
+stop_day <- stop_by %>%
+  ggvis(x = ~Day, fill = ~Race) %>%
+  layer_bars() %>%
+  add_tooltip(function(stop_by) stop_by$Race)+
+  labs(title = "Counts of people recorded in every day in a week")
+stop_day
+```
 
-(3) Among white people and black people recorded by the BPD, people aged greater than 40  contribute the largest percentage, which are 43.44% and 37.66% respectively.
+####Picture: counts of people recorded in every day in a week
+Day   | Count | A (Asian)| B (Black)| H (Hispanic) | O (Other)| W (White)
+----  |------ |----------|----------|--------------|----------|------
+1     | 1692  | 116      | 653      | 221          | 159      | 543
+2     | 1665  | 149      | 543      | 182          | 166      | 625
+3     | 2344  | 189      | 645      | 231          | 236      | 1043
+4     | 2287  | 206      | 616      | 251          | 244      | 970
+5     | 1941  | 152      | 592      | 223          | 168      | 806
+6     | 2231  | 186      | 764      | 268          | 203      | 810
+7     | 2131  | 143      | 823      | 300          | 208      | 657
 
-(4) Among Asian, Hispanic and other people recorded by the BPD, people aged from 18 to 29 contribute the largest percentage, which are 43.37%, 46.12% and 43.28%, respectively.
+(1) Among all the day in a week, the number of incidents happened on Sunday and Monday are much less than those happened through Tuesday and Saturday.
 
-(5) Black people contribute a especially high percentage of the incidents at night, while the white people contribute a especially high percentage of the incidents at the noon.
+(2) The count of records of Asian people is much less than that of any other race of people in each day and the whole week.
+
+(3) Asian people are less liable to commit an incident on Sunday, and more liable to commit an incident on Wednesday.
+
+(4) Black and Hispanic people are less liable to commit an incident on Monday, and more liable to commit an incident on Saturday.
+
+(5) White people are less liable to commit an incident on Sunday, and more liable to commit an incident on Tuesday and Wednesday, which is similar to Asian people.
 
@@
 Note: you can change the value of the input slider to see the number of incidents happened in specific time span of the day.
-
-```{r, eval = FALSE, echo=FALSE}
-stop_hour <- stop_by %>%
-  group_by(Hour) %>%
-  summarise(count = n())
-stop_hour
-```
 
@@
 #### Table: Probability of arrested by BPD of a specific age range for a specific reason
-P     | I     | K     | R     | T     | W     
-------|-------|-------|-------|-------|-------
-0-18  | 17.95%| 67.57%| 13.95%| 3.24% | 50.00%  
-18-29 | 8.45% | 12.50%| 3.87% | 1.82% | 32.35%  
-30-39 | 6.32% | 25.00%| 5.04% | 1.49% | 26.92%  
-40+   | 7.87% | 23.86%| 6.13% | 1.41% | 31.82%
+P     | I (Investigation) |K (Probation/ Parole)| R (Reasonable Suspicion)|T (Traffic)| W (Wanted) 
+----  |------------------ |---------------------|-------------------------|-----------|-------
+0-18  | 17.95%            | 67.57%              | 13.95%                  | 3.24%     | 50.00%  
+18-29 | 8.45%             | 12.50%              | 3.87%                   | 1.82%     | 32.35%  
+30-39 | 6.32%             | 25.00%              | 5.04%                   | 1.49%     | 26.92%  
+40+   | 7.87%             | 23.86%              | 6.13%                   | 1.41%     | 31.82%
 
@@
 
-```{r, echo=FALSE}
-## Read in the Data:
-#finaldf <- readRDS( 'StopData_finaldf.rds' )
-mergedf <- readRDS( 'StopData_merged.rds' )
-```
+## Conclusion
 
-```{r, echo=FALSE}
-## Location info:
-latmax <- max( mergedf$lat, na.rm = TRUE ) 
-latmin <- min( mergedf$lat, na.rm = TRUE )
-lonmax <- max( mergedf$long, na.rm = TRUE ) 
-lonmin <- min( mergedf$long, na.rm = TRUE )
-latvals <- c( latmin, latmax )
-lonvals <- c( lonmin, lonmax )
-```
-
-```{r, echo=FALSE}
-## Make the variables factors:
-mergedf$CarSearch <- factor(mergedf$CarSearch)
-levels(mergedf$CarSearch) <- c("No Search", "Search")
-
-mergedf$Race <- factor(mergedf$Race)
-levels(mergedf$Race) <- c("Asian", "Black", "Hispanic", "Other", "White")
-
-mergedf$Enforcement <- factor(mergedf$Enforcement)
-levels(mergedf$Enforcement) <- c("Arrest", "Citation", "Other", "Warning")
-
-mergedf$Reason <- factor(mergedf$Reason)
-levels(mergedf$Reason) <- c("Other", "Investigation", "Probation/Parole", "Reasonable Suscipcion", "Traffic", "Wanted")
-
-mergedf <- mergedf %>%
-  mutate(Arrest = ifelse( 
-    as.character(Enforcement) == "Arrest", 
-    "Arrested",
-    "Not Arrested") ) %>%
-  mutate(Arrest = factor(Arrest))
-
-v <- factor(mergedf$Other)
-levels(v)
-
-mergedf <- mergedf %>%
-  mutate(Emergency.Psych.Eval = ifelse( str_detect(as.character(Other), "MH"), 
-                                        yes = "Yes",
-                                        no = "No") )
-
-mergedf$Emergency.Psych.Eval <- factor(mergedf$Emergency.Psych.Eval)
-
-berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14)
-
-```
-
-## MAP 1:All BPD Stops Density, 2015-2016
-```{r, fig.width=10, fig.asp=0.8, echo=FALSE}
-# a contour plot
-ggmap(berkMap) +
-  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
-    size = 2, bins = 5, data = mergedf, geom = "polygon") +
-  scale_fill_gradient(low = "black", high = "red") +
-    theme (panel.grid.major = element_blank (), # remove major grid
-		       panel.grid.minor = element_blank (),  # remove minor grid
-		       axis.text = element_blank (), 
-		       axis.title = element_blank (),
-	         axis.ticks = element_blank ()) + 
-  ggtitle("All BPD Stops Density, 2015-2016") +
-  labs(alpha = element_blank())+
-  guides(alpha = FALSE)
-
-```
-
-## MAP 2: BPD Stop Contour Map of Berkeley by Race
-```{r, fig.height = 5, fig.width = 10, echo=FALSE}
-ggmap(berkMap) +
-  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
-    bins = I(5), geom = "polygon", data = mergedf ) +
-  scale_fill_gradient2( "StopDensity",
-    low = "white", mid = "orange", high = "red", midpoint = 100) +
-  labs(x = "Longitude", y = "Latitude") + facet_wrap( ~ Race ) +
-  scale_alpha(range = c(.2, .55), guide = FALSE) +
-  ggtitle("BPD Stop Contour Map of Berkeley by Race") +
-  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
-```
-
-## MAP 3: BPD Stop Contour Map of Berkeley by Age Range
-```{r, fig.height = 8, fig.width = 10, echo=FALSE}
-ggmap(berkMap) +
-  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
-    bins = I(5), geom = "polygon", data = mergedf ) +
-  scale_fill_gradient2( "Stop Density",
-    low = "white", mid = "orange", high = "red", midpoint = 25) +
-  labs(x = "Longitude", y = "Latitude") + facet_wrap(~ AgeRange) +
-  scale_alpha(range = c(.2, .55), guide = FALSE) +
-  ggtitle("BPD Stop Contour Map of Berkeley by Age Range") +
-  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
-```
-
-We found that the
+1.
+From analyzing the relationship between population density and stop data density, we found that he place that are the most likely to be stopped is not the place where most people live, but the transportation hub.
+2.
+From various analysis of the data by race, such as the stop time by race, car search by race, we found that black people are the number one to be arrested and even put into jail.
+Surprisingly, they are not the number one to be stopped.
+White people are the largest group to be stopped.
+3. During daytime, young people and white people are more likely to be stopped. During nightime, people over 40 and black people are more likely to be stopped by police. 
+4. Asian and female are more likely to be arrested by the BDP for probation/parole.

### 2016-08-03 final_paper/draft.Rmd
draft paper

@@
  For the final project, our group cleaned, explored, and analyzed four different data sets from the City of Berkeley that contained information on stops, calls for service, arrests, and jail bookings made by the the police department in 2016 (and 2015 as well for the stop data).
-Rather than limiting the project to one set of data, four sets were chosen in order to gain a more holistic and comprehensive understanding of the data.
+Rather than limiting the project to one set of data, three sets were chosen in order to gain a more holistic and comprehensive understanding of the data.
 The additional variables included expanded the project’s capacity for manipulating data, examining relationships, and improving result reliability.
@@
 * [Jail Bookings (250)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Jail-Bookings/7ykt-c32j)
-* [Calls for Service (4,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Calls-for-Service/k2nh-s5h5)
 * [Berkeley Census Data](https://data.cityofberkeley.info/)
@@
 In order to prepare the data for exploration, the cleaning process included changing the call date and time to the lubridate format, transforming the addresses into longitude and latitude referencing Google Maps, and separating the information on dispositions into separate row entries for each individual assessed in a case and splitting it further by isolating “other” and six character dispositions into different columns.
-
-	The calls for service data (4913 offenses) contained information on the case number, the offense type with a description of said offense, the date and start and end time for each case and the date of case entry into the database, and the location.
-Because the data set contained little to no information on the subjects apprehended in each case, it was largely used for mapping purposes.
 

### 2016-08-03 Stop_Data_Analysis.Rmd
starting stories

@@
 
+Story:  85.27% of stop dispositions record information about people involved, and only 19.23% of stop disposition record information about additional information.
+This fact can be caused by 2 reason: One is that the only small amount case need additional information to describe it, such as this is a primary cases.
+Most cases are just ordinary cases.
+The other one is that the Police Department may care more about the people involved rather than the case itself.
+
 ### Analysis of Race
@@
 
+story: 
+Comparing to any other race, Asian people contributes the lowest percentage of the incidents.
+This corresponds to the "model-minority" stereotype of the Asian poeple - seldom causing incidents.
+In addition, top 2 races of causing incidents are Black and White.
+This fact also makes sense - the majority of race of Berkeley residents is black and white.
+
+The percentage of people aged from 0 to 18 of each race is less than 9%.
+People aged from 0 to 18 contributes the smallest percentage in each race.
+A reason for this fact is that young teenager contributes minority of the population of Berkeley.
+And we can also induce that Berkeley teenger eduation is qualified.
+
+The percentage of people aged from 18 to 29 of each race is about or greater than 30, which is a important factor.
+One reason for this fact is that Bekeley is a university town.
+Most university students(undergraduates, master and PhD students) are aged from 18 to 29.
+Therefore, the ratio of people aged from 18 to 29 to all the residents in Berkeley should be greater than the average ratio of other cities in the US.
+Another reason is that young people maybe more impulsive to cause an incidents.
+The fact that quite a few of incidents are caused by people aged from 18 to 29 is reasonable here at Berkeley.
+Another interesting fact about people aged from 18 to 29 is that they contributes a majority of incidents at night.
+A reason for this fact is that these people are more liable to stay up late or all night and enjoy their nightlife outside and cause incidents.
+
 ### Analysis of Day
@@
 
-(1) Among all the day in a week, the number of incidents happened on Sunday and Monday are much less than those happened through Tuesday and Saturday.
+(1) Among all the day in a week, the number of incidents happened on Monday are much less than those happened through Tuesday and Saturday.
 
@@
 (5) White people are less liable to commit an incident on Sunday, and more liable to commit an incident on Tuesday and Wednesday, which is similar to Asian people.
+
+Story:
+Number of incidents on Monday is much more less than that of any other days in a week.
+An reasonable guess for this fact is that people are more careful on Monday as Monday is the beginning of the week.
+In addition, after the rest on the weekends, people are more energetic to focus and reduce the possibility of cause an incidents.
+
+An interesting fact is that number of incidents on Sunday is much more less than other 5 days in a week also.
+There are two guess for this fact: The first reason is that part of people tend to rest at home on Sunday, so these people reduce the amount of incidents happened on Sunday.
+(A senerio fits for the reason is that you can seldom see large amount of people on Berkeley campus on Sunday.) The second reason is that the police may want to rest on Sunday, so the number of policemen working on Sunday is less.
+Therefore, the Berkeley Police Department stop for less incidents on Sunday.
+
+Another interesing fact is that the number of incidents happened on Friday and Saturday is high.
+A reasonable deduction of the fact is that people tend to participate some entertainment activities on the weekends, so the additional activities increase the probability of causing an error.
 
@@
 (3) After the evening until the midnight, the incidents number increases gradually again in a day. The tally of the statistics and observation in daily life is in a good agreement.
+
+Story:
 

### 2016-08-03 final_paper/Stop_Data_Analysis.Rmd

--- 2016-08-02 Stop_Data_Analysis.Rmd
@@
 date: "July 29, 2016"
-output: html_document
+output: pdf_document
 ---
@@
 knitr::opts_chunk$set(echo = TRUE)
+```
+
+```{r}
+#setwd("~/Desktop/Stats R/PROJECT/finalproject_angry_ladies/final_paper")
 ```

### 2016-08-04 final_paper/Shinyapp.Rmd
update

@@
 ```
+
+```{r}
+
+############################## stop data ############################## 
+stop_by <- read_csv("cleaned_stop_data.csv")
+
+stop_by$Call.Date.Time <- mdy_hm(stop_by$Call.Date.Time)
+stop_by$AgeRange <- as.factor(stop_by$AgeRange)
+
+stop_by <- stop_by %>%
+  mutate(Hour = as.integer(hour(Call.Date.Time))) %>%
+  mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))
+
+stop_by$Race<- stop_by$Race %>% 
+  str_replace("A", "Asian") %>% 
+  str_replace("B", "Black") %>% 
+  str_replace("H", "Hispanic") %>% 
+  str_replace("O", "Other") %>% 
+  str_replace("W", "White")
+
+stop_by$Gender<- stop_by$Gender %>% 
+  str_replace("F", "Female") %>% 
+  str_replace("M", "Male")
+
+stop_by$Reason <- stop_by$Reason %>% 
+  str_replace("I", "Investigation") %>% 
+  str_replace("T", "Traffic") %>% 
+  str_replace("R", "Reasonable Suspicion") %>% 
+  str_replace("R", "Probation/ Parole") %>% 
+  str_replace("W", "Wanted")
+
+stop_by$Enforcement <- stop_by$Enforcement %>% 
+  str_replace("A", "Arrest") %>% 
+  str_replace("C", "Citation") %>% 
+  str_replace("O", "Other") %>% 
+  str_replace("W", "Warning")
+
+stop_by$CarSearch <- stop_by$CarSearch %>% 
+  str_replace("S", "Search") %>% 
+  str_replace("N", "No Search")
+
+write.csv(stop_by, "stop_data_jenny.csv", row.names = F)
+
+```
+
+```{r}
+p1 <- ggplot(stop_by)+
+  geom_bar(aes(x=Race),
+           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))+
+  labs(title = "Stop data by race",
+       x = "Race",
+       y = "Counts")
+
+p2 <- ggplot(race)+
+  geom_bar(aes(x=Description, 
+               y=Counts),
+           stat = "identity", 
+           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))+
+  labs(title = "Demographic race data",
+       x = "Race",
+       y = "Counts")
+
+multiplot(p1, p2,cols=2)
+
+## asian #FFCC00
+## black: #000000
+## hispanic: #330099
+## other: #66FFCC
+## white: #FFFFFF
+
+c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF")
+
+```
+
+```{r}
+stop_by
+m <- subset(stop_by, Hour = "7") %>% 
+  group_by(Race) %>% 
+  tally() %>% 
+  arrange(desc(n))
+
+rrr <- subset(stop_by, Hour = "7") %>% 
+  group_by(Race) %>% 
+  tally() 
+rrr$Race<-as.factor(rrr$Race)
+
+ggplot(rrr)+
+  geom_bar(aes(x=Race, y = n),
+           stat = "identity",
+           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))
+
+rrr <- subset(stop_by, Hour = "7") %>% 
+  group_by(Race) %>% 
+  tally()
+
+           fill = c("red", "yellow", "white", "blue"))
+           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))
+
+m[sort(m$n, decreasing=T),]
+m
+m[order(-m[,3],m[,1]),]
+
+arrange(m,desc(n))
+
+```

### 2016-08-04 final_paper/Some thoughts.Rmd
update

--- 2016-08-02 final_paper/BDP_Paperdraft.Rmd
@@
 ---
-title: "BPD_draftpaper"
-author: "Amy Zhu; Mengyu Li; Rebecca Reus; Shangjun Jiang"
-date: "2 August 2016"
+title: "Some ideas"
+author: "Shangjun Jiang"
+date: "4 August 2016"
 output: html_document
-runtime: shiny
 ---
@@
 library(tidyr)
-library(readr)
 library(plyr)
 library(dplyr)
-library(ggvis)
-library(tidyr)
 library(stringr)
@@
 library(RColorBrewer)
+setwd("~/Desktop/Stats R/PROJECT/finalproject_angry_ladies/final_paper")
 ```
 
-## Introduction
-
- For the final project, our group cleaned, explored, and analyzed four different data sets from the City of Berkeley that contained information on stops, calls for service, arrests, and jail bookings made by the the police department in 2016 (and 2015 as well for the stop data).
-Rather than limiting the project to one set of data, four sets were chosen in order to gain a more holistic and comprehensive understanding of the data.
-The additional variables included expanded the project’s capacity for manipulating data, examining relationships, and improving result reliability.
-
-#### Data Collection
-
-* [Stop Data (16,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Stop-Data/6e9j-pj9p)
-* [Arrest (200)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Arrests/xi7q-nji6)
-* [Jail Bookings (250)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Jail-Bookings/7ykt-c32j)
-* [Calls for Service (4,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Calls-for-Service/k2nh-s5h5)
-
- With the variety of resources and information made available by collating multiple data sets, the objective for the project was to study differences in police activity (in terms of call requests and patrols) and intensity of assessed offenses based on time, race, gender, age, and mental health.
-As a challenge, another project target was to create a map applet depicting the density of police activity in Berkeley with an interactive component allowing the visitor to input an address and observe their proximity and observe the types of incidences that occurred most commonly in the area.
-
-	For the stop data, there were 16,255 incidents assessed by the Berkeley police.
-In each case, the call date and time, location, incident type, and disposition(s) were recorded.
-The cases typically had a six character disposition, with each character conveying race, gender, age, reason, enforcement, and car search,  respectively, for each subject involved in the incident, although there were additional dispositions, ranging from one to three characters,  that could be input and conveyed other information.
-In order to prepare the data for exploration, the cleaning process included changing the call date and time to the lubridate format, transforming the addresses into longitude and latitude referencing Google Maps, and separating the information on dispositions into separate row entries for each individual assessed in a case and splitting it further by isolating “other” and six character dispositions into different columns.
-To improve readability, the acronyms and abbreviations used in the data set were mutated to convey their meanings  (will do soon).
-
-	The calls for service data (4913 offenses) contained information on the case number, the offense type with a description of said offense, the date and start and end time for each case and the date of case entry into the database, and the location.
-Because the data set contained little to no information on the subjects apprehended in each case, it was largely used for mapping purposes.
-
-	The data for arrests (205), and jail bookings (223) contained similar information on case/arrest/booking number, date and time, type, and subject information (name, race, sex, D.O.B., age, height, weight, hair, eyes, and occupation) and statute information (type, description, agency, and disposition).
-Cleaning required the dates and times to be put into lubridate format, for the two data sets to be compiled in a reasonable, and for other needed adjustments.
-The same code created for the stop data would be used again to convert address information to longitudinal and latitudinal coordinates.
-
-## Data Analysis
+## Data Loading
 
@@
   mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))
-```
 
-### About the Berkeley PD Stop Data
+stop_by$Race<- stop_by$Race %>% 
+  str_replace("A", "Asian") %>% 
+  str_replace("B", "Black") %>% 
+  str_replace("H", "Hispanic") %>% 
+  str_replace("O", "Other") %>% 
+  str_replace("W", "White")
 
-The Berkeley PD stop data reflects accidents that have been visited by the Berkeley Police Department.
-Raw data set is downloaded from City of Berkeley Open Data website and provided by Berkeley Police Department.
+stop_by$Gender<- stop_by$Gender %>% 
+  str_replace("F", "Female") %>% 
+  str_replace("M", "Male")
 
-Analysis in this part focuses on the information provided by the dispositions variable.
-The dispositions variable can be divided into three types.
-The first type only includes more than one six-character person message.
-Each character in each person message represents the race, gender and age range of the person, reason and enforcement of the stop, and the car search information during the stop.
-The second type of the dispositions includes only additional dispositions, such as P for primary cases.
-The third type of dispositions includes both person information and additional dispositions.
+stop_by$Reason <- stop_by$Reason %>% 
+  str_replace("I", "Investigation") %>% 
+  str_replace("T", "Traffic") %>% 
+  str_replace("R", "Reasonable Suspicion") %>% 
+  str_replace("R", "Probation/ Parole") %>% 
+  str_replace("W", "Wanted")
 
-For the analysis in this part, only person information for all the people in each incident and the Call Time/Date of each incident are analyzed.
+stop_by$Enforcement <- stop_by$Enforcement %>% 
+  str_replace("A", "Arrest") %>% 
+  str_replace("C", "Citation") %>% 
+  str_replace("O", "Other") %>% 
+  str_replace("W", "Warning")
 
-### Analysis of Race
-#### Picture: count of person recorded in each race
-```{r, echo=FALSE}
-stop_race <- stop_by %>%
-  ggvis(x = ~Race , fill = ~AgeRange) %>%
-  layer_bars() %>%
-  add_tooltip(function(stop_by) stop_by$AgeRange)
+stop_by$CarSearch <- stop_by$CarSearch %>% 
+  str_replace("S", "Search") %>% 
+  str_replace("N", "No Search")
 
-stop_race
-```
+############################## location data ############################## 
+mergedf <- readRDS( 'StopData_merged.rds' )
 
-```{r, echo=FALSE, eval=FALSE}
-stop_race_total_table <- stop_by %>%
-    group_by(Race) %>%
-    summarise(totaLCount = n())
-stop_race_agerange_table <- stop_by %>%
-    group_by(Race, AgeRange) %>%
-    summarise(count = n())
-stop_race_table <- left_join(stop_race_agerange_table, stop_race_total_table, by = "Race")
-stop_race_table <- stop_race_table %>%
-    mutate(percentage = count/totaLCount) %>%
-    select(-count)
-stop_race_table <- stop_race_table %>%
-    spread(key = AgeRange, value = percentage)
-stop_race_table
-```
+## Location info:
+latmax <- max(mergedf$lat, na.rm = TRUE) 
+latmin <- min(mergedf$lat, na.rm = TRUE)
+lonmax <- max(mergedf$long, na.rm = TRUE) 
+lonmin <- min(mergedf$long, na.rm = TRUE)
+latvals <- c(latmin, latmax)
+lonvals <- c(lonmin, lonmax)
 
-####Table: count of person recorded in each race
-
-Race  | Count | 0-18 | 18-29 | 30-39 | 40+
-----  |------ |------|-------|-------|-------
-A     | 1141  | 8.50%| 43.47%| 21.03%| 26.99%
-B     | 4636  | 2.29%| 34.97%| 25.09%| 37.66%
-H     | 1676  | 3.22%| 46.12%| 29.18%| 21.47%
-O     | 1384  | 2.67%| 43.28%| 29.17%| 26.87%
-W     | 5454  | 1.72%| 29.45%| 25.39%| 43.44%
-
-#### Picture: count of person recorded of each race in each hour
-```{r, echo=FALSE}
-stop_time_percentage <- ggplot(stop_by, aes(x = Hour, fill = AgeRange)) + 
-  geom_bar(aes(y = ..count..)) +
-  labs(y = "Count") +
-  facet_wrap(~Race)
-stop_time_percentage
-```
-
-(1) Among the 14291 personnel information recorded, white people contributes the largest percentage.
-38.16% of all records, namely 5454 people are white.
-black people contributes the second largest percentage.
-32.44% of all records, namely 4636 people are black.
-
-(2) The percentage of people aged from 0 to 18 of each race is less than 9%. People aged from 0 to 18 contributes the smallest percentage in each race, which makes sense in daily life.
-
-(3) Among white people and black people recorded by the BPD, people aged greater than 40  contribute the largest percentage, which are 43.44% and 37.66% respectively.
-
-(4) Among Asian, Hispanic and other people recorded by the BPD, people aged from 18 to 29 contribute the largest percentage, which are 43.37%, 46.12% and 43.28%, respectively.
-
-(5) Black people contribute a especially high percentage of the incidents at night, while the white people contribute a especially high percentage of the incidents at the noon.
-
-## Analysis of incidents of each hour in a day
-#### Picture: count of person recorded in each hour in a day
-```{r, echo=FALSE}
-stop_time <- stop_by %>%
-  ggvis(x = ~Hour, fill := "lightpink") %>%
-  layer_histograms(width = input_slider(1, 12, value = 1, step = 2, label = "width"))
-stop_time
-```
-
-Note: you can change the value of the input slider to see the number of incidents happened in specific time span of the day.
-
-```{r, eval = FALSE, echo=FALSE}
-stop_hour <- stop_by %>%
-  group_by(Hour) %>%
-  summarise(count = n())
-stop_hour
-```
-
-#### Table: count of person recorded in each hour in a day
-Hour  | Count | Hour | Count| Hour | Count| Hour | Count
-----  |------ |------|------|------|------|------|------
-0     | 823   | 6    | 157  | 12   | 802  | 18   | 544
-1     | 618   | 7    | 341  | 13   | 695  | 19   | 685
-2     | 489   | 8    | 489  | 14   | 515  | 20   | 834
-3     | 289   | 9    | 582  | 15   | 542  | 21   | 971
-4     | 135   | 10   | 682  | 16   | 612  | 22   | 1098
-5     | 61    | 11   | 678  | 17   | 578  | 23   | 1071
-
-(1) time span of 22:00 to 22:59 in a day has the highest incidents count, and 23:00 to 23:59 has the second highest, which means that Berkeley area is most dangerous from 10pm to 12am in a day, which corresponds to daily life experience.
-(2) From 12am to 5am, incidents count decreases gradually.
-A reasonable conjecture of this fact is that more and more people choose to sleep as time goes by in this time span.
-For the similar reason, incidents count increases gradually from 5am to 11am.
-(3) After the evening until the midnight, the incidents number increases gradually again in a day. The tally of the statistics and observation in daily life is in a good agreement.
-
-#### Picture: Probability of BPD stop for a specific age range in a specific hour
-
-```{r, echo=FALSE}
-stop_by_hourCount <- stop_by %>%
-  group_by(Hour) %>%
-  summarise(hourcount = n())
-stop_by_ageRangeCount <- stop_by %>%
-  group_by(Hour, AgeRange) %>%
-  summarise(ageCount = n())
-stop_by_percentage <- left_join(stop_by_ageRangeCount, stop_by_hourCount, by = "Hour")
-stop_by_percentage <- stop_by_percentage %>%
-  mutate(percentage = ageCount/hourcount)
-```
-
-```{r, echo=FALSE}
-stop_time_percentage2 <- ggplot(stop_by_percentage, aes(x = Hour, fill = AgeRange)) + 
-  geom_bar(aes(y = percentage), stat="identity") +
-  labs(y = "Density") +
-  facet_wrap(~AgeRange)
-stop_time_percentage2
-```
-
-(1) People aged from 18 to 29 are obviously liable to commit incidents at night.
-The average ratio of people aged from 18 to 29 to all people stoped by the Berkeley Police Department at night is greater than 40%.
-
-(2) People aged greater than 40 are obviously liable to commit incidents in the daytime.
-The average ratio of people aged greater than 40 to all people stoped by the Berkeley Police Deparment during the day is greater than 40%.
-
-(3) Ratio of incidents commited by people aged between 0 and 18 and people aged from 30 to 39 fluctuates during the daytime and the night, with an average ratio of 2.5% and 25% respectively.
-
-#### Probability of BPD stop for a specific race in a specific hour
-
-```{r, echo=FALSE}
-stop_by_hourCount <- stop_by %>%
-  group_by(Hour) %>%
-  summarise(hourcount = n())
-stop_by_raceRangeCount <- stop_by %>%
-  group_by(Hour, Race) %>%
-  summarise(raceCount = n())
-stop_by_percentage1 <- left_join(stop_by_raceRangeCount, stop_by_hourCount, by = "Hour")
-stop_by_percentage1 <- stop_by_percentage1 %>%
-  mutate(percentage = raceCount/hourcount)
-```
-```{r, echo=FALSE}
-stop_time_percentage3 <- ggplot(stop_by_percentage1, aes(x = Hour, fill = Race)) + 
-  geom_bar(aes(y = percentage), stat="identity") +
-  labs(y = "Density") +
-  facet_wrap(~Race)
-stop_time_percentage3
-```
-
-(1) Black people are obviously liable to commit incidents at night. The average ratio of black people to all people stoped by the Berkeley Police Department at night is about 40%.
-
-(2) White people are obviously liable to commit incidents in the daytime. The average ratio of white people to all people stoped by the Berkeley Police Deparment during the day is about 45%.
-
-(3) Ratio of incidents commited by Asian people, Hispanic people and other people fluctuate during the daytime and the night, with an average ratio of 8%, 11% and 9% respectively.
-
-### Analysis of preference of Berkeley Police Department arresting people
-#### Picture: Probability of arrested by BPD of a specific race for a specific reason
-```{r,echo=FALSE}
-stop_by_reason_race <- stop_by %>%
-  group_by(Race, Reason) %>%
-  summarise(totalCount = n())
-stop_by_arrest <- stop_by %>%
-  filter(Enforcement == "A") %>%
-  group_by(Race, Reason) %>%
-  summarise(arrestCount = n())
-stop_by_percentage_race <- left_join(stop_by_reason_race, stop_by_arrest, by = c("Race", "Reason"))
-stop_by_percentage_race <- stop_by_percentage_race %>%
-  mutate(percentage = arrestCount / totalCount)
-stop_arrest <- ggplot(stop_by_percentage_race, aes(x = Race, fill = Race)) + 
-  geom_bar(aes(y = percentage), stat="identity") +
-  labs(y = "Density") +
-  facet_wrap(~Reason)
-stop_arrest
-```
-
-#### Table: Probability of arrested by BPD of a specific race for a specific reason
-P   | I      | K     | R     | T     | W     |
-----|------- |-------|-------|-------|-------|
-A   | 10.34% | 75%   | 10.34%| 1.24% | 17.60%|
-B   | 8.31%  | 15.79%| 6.44% | 2.28% | 50.00%|  
-H   | 6.98%  | 16.67%| 9.09% | 2.90% | 25.00%| 
-O   | 10.14% | 18.18%| 5.80% | 0.41% | 41.67%| 
-W   | 8.14%  | 26.79%| 3.40% | 1.04% | 32.56%|
-
-Note: in the picture and table above, I for Investigation, T for Traffic, R for Reasonable Suspicion, K for Probation/Parole, W for wanted.
-
-(1) The probability of arrested in a stop with the reason Traffic is much lower than other reasons. The average conditional probability of arrested given reason is traffic is 1.58%.
-
-(2) The probability of arrested in a stop with the reason Wanted is much higher than other reason. The average conditional probability of arrested given reason is wanted is 33.33%.
-
-(3) An interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and race Asian is 75%, which is much higher than that of any other race.
-Asian people is much more liable to be arrested by Berkeley Police Department during Probation or Parole if stopped by the BPD.
-
-#### Picture: Probability of arrested by BPD of a specific age range for a specific reason
-```{r, echo=FALSE}
-stop_by_reason_age <- stop_by %>%
-  group_by(AgeRange, Reason) %>%
-  summarise(totalCount = n())
-stop_by_arrest_age <- stop_by %>%
-  filter(Enforcement == "A") %>%
-  group_by(AgeRange, Reason) %>%
-  summarise(arrestCount = n())
-stop_by_percentage_age <- left_join(stop_by_reason_age, stop_by_arrest_age, by = c("AgeRange", "Reason"))
-stop_by_percentage_age <- stop_by_percentage_age %>%
-  mutate(percentage = arrestCount / totalCount)
-
-stop_arrest_age <- ggplot(stop_by_percentage_age, aes(x = AgeRange, fill = AgeRange)) + 
-  geom_bar(aes(y = percentage), stat="identity") +
-  labs(y = "Density") +
-  facet_wrap(~Reason)
-stop_arrest_age
-```
-
-#### Table: Probability of arrested by BPD of a specific age range for a specific reason
-P     | I     | K     | R     | T     | W     
-------|-------|-------|-------|-------|-------
-0-18  | 17.95%| 67.57%| 13.95%| 3.24% | 50.00%  
-18-29 | 8.45% | 12.50%| 3.87% | 1.82% | 32.35%  
-30-39 | 6.32% | 25.00%| 5.04% | 1.49% | 26.92%  
-40+   | 7.87% | 23.86%| 6.13% | 1.41% | 31.82%
-
-(1) Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.
-
-(2) Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and age range 0-18 is 67.56%, which is much higher than that of any other age range.
-
-(3) The conditional probability of arrested given person involved aged from 0 to 18 in a stop given any reason is more than that of given person with any other age range.
-Therefore, in a stop, teenagers is more liable to be arrested by Berkeley Police Department.
-
-#### Picture: Probability of arrested by BPD of a specific gender for a specific reason
-```{r, echo=FALSE}
-stop_by_reason_gender <- stop_by %>%
-  group_by(Gender, Reason) %>%
-  summarise(totalCount = n())
-stop_by_arrest_gender <- stop_by %>%
-  filter(Enforcement == "A") %>%
-  group_by(Gender, Reason) %>%
-  summarise(arrestCount = n())
-stop_by_percentage_gender <- left_join(stop_by_reason_gender, stop_by_arrest_gender, by = c("Gender", "Reason"))
-stop_by_percentage_gender <- stop_by_percentage_gender %>%
-  mutate(percentage = arrestCount / totalCount)
-
-stop_arrest_gender <- ggplot(stop_by_percentage_gender, aes(x = Gender, fill = Gender)) + 
-  geom_bar(aes(y = percentage), stat="identity") +
-  labs(y = "Density") +
-  facet_wrap(~Reason)
-stop_arrest_gender
-```
-
-#### Table: Probability of arrested by BPD of a specific gender range for a specific reason
-P     | I (Investigation) |K (Probation/ Parole)| R (Reasonable Suspicion)|T (Traffic)| W (Wanted) 
-----  |------------------ |---------------------|-------------------------|-----------|-------
-F     | 9.05%             | 50.00%              | 5.00%                   | 1.10%     | 28.57%  
-M     | 8.26%             | 20.11%              | 5.42%                   | 1.85%     | 37.04%
-
-(1) Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.
-
-(2) Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole of female is 50.00%, which is much higher than that of male, which is 20.11%.
-
-```{r, echo=FALSE}
-## Read in the Data:
-#finaldf <- readRDS( 'StopData_finaldf.rds' )
-mergedf <- readRDS( 'StopData_merged.rds' )
-```
-
-```{r, echo=FALSE}
-## Location info:
-latmax <- max( mergedf$lat, na.rm = TRUE ) 
-latmin <- min( mergedf$lat, na.rm = TRUE )
-lonmax <- max( mergedf$long, na.rm = TRUE ) 
-lonmin <- min( mergedf$long, na.rm = TRUE )
-latvals <- c( latmin, latmax )
-lonvals <- c( lonmin, lonmax )
-```
-
-```{r, echo=FALSE}
 ## Make the variables factors:
@@
 v <- factor(mergedf$Other)
-levels(v)
 
@@
 
-berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14)
+### BerkeleyMap 
+berkMap = map = get_map(location = c(lon = mean(lonvals), lat = mean(latvals)), zoom = 14)
+
+############################## census data Rebecca ##############################
+
+blocks <- readOGR("Census_Tract_Polygons2010","Census_tracts_2010", verbose = TRUE)
+
+locationCensusFiles <- "Census_Tract_Polygons2010"
+blocks <- readOGR(locationCensusFiles,"Census_tracts_2010", verbose = TRUE)
+b2 <- spTransform(blocks, CRS("+proj=longlat +datum=WGS84"))
+b3 <- fortify(b2)
+
+############################## census data JENNY ############################## 
+census <- read.csv("Census_Data_2000_And_2010.csv")
+census2010<- census[census$Year==2010,]
+
+racecensus2010 <- census2010[census2010$Heading=="Not Hispanic or Latino"|census2010$Heading=="HISPANIC OR LATINO AND RACE",]
+sexcensus2010 <- census2010[census2010$Heading=="Sex",]
+agecensus2010 <- census2010[census2010$Heading=="Age",]
+
+race <- select(racecensus2010, Description, Amount) 
+race$Description <- race$Description %>% 
+  str_replace_all("Hispanic or Latino.*", "Hispanic") %>% 
+  str_replace("Black or African American", "Black") %>% 
+  str_replace("American Indian and Alaska Native", "Other") %>% 
+  str_replace("Native Hawaiian and Other Pacific Islander", "Other") %>% 
+  str_replace("Some other race", "Other") %>% 
+  str_replace("Two or more races", "Other")
+
+race <- race[race$Description !="Not Hispanic",]
+
+race <- race %>% 
+  group_by(Description) %>% 
+  tally(Amount)
+
+names(race) <- c("Description", "Counts")
+
+sex <- select(sexcensus2010, Description, Amount)
+
+age <- select(agecensus2010, Description, Amount) 
+age <- age[age$Description != "Median age",]
+age$Description <- c("0-18", "0-18", "18-64", "65+")
+
+age<- age %>% 
+  group_by(Description) %>% 
+  tally(Amount)
+names(age) <- c("Description", "Counts")
+
+############################## arrest data ############################## 
+arrest <- read.csv("Berkeley_PD_Log_-_Arrests.csv")
+
+############################## jail data ############################## 
+jail <- read.csv("Berkeley_PD_Log_-_Jail_Bookings.csv")
+
+############################## call for service data ##############################
+
+callservice<- read.csv("Berkeley_PD_-_Calls_for_Service.csv")
+
+callservice<- callservice %>% 
+  mutate(latlong= 
+           str_replace_all(callservice$Block_Location,"[0-9]* [A-Za-z]*", "") %>%
+           str_replace_all("\nBerkeley,\n", "") %>% 
+           str_replace("[A-Za-z]*", "") %>% 
+           str_replace(";","") %>% 
+           str_replace("&","") %>% 
+           str_replace("[A-Za-z]*", ""))
+
+callservice$latlong<- str_replace(callservice$latlong, "\\(", "") %>% 
+                    str_replace("\\)", "")
+
+callservice<- callservice %>% 
+  separate(latlong, c("lat", "long"), sep=",")
+
+callservice$lat <- as.numeric(callservice$lat)
+callservice$long <- as.numeric(callservice$long)
+
+############################## multiplot function ############################## 
+multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
+  library(grid)
+
+  # Make a list from the ... arguments and plotlist
+  plots <- c(list(...), plotlist)
+
+  numPlots = length(plots)
+
+  # If layout is NULL, then use 'cols' to determine layout
+  if (is.null(layout)) {
+    # Make the panel
+    # ncol: Number of columns of plots
+    # nrow: Number of rows needed, calculated from # of cols
+    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
+                    ncol = cols, nrow = ceiling(numPlots/cols))
+  }
+
+ if (numPlots==1) {
+    print(plots[[1]])
+
+  } else {
+    # Set up the page
+    grid.newpage()
+    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
+
+    # Make each plot, in the correct location
+    for (i in 1:numPlots) {
+      # Get the i,j matrix positions of the regions that contain this subplot
+      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
+
+      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
+                                      layout.pos.col = matchidx$col))
+    }
+  }
+}
 
@@
 
-## MAP 1:All BPD Stops Density, 2015-2016
-```{r, fig.width=10, fig.asp=0.8, echo=FALSE}
-# a contour plot
-ggmap(berkMap) +
-  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
-    size = 2, bins = 5, data = mergedf, geom = "polygon") +
-  scale_fill_gradient(low = "black", high = "red") +
-    theme (panel.grid.major = element_blank (), # remove major grid
-		       panel.grid.minor = element_blank (),  # remove minor grid
-		       axis.text = element_blank (), 
-		       axis.title = element_blank (),
-	         axis.ticks = element_blank ()) + 
-  ggtitle("All BPD Stops Density, 2015-2016") +
-  labs(alpha = element_blank())+
-  guides(alpha = FALSE)
+## PART I: Racial Discrimination??
+
+```{r}
+p1 <- ggplot(stop_by)+
+  geom_bar(aes(x=Race),
+           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))+
+  labs(title = "Stop data by race",
+       x = "Race",
+       y = "Counts")
+
+p2 <- ggplot(race)+
+  geom_bar(aes(x=Description, 
+               y=Counts),
+           stat = "identity", 
+           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))+
+  labs(title = "Demographic race data",
+       x = "Race",
+       y = "Counts")
+
+multiplot(p1, p2,cols=2)
 
@@
 
-## MAP 2: BPD Stop Contour Map of Berkeley by Race
-```{r, fig.height = 5, fig.width = 10, echo=FALSE}
-ggmap(berkMap) +
-  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
-    bins = I(5), geom = "polygon", data = mergedf ) +
-  scale_fill_gradient2( "StopDensity",
-    low = "white", mid = "orange", high = "red", midpoint = 100) +
-  labs(x = "Longitude", y = "Latitude") + facet_wrap( ~ Race ) +
-  scale_alpha(range = c(.2, .55), guide = FALSE) +
-  ggtitle("BPD Stop Contour Map of Berkeley by Race") +
-  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
-```
+#### arrest + jail data
 
-## MAP 3: BPD Stop Contour Map of Berkeley by Age Range
-```{r, fig.height = 8, fig.width = 10, echo=FALSE}
-ggmap(berkMap) +
-  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
-    bins = I(5), geom = "polygon", data = mergedf ) +
-  scale_fill_gradient2( "Stop Density",
-    low = "white", mid = "orange", high = "red", midpoint = 25) +
-  labs(x = "Longitude", y = "Latitude") + facet_wrap(~ AgeRange) +
-  scale_alpha(range = c(.2, .55), guide = FALSE) +
-  ggtitle("BPD Stop Contour Map of Berkeley by Age Range") +
-  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
-```
+## PART II: time of being stopped
 
-We found that the
+#### by race
+#### by age
+
+#### location of being stopped
+#### census data + stop data map + call for service map
+
+## PART IV: Exploration
+#### pick location + time
+#### time you top 3 stop reasons + top 2 race + top 2 age groups

### 2016-08-04 final_paper/draft.Rmd
draft
update

@@
 * [Jail Bookings (250)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Jail-Bookings/7ykt-c32j)
-* [Berkeley Census Data](https://data.cityofberkeley.info/)
+* [Calls for Service (4,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Calls-for-Service/k2nh-s5h5)
+* [Berkeley Census Data](https://data.cityofberkeley.info/Demographics/Census-Data-2000-And-2010/bnkq-f2ge)
 
@@
   mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))
+
+stop_by$Race<- stop_by$Race %>% 
+  str_replace("A", "Asian") %>% 
+  str_replace("B", "Black") %>% 
+  str_replace("H", "Hispanic") %>% 
+  str_replace("O", "Other") %>% 
+  str_replace("W", "White")
+
+stop_by$Gender<- stop_by$Gender %>% 
+  str_replace("F", "Female") %>% 
+  str_replace("M", "Male")
+
+stop_by$Reason <- stop_by$Reason %>% 
+  str_replace("I", "Investigation") %>% 
+  str_replace("T", "Traffic") %>% 
+  str_replace("R", "Reasonable Suspicion") %>% 
+  str_replace("R", "Probation/ Parole") %>% 
+  str_replace("W", "Wanted")
+
+stop_by$Enforcement <- stop_by$Enforcement %>% 
+  str_replace("A", "Arrest") %>% 
+  str_replace("C", "Citation") %>% 
+  str_replace("O", "Other") %>% 
+  str_replace("W", "Warning")
+
+stop_by$CarSearch <- stop_by$CarSearch %>% 
+  str_replace("S", "Search") %>% 
+  str_replace("N", "No Search")
 
@@
 ### BerkeleyMap 
-berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14)
+berkMap = map = get_map(location = c(lon = mean(lonvals), lat = mean(latvals)), zoom = 14)
 
@@
                fill = Race))+
-  facet_wrap(~Race)
+  facet_wrap(~Race)+
+ labs(x= "Age",
+       y = "Number of people",
+       title = "Jail Data")
 

### 2016-08-04 final_paper/server.R

@@ -1,0 +1,35 @@
+
+# load other packages
+library(shiny)
+library(dplyr)
+library(readr)
+library(ggplot2)
+library(plyr)
+
+stop_by <- read_csv("stop_data_jenny.csv")
+
+##Problems:
+#1. add_legend
+#2. scale
+#4. change size of the point 
+
+shinyServer(function(input, output) {
+  
+  output$Plot <- renderPlot({
+  
+  test <- subset(stop_by, Hour = input$hour) %>% 
+    group_by(Race) %>% 
+    tally() 
+
+  rrr$Race<-as.factor(rrr$Race)
+  
+  ggplot(rrr)+
+    geom_bar(aes(x=Race, y = n),
+             stat = "identity",
+             fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))
+  
+  })
+  
+
+})
+

### 2016-08-04 final_paper/ui.R

@@ -1,0 +1,28 @@
+library(shiny)
+library(ggplot2)
+
+shinyUI(fluidPage(
+  
+  # Application title
+  titlePanel("Hour and the top three group"),
+  
+  # Sidebar layout
+  sidebarLayout(
+    
+    # Sidebar panel
+    sidebarPanel(
+      sliderInput('hour',
+                  'HOUR: ',
+                  0, 24, value=7,
+                  animate = T)
+      ),
+    
+    
+    # Main panel
+    mainPanel(
+      plotOutput("Plot")
+      )
+      )
+      ))
+
+

### 2016-08-04 eda/Heatmap/StopData_clean_merged_rds.R
creating required folders

@@ -1,0 +1,54 @@
+# StopData_clean_merged_rds.R
+# Purpoe: This script does some minor necessary cleaning to StopData_merged.rds data that was created by the location cleaning script.
+
+library(sp)
+library(ggmap)
+library(tidyr)
+library(dplyr)
+library(stringr)
+library(lubridate)
+library(ggplot2)
+library(readr)
+library(maps)
+library(mapdata)
+library(maptools)
+library(rgdal)
+library(rgeos)
+
+# data file names:
+data_file <- "StopData_merged.rds" # we will be modifying this file.
+new_data_file <- "StopData_merged.rds"
+
+# read in the data to modify:
+df <- readRDS( data_file )
+
+# make the changes necessary for plotting and mapping the data:
+df$CarSearch <- factor(df$CarSearch)
+levels(df$CarSearch) <- c("No Search", "Search")  
+df$Race <- factor(df$Race)
+levels(df$Race) <- c("Asian", "Black", "Hispanic", "Other", "White")
+df$Enforcement <- factor(df$Enforcement)
+levels(df$Enforcement) <- c("Arrest", "Citation", "Other", "Warning")
+df$Reason <- factor(df$Reason)
+levels(df$Reason) <- c("Other", "Investigation", "Probation/Parole", "Reasonable Suscipcion", "Traffic", "Wanted")
+df <- df %>%
+  mutate(Arrest = ifelse( 
+    as.character(Enforcement) == "Arrest", 
+    "Arrested",
+    "Not Arrested") ) %>%
+  mutate(Arrest = factor(Arrest))
+v <- factor(df$Other)
+df <- df %>%
+  mutate(Emergency.Psych.Eval = ifelse( str_detect(as.character(Other), "MH"), 
+                                        yes = "Yes",
+                                        no = "No") )
+df$Emergency.Psych.Eval <- factor(df$Emergency.Psych.Eval)
+
+# save the changed data as a new RDS (RDS file saves more space than  CVS file):
+saveRDS(object = df, file = new_data_file )
+
+# remove unecessary variables:
+rm(v)
+rm(data_file)
+rm(new_data_file)
+rm(df)

### 2016-08-04 eda/Heatmap/heatmap.R

@@ -1,0 +1,72 @@
+# Title: heatmap.R
+# Purpose: fix the heat map for the Berkeley sensus data
+# Author: Rebecca Reus
+# Date: 8-4
+
+library(sp)
+library(ggmap)
+library(tidyr)
+library(dplyr)
+library(stringr)
+library(lubridate)
+library(ggplot2)
+library(readr)
+library(maps)
+library(mapdata)
+library(maptools)
+library(rgdal)
+library(rgeos)
+
+# data file names:
+data_file <- "StopData_merged2.rds" # use the cleaned version of StopData_merged.rds for plotting. See StopData_clean_merged_rds.R for more information.
+
+# read in the data to modify:
+df <- readRDS( data_file )
+
+# Location info (to set lim values for coordinates):
+latmax <- max( df$lat, na.rm = TRUE )
+latmin <- min( df$lat, na.rm = TRUE )
+lonmax <- max( df$long, na.rm = TRUE )
+lonmin <- min( df$long, na.rm = TRUE )
+latvals <- c( latmin, latmax )
+lonvals <- c( lonmin, lonmax )
+######################################
+
+# get the plain Berkeley map from Google:
+berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14)
+
+# get the census shape files:
+##locationCensusFiles <- "C:/Users/Rebecca/Dropbox/School/2016_su_School/stat133/finalproject_angry_ladies/Rebeccas_location_Code2/Census_Block_Group_Polygons2010"
+blocks <- readOGR("Census_Block_Group_Polygons2010","Census_blockgroups_2010", verbose = TRUE)
+
+blocks.polygons <- blocks@polygons
+blocks.data <- blocks@data
+blocks.plot.order <- blocks@plotOrder
+b2 <- spTransform(blocks, CRS("+proj=longlat +datum=WGS84"))
+b3 <- fortify(b2)
+
+ggmap(berkMap) +
+  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
+                 size = 2, bins = 5, data = b3, geom = "polygon") +
+  scale_fill_gradient(low = "black", high = "red") +
+  theme (panel.grid.major = element_blank (), # remove major grid
+         panel.grid.minor = element_blank ()  # remove minor grid
+  ) + 
+  ggtitle("Population Density") +
+  labs(alpha = element_blank())+
[…]

### 2016-08-04 clean_data/StopData_clean_merged_rds.R
adding some more data

@@ -1,5 +1,5 @@
 # StopData_clean_merged_rds.R
-# Purpoe: This script does some minor necessary cleaning to StopData_merged.rds data that was created by the location cleaning script.
+# Purpose: This script does some minor necessary cleaning to StopData_merged.rds data that was created by the location cleaning script.
 
 library(sp)
 library(ggmap)

### 2016-08-04 eda/Heatmap/heatmap.R

@@ -18,7 +18,7 @@ library(rgdal)
 library(rgeos)
 
 # data file names:
-data_file <- "StopData_merged2.rds" # use the cleaned version of StopData_merged.rds for plotting. See StopData_clean_merged_rds.R for more information.
+data_file <- "../../clean_data/StopData_merged2.rds" # use the cleaned version of StopData_merged.rds for plotting. See StopData_clean_merged_rds.R for more information.
 
 # read in the data to modify:
 df <- readRDS( data_file )

### 2016-08-05 eda/Heatmap/heatmap.R
heatmap

@@ -13,12 +13,19 @@ library(ggplot2)
 library(readr)
 library(maps)
 library(mapdata)
-library(maptools)
 library(rgdal)
 library(rgeos)
+library(gpclib)       # loads polygon clipping library
+library(maptools)     # loads sp library too
+library(RColorBrewer) # creates nice color schemes
+library(classInt)     # finds class intervals for continuous variables
 
 # data file names:
 data_file <- "../../clean_data/StopData_merged2.rds" # use the cleaned version of StopData_merged.rds for plotting. See StopData_clean_merged_rds.R for more information.
+census_tract_folder <- "../../clean_data/Census_Tract_Polygons2010"
+census_tract_shp <- "Census_tracts_2010"
+census_tract_csv <- "../../clean_data/Census_tracts_2010.csv"
+
 
 # read in the data to modify:
 df <- readRDS( data_file )
@@ -33,28 +40,74 @@ lonvals <- c( lonmin, lonmax )
 ######################################
 
 # get the plain Berkeley map from Google:
-berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14)
+berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 13)
 
 # get the census shape files:
 ##locationCensusFiles <- "C:/Users/Rebecca/Dropbox/School/2016_su_School/stat133/finalproject_angry_ladies/Rebeccas_location_Code2/Census_Block_Group_Polygons2010"
-blocks <- readOGR("Census_Block_Group_Polygons2010","Census_blockgroups_2010", verbose = TRUE)
+#blocks <- readOGR("Census_Block_Group_Polygons2010","Census_blockgroups_2010", verbose = TRUE)
+census <- readOGR(census_tract_folder,census_tract_shp, verbose = TRUE)
 
-blocks.polygons <- blocks@polygons
-blocks.data <- blocks@data
-blocks.plot.order <- blocks@plotOrder
-b2 <- spTransform(blocks, CRS("+proj=longlat +datum=WGS84"))
-b3 <- fortify(b2)
+# blocks.polygons <- blocks@polygons
+# blocks.data <- blocks@data
+# blocks.plot.order <- blocks@plotOrder
+census2 <- spTransform(census, CRS("+proj=longlat +datum=WGS84"))
+census3 <- fortify(census2)
[…]

### 2016-08-05 eda/Heatmap/HeatmapExploration2.Rmd
blah

---
title: "Heat Map Exploration"
author: "Rebecca Reus"
date: "August 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sp)
library(ggmap)
library(tidyr)
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(maps)
library(mapdata)
library(rgdal)
library(rgeos)
#library(gpclib)       # loads polygon clipping library
library(maptools)     # loads sp library too
library(RColorBrewer) # creates nice color schemes
library(classInt)     # finds class intervals for continuous variables
library(gdata)
library(XLConnect)
library(data.table)
```

### Set data file names:
```{r}
# data file names:
data_file <- "../../clean_data/StopData_merged2.rds" # use the cleaned version of StopData_merged.rds for plotting. See StopData_clean_merged_rds.R for more information.
census_tract_folder <- "../../raw_data/Census_Tract_Polygons2010"
census_tract_shp <- "Census_tracts_2010"
census_tract_csv <- "../../raw_data/Census_tracts_2010.csv"
alameda_xls <- "../../raw_data/2010_Pop_Block_County.xls"
alameda_tract_nums_xls <- "../../raw_data/census_tract_numbers.xls"

```

### Read in the data:
```{r}
# read in the data to modify:
df <- readRDS( data_file )
wb <- loadWorkbook(alameda_xls)
alameda <- readWorksheet(wb, sheet=1,startRow = 5) 
```

### Get some location info from the data:
```{r}
# Location info (to set lim values for coordinates):
latmax <- max( df$lat, na.rm = TRUE )
latmin <- min( df$lat, na.rm = TRUE )
lonmax <- max( df$long, na.rm = TRUE )
lonmin <- min( df$long, na.rm = TRUE )
latvals <- c( latmin, latmax )
lonvals <- c( lonmin, lonmax )
```

### Get the regular Berkeley map:
```{r}
# get the plain Berkeley map from Google:
berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 12)
```

### Draw the regular Berkeley Map:
```{r}
ggmap(berkMap)
```

### Get the census shape files:
```{r}
##locationCensusFiles <- "C:/Users/Rebecca/Dropbox/School/2016_su_School/stat133/finalproject_angry_ladies/Rebeccas_location_Code2/Census_Block_Group_Polygons2010"
#blocks <- readOGR("Census_Block_Group_Polygons2010","Census_blockgroups_2010", verbose = TRUE)
census <- readOGR(census_tract_folder,census_tract_shp, verbose = TRUE)

# blocks.polygons <- blocks@polygons
# blocks.data <- blocks@data
# blocks.plot.order <- blocks@plotOrder
censusT <- spTransform(census, CRS("+proj=longlat +datum=WGS84"))
censusT@data$ID <- as.numeric(censusT@data$ID)
censusTfortified <- fortify(censusT)
#census4 <- census2@data
#pop <- census4$TotalPop
# census3 <- census3 %>%
#   mutate(id = as.factor(id))
# popdf <- as.vector(data.frame(id = as.character(unique(census3$id)), pop = as.character(pop), stringsAsFactors = FALSE, mode = "list" ))
# census5 <- merge(census3, popdf, by = id)

censusCSV <- read.csv(file=census_tract_csv, header = TRUE)

#census3 <- fortify(census2)
censusTmergedCSV <- merge(censusT, censusCSV, by='ID')

# census4$id <- 0:32
# census5 <- merge(census3, census4, by='id')
```

### Trying some things:
```{r}
# library(rgdal)     # R wrapper around GDAL/OGR
# library(ggplot2)   # for general plotting
# library(ggmaps)    # for fortifying shapefiles

# First read in the shapefile, using the path to the shapefile and the shapefile name minus the
# extension as arguments
# Next the shapefile has to be converted to a dataframe for use in ggplot2
shapefile_df <- fortify(censusT)

# Now the shapefile can be plotted as either a geom_path or a geom_polygon.
# Paths handle clipping better. Polygons can be filled.
# You need the aesthetics long, lat, and group.

```

```{r}
ggmap(berkMap) +
geom_path(data = shapefile_df, 
          aes(x = long, y = lat, group = group),
          color = 'black', size = .2) 
```

```{r}
ggmap(berkMap) +
geom_polygon(data = shapefile_df, 
          aes(x = long, y = lat, group = group),
          color = 'black', size = .2) 
```

### Ditch the axes theme:
```{r}
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )
```

### WORKs:
```{r}
# READ IN DATA RIGHT!!!
census.df <- fortify(census2, region = 'GEOID10')
pop <- read.csv(file=census_tract_csv, header = TRUE)
names(pop)[13] <- "id"
pop <- pop %>%
  select(GEOID10, id, TotalPop) %>% 
  mutate( id = str_c("0", as.character(id)) ) 
census.df_pop <- left_join(census.df, pop, by = "id")
```

```{r}
popdenmap <- ggmap(berkMap) + 
  geom_polygon(data = census.df_pop, aes(x = long, y = lat, group = group, fill = TotalPop), color = "white", alpha = .7) 
popdenmap
```

```{r}
popdenmap <- popdenmap + theme_bw()
```

```{r}
popdenmap  +
  coord_fixed(1.3) 
```

```{r}
popdenmap <- popdenmap +
  ditch_the_axes 
popdenmap
```

```{r}
popdenmap + scale_fill_gradient(trans = "log10")
```

```{r}
popamts <- censusCSV$TotalPop
popmin <- min(popamts)
popmin
popmax <- max(popamts)
popmax

#popints <- diff(range(popamts))

popints <- (popamts - min(popamts))/diff(range(popamts)) # recorts the percentage difference from the minimum population

popints_df

```

```{r}
pd2 <- elbow_room1 + 
    scale_fill_gradientn(colours = rev(rainbow(7)),
                         breaks = c(2, 4, 10, 100, 1000, 10000),
                         trans = "log10") 
pd2
```

### All BPD Stops Density, 2015-2016
```{r}
# a contour plot
ggmap(berkMap) +
  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
                 size = 2, bins = 5, data = df, geom = "polygon") +
  scale_fill_gradient(low = "black", high = "red") +
  theme (panel.grid.major = element_blank (), # remove major grid
         panel.grid.minor = element_blank ()  # remove minor grid
  )+ 
  ggtitle("All BPD Stops Density, 2015-2016") +
  labs(alpha = element_blank())+
  guides(alpha = FALSE)

```

### Works, but no google maps:
```{r}
### Read in .shp file only for Census:
shpfile <- str_c(census_tract_folder, "/", census_tract_shp, ".shp")
sh <- readShapePoly(shpfile)
```

```{r}
### Plot the census shape file (polygon outline only)
plot(sh)
```

```{r}
### Modify the shape file to plot the population info:
sh@data$ID <- as.numeric(sh@data$ID)

# Read in the demographic data and merge on Neighbourhood Id
census5_2 <- read.csv(file=census_tract_csv, header = TRUE)
sh2 <- merge(sh, census5_2, by='ID')

# Set the palette
p <- colorRampPalette(c("white", "red"))(128)
palette(p)

# Scale the total population to the palette
popamts <- sh2@data$TotalPop.x

# create the colors:
cols <- (popamts - min(popamts))/diff(range(popamts))*127+1
```

```{r}
### Plot the census shape data with colors for population:
plot(sh, col=cols)
```

### Dealing with population info:
```{r}
#write.csv(alameda, file = "alameda_population_by_tract.csv")
alameda = alameda[-1,]
rownames(alameda) <- 1:nrow(alameda)
names(alameda)[1] <- "GEO"

alameda2 <- alameda %>%
  separate(GEO, c("Block","Group", "Tract", "County", "State"), ", ")

names(alameda2)[3] <- "NAME10"

alameda2 <- alameda2 %>%
  mutate(NAME10 = str_replace(NAME10, "Census Tract ", ""))

alameda2 <- alameda2 %>%
  select(-Block,-Group,-County,-State)

a3 <- alameda2 %>%
        group_by(NAME10)%>% 
        summarise(Total.population = sum(Total.population),
                  Hispanic.or.Latino = sum(Hispanic.or.Latino),
                  Total.population..not.Hispanic.or.Latino = sum(Total.population..not.Hispanic.or.Latino),
                  One.race.total = sum(One.race.total),
                  White = sum(White),
                  American.Indian.and.Alaska.Native = sum(American.Indian.and.Alaska.Native),
                  Asian = sum(Asian),
                  Native.Hawaiian.and.Other.Pacific.Islander = sum(Native.Hawaiian.and.Other.Pacific.Islander),
                  Some.Other.Race = sum(Some.Other.Race),
                  Two.or.More.Races = sum(Two.or.More.Races))

#census_alameda <- left_join(censusCSV, a3, by = "NAMELSAD10")
```

```{r}
# READ IN DATA RIGHT!!!
census.df <- fortify(census2, region = 'NAME10')
pop <- censusCSV
pop <- pop %>%
  select(NAME10, TotalPop) %>%
  mutate(NAME10 = as.character(NAME10))

names(a3)[1] <- "id"
names(pop)[1] <- "id"

pop <- left_join(pop, a3, by = "id")
census.df <- left_join(census.df, pop, by = "id")
write.csv2(a3,"alameda_clean_census.csv")
write.csv2(pop, "berkeley_census_clean.csv")
```

```{r}
popdenmap <- ggmap(berkMap) + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = TotalPop), color = "white", alpha = .7) 
popdenmap
```

### 2016-08-06 final_paper/Shinyapp.Rmd
updated version

@@
 setwd("~/Desktop/Stats R/PROJECT/finalproject_angry_ladies/final_paper")
+```
+
+```{r}
+
+             fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF")
 ```
@@
 rrr <- subset(stop_by, Hour = "7") %>% 
-  group_by(Race) %>% 
+  group_by(AgeRange) %>% 
   tally() 
@@
 ggplot(rrr)+
-  geom_bar(aes(x=Race, y = n),
+  geom_bar(aes(x=AgeRange, y = n),
+           stat = "identity")
+
+ggplot(rrr)+
+  geom_bar(aes(x=AgeRange, y = n),
            stat = "identity",
@@
 ```
+
+```{r}
+subset(stop_by, Hour = "7") %>% 
+    group_by(Agerange) %>% 
+    tally()
+
+  test <- subset(stop_by, Hour = input$hour) %>% 
+    group_by(input$dataset) %>% 
+    tally()
+
+  test$Race<-as.factor(test$Race)
+
+     sidebarPanel(
+      sliderInput('hour',
+                  'HOUR: ',
+                  0, 24, value=7,
+                  animate = T),
+
+```
+
+```{r}
+ggmap(berkMap) +
+  stat_density2d(aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
+    bins = I(5), geom = "polygon", data = mergedf ) +
+  scale_fill_gradient2( "Stop Density",
+    low = "white", mid = "orange", high = "red", midpoint = 25) +
+  labs(x = "Longitude", y = "Latitude") + facet_wrap(~ AgeRange) +
+  scale_alpha(range = c(.2, .55), guide = FALSE) +
+  ggtitle("BPD Stop Contour Map of Berkeley by Age Range") +
+  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
+
+```

### 2016-08-06 final_paper/server.R

@@ -8,6 +8,11 @@ library(plyr)
 
 stop_by <- read_csv("stop_data_jenny.csv")
 
+mapdf <- read_csv("mergedf.csv")
+
+berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14)
+
+
 ##Problems:
 #1. add_legend
 #2. scale
@@ -17,16 +22,15 @@ shinyServer(function(input, output) {
   
   output$Plot <- renderPlot({
   
-  test <- subset(stop_by, Hour = input$hour) %>% 
-    group_by(Race) %>% 
-    tally() 
-
-  rrr$Race<-as.factor(rrr$Race)
-  
-  ggplot(rrr)+
-    geom_bar(aes(x=Race, y = n),
-             stat = "identity",
-             fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))
+  ggmap(berkMap) +
+    stat_density2d(aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
+                   bins = I(5), geom = "polygon", data = mapdf ) +
+    scale_fill_gradient2( "Stop Density",
+                          low = "white", mid = "orange", high = "red", midpoint = 25) +
+    labs(x = "Longitude", y = "Latitude") + facet_wrap(~ input$dataset) +
+    scale_alpha(range = c(.2, .55), guide = FALSE) +
+    ggtitle("BPD Stop Contour Map of Berkeley by Age Range") +
+    guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
   
   })
   

### 2016-08-06 final_paper/ui.R

@@ -10,12 +10,8 @@ shinyUI(fluidPage(
   sidebarLayout(
     
     # Sidebar panel
-    sidebarPanel(
-      sliderInput('hour',
-                  'HOUR: ',
-                  0, 24, value=7,
-                  animate = T)
-      ),
+      selectInput("dataset", "Choose a dataset:", 
+                  choices = c("Race", "AgeRange", "Gender")),
     
     
     # Main panel
@@ -24,5 +20,6 @@ shinyUI(fluidPage(
       )
       )
       ))
+
 
 

### 2016-08-06 eda/Heatmap/HeatmapExploration.Rmd

--- 2016-08-05 eda/Heatmap/HeatmapExploration2.Rmd
@@
 library(classInt)     # finds class intervals for continuous variables
-library(gdata)
-library(XLConnect)
-library(data.table)
 ```
@@
 data_file <- "../../clean_data/StopData_merged2.rds" # use the cleaned version of StopData_merged.rds for plotting. See StopData_clean_merged_rds.R for more information.
-census_tract_folder <- "../../raw_data/Census_Tract_Polygons2010"
+census_tract_folder <- "../../clean_data/Census_Tract_Polygons2010"
 census_tract_shp <- "Census_tracts_2010"
-census_tract_csv <- "../../raw_data/Census_tracts_2010.csv"
-alameda_xls <- "../../raw_data/2010_Pop_Block_County.xls"
-alameda_tract_nums_xls <- "../../raw_data/census_tract_numbers.xls"
-
+census_tract_csv <- "../../clean_data/Census_tracts_2010.csv"
 ```
@@
 df <- readRDS( data_file )
-wb <- loadWorkbook(alameda_xls)
-alameda <- readWorksheet(wb, sheet=1,startRow = 5) 
 ```
@@
 # get the plain Berkeley map from Google:
-berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 12)
+berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 13)
 ```
@@
 # extension as arguments
+shapefile <- censusT
+
 # Next the shapefile has to be converted to a dataframe for use in ggplot2
-shapefile_df <- fortify(censusT)
+shapefile_df <- fortify(shapefile)
 
@@
 
+### Some things that didn't work:
+```{r}
+#usa <- map_data("usa") # we already did this, but we can do it again
+# ggmap(berkMap) + geom_polygon(data = census3, aes(x=long, y = lat, group = group)) + 
+#   coord_fixed(1.3)
+
+# mappp <- ggmap(berkMap) + 
+#   geom_polygon(data = census5, aes(x=lofill = TotalPop), color = "white") +
+#   geom_polygon(color = "black", fill = NA) +
+#   theme_bw() + 
+#   scale_fill_gradient(trans = "log10")
+# mappp
+
+# 
+# ca_base <- ggmap(berkMap) + ggplot(data = census5, mapping = aes(x = long, y = lat, group = group)) + 
+#   coord_fixed(1.3) + 
+#   geom_polygon(color = "black", fill = "gray")
+# ca_base + theme_nothing()
+
+# ggmap(berkMap) + get_map(census5)
+# 
+# 
+#   stat_density2d(aes(x = long, y = lat, color = TotalPop),
+#                  size = 2, bins = 5, data = census5, geom = "polygon") +
+#   theme (panel.grid.major = element_blank (), # remove major grid
+#          panel.grid.minor = element_blank ()  # remove minor grid
+#   ) + 
+#   ggtitle("Population Density by Census 2010 Tract Polygons") +
+#   labs(alpha = element_blank())
+
+```
+
 ### Ditch the axes theme:
@@
 
-### WORKs:
+### Sort of works:
+```{r}
+# census2 <- spTransform(census, CRS("+proj=longlat +datum=WGS84"))
+# census3 <- fortify(census2)
+# census4 <- read.csv(file=census_tract_csv, header = TRUE)
+# census4$id <- 0:32
+# census5 <- merge(census3, census4, by='id')
+```
+
 ```{r}
@@
 ```{r}
-popdenmap <- ggmap(berkMap) + 
-  geom_polygon(data = census.df_pop, aes(x = long, y = lat, group = group, fill = TotalPop), color = "white", alpha = .7) 
-popdenmap
-```
-
-```{r}
-popdenmap <- popdenmap + theme_bw()
-```
-
-```{r}
-popdenmap  +
-  coord_fixed(1.3) 
-```
-
-```{r}
-popdenmap <- popdenmap +
-  ditch_the_axes 
-popdenmap
-```
-
-```{r}
-popdenmap + scale_fill_gradient(trans = "log10")
-```
-
-```{r}
-popamts <- censusCSV$TotalPop
-popmin <- min(popamts)
-popmin
-popmax <- max(popamts)
-popmax
-
-#popints <- diff(range(popamts))
-
-popints <- (popamts - min(popamts))/diff(range(popamts)) # recorts the percentage difference from the minimum population
-
-popints_df
 
@@
 ```{r}
-pd2 <- elbow_room1 + 
+elbow_room1 <- ggmap(berkMap) + 
+  geom_polygon(data = census.df_pop, aes(x = long, y = lat, group = group, fill = TotalPop), color = "white", alpha = .7) 
+elbow_room1 
+```
+
+```{r}
+elbow_room1 <- ggmap(berkMap) + 
+  geom_polygon(data = census.df_pop, aes(x = long, y = lat, group = group, fill = TotalPop), color = "white") +
+  coord_fixed(1.3) 
+elbow_room1 
+```
+
+```{r}
+elbow_room1 + scale_fill_gradient(trans = "log10")
+```
+
+```{r}
+eb2 <- elbow_room1 + 
     scale_fill_gradientn(colours = rev(rainbow(7)),
                          breaks = c(2, 4, 10, 100, 1000, 10000),
-                         trans = "log10") 
-pd2
+                         trans = "log10") +
+  labs(fill = "Total Population")
+eb2
 ```
@@
 ```
-
-### Dealing with population info:
-```{r}
-#write.csv(alameda, file = "alameda_population_by_tract.csv")
-alameda = alameda[-1,]
-rownames(alameda) <- 1:nrow(alameda)
-names(alameda)[1] <- "GEO"
-
-alameda2 <- alameda %>%
-  separate(GEO, c("Block","Group", "Tract", "County", "State"), ", ")
-
-names(alameda2)[3] <- "NAME10"
-
-alameda2 <- alameda2 %>%
-  mutate(NAME10 = str_replace(NAME10, "Census Tract ", ""))
-
-alameda2 <- alameda2 %>%
-  select(-Block,-Group,-County,-State)
-
-a3 <- alameda2 %>%
-        group_by(NAME10)%>% 
-        summarise(Total.population = sum(Total.population),
-                  Hispanic.or.Latino = sum(Hispanic.or.Latino),
-                  Total.population..not.Hispanic.or.Latino = sum(Total.population..not.Hispanic.or.Latino),
-                  One.race.total = sum(One.race.total),
-                  White = sum(White),
-                  American.Indian.and.Alaska.Native = sum(American.Indian.and.Alaska.Native),
-                  Asian = sum(Asian),
-                  Native.Hawaiian.and.Other.Pacific.Islander = sum(Native.Hawaiian.and.Other.Pacific.Islander),
-                  Some.Other.Race = sum(Some.Other.Race),
-                  Two.or.More.Races = sum(Two.or.More.Races))
-
-#census_alameda <- left_join(censusCSV, a3, by = "NAMELSAD10")
-```
-
-```{r}
-# READ IN DATA RIGHT!!!
-census.df <- fortify(census2, region = 'NAME10')
-pop <- censusCSV
-pop <- pop %>%
-  select(NAME10, TotalPop) %>%
-  mutate(NAME10 = as.character(NAME10))
-
-names(a3)[1] <- "id"
-names(pop)[1] <- "id"
-
-pop <- left_join(pop, a3, by = "id")
-census.df <- left_join(census.df, pop, by = "id")
-write.csv2(a3,"alameda_clean_census.csv")
-write.csv2(pop, "berkeley_census_clean.csv")
-```
-
-```{r}
-popdenmap <- ggmap(berkMap) + 
-  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = TotalPop), color = "white", alpha = .7) 
-popdenmap
-```

### 2016-08-06 eda/Heatmap/HeatmapExploration3.Rmd

---
title: "Heat Map Exploration"
author: "Rebecca Reus"
date: "August 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r, include = FALSE, echo = FALSE}
library(sp)
library(ggmap)
library(tidyr)
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(maps)
library(mapdata)
library(rgdal)
library(rgeos)
#library(gpclib)       # loads polygon clipping library
library(maptools)     # loads sp library too
library(RColorBrewer) # creates nice color schemes
library(classInt)     # finds class intervals for continuous variables
library(gdata)
library(XLConnect)
#library(data.table)
```

### Set data file names:
```{r}
# data file names:
data_file <- "../../clean_data/StopData_merged2.rds" # use the cleaned version of StopData_merged.rds for plotting. See StopData_clean_merged_rds.R for more information.
census_tract_folder <- "../../raw_data/Census_Tract_Polygons2010"
census_tract_shp <- "Census_tracts_2010"
census_tract_csv <- "../../raw_data/Census_tracts_2010.csv"
alameda_xls <- "../../raw_data/2010_Pop_Block_County.xls"
alameda_tract_nums_xls <- "../../raw_data/census_tract_numbers.xls"
alameda_clean_csv <- "../../raw_data/alameda_clean_census.csv"
berkeley_clean_csv <- "../../clean_data/berkeley_census_clean.csv"

```

### Read in the data:
```{r}
# read in the data to modify:
df <- readRDS( data_file )
berkpop <- read.csv2(berkeley_clean_csv)
berkpop <- berkpop[,-1]
a3 <- read.csv2(alameda_clean_csv)
a3 <- a3[,-1]
```

```{r, eval = F, include = F}
# Used to clean the alameda data, do not run.
wb <- loadWorkbook(alameda_xls)
alameda <- readWorksheet(wb, sheet=1,startRow = 5) 
```

```{r}
# Location info (to set lim values for coordinates):
latmax <- max( df$lat, na.rm = TRUE )
latmin <- min( df$lat, na.rm = TRUE )
lonmax <- max( df$long, na.rm = TRUE )
lonmin <- min( df$long, na.rm = TRUE )
latvals <- c( latmin, latmax )
lonvals <- c( lonmin, lonmax )
```

### Get the regular Berkeley map:
```{r,include = F, echo = FALSE}
# get rid of the axes:
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )
```
```{r,include = F, echo = FALSE}
# get the plain Berkeley map from Google:
berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 12) 
```

```{r}
berkgg <-ggmap(berkMap) +
  xlim(-122.335, lonmax) + ylim(latvals) +
  ditch_the_axes 
berkgg
```

### Get the census shape files:
```{r, include=FALSE, echo = FALSE}

census <- readOGR(census_tract_folder,census_tract_shp, verbose = TRUE)

censusT <- spTransform(census, CRS("+proj=longlat +datum=WGS84"))
censusT@data$ID <- as.numeric(censusT@data$ID)

censusCSV <- read.csv(file=census_tract_csv, header = TRUE)
```

### Trying some things:
```{r, eval = F, include = F}
# library(rgdal)     # R wrapper around GDAL/OGR
# library(ggplot2)   # for general plotting
# library(ggmaps)    # for fortifying shapefiles

# First read in the shapefile, using the path to the shapefile and the shapefile name minus the
# extension as arguments
# Next the shapefile has to be converted to a dataframe for use in ggplot2
shapefile_df <- fortify(censusT)

# Now the shapefile can be plotted as either a geom_path or a geom_polygon.
# Paths handle clipping better. Polygons can be filled.
# You need the aesthetics long, lat, and group.

```

```{r, eval = F, include = F}
ggmap(berkMap) +
geom_path(data = shapefile_df, 
          aes(x = long, y = lat, group = group),
          color = 'black', size = .2) 
```

```{r, eval = F, include = F}
ggmap(berkMap) +
geom_polygon(data = shapefile_df, 
          aes(x = long, y = lat, group = group),
          color = 'black', size = .2) 
```

### Examine total population by census block:
```{r}
# READ IN DATA RIGHT!!!
census.df <- fortify(censusT, region = 'GEOID10')
pop <- read.csv(file=census_tract_csv, header = TRUE)
names(pop)[13] <- "id"
pop <- pop %>%
  select(GEOID10, id, TotalPop) %>% 
  mutate( id = str_c("0", as.character(id)) ) 
census.df_pop <- left_join(census.df, pop, by = "id")
```

```{r}
popdenmap <- ggmap(berkMap) + 
  geom_polygon(data = census.df_pop, aes(x = long, y = lat, group = group, fill = TotalPop), color = "white", alpha = .7) 
popdenmap
```

```{r, eval = F, include = F}
popdenmap <- popdenmap + theme_bw()
```

```{r, eval = F, include = F}
popdenmap  +
  coord_fixed(1.3) 
```

```{r, eval = F, include = F}
popdenmap <- popdenmap +
  ditch_the_axes 
popdenmap
```

```{r, eval = F, include = F}
popdenmap + scale_fill_gradient(trans = "log10")
```

```{r, eval = F, include = F}
popamts <- censusCSV$TotalPop
popmin <- min(popamts)
popmin
popmax <- max(popamts)
popmax

#popints <- diff(range(popamts))

popints <- (popamts - min(popamts))/diff(range(popamts)) # recorts the percentage difference from the minimum population

#popints_df 
```

```{r, eval = F, include = F}
pd2 <- popdenmap + 
    scale_fill_gradientn(colours = rev(rainbow(7)),
                         breaks = c(2, 4, 10, 100, 1000, 10000),
                         trans = "log10") 
pd2
```

```{r, include = F, echo = F}
bpop2 <- berkpop %>%
  select(-Total.population..not.Hispanic.or.Latino) %>%
   select(-Two.or.More.Races) %>%
  select(-One.race.total)

bpop2$Other = bpop2$American.Indian.and.Alaska.Native +
                  bpop2$Native.Hawaiian.and.Other.Pacific.Islander +
                  bpop2$Some.Other.Race

bpop2 <- bpop2 %>%
  mutate(Percent.Black = Black.or.African.American/Total.population) %>%
  mutate(Percent.Other = Other/Total.population) %>%
  mutate(Percent.White = White/Total.population) %>%
  mutate(Percent.Asian = Asian/Total.population) %>%
  mutate(Percent.Hispanic = Hispanic.or.Latino/Total.population) %>%
  mutate(Percent.Berkeley = Total.population/sum(Total.population)) %>%
  select(-TotalPop) %>%
   select(-American.Indian.and.Alaska.Native,-Native.Hawaiian.and.Other.Pacific.Islander, -Some.Other.Race)

# %>%
#   mutate(Relative.Per.Black = Black.or.African.American/sum(Black.or.African.American)) %>%
#   mutate(Relative.Per.Other = Other/sum(Other)) %>%
#   mutate(Relative.Per.White = White/sum(White)) %>%
#   mutate(Realtive.Per.Asian = Asian/sum(Asian)) %>%
#   mutate(Relative.Per.Hispanic = Hispanic.or.Latino/sum(Hispanic.or.Latino)) %>%

names(bpop2)[2] <- "Population"
names(bpop2)[3] <- "Hispanic"
names(bpop2)[5] <- "Black"
names(bpop2)

```

```{r, include = FALSE, echo = FALSE}
# read in the data:
census.df <- fortify(censusT, region = 'NAME10')
census.df <- left_join(census.df, bpop2, by = "id")
```

Population by Race:
```{r}
p <- colorRampPalette(c("white", "red"))(128)
palette(p)

# Scale the total population to the palette
#popamts <- sh2@data$TotalPop.x

# create the colors:
# cols <- (popamts - min(popamts))/diff(range(popamts))*127+1

black1 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Black.or.African.American), color = "white", alpha = .5)  +
  ggtitle("Black Population") + 
  scale_fill_gradientn(trans = "log10", colors = p) +
  labs(fill = "Population\nLog Scale")
black1

# +
#   labs(fill = element_blank())+
#   guides(fill = FALSE)

black2 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Percent.Black), color = "white", alpha = .7) +
  ggtitle("Percent Black") +
  scale_fill_gradientn(trans = "log10", colors = p) +
  labs(fill = "Population\nDensity\nLog Scale")
black2
```

View side by side:
```{r}
require(gridExtra)
grid.arrange(black1, black2, ncol=2)
```

### BPD Stops:
```{r}
# edit bpd stops:
# nblacks <- sum(df$Race == "Black")
# nasians <- sum(df$Race == "Asian")
# nwhite <- sum(df$Race == "White")
# nother <- sum(df$Race == "Other")
# nhisp <- sum(df$Race == "Hispanic")
# totalpop <- sum(berkpop$TotalPop)
# stop.summary <- data.frame(Race = unique(df$Race))

stop.summary <- df %>%  
   group_by(Race) %>%
   summarise(Stopped=n() ) %>%
  mutate(Percent.Stopped = Stopped/nrow(df))

racepops <- c(sum(bpop2$Asian), sum(bpop2$Black), sum(bpop2$Hispanic), sum(bpop2$Other), sum(bpop2$White) )

stop.summary$Race.Population = racepops
stop.summary$Percent.Race.Stopped = stop.summary$Stopped/stop.summary$Race.Population

```

```{r}
# all stop density:
allbpd_density <- berkgg +
  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = 2* ..level..),
                 size = 2, bins = 5, data = df, geom = "polygon") +
  scale_fill_gradientn(trans = "log10", colors = p) +
  theme (panel.grid.major = element_blank (), # remove major grid
         panel.grid.minor = element_blank ()  # remove minor grid
  ) + 
  ggtitle("All BPD Stops Density, 2015-2016") +
  labs(alpha = element_blank()) +
  guides(alpha = FALSE)
allbpd_density

# by race:
bpd_race <- berkgg +
  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha =  ..level..),
                 size = 2, bins = 5, data = df, geom = "polygon") +
  scale_fill_gradientn(trans = "log10", colors = p) +
  theme (panel.grid.major = element_blank (), # remove major grid
         panel.grid.minor = element_blank ()  # remove minor grid
  ) + 
  guides(alpha = FALSE) +
  facet_grid( ~ Race)
bpd_race

```

### Bar Charts:

```{r}
p <- ggplot(df, aes(Enforcement, fill = Enforcement)) + 
    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  facet_grid(~Race) +
  scale_y_continuous(labels=percent) +
  labs(y = "Percent Stopped") +
  ggtitle("Results of Stops by Race")
p
```

```{r}
p1 <- ggplot(df, aes(Enforcement, fill = Enforcement)) + 
    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(y = "Percent Stopped") + ggtitle("Results of Stops") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )   + guides(fill = FALSE)

p1race <- ggplot(df, aes(Enforcement, fill = Enforcement)) + 
    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(y = "Percent Stopped") +
  ggtitle("Results of Stops by Race") +
  facet_grid( ~ Race) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

```

```{r}
grid.arrange(p1,p1race,ncol=2)
```

```{r}

p2 <- ggplot(df, aes(Reason, fill = Reason)) + 
    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
+  
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(y = "Percent") + ggtitle("Reason for Stops") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )   + guides(fill = FALSE)

p2race <- ggplot(df, aes(Reason, fill = Reason)) + 
    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(y = "Percent Stopped") +
  ggtitle("Reason for Stops") +
  facet_grid( ~ Race) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

p3 <- ggplot(df, aes(CarSearch, fill = CarSearch)) + 
    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(y = "Percent Stopped") +
  ggtitle("Car Search by Race") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )   + guides(fill = FALSE)

p3race <- ggplot(df, aes(Reason, fill = Reason)) + 
    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(y = "Percent Stopped") +
  ggtitle("Car Search by Race") +
  facet_grid( ~ Race) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

df$AgeRange = factor(df$AgeRange)
levels(df$AgeRange) <- c("10-19","20-29","30-39","40-49")

p4 <- ggplot(df, aes(AgeRange, fill = AgeRange )) + 
    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(y = "Percent Stopped") +
  ggtitle("Age Range of Car Stops by Race") +
  labs(x = "Age Range") +
  labs(fill = "Age Range")

```

```{r}

```

### 2016-08-06 eda/Location/StopData_Maps.Rmd

--- 2016-08-03 Rebeccas_location_Code2/StopData_Maps.Rmd
@@
 library(RgoogleMaps)
-library(tmap)
+#library(tmap)
 library(sp)

### 2016-08-07 eda/Heatmap/HeatmapExploration4.Rmd
heatmaps

---
title: "Heat Map Exploration"
author: "Rebecca Reus"
date: "August 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r, include = FALSE, echo = FALSE, warning = FALSE}
library(sp)
library(ggmap)
library(tidyr)
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(maps)
library(mapdata)
library(rgdal)
library(rgeos)
#library(gpclib)       # loads polygon clipping library
library(maptools)     # loads sp library too
library(RColorBrewer) # creates nice color schemes
library(classInt)     # finds class intervals for continuous variables
library(gdata)
library(XLConnect)
library(data.table)
library(gridExtra)
```

### Set data file names:
```{r}
# data file names:
data_file <- "../../clean_data/StopData_merged2.rds" # use the cleaned version of StopData_merged.rds for plotting. See StopData_clean_merged_rds.R for more information.
census_tract_folder <- "../../raw_data/Census_Tract_Polygons2010"
census_tract_shp <- "Census_tracts_2010"
census_tract_csv <- "../../raw_data/Census_tracts_2010.csv"
alameda_xls <- "../../raw_data/2010_Pop_Block_County.xls"
alameda_tract_nums_xls <- "../../raw_data/census_tract_numbers.xls"
alameda_clean_csv <- "../../raw_data/alameda_clean_census.csv"
berkeley_clean_csv <- "../../clean_data/berkeley_census_clean.csv"
```

### Read in the data:
```{r}
# read in the data to modify:
df <- readRDS( data_file )
berkpop <- read.csv2(berkeley_clean_csv)
berkpop <- berkpop[,-1]
a3 <- read.csv2(alameda_clean_csv)
a3 <- a3[,-1]
```

```{r, eval = F, include = F}
# Used to clean the alameda data, do not run.
wb <- loadWorkbook(alameda_xls)
alameda <- readWorksheet(wb, sheet=1,startRow = 5) 
```

```{r}
# Location info (to set lim values for coordinates):
latmax <- max( df$lat, na.rm = TRUE )
latmin <- min( df$lat, na.rm = TRUE )
lonmax <- max( df$long, na.rm = TRUE )
lonmin <- min( df$long, na.rm = TRUE )
latvals <- c( latmin, latmax )
lonvals <- c( lonmin, lonmax )
```

### Get the regular Berkeley map:
```{r,include = F, echo = FALSE}
# get rid of the axes:
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )

# get the plain Berkeley map from Google:
berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 12)

berkgg <-ggmap(berkMap) +
  xlim(-122.335, lonmax) + ylim(latvals) +
  ditch_the_axes 
berkgg
```

### Get the census shape files:
```{r, include=FALSE, echo = FALSE}

census <- readOGR(census_tract_folder,census_tract_shp, verbose = TRUE)

censusT <- spTransform(census, CRS("+proj=longlat +datum=WGS84"))
censusT@data$ID <- as.numeric(censusT@data$ID)

censusCSV <- read.csv(file=census_tract_csv, header = TRUE)
```

### Examine total population by census block:
```{r, include = F, message = F}
# READ IN DATA RIGHT!!!
census.df <- fortify(censusT, region = 'GEOID10')
pop <- read.csv(file=census_tract_csv, header = TRUE)
names(pop)[13] <- "id"
pop <- pop %>%
  select(GEOID10, id, TotalPop) %>% 
  mutate( id = str_c("0", as.character(id)) ) 
census.df_pop <- left_join(census.df, pop, by = "id")
```

```{r, include = F}
popdenmap <- berkgg + 
  geom_polygon(data = census.df_pop, aes(x = long, y = lat, group = group, fill = TotalPop), color = "white", alpha = .7) 
popdenmap
```

```{r,include = F}
pd2 <- popdenmap + 
    scale_fill_gradientn(colours = rev(rainbow(7) ) ) 
pd2
```

## Population density maps:
```{r}
grid.arrange(popdenmap,pd2, ncol = 2)
```

```{r, include = F, echo = F}
bpop2 <- berkpop %>%
  select(-Total.population..not.Hispanic.or.Latino) %>%
   select(-Two.or.More.Races) %>%
  select(-One.race.total)

bpop2$Other = bpop2$American.Indian.and.Alaska.Native +
                  bpop2$Native.Hawaiian.and.Other.Pacific.Islander +
                  bpop2$Some.Other.Race

bpop2 <- bpop2 %>%
  mutate(Percent.Black = Black.or.African.American/Total.population) %>%
  mutate(Percent.Other = Other/Total.population) %>%
  mutate(Percent.White = White/Total.population) %>%
  mutate(Percent.Asian = Asian/Total.population) %>%
  mutate(Percent.Hispanic = Hispanic.or.Latino/Total.population) %>%
  mutate(Percent.Berkeley = Total.population/sum(Total.population)) %>%
  select(-TotalPop) %>%
   select(-American.Indian.and.Alaska.Native,-Native.Hawaiian.and.Other.Pacific.Islander, -Some.Other.Race)

# %>%
#   mutate(Relative.Per.Black = Black.or.African.American/sum(Black.or.African.American)) %>%
#   mutate(Relative.Per.Other = Other/sum(Other)) %>%
#   mutate(Relative.Per.White = White/sum(White)) %>%
#   mutate(Realtive.Per.Asian = Asian/sum(Asian)) %>%
#   mutate(Relative.Per.Hispanic = Hispanic.or.Latino/sum(Hispanic.or.Latino)) %>%

names(bpop2)[2] <- "Population"
names(bpop2)[3] <- "Hispanic"
names(bpop2)[5] <- "Black"
names(bpop2)

# write the clean berkeley census 2010 data to file:
#write_csv(x = bpop2, path = "../../clean_data/berk_census2010_clean.csv")

```

```{r, include = FALSE, echo = FALSE, warning=FALSE}
# read in the data:
census.df <- fortify(censusT, region = 'NAME10')
census.df <- left_join(census.df, bpop2, by = "id")
#write_csv(x = census.df, path = "../../clean_data/berk_census2010_mapdata.csv")
```

Population by Race:
```{r, include = F, echo = F, warning = FALSE}

# create color palette:

p <- colorRampPalette(c("white", "red"))(128)
palette(p)

# black population:

black1 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Black), color = "white", alpha = .5)  +
  ggtitle("Berkeley Census 2010:\nBlack Population") + 
  scale_fill_gradientn(colors = p) +
  labs(fill = "Count")

black2 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Percent.Black), color = "white", alpha = .7) +
  ggtitle("Berkeley Census 2010:\nPercent Black") +
  scale_fill_gradientn(colors = p) +
  labs(fill = "Density")

# white population:

w1 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = White), color = "white", alpha = .5)  +
  ggtitle("Berkeley Census 2010:\nWhite Population") + 
  scale_fill_gradientn(colors = p) +
  labs(fill = "Count")

w2 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Percent.White), color = "white", alpha = .7) +
  ggtitle("Berkeley Census 2010:\nPercent White") +
  scale_fill_gradientn(colors = p) +
  labs(fill = "Density")

# asian population:

a1 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Asian), color = "white", alpha = .5)  +
  ggtitle("Berkeley Census 2010:\nAsian Population") + 
  scale_fill_gradientn(colors = p) +
  labs(fill = "Count")

a2 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Percent.Asian), color = "white", alpha = .7) +
  ggtitle("Berkeley Census 2010:\nPercent Asian") +
  scale_fill_gradientn(colors = p) +
  labs(fill = "Density")

# hispanic population:

h1 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Hispanic), color = "white", alpha = .5)  +
  ggtitle("Berkeley Census 2010:\nHispanic Population") + 
  scale_fill_gradientn(colors = p) +
  labs(fill = "Count")

h2 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Percent.Hispanic), color = "white", alpha = .7) +
  ggtitle("Berkeley Census 2010:\nPercent Hispanic") +
  scale_fill_gradientn(colors = p) +
  labs(fill = "Density")

# other population:

o1 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Other), color = "white", alpha = .5)  +
  ggtitle("Berkeley Census 2010:\nOther Population") + 
  scale_fill_gradientn(colors = p) +
  labs(fill = "Count")

o2 <- berkgg + 
  geom_polygon(data = census.df, aes(x = long, y = lat, group = group, fill = Percent.Other), color = "white", alpha = .7) +
  ggtitle("Berkeley Census 2010:\nPercent Other") +
  scale_fill_gradientn(colors = p) +
  labs(fill = "Density")

```

View side by side:
```{r, fig.width = 15}
grid.arrange(w1, black1, h1, a1, o1, ncol = 5)
grid.arrange(w2, black2, h2, a2, o2, ncol = 5)
```

### BPD Stops:
```{r}
# edit bpd stops:
# nblacks <- sum(df$Race == "Black")
# nasians <- sum(df$Race == "Asian")
# nwhite <- sum(df$Race == "White")
# nother <- sum(df$Race == "Other")
# nhisp <- sum(df$Race == "Hispanic")
# totalpop <- sum(berkpop$TotalPop)
# stop.summary <- data.frame(Race = unique(df$Race))

stop.summary <- df %>%  
   group_by(Race) %>%
   summarise(Stopped=n() ) %>%
  mutate(Percent.Stopped = Stopped/nrow(df))

racepops <- c(sum(bpop2$Asian), sum(bpop2$Black), sum(bpop2$Hispanic), sum(bpop2$Other), sum(bpop2$White) )

stop.summary$Race.Population = racepops
stop.summary$Percent.Race.Stopped = stop.summary$Stopped/stop.summary$Race.Population

#write.csv(x = stop.summary, file = "../../clean_data/StopData_summary.csv")

```

```{r, include = F, echo = F, warning = F, message = F}
# all stop density:
allbpd_density <- berkgg +
  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha =  ..level../100),
                 size = 2, bins = 5, data = df, geom = "polygon") +
  scale_fill_gradientn( colors = p) +
  theme (panel.grid.major = element_blank (), # remove major grid
         panel.grid.minor = element_blank ()  # remove minor grid
  ) + 
  ggtitle("All BPD Stops Density, 2015-2016") +
  labs(fill = "Stops") +
  guides(alpha = FALSE)
allbpd_density

# by race:
bpd_race <- berkgg +
  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha =  ..level../100),
                 size = 2, bins = 5, data = df, geom = "polygon") +
  scale_fill_gradientn(trans = "log10", colors = p) +
  theme (panel.grid.major = element_blank (), # remove major grid
         panel.grid.minor = element_blank ()  # remove minor grid
  ) + 
  guides(alpha = FALSE) +
  facet_grid( ~ Race)
bpd_race
```

```{r, eval = T}
allbpd_density
bpd_race
```

```{r. eval = F}
# data file names:
data_file <- "../../raw_data/StopData_merged2.rds" # use the cleaned version of StopData_merged.rds for plotting. See StopData_clean_merged_rds.R for more information.
census_tract_folder <- "../../raw_data/Census_Tract_Polygons2010"
census_tract_shp <- "Census_tracts_2010"
census_tract_csv <- "../../raw_data/Census_tracts_2010.csv"
berkeley_clean_csv <- "../../raw_data/berkeley_census_clean.csv"
census <- readOGR(census_tract_folder,census_tract_shp, verbose = TRUE)

censusT <- spTransform(census, CRS("+proj=longlat +datum=WGS84"))
censusT@data$ID <- as.numeric(censusT@data$ID)

censusCSV <- read.csv(file=census_tract_csv, header = TRUE)
census.df <- fortify(censusT, region = 'NAME10')

#berkcensus2010map <- census.df

df <- readRDS( data_file )
berkpop <- read.csv2(berkeley_clean_csv)
berkpop <- berkpop[,-1]

bpop2 <- berkpop %>%
  select(-Total.population..not.Hispanic.or.Latino) %>%
   select(-Two.or.More.Races) %>%
  select(-One.race.total)

bpop2$Other = bpop2$American.Indian.and.Alaska.Native +
                  bpop2$Native.Hawaiian.and.Other.Pacific.Islander +
                  bpop2$Some.Other.Race

bpop2 <- bpop2 %>%
  mutate(Percent.Black = Black.or.African.American/Total.population) %>%
  mutate(Percent.Other = Other/Total.population) %>%
  mutate(Percent.White = White/Total.population) %>%
  mutate(Percent.Asian = Asian/Total.population) %>%
  mutate(Percent.Hispanic = Hispanic.or.Latino/Total.population) %>%
  mutate(Percent.Berkeley = Total.population/sum(Total.population)) %>%
  select(-TotalPop) %>%
   select(-American.Indian.and.Alaska.Native,-Native.Hawaiian.and.Other.Pacific.Islander, -Some.Other.Race)

names(bpop2)[2] <- "Population"
names(bpop2)[3] <- "Hispanic"
names(bpop2)[5] <- "Black"
names(bpop2)

#berkcensus2010$id <- as.character(berkcensus2010$id)
census.df <- left_join(census.df, bpop2, by = "id")
berkcensus2010map <- census.df

saveRDS(berkcensus2010map, "../../clean_data/berk_census2010_mapdata.rds")

```

### 2016-08-07 eda/Heatmap/Results.Rmd
heatmaps

---
title: "Results"
author: "Rebecca Reus"
date: "August 7, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, echo = FALSE, warning = FALSE}
library(sp)
library(ggmap)
library(tidyr)
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(maps)
library(mapdata)
library(rgdal)
library(rgeos)
#library(gpclib)       # loads polygon clipping library
library(maptools)     # loads sp library too
library(RColorBrewer) # creates nice color schemes
library(classInt)     # finds class intervals for continuous variables
library(gdata)
library(XLConnect)
library(data.table)
library(gridExtra)
library(pander)
library(formattable)

# data file names:

file_stops <- "../../clean_data/StopData_clean.csv" 
file_berkcensus2010 <- "../../clean_data/berk_census2010_clean.csv" 
file_berkcensus2010map <-"../../clean_data/berk_census2010_mapdata.rds" 
file_stopsummary <- "../../clean_data/StopData_summary.csv"

# read in the data:

stops <- read_csv(file = file_stops)
berkcensus2010 <- read_csv(file = file_berkcensus2010)
berkcensus2010map <- readRDS(file_berkcensus2010map)
stopsummary <- read.csv( file_stopsummary )
stopsummary <- select(stopsummary, -X)

# Location info (to set lim values for coordinates):

latmax <- max( stops$lat, na.rm = TRUE )
latmin <- min( stops$lat, na.rm = TRUE )
lonmax <- max( stops$long, na.rm = TRUE )
lonmin <- min( stops$long, na.rm = TRUE )
latvals <- c( latmin, latmax )
lonvals <- c( lonmin, lonmax )

# get rid of the axes theme:

ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )

# get the plain Berkeley map from Google:
berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 12)

berkgg <-ggmap(berkMap) +
  xlim(-122.335, lonmax) + ylim(latvals) +
  ditch_the_axes 
#berkgg

```

```{r, include = FALSE, echo = FALSE, warning = FALSE}

# create color palette:

p <- colorRampPalette(c("white", "red"))(128)
palette(p)

# create the maps:
popdenmap <- berkgg + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Berkeley), color = "white", alpha = .7) +
  ggtitle("2010 Population Density") +
  labs(fill = "Percent") +
  scale_fill_gradientn(colors = p)
popdenmap

# pd2 <- popdenmap + 
#   scale_fill_gradientn(colours = rev(rainbow(7) ) ) +
#   ggtitle("Berkeley Census 2010:\nPopulation")
# pd2

# black population:

black2 <- berkgg + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Black), color = "white", alpha = .7) +
  scale_fill_gradientn(colors = p, limits = c(0,1)) +
  guides( fill = "none" ) +
  ggtitle("% Black")

# white population:

w2 <- berkgg + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.White), color = "white", alpha = .7) +
  scale_fill_gradientn(colors = p, limits = c(0,1)) +
  guides( fill = "none" ) +
  ggtitle("% White")

# asian population:

a2 <- berkgg + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Asian), color = "white", alpha = .7) +
  scale_fill_gradientn(colors = p, limits = c(0,1)) +
  guides( fill = "none" ) +
  ggtitle("% Asian")

# hispanic population:

h2 <- berkgg + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Hispanic), color = "white", alpha = .7) +
  scale_fill_gradientn(colors = p, limits = c(0,1)) +
  guides( fill = "none" ) +
  ggtitle( "% Hispanic")

# other population:

o2 <- berkgg + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Other), color = "white", alpha = .7) +
  scale_fill_gradientn(colors = p, limits = c(0,1)) +
  guides( fill = "none" ) +
  ggtitle("% Other")
```

## Population Density Maps

```{r, warning = FALSE, echo = F, fig.width = 5, fig.height = 5}
popdenmap
```

```{r, warning = FALSE, echo = FALSE,  fig.width = 15, fig.height = 5}
grid.arrange(w2,black2,a2,h2,o2, ncol = 5)
```

## Comparing Stop Density to Population Density

```{r, message = F, echo = FALSE}
ss <- stopsummary %>%
  select(Race, Percent.Stopped, Percent.Race.Stopped)

ss$Percent.Stopped <- percent(ss$Percent.Stopped)
ss$Percent.Race.Stopped <- percent(ss$Percent.Race.Stopped)

names(ss)[2:3] <- c("Percent Stopped", "Percent Population Stopped")

panderOptions("digits", 2)
pander(ss)

bargraph1 <- ggplot(ss, aes(Race, `Percent Stopped`, fill = Race))+
  geom_bar(stat = "identity") +
  guides( fill = FALSE)
bargraph2 <- ggplot(ss, aes(Race, `Percent Population Stopped`, fill = Race))+
  geom_bar(stat = "identity") +
  guides( fill = FALSE)
grid.arrange(bargraph1, bargraph2, ncol = 2)
```

### 2016-08-08 eda/PaperDrafts/draft2.Rmd
test

---
title: "STAT 133 Final Project: Visualizing Berkeley Police Data"
author: "Amy Zhu, Mengyu Li, Rebecca Reus and Shangjun Jiang"
date: "August 7, 2016"
output: html_document
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sp)
library(ggmap)
library(tidyr)
library(readr)
library(plyr)
library(dplyr)
library(ggvis)
library(tidyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(data.table)
library(maps)
library(mapdata)
library(maptools)
library(scales)
library(RgoogleMaps)
library(rgdal)
library(rgeos)
library(RColorBrewer)
library(grid)
library(gridExtra)
library(pander)
library(formattable)
```

```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
# read in the data and define parameters:

# data file names:

file_stops <- "../../clean_data/StopData_clean.csv" 
file_berkcensus2010 <- "../../clean_data/berk_census2010_clean.csv" 
file_berkcensus2010map <-"../../clean_data/berk_census2010_mapdata.rds" 
file_tables <-  "../../clean_data/Tables_AgeSexRacePopStop.rds"
file_calls <-  "../../clean_data/CallsForService.rds"
file_arrests <- "../../raw_data/Berkeley_PD_Log_-_Arrests.csv"
file_jail <- "../../raw_data/Berkeley_PD_Log_-_Jail_Bookings.csv"

# read in the data:

# Police Stop Data:
stops <- read_csv(file = file_stops)

# Census 2010 Data by Polygonal Tract Number:
berkcensus2010 <- read_csv(file = file_berkcensus2010)

# Census 2010 Data by Polygonal Tract Number, ready for mapping:
berkcensus2010map <- readRDS(file_berkcensus2010map)

# Police Calls for service:
callservice <- read_csv(file = file_calls)

# Tables to use in paper:
tables <- readRDS(file = file_tables)
ss <- as.data.frame(tables[5]) # stop data by race proportional to population
pop1 <- as.data.frame(tables[4]) # census data population count table.
race <- as.data.frame(tables[3]) # race population count
sex <- as.data.frame(tables[2]) # sex population count
age <- as.data.frame(tables[1]) # age population count

# arrest data:
arrest <- read_csv(file_arrests)

# jail data:
jail <- read.csv(file_jail)

## Location info:

latmax <- max(stops$lat, na.rm = TRUE) 
latmin <- min(stops$lat, na.rm = TRUE)
lonmax <- max(stops$long, na.rm = TRUE) 
lonmin <- min(stops$long, na.rm = TRUE)
latvals <- c(latmin, latmax)
lonvals <- c(lonmin, lonmax)

# get rid of the axes theme:

ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )

# get the plain Berkeley map from Google:

# zoomed out berkeley map:
berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 12) 
berkgg_zoom1 <-ggmap(berkMap) +
  xlim(-122.335, lonmax) + ylim(latvals) +
  ditch_the_axes

# zoomed in berkeley map:
berkMap2 = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14) 
berkgg_zoom2 <-ggmap(berkMap2) +
  ditch_the_axes

### CENSUS DATA MAPS:

# create color palette:

p <- colorRampPalette(c("white", "red"))(128)
palette(p)

# create the maps:

# stop data map
map_stop1 <- berkgg_zoom2 +
  geom_point(data = stops, aes(x=long,y=lat), alpha = .1) +
  ggtitle("Berkeley Police Stops, 2015-2016")

map_stop1

# populatrion density map
popdenmap <- berkgg_zoom1 + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Berkeley), color = "white", alpha = .7) +
  ggtitle("2010 Population Density") +
  labs(fill = "Percent") +
  scale_fill_gradientn(colors = p)

# black population:

black2 <- berkgg_zoom1 + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Black), color = "white", alpha = .7) +
  scale_fill_gradientn(colors = p, limits = c(0,1)) +
  guides( fill = "none" ) +
  ggtitle("% Black")

# white population:

w2 <- berkgg_zoom1 + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.White), color = "white", alpha = .7) +
  scale_fill_gradientn(colors = p, limits = c(0,1)) +
  guides( fill = "none" ) +
  ggtitle("% White")

# asian population:

a2 <- berkgg_zoom1 + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Asian), color = "white", alpha = .7) +
  scale_fill_gradientn(colors = p, limits = c(0,1)) +
  guides( fill = "none" ) +
  ggtitle("% Asian")

# hispanic population:

h2 <- berkgg_zoom1 + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Hispanic), color = "white", alpha = .7) +
  scale_fill_gradientn(colors = p, limits = c(0,1)) +
  guides( fill = "none" ) +
  ggtitle( "% Hispanic")

# other population:

o2 <- berkgg_zoom1 + 
  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Other), color = "white", alpha = .7) +
  scale_fill_gradientn(colors = p, limits = c(0,1)) +
  guides( fill = "none" ) +
  ggtitle("% Other")

# bar graphs based on stop data:
names(ss)[2:3] <- c("Percent Stopped", "Percent Population Stopped")
bargraph1 <- ggplot(ss, aes(Race, `Percent Stopped`, fill = Race))+
  geom_bar(stat = "identity") +
  guides( fill = FALSE)
bargraph2 <- ggplot(ss, aes(Race, `Percent Population Stopped`, fill = Race))+
  geom_bar(stat = "identity") +
  guides( fill = FALSE)

# STOP DENSITY MAP:
stopdensitymap <- berkgg_zoom2 +
  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
    size = 2, bins = 5, data = stops, geom = "polygon") +
  scale_fill_gradient(low = "black", high = "red") +
    theme (panel.grid.major = element_blank (), # remove major grid
		       panel.grid.minor = element_blank ()  # remove minor grid
		       )+ 
  ggtitle("All BPD Stops Density, 2015-2016") +
  labs(alpha = element_blank())+
  guides(alpha = FALSE)

# stop density 
stopdensitymap4 <- berkgg_zoom2 +
  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
    bins = I(5), geom = "polygon", data = stops ) +
  scale_fill_gradient2( "StopDensity",
    low = "white", mid = "orange", high = "red", midpoint = 100) +
  labs(x = "Longitude", y = "Latitude") +
  scale_alpha(range = c(.2, .55), guide = FALSE) +
  ggtitle("Berkeley Police Stop Density Map, 2015-2016") +
  guides(fill = "none")
stopdensitymap4

# stop density by race
stopdensitymap3 <- berkgg_zoom2 +
  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
    bins = I(5), geom = "polygon", data = stops ) +
  scale_fill_gradient2( "StopDensity",
    low = "white", mid = "orange", high = "red", midpoint = 100) +
  labs(x = "Longitude", y = "Latitude") + facet_grid( ~ Race ) +
  scale_alpha(range = c(.2, .55), guide = FALSE) +
  ggtitle("Berkeley Police Stop Density Map by Race") +
  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10)) +
  guides(fill = FALSE)

stopdensitymap3

# stops leading to arrests by race:
df <- subset(stops, as.character(Enforcement) == "Arrest")
map_arrested1 <- berkgg_zoom1 +
  geom_point(aes(x = long, y = lat, colour = Race), data = df, alpha = 0.5) + 
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + 
  ggtitle("Police Stops Leading to Arrests") +
  facet_grid(~ Race) +
  guides( colour = FALSE)

# stops by reason:
map_reasons1 <- berkgg_zoom2 +
  geom_point(aes(x = long, y = lat, colour = Reason), data = stops, alpha = .7) +
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + 
  ggtitle("Police Stops by Reason")

# arrest data histogram by race
histogram_arrest1 <- ggplot(na.omit(arrest))+
  geom_bar(aes(x=Age, 
               fill = Race))+
  facet_grid(~Race)+
 labs(x= "Age",
       y = "Number of people",
       title = "Arrest Data") +
  guides( fill = FALSE)

# jail data histogram by race
histogram_jail1 <- ggplot(na.omit(jail))+
  geom_bar(aes(x=Age,
               fill = Race))+
  facet_grid(~Race)+
 labs(x= "Age",
       y = "Number of people",
       title = "Jail Data") + 
  guides(fill = FALSE)

# histogram of stop data by race and hour:
histogram_stoptime1 <- ggplot(stops, aes(x = Hour, fill = Race)) + 
  geom_bar(aes(y = ..count..)) +
  labs(y = "Count") +
  facet_grid(~Race)+
  labs(title = "Number of people each race in every hour") +
  guides( fill = FALSE )
```

## Introduction

For the STAT 133 final project, our group cleaned and explored the four police data sets available on the Berkeley Open Data website. We examined police stops, calls for service, arrests, and jail bookings made by the the police department over the past 3-6 month period. Rather than limiting the project to one set of data, these four sets were chosen in order to gain a more holistic and comprehensive understanding of the data. The additional resources of information each data set provided expanded our ability to understand and visualize the data.

One of our main goals was to find some ways of visualizing the relationship between the Berkeley Police Department and different groups of people. In order to compare the proportion of each group represented in the Berkeley police data relative to the population of those groups in Berkeley, we also considered three census data sets that are also available through Berkeley's open data site: first, we included the Berkeley 2000-2010 census data, which gave us the total population counts and percentages of each group of people in Berkeley for the 2010 census; next, we considered the Alameda County 2010 census tract population data, which allowed us to obtain the demographic information for 33 different neighborhoods (or "tracts") inside Berkeley; and finally, the Berkeley 2010 census tract polygons shapefile (.shp), which provided the information we needed to visually display the 33 neighborhoods.

After detailing our process of cleaning the police and census data, we will provided some visualizations of the data and some preliminary findings about the police data. Afterwards, we will discuss problems with the data and some ideas for future statistical analysis.

## Berkeley Stop Data Cleaning Process

The original stop data set included 16,255 stops recorded by the Berkeley police over a 6 month period. For each stop, the police recorded the date and time, location (as a character string, usually in the form of a street block or a street intersection), incident type (referring to whether the stop was traffic, pedestrian, bike, or suspicious vehicle), and a column called disposition. The disposition column was a character string typically including 6 characters per individual as follows: the first character was race, the second gender, the third age range, the fourth reason (for the stop), the fifth enforcement (the result of the stop), and the sixth car search (yes or no). Each individual that was involved with the stop was recorded using comma separation in the same column followed by an additional 6 characters. Additional optional dispositions were sometimes included, again using comma-separation from the 6-character strings (though not necessarily in any order), ranging from one to three characters. These additional disposition abbreviations represented the following options: Primary Case Report, MDT Narrative Only, Arrest Reporty Only (No Case Report Submitted), Incident Report, Field Card, Collision Investigation Report, Emergency Psychiatric Evaluation, Impounded Vehicle, and Officer Made a Stop of More Than 5 Persons. We found it reasonable to assume that most of these additional dispositions were related to police paperwork procedures, and since they were used so sparingly, we decided to not consider most of them. However, in spite of only being used less than 20 times, we found the Emergency Psychiatric Evaluation disposition to be of interest, and so we chose to investigate this one in addition to the main 6 character set of dispostions, though we chose not to focus on it during the limited scope of this paper.

Cleaning the stop data was relatively time-consuming compared to the other data sets. First, the dispositions column in the original data set contained many columns and sometimes rows worth of information, though the comma-separated pieces were without order (for example, sometimes the optional disposition letters came before the 6-character string instead).  We handled this problem using data tidying techniques, such as separating disposition information into separate row entries for each individual assessed, in the case of multiple persons stopped, and splitting the column by isolating the six character dispositions into different columns. The optional disposition character strings were moved and separated into more columns.

Additional cleaning for the stop data included changing the stop date and time variable to the `lubridate` date format. We also created an hour column in order to examine the stop occurrences by hour over a 24-hour period.

Turning the stop data's character locations into location coordinates using a Google Maps geolocation library proved difficult because of common data entry mistakes. For example, there were common mispellings of street names and the use of several different abbreviations to mean the same thing. Additionally, we found that the geolocation provided by Google does not accept forward slash characters the data used in the colum to indicate street intersections, so these forward slash characters were changed to `and` for the location processing. Google also does not accept the term "block" to indicate a range of addresses, and so the word "block" was removed from the data set for location processing. Most of these types of errors were delt with on an automated basis before geolocation, using `str_detect` and `str_replace` from the `stringr` package. The number of problems with the location data that were left over afterwards were small enough to handle on a case-by-case basis. Further, because Google limits geolocation to 2500 queries per 24-hour period, it took about a week to process all the locations into coordinates from the stop data (this includes the time it took to fix spelling errors).

At the end of the cleaning process, the stop data was whittled down from 16,255 observations to 14,291 observations (due to missing disposition values, etc.) that we will use in the following analysis.

## Berkeley Arrest and Jail Bookings Cleaning Process

The number of observations available for the arrest data set and the jail data set were significantly smaller than the number of observations for the stop data set. The arrest data had 205 observations and the jail bookings data had 223. These observations were obtained by the Berkeley Police department over a 3-month period, rather than a 6-month period. They both contained similar variables, including a case/arrest/booking number, date and time, type, and subject information (name, race, sex, D.O.B., age, height, weight, hair, eyes, and occupation) and statute information (type, description, agency, and disposition). Cleaning required the dates and times to be put into `lubridate` format, and we also created an hour column to consider the number of occurences during each hour of the day.

## Berkeley Census 2010 Data Cleaning Process

The Berkeley census 2010 data was utilized from 3 different data sets. We first considered the summary information provided by the city website in an excel document, which was easy to clean and use in our analysis, since the information was already summarized. However, in order to understand the demographics of populations within Berkeley in a way that could provide a spatial visualization, we needed to use the census 2010 Alameda County population tract data, which provided the population and race data by tract number. For example, there are 33 census 2010 tracts within the city of Berkeley. Finally, to get the map information of these 33 tracts, we also needed the census 2010 Berkeley tract polygons shapefile. After converting the shapefile to a dataframe and changing the coordinate system to the one used by Google, we were able to `left_join` this new polygonal mapping data frame to the data frame containing the population information for the Berkeley tracts by the unique census 2010 tract number. Combining the tract map information with the tract population information gave us a data frame containing population information for 33 different areas that we could represent spatially. The goal was to visualize the population density of different groups of people within the city by neighborhood, which provided us with a context for the stop data.

## Berkeley Police Stop Data Analysis

Below is the stop data points mapped by location. We have excluded some observations on the outskirts of Berkeley in order to zoom in to the street level. Note that the `alpha` variable in `ggplot` is set to a small number in order to show the incidents that occur multiple times in the same place. In other words, the darker the point, the more police stops have occured at that location.

```{r, warning = FALSE, echo = F, fig.width = 5, fig.height = 5, message = FALSE, fig.align="center"}
map_stop1
```

A more interesting visualization of this data is the following two-dimensional density map of the police stop data, displayed below. One of the most interesting aspects of the stop data in this image is that most police stops seem to be clustered west of UC Berkeley Campus in the Downtown area. This was contrary to what some of us might have thought before, which was that the area with the higher police stops would be corrolated to the area with the higher population. As you will see, however, this is not necessarily true.

```{r, warning = FALSE, echo = F, fig.width = 5, fig.height = 5, message = FALSE, fig.align="center"}
stopdensitymap4
```

To visualize the relationship between police stops and population density, consider the following map, which visualizes the percent population of Berkeley for each of the 33 census 2010 tracts. As you can see, the least densely populated area is the census 2010 tract containing the University of California. This makes sense because the area of this tract includes a campus and some student housing areas, where permanent residents are less likeley to live. One of the most population-dense areas is Southside, the portion of Berkeley directly south of the UC campus. Note that even though the tract including the Southside area is the most densely populated of the 33 census 2010 tracts in Berkeley, it is not the area where most police stops occur.

```{r, warning = FALSE, echo = F, fig.width = 5, fig.height = 5, message = FALSE, fig.align = "center"}
popdenmap # print population density map
```

In order to break down the stop data further, we mapped the data by each of the disposition variables. We found race to be one of the most interesting variables. Consider the Berkeley stop data density map faceted by race:

```{r, fig.height = 3, fig.width = 12, echo=FALSE, warning = FALSE, message = FALSE, fig.align="center"}
stopdensitymap3 
```

The image above seems to imply that the group of people stopped the most are African Americans in Berkeley. For instance, in Centeral Berkeley, the stops for the races defined by the Berkeley Police Department as Asian, Hispanic, White, and Other seem comparatively lower than for Blacks in the same area. In order to consider the context of these stop incidents, we considered the population distribution for these areas, as shown below. Note that the population density visualized by the following map is conveyed by the range between white (0%) and red (100%). As you can see, White people make up a striking majority in all census 2010 tract areas, which seems to imply that the population of Black people in Berkeley is being disproportionally affected by police stops.

```{r, warning = FALSE, echo = FALSE,  fig.width = 15, fig.height = 3, message = FALSE, fig.align = "center"}
grid.arrange(a2,black2,h2,o2,w2, ncol = 5) # print population density by race.
```

Now, with the population percentages of each group in hand, we were able to compare the results of our stop data analysis to the percent population of each group. The `Percent Stopped` column in the table below represents the percent of each group compared to the total number of people stopped, while the `Percent Population Stopped` column represents the proportion of people in each group stopped relative to the population of that group living in Berkeley. Even though the people stopped by police are not necessarily Berkeley residents, the `Percent Population Stopped` gives us some perspective of which groups of people in Berkeley might be more likely to be affected by police stops than others.

```{r, message = F, echo = FALSE, warning = FALSE, fig.align="center"}
panderOptions("digits", 3)
pander(ss) # table for stop data percentages by race relative to population
#grid.arrange(bargraph1, bargraph2, ncol = 2)
```

A visualization of the groups of people stopped in proportion to their population is provided by the bar chart below.

!!!!!!!!!!!!!!!!!!!!!INCLUDE BAR GRAPH HERE of % race stopped vs. percent race population, as described above!!!!!!!!!!!!!!!!!!!!!!!!

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.align = "center"}
#### !!!!!!!!!!!!!!!!!!!!!INCLUDE BAR GRAPH HERE of % race stopped vs. relative percent, as described above!!!!!!!!!!!!!!!!!!!!!!!!
```

Another disposition variable we found to be of interest was the enforcement variable, which has the options Arrest, Citation, Warning, and Other. We were interested to see how these enforcements were applied to different groups of people. Consider the map below, which appears to suggest that the enforcement of arrest is applied differently to different groups of people.

```{r, echo=FALSE, warning=F,message=F, fig.align="center", fig.width = 12, fig.height=3}
map_arrested1 
```

!!!!!!!!!!!!!!!!!!!!!INCLUDE BAR GRAPH & TABLE OF ENFORCEMENT BY RACE HERE (Use percentage %, not count) !!!!!!!!!!!!!!!!!!!!!!!!

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.align = "center"}
#### INCLUDE BAR GRAPH & TABLE OF ENFORCEMENT BY RACE HERE(Use percentage, not count)!!!!!!!!!!!!!!!!!!!!!!!!
```

A third interesting disposition variable was the reason disposition, which indicates the reason for the police stop in each case. The reasons available for police stops are Investigation, Other, Probation/Parole, Reasonable Suspicion, Traffic, and Wanted. One way we explored the data initially was to see how these reasons affected different groups of people, and we would recommend exploring this idea in depth at a later time. Due to the length and scope of this paper however, we decided to end our discussion of the stop data by showing that most of the Berkeley police stops are traffic stops. Mapping the coordinates seems to suggest this fact, and it is confirmed when analyzing the data.

```{r, fig.height = 5, fig.width = 5, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
map_reasons1
```

!!!!!!!!!!!!!!!!!!!!!INCLUDE BAR GRAPH & TABLE OF POLICE STOPS BY REASON HERE (Use percentage, not count) !!!!!!!!!!!!!!!!!!!!!!!!

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.align="center"}
#### INCLUDE BAR GRAPH & TABLE OF POLICE STOPS BY REASON HERE(Use percentage, not count)!!!!!!!!!!!!!!!!!!!!!!!!
```

## Berkeley Police Jail and Arrest Data Set Analysis

Moving on from the stop data, we considered the available jail data and arrest data for the city of Berkeley. Two of the most interesting variables to consider in these two data sets were race and age. We decided the examine the percent of people arrested and the percent of people jailed in Berkeley by both age and race. Below is a visualization. Note that the following percentages are related to the number of total people jailed or the number of people arrested, respectively, and we have not adjusted the data relative to the population of Berkeley.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 12, fig.height=3, fig.align="center"}
histogram_jail1
histogram_arrest1
# I decided to take these out of the multi.plot function because I was trying to adjust the format a certain way, I'm sorry!
```

One of the most interesting things about the above histograms is the number of black people arrested in Berkeley over the past 3 month period is higher than the numbers of people arrested in Berkeley over the past 3 months for other groups of people. Note that this statement is even stronger for young black people. Even though most of the population in Berkeley is white, most of the people recently arrested are black.

Another idea we had was to visualize the arrest data for each community by hour. 
!!!!!!!!!!!PLEASE FIX LABEL SO THAT IT IS PERCENT ARRESTED, NOT COUNT!!!! ALSO, PLEASE DISPLAY A SMALL TABLE OF THIS DATA FOLLOWING THE HISTOGRAM.!!!!!!!!!!!!!

```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.width = 12, fig.height=3, fig.align = "center"}
histogram_stoptime1
#!!!!!!!!!!!PLEASE FIX LABEL SO THAT IT IS PERCENT ARRESTED, NOT COUNT!!!! ALSO, PLEASE DISPLAY A SMALL TABLE OF THIS DATA FOLLOWING THE HISTOGRAM.!!!!!!!!!!!!!
```

!!!!!!!!!!!!!! 1-2 sentence: ANALYZE FINDINGS FROM chart & TABLE ABOVE HERE!!!!!!!!!!

## Problems

The data sets provided by the police (stops, arrests, and jail bookings) have several problems. First, concerning the methods for recording race, especially for police stops that do not result in an arrest, it seems unreasonable to assume that racial information would have been provided by the individual stopped. It seems more likely that police officers are judging the race of the people they stop during stops for themselves in order to record it in the system that ultimately provided us with that data set. The problem with that method of data collection is that it opens up an opportunity for error; for example, if the data is recorded by the officers through their judgment of the individuals physical appearance, they may unknowingly classify that individual as a race that the individual does not identify themselves as. Even though there doesn't seem to be a means to get around this problem with the actual method of collecting racial data for police stops, it is still important to address it as something that could induce error into a statistical analysis.

In further regards to the racial data available, it is important to note that the census data we used to compare the stop data to included many more categories for race than just the five used by police (Asian, Black, Hispanic, Other, and White). To deal with this problem, other races included by the census were moved to the category of other. Additionally, the census data had a category for Two or More Races, which we also included as Other in order to compare the data sets. While the census data allowed for overlap between racial categories, there was no overlap for the racial categories in the police data. This creates a problem when comparing the two data sets, and so it should be acknowledged.

Secondly, the police data used in this paper comes from a data set maintained by the City of Berkeley that is updated every 3 months (as is the case for arrests and jail bookings) or every 6 months (as is the case for stops). A more thorough analysis of this data may want to include an analysis of the data over a longer time frame.

Next, with regard to the stop data in particular, the many data-entry problems we faced in the character string location column (which sometimes included only 1 street), there is likely some stops that are not mapped correctly due to approximations made by the Google geolocation service (for example, in some cases, only having one street name instead of a corner caused Google to estimate the exact location based on few conditions). Though most of the location problems were fixed using string detection, there is likely some error in our visual representation of the stop data maps.

Finally, with regard to the calls for service, we were unfortunately unable to incorporate that particular data set due to constraints on time. However, a more full analysis of police interactions with the community might consider incorporating the calls for service data.

## Conclusion

The goal of this paper was to investigate and visualize Berkeley Police interactions with the community through the available police data. Although our preliminary findings unfortunately seem to suggest that certain groups of people, especially black people, are disproportionately likely to be stopped, arrested, and jailed in the city of Berkeley, we recommend a more rigorous statistical analysis of the data be done in order to assess this claim. We also recommend an idea we had for this project to future researchers that could make the Berkeley police data more accessible. For example, one could make the data more accessible by creating a `shiny` app that allows the user to toggle between the different dispositions of the data (race, age range, enforcement, reason, etc.) and between modes of view (map or bar chart, for example) or order to play with the data in a visual way. Another `shiny` app idea we had was to filter the calls for service data by proximity to certain locations so that users could use the app to select a location and find a visual displaying the most common types of calls in that area. It would also be useful to find a way to create an app can easily be updated when the new data is posted (every 3-6 months), allowing users of the app to get the most up-to-date information.

## Sources

* [Stop Data (16,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Stop-Data/6e9j-pj9p)
* [Arrest (200)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Arrests/xi7q-nji6)
* [Jail Bookings (250)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Jail-Bookings/7ykt-c32j)
* [Calls for Service (4,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Calls-for-Service/k2nh-s5h5)
* [Berkeley Census Data](https://data.cityofberkeley.info/Demographics/Census-Data-2000-And-2010/bnkq-f2ge)
*[Census 2010 Population and race data by county tract polygons](http://www.bayareacensus.ca.gov/small/small.htm)
*[Berkeley Census 2010 Tract Polygons](https://data.cityofberkeley.info/Demographics/Census-Tract-Polygons-2010/peq3-2arw)

### 2016-08-08 final_paper/draft.Rmd
test

@@
   layer_bars() %>%
-  add_tooltip(function(stop_by) stop_by$Race)+
+  add_tooltip(function(stop_by) stop_by$Race) 
   labs(title = "Counts of people recorded in every day in a week")
@@
 4. Asian and female are more likely to be arrested by the BDP for probation/parole.
+
+```{r}
+
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), 
+        panel.background = element_blank())
+
+  theme(plot.title=element_text(family="xkcd-Regular"), 
+        text=element_text(family="xkcd-Regular"), 
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
+p1 <- ggplot(stop_by)+
+  geom_bar(aes(x=Race),
+           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))+
+  labs(title = "Stop data by race",
+       x = "Race",
+       y = "Counts")
+
+p2 <- ggplot(race)+
+  geom_bar(aes(x=Description, 
+               y=Counts),
+           stat = "identity", 
+           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))+
+  labs(title = "Demographic race data",
+       x = "Race",
+       y = "Counts")
+
+multiplot(p1, p2,cols=2)
+
+ggplot(mtcars, aes(factor(cyl))) + 
+          geom_bar(aes(y = (..count..)/sum(..count..))) + 
+          scale_y_continuous(labels=percent)
+
+ggplot(pop)+
+  geom_bar(aes(x=Race,
+               y=(stop/census)), 
+           stat = "identity",
+           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))+
+  geom_text(aes(x = Race, 
+                y = (stop/census), 
+                label = paste0((stop/census),"%")), 
+            size=3)+
+  scale_y_continuous(labels=percent)+
+  labs(title = "The percentage of being stopped by police",
+       x = "Race",
+       y = "Percentage")
+
+```
+
+```{r}
+############################## stop data ############################## 
+stop_by <- read_csv("cleaned_stop_data.csv")
+
+stop_by$Call.Date.Time <- mdy_hm(stop_by$Call.Date.Time)
+stop_by$AgeRange <- as.factor(stop_by$AgeRange)
+
+stop_by <- stop_by %>%
+  mutate(Hour = as.integer(hour(Call.Date.Time))) %>%
+  mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))
+
+stop_by$Race<- stop_by$Race %>% 
+  str_replace("A", "Asian") %>% 
+  str_replace("B", "Black") %>% 
+  str_replace("H", "Hispanic") %>% 
+  str_replace("O", "Other") %>% 
+  str_replace("W", "White")
+
+stop_by$Gender<- stop_by$Gender %>% 
+  str_replace("F", "Female") %>% 
+  str_replace("M", "Male")
+
+stop_by$Reason <- stop_by$Reason %>% 
+  str_replace("I", "Investigation") %>% 
+  str_replace("T", "Traffic") %>% 
+  str_replace("R", "Reasonable Suspicion") %>% 
+  str_replace("R", "Probation/ Parole") %>% 
+  str_replace("W", "Wanted")
+
+stop_by$Enforcement <- stop_by$Enforcement %>% 
+  str_replace("A", "Arrest") %>% 
+  str_replace("C", "Citation") %>% 
+  str_replace("O", "Other") %>% 
+  str_replace("W", "Warning")
+
+stop_by$CarSearch <- stop_by$CarSearch %>% 
+  str_replace("S", "Search") %>% 
+  str_replace("N", "No Search")
+```

### 2016-08-08 eda/PaperDrafts/OtherDataCleaning.R

@@ -1,0 +1,147 @@
+library(sp)
+library(ggmap)
+library(tidyr)
+library(plyr)
+library(dplyr)
+library(stringr)
+library(lubridate)
+library(ggplot2)
+library(readr)
+library(data.table)
+library(rworldmap)
+library(maps)
+library(mapdata)
+library(maptools)
+library(rgeos)
+library(scales)
+library(RgoogleMaps)
+library(tmap)
+library(sp)
+library(rgdal)
+library(rgeos)
+library(classInt)
+library(gdata)
+library(XLConnect)
+library(RColorBrewer)
+
+
+############################## census data JENNY ############################## 
+# data file names:
+file_stops <- "../../clean_data/StopData_clean.csv" 
+file_stopsummary <- "../../clean_data/StopData_summary.csv" 
+mergedf <- read_csv(file = file_stops)
+census <- read.csv("../../raw_data/Census_Data_2000_And_2010.csv")
+census2010 <- census[census$Year==2010,]
+totalpop <- census2010[1,]$Amount
+stopsummary <- read.csv( file_stopsummary )
+stopsummary <- select(stopsummary, -X)
+
+racecensus2010 <- census2010[census2010$Heading=="Not Hispanic or Latino"|census2010$Heading=="HISPANIC OR LATINO AND RACE",]
+sexcensus2010 <- census2010[census2010$Heading=="Sex",]
+agecensus2010 <- census2010[census2010$Heading=="Age",]
+
+
+race <- select(racecensus2010, Description, Amount) 
+race$Description <- race$Description %>% 
+  str_replace_all("Hispanic or Latino.*", "Hispanic") %>% 
+  str_replace("Black or African American", "Black") %>% 
+  str_replace("American Indian and Alaska Native", "Other") %>% 
+  str_replace("Native Hawaiian and Other Pacific Islander", "Other") %>% 
+  str_replace("Some other race", "Other") %>% 
+  str_replace("Two or more races", "Other")
+
+race <- race[race$Description !="Not Hispanic",]
+
+race <- race %>% 
+  group_by(Description) %>% 
+  tally(Amount)
+
+names(race) <- c("Description", "Counts")  
+
+sex <- select(sexcensus2010, Description, Amount) 
+
+age <- select(agecensus2010, Description, Amount) 
+age <- age[age$Description != "Median age",]
[…]

### 2016-08-08 clean_data/CleanDataScripts/CleanData_MakePercentageTables.R
uploading table data and cleaning script

@@ -1,0 +1,122 @@
+# CleanData_MakePercentageTables.R
+
+# Purpose: Manipulates census and stop data to produce small tables of percentages.
+#          Please note that this code has already been run, so it doesn't need to be run again.
+#          Look in the clean data folder for the .rds file containing the tables.
+#          Please look at the documentation below to understand how to read the tables in from the files.
+
+# how to create the percentage tables for bar charts:
+library(sp)
+library(ggmap)
+library(tidyr)
+library(plyr)
+library(dplyr)
+library(stringr)
+library(lubridate)
+library(ggplot2)
+library(readr)
+library(data.table)
+library(rworldmap)
+library(maps)
+library(mapdata)
+library(maptools)
+library(rgeos)
+library(scales)
+library(RgoogleMaps)
+library(sp)
+library(rgdal)
+library(rgeos)
+library(classInt)
+library(gdata)
+library(XLConnect)
+library(RColorBrewer)
+
+
+# data files:
+
+file_stops <- "../../clean_data/StopData_clean.csv" 
+file_tract_populations <- "../../clean_data/berk_census2010_clean.csv"
+file_saveAs <- "../../clean_data/StopData_tbls.rds" 
+mergedf <- read_csv(file = file_stops)
+bpop2 <- read_csv(file = file_tract_populations )
+
+# stop data sumamary by race:
+
+tbl_summary <- mergedf %>%  
+  group_by(Race) %>%
+  summarise(Stopped=n() ) %>%
+  mutate(Percent.Stopped = Stopped/nrow(mergedf)) 
+temp <- c(sum(bpop2$Asian), sum(bpop2$Black), sum(bpop2$Hispanic), sum(bpop2$Other), sum(bpop2$White) )
+tbl_summary$Race.Population = temp
+tbl_summary$Percent.Race.Stopped = tbl_summary$Stopped/tbl_summary$Race.Population
+tbl_summary <- tbl_summary %>%
+  select(-Stopped, -Race.Population)
+tbl_summary
+
+# percent reason:
+
+tbl_reason <- mergedf %>%  
+  group_by( Reason) %>%
+  summarise(Stopped=n() ) %>%
+  mutate(Percent.Stopped = Stopped/nrow(mergedf)) 
+
[…]

### 2016-08-08 clean_data/CleanDataScripts/CleanData_OtherCleaning.R

@@ -144,4 +144,32 @@ mergedf <- mergedf %>%
 # write.csv(mergedf, "../../clean_data/StopData_clean.csv" )
 
 
+# how to create the tables:
 
+tbl_summary <- mergedf %>%  
+  group_by(Race) %>%
+  summarise(Stopped=n() ) %>%
+  mutate(Percent.Stopped = Stopped/nrow(df)) 
+
+tbl_summary
+
+# racepops <- c(sum(bpop2$Asian), sum(bpop2$Black), sum(bpop2$Hispanic), sum(bpop2$Other), sum(bpop2$White) )
+# stop.summary$Race.Population = racepops
+# stop.summary$Percent.Race.Stopped = stop.summary$Stopped/stop.summary$Race.Population
+
+# percent reason:
+
+tbl_reason <- mergedf %>%  
+  group_by( Reason) %>%
+  summarise(Stopped=n() ) %>%
+  mutate(Percent.Stopped = Stopped/nrow(df)) 
+
+tbl_enforcement <- mergedf %>% 
+  group_by(Enforcement) %>%
+  summarise(Stopped = n()) %>%
+  mutate(Percent.Stopped = Stopped/nrow(df))
+
+tbl_enforcement_W <- mergedf[as.character(Race) == "White"] %>%
+  group_by(Reason) %>%
+  summarise(Stopped=n() ) %>%
+  mutate(Percent.Stopped = Stopped/nrow(df)) 

### 2016-08-08 eda/CensusDataExploration/HeatmapExploration3.Rmd
organizing folders

--- 2016-08-06 eda/Heatmap/HeatmapExploration3.Rmd
@@
 library(XLConnect)
-#library(data.table)
+library(data.table)
 ```
@@
   )
-```
-```{r,include = F, echo = FALSE}
+
 # get the plain Berkeley map from Google:
 berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 12)
-```
 
-```{r}
 berkgg <-ggmap(berkMap) +
@@
 ```{r}
-popdenmap <- ggmap(berkMap) + 
+popdenmap <- berkgg + 
   geom_polygon(data = census.df_pop, aes(x = long, y = lat, group = group, fill = TotalPop), color = "white", alpha = .7) 

### 2016-08-08 eda/JailAndArrestExploration/Arrest and Jail Bookings.Rmd
organizing folders

---
title: "Arrest and Jail Bookings"
author: "Shangjun Jiang"
date: "22 July 2016"
output: html_document
---

```{r}
setwd("~/Desktop/Stats R/PROJECT/finalproject_angry_ladies/Jenny_Analysis")
```

#load packages
```{r}
library(xml2)
library(rvest)
library(ggplot2)
library(lubridate)
library(stringr)
library(dplyr)
library(tidyr)
library(readr)
library(shiny)
library(ggvis)
```

#load datasets
```{r}
jail <- read.csv("Berkeley_PD_Log_-_Jail_Bookings.csv")
jail$Age <- as.integer(jail$Age)
jail$Age[jail$Age<20] <- "0-19"
jail$Age[jail$Age>=20 & jail$Age<30] <- "20-29"
jail$Age[jail$Age>=30 & jail$Age<40] <- "30-39"
jail$Age[jail$Age>=40 & jail$Age<50] <- "40-49"
jail$Age[jail$Age>=50 & jail$Age<60] <- "50-59"
jail$Age[jail$Age>=60 & jail$Age<70] <- "60-69"
jail$Age[jail$Age>=70 & jail$Age<80] <- "70-79"
jail$Age <- as.factor(jail$Age)

arrest <- read.csv("Berkeley_PD_Log_-_Arrests.csv")
arrest$Age <- as.integer(arrest$Age)
arrest$Age[arrest$Age<20] <- "0-19"
arrest$Age[arrest$Age>=20 & arrest$Age<30] <- "20-29"
arrest$Age[arrest$Age>=30 & arrest$Age<40] <- "30-39"
arrest$Age[arrest$Age>=40 & arrest$Age<50] <- "40-49"
arrest$Age[arrest$Age>=50 & arrest$Age<60] <- "50-59"
arrest$Age[arrest$Age>=60 & arrest$Age<70] <- "60-69"
arrest$Age[arrest$Age>=70 & arrest$Age<80] <- "70-79"
arrest$Age <- as.factor(arrest$Age)

jail[jail == ""] <- NA

arrest[arrest == ""] <- NA

#str(jail) # 'data.frame':	249 obs. of  19 variables:

#str(arrest) # 232 obs. of  16 variables:
```

# differences between jail and arreasts   match 28/29  
# delete status desription
# subject, race, age, height, weight, hair, eyes, status, status.type, status.description, case.number

#graph the race data  1. Black 2. White 
# status. type 
#graph the age data

```{r}
## Disposition

length(which(jail$Disposition == "NO COMPLAINT"))   #28
length(jail[which(jail$Disposition == "NO COMPLAINT"),]) #19
jail[which(jail$Disposition == "NO COMPLAINT"),]

jail[jail$Disposition == "BAILED",]  #19

length(which(jail$Disposition == "BAILED"))  #14

length(jail[jail$Disposition == "CITE-JAIL",])  #19

#SANTA RITA JAIL
length(jail[jail$Disposition == "SANTA RITA JAIL",])  #19

length(jail[jail$Disposition == "849(b) PC-RELEASEL",])  #19

jail$Disposition == "849(b) PC-RELEASEL"

```

#clean jail dataset and arrest dataset in order to combine them together
```{r}

cleanedjail <- jail
cleanedjail$Statute <- NULL
cleanedjail$Statute.Description <- NULL
cleanedjail$Height <- NULL
cleanedjail$Weight <- NULL
cleanedjail$Hair <- NULL
cleanedjail$Eyes <- NULL
cleanedjail$Date.Of.Birth <- NULL
cleanedjail$Subject <- tolower(cleanedjail$Subject)
names(cleanedjail) <- c("Booking_Number","Booking_Date_Time", "Subject", "Race", "Sex","Age", "Occupation","Statute_Type", "Arrest_Date_Time","Case_Number","Booking_Agency", "Disposition")

cleanedarrest <- arrest
cleanedarrest$Statute <- NULL
cleanedarrest$Statute.Description <- NULL
cleanedarrest$Height <- NULL
cleanedarrest$Weight <- NULL
cleanedarrest$Hair <- NULL
cleanedarrest$Eyes <- NULL
cleanedarrest$Date.of.Birth <- NULL
cleanedarrest$Subject <- tolower(cleanedarrest$Subject)
names(cleanedarrest) <- c("Arrest_Number", "Date_Time", "Arrest_Type", "Subject", "Race", "Sex", "Age", "Statute_Type", "Case_Number")

names(cleanedarrest)
names(cleanedjail)

comb <- inner_join(cleanedarrest,cleanedjail, by=c("Case_Number", "Race", "Sex", "Age", "Subject", "Statute_Type"))
names(comb)

#join two lists together
ja <- inner_join(arrest,jail, by="Case.Number")
left_join
str(ja) #'data.frame':	28 obs. of  23 variables:

```

```{r}
ggplot(cleanedjail)+
  geom_bar(aes(x=Booking_Agency))

#CA0010300 Berkeley Police Department
#CA0019700 UC Berkeley Police Department

```

# Disposition
```{r}
ggplot(jail)+
  geom_bar(aes(x=Disposition))

```

#people who are in jail
```{r}

ggplot(na.omit(jail))+
  geom_bar(aes(x=Age, 
               fill = Race))+
  facet_wrap(~Sex)+
  labs(x= "Age",
       y = "Number of people",
       title = "Jail Data")

ggplot(na.omit(jail))+
  geom_bar(aes(x=Occupation, 
               fill=Race))

```

## Arrest dataset

```{r,fig.width=10, fig.asp=0.7}
ggplot(na.omit(arrest))+
  geom_bar(aes(x=Age, fill = Sex))+
  facet_wrap(~Race)+
 labs(x= "Age",
       y = "Number of people",
       title = "Arrest Data")
```

#people who are arrest
```{r}
ggplot(na.omit(arrest))+
  geom_bar(aes(x=Age, fill = Sex))+
  facet_wrap(~Race)+
 labs(x= "Age",
       y = "Number of people",
       title = "Arrest Data")

ggplot(na.omit(arrest))+
  geom_bar(aes(x=Age, fill = Race))+
  facet_wrap(~Sex)

arrest$Age <- as.factor(arrest$Age)

ggplot(na.omit(arrest))+
  geom_bar(aes(x=Race, fill=Age))+
  facet_wrap(~Sex)

```

#occupation
```{r}
ggplot(na.omit(jail))+
  geom_bar(aes(x=Occupation))

```

### 2016-08-08 eda/JailAndArrestExploration/Jail.Rmd
organizing folders

--- 2016-08-02 Jenny_Analysis/Jail.Rmd
@@
 
+najail$Disposition <- str_replace(najail$Disposition, "NO COMPLAINT", "NC") %>% 
+  str_replace("SANTA RITA JAIL", "SRC") %>% 
+  str_replace("")
+
 ggplot(najail)+
@@
                fill = Race))+
-  facet_wrap(~Sex)
+  facet_wrap(~Sex)+
+   scale_x_discrete(labels = function(x) str_wrap(x, width = 5))
 

### 2016-08-08 eda/StopDataExploration/Berkeley_PD_Stop.Rmd
organizing folders

--- 2016-07-31 Berkeley_PD_Stop.Rmd
@@
 
+totalIncidentsNumber <- nrow(stop_df)
+typeone <- stop_df %>%
+    filter(Other == "" & Dispositions != "")
+typeoneNumber = nrow(typeone)
+typetwo <- stop_df %>%
+    filter(Other != "" & Dispositions == "")
+typetwoNumber = nrow(typetwo)
+typethree <- stop_df %>%
+    filter(Other != "" & Dispositions != "")
+typethreeNumber = nrow(typethree)
+
 # Arranges information in the Dispositions column into separate objects for each individual involved 

### 2016-08-08 eda/PaperDrafts/BDP_Paperdraft.Rmd
still organizing

--- 2016-08-02 final_paper/BDP_Paperdraft.Rmd
@@
   mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))
+
+############################## location data ############################## 
+mergedf <- readRDS( 'StopData_merged.rds' )
+
+## Location info:
+latmax <- max(mergedf$lat, na.rm = TRUE) 
+latmin <- min(mergedf$lat, na.rm = TRUE)
+lonmax <- max(mergedf$long, na.rm = TRUE) 
+lonmin <- min(mergedf$long, na.rm = TRUE)
+latvals <- c(latmin, latmax)
+lonvals <- c(lonmin, lonmax)
+
+## Make the variables factors:
+mergedf$CarSearch <- factor(mergedf$CarSearch)
+levels(mergedf$CarSearch) <- c("No Search", "Search")
+
+mergedf$Race <- factor(mergedf$Race)
+levels(mergedf$Race) <- c("Asian", "Black", "Hispanic", "Other", "White")
+
+mergedf$Enforcement <- factor(mergedf$Enforcement)
+levels(mergedf$Enforcement) <- c("Arrest", "Citation", "Other", "Warning")
+
+mergedf$Reason <- factor(mergedf$Reason)
+levels(mergedf$Reason) <- c("Other", "Investigation", "Probation/Parole", "Reasonable Suscipcion", "Traffic", "Wanted")
+
+mergedf <- mergedf %>%
+  mutate(Arrest = ifelse( 
+    as.character(Enforcement) == "Arrest", 
+    "Arrested",
+    "Not Arrested") ) %>%
+  mutate(Arrest = factor(Arrest))
+
+v <- factor(mergedf$Other)
+levels(v)
+
+mergedf <- mergedf %>%
+  mutate(Emergency.Psych.Eval = ifelse( str_detect(as.character(Other), "MH"), 
+                                        yes = "Yes",
+                                        no = "No") )
+
+mergedf$Emergency.Psych.Eval <- factor(mergedf$Emergency.Psych.Eval)
+
+### BerkeleyMap 
+berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14)
+
 ```
+
+## Population by census block group:
+
+We found the census data online and map the Berkeley population in block groups.
+![](mygraphic.png)
 
@@
 
-For the analysis in this part, only person information for all the people in each incident and the Call Time/Date of each incident are analyzed.
+## MAP 1:All BPD Stops Density, 2015-2016
+```{r, fig.width=10, fig.asp=0.8, echo=FALSE}
+# a contour plot
+ggmap(berkMap) +
+  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
+    size = 2, bins = 5, data = mergedf, geom = "polygon") +
+  scale_fill_gradient(low = "black", high = "red") +
+    theme (panel.grid.major = element_blank (), # remove major grid
+		       panel.grid.minor = element_blank (),  # remove minor grid
+		       axis.text = element_blank (), 
+		       axis.title = element_blank (),
+	         axis.ticks = element_blank ()) + 
+  ggtitle("All BPD Stops Density, 2015-2016") +
+  labs(alpha = element_blank())+
+  guides(alpha = FALSE)
+
+```
+
+## MAP 2: BPD Stop Contour Map of Berkeley by Age Range
+```{r, fig.height = 8, fig.width = 10, echo=FALSE}
+ggmap(berkMap) +
+  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
+    bins = I(5), geom = "polygon", data = mergedf ) +
+  scale_fill_gradient2( "Stop Density",
+    low = "white", mid = "orange", high = "red", midpoint = 25) +
+  labs(x = "Longitude", y = "Latitude") + facet_wrap(~ AgeRange) +
+  scale_alpha(range = c(.2, .55), guide = FALSE) +
+  ggtitle("BPD Stop Contour Map of Berkeley by Age Range") +
+  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
+```
 
@@
    Race     | Count | 0-18 | 18-29 | 30-39 | 40+
-----  |------ |------|-------|-------|-------
-A     | 1141  | 8.50%| 43.47%| 21.03%| 26.99%
-B     | 4636  | 2.29%| 34.97%| 25.09%| 37.66%
-H     | 1676  | 3.22%| 46.12%| 29.18%| 21.47%
-O     | 1384  | 2.67%| 43.28%| 29.17%| 26.87%
-W     | 5454  | 1.72%| 29.45%| 25.39%| 43.44%
+----------  |------ |------|-------|-------|-------
+A (Asian)   | 1141  | 8.50%| 43.47%| 21.03%| 26.99%
+B (Black)   | 4636  | 2.29%| 34.97%| 25.09%| 37.66%
+H (Hispanic)| 1676  | 3.22%| 46.12%| 29.18%| 21.47%
+O (Other)   | 1384  | 2.67%| 43.28%| 29.17%| 26.87%
+W (White)   | 5454  | 1.72%| 29.45%| 25.39%| 43.44%
 
 #### Picture: count of person recorded of each race in each hour
+
 ```{r, echo=FALSE}
@@
   labs(y = "Count") +
-  facet_wrap(~Race)
+  facet_wrap(~Race)+
+  labs(title = "Count of person recorded of each race in each hour")
 stop_time_percentage
+```
+
+## MAP 3: BPD Stop Contour Map of Berkeley by Race
+```{r, fig.height = 5, fig.width = 10, echo=FALSE}
+ggmap(berkMap) +
+  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
+    bins = I(5), geom = "polygon", data = mergedf ) +
+  scale_fill_gradient2( "StopDensity",
+    low = "white", mid = "orange", high = "red", midpoint = 100) +
+  labs(x = "Longitude", y = "Latitude") + facet_wrap( ~ Race ) +
+  scale_alpha(range = c(.2, .55), guide = FALSE) +
+  ggtitle("BPD Stop Contour Map of Berkeley by Race") +
+  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
 ```
@@
 (5) Black people contribute a especially high percentage of the incidents at night, while the white people contribute a especially high percentage of the incidents at the noon.
+
+### Analysis of Day
+####Picture: count of person recorded in each day in a week
+```{r, echo=FALSE}
+stop_day <- stop_by %>%
+  ggvis(x = ~Day, fill = ~Race) %>%
+  layer_bars() %>%
+  add_tooltip(function(stop_by) stop_by$Race)
+stop_day
+```
+
+```{r, eval=FALSE, echo=FALSE}
+stop_day_total_table <- stop_by %>%
+    group_by(Day) %>%
+    summarise(totaLCount = n())
+stop_day_race_table <- stop_by %>%
+    group_by(Day, Race) %>%
+    summarise(count = n())
+stop_day_table <- left_join(stop_day_race_table, stop_day_total_table, by = "Day")
+stop_day_table <- stop_day_table %>%
+    spread(key = Race, value = count)
+stop_day_table
+```
+####Picture: count of person recorded in each day in a week
+Day   | Count | A    | B    | H    | O    | W
+----  |------ |------|------|------|------|------
+1     | 1692  | 116  | 653  | 221  | 159  | 543
+2     | 1665  | 149  | 543  | 182  | 166  | 625
+3     | 2344  | 189  | 645  | 231  | 236  | 1043
+4     | 2287  | 206  | 616  | 251  | 244  | 970
+5     | 1941  | 152  | 592  | 223  | 168  | 806
+6     | 2231  | 186  | 764  | 268  | 203  | 810
+7     | 2131  | 143  | 823  | 300  | 208  | 657
+
+(1) Among all the day in a week, the number of incidents happened on Sunday and Monday are much less than those happened through Tuesday and Saturday.
+
+(2) The count of records of Asian people is much less than that of any other race of people in each day and the whole week.
+
+(3) Asian people are less liable to commit an incident on Sunday, and more liable to commit an incident on Wednesday.
+
+(4) Black and Hispanic people are less liable to commit an incident on Monday, and more liable to commit an incident on Saturday.
+
+(5) White people are less liable to commit an incident on Sunday, and more liable to commit an incident on Tuesday and Wednesday, which is similar to Asian people.
 
@@
 
-```{r, echo=FALSE}
-## Read in the Data:
-#finaldf <- readRDS( 'StopData_finaldf.rds' )
-mergedf <- readRDS( 'StopData_merged.rds' )
-```
-
-```{r, echo=FALSE}
-## Location info:
-latmax <- max( mergedf$lat, na.rm = TRUE ) 
-latmin <- min( mergedf$lat, na.rm = TRUE )
-lonmax <- max( mergedf$long, na.rm = TRUE ) 
-lonmin <- min( mergedf$long, na.rm = TRUE )
-latvals <- c( latmin, latmax )
-lonvals <- c( lonmin, lonmax )
-```
-
-```{r, echo=FALSE}
-## Make the variables factors:
-mergedf$CarSearch <- factor(mergedf$CarSearch)
-levels(mergedf$CarSearch) <- c("No Search", "Search")
-
-mergedf$Race <- factor(mergedf$Race)
-levels(mergedf$Race) <- c("Asian", "Black", "Hispanic", "Other", "White")
-
-mergedf$Enforcement <- factor(mergedf$Enforcement)
-levels(mergedf$Enforcement) <- c("Arrest", "Citation", "Other", "Warning")
-
-mergedf$Reason <- factor(mergedf$Reason)
-levels(mergedf$Reason) <- c("Other", "Investigation", "Probation/Parole", "Reasonable Suscipcion", "Traffic", "Wanted")
-
-mergedf <- mergedf %>%
-  mutate(Arrest = ifelse( 
-    as.character(Enforcement) == "Arrest", 
-    "Arrested",
-    "Not Arrested") ) %>%
-  mutate(Arrest = factor(Arrest))
-
-v <- factor(mergedf$Other)
-levels(v)
-
-mergedf <- mergedf %>%
-  mutate(Emergency.Psych.Eval = ifelse( str_detect(as.character(Other), "MH"), 
-                                        yes = "Yes",
-                                        no = "No") )
-
-mergedf$Emergency.Psych.Eval <- factor(mergedf$Emergency.Psych.Eval)
-
-berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14)
-
-```
-
-## MAP 1:All BPD Stops Density, 2015-2016
-```{r, fig.width=10, fig.asp=0.8, echo=FALSE}
-# a contour plot
-ggmap(berkMap) +
-  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
-    size = 2, bins = 5, data = mergedf, geom = "polygon") +
-  scale_fill_gradient(low = "black", high = "red") +
-    theme (panel.grid.major = element_blank (), # remove major grid
-		       panel.grid.minor = element_blank (),  # remove minor grid
-		       axis.text = element_blank (), 
-		       axis.title = element_blank (),
-	         axis.ticks = element_blank ()) + 
-  ggtitle("All BPD Stops Density, 2015-2016") +
-  labs(alpha = element_blank())+
-  guides(alpha = FALSE)
-
-```
-
-## MAP 2: BPD Stop Contour Map of Berkeley by Race
-```{r, fig.height = 5, fig.width = 10, echo=FALSE}
-ggmap(berkMap) +
-  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
-    bins = I(5), geom = "polygon", data = mergedf ) +
-  scale_fill_gradient2( "StopDensity",
-    low = "white", mid = "orange", high = "red", midpoint = 100) +
-  labs(x = "Longitude", y = "Latitude") + facet_wrap( ~ Race ) +
-  scale_alpha(range = c(.2, .55), guide = FALSE) +
-  ggtitle("BPD Stop Contour Map of Berkeley by Race") +
-  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
-```
-
-## MAP 3: BPD Stop Contour Map of Berkeley by Age Range
-```{r, fig.height = 8, fig.width = 10, echo=FALSE}
-ggmap(berkMap) +
-  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
-    bins = I(5), geom = "polygon", data = mergedf ) +
-  scale_fill_gradient2( "Stop Density",
-    low = "white", mid = "orange", high = "red", midpoint = 25) +
-  labs(x = "Longitude", y = "Latitude") + facet_wrap(~ AgeRange) +
-  scale_alpha(range = c(.2, .55), guide = FALSE) +
-  ggtitle("BPD Stop Contour Map of Berkeley by Age Range") +
-  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
-```
-
-We found that the
+## Conclusion

### 2016-08-08 eda/PaperDrafts/BDPpaper.Rmd
still organizing

--- 2016-08-02 Stop_Data_Analysis.Rmd
@@
 ---
-title: "Stop Data Analysis"
-author: "Mengyu Li"
-date: "July 29, 2016"
+title: "paper1"
+author: "Amy Zhu; Mengyu Li; Rebecca Reus; Shangjun Jiang"
+date: "2 August 2016"
 output: html_document
+runtime: shiny
 ---
@@
 knitr::opts_chunk$set(echo = TRUE)
+library(sp)
+library(ggmap)
+library(tidyr)
+library(readr)
+library(plyr)
+library(dplyr)
+library(ggvis)
+library(tidyr)
+library(stringr)
+library(lubridate)
+library(ggplot2)
+library(readr)
+library(data.table)
+library(rworldmap)
+library(maps)
+library(mapdata)
+library(maptools)
+library(scales)
+library(RgoogleMaps)
+library(tmap)
+library(sp)
+library(rgdal)
+library(rgeos)
+library(RColorBrewer)
 ```
@@
 ```{r, echo=FALSE}
-library(readr)
-library(ggplot2)
-library(dplyr)
-library(lubridate)
-library(stringr)
-library(ggvis)
-library(tidyr)
+############################## location data ############################## 
+mergedf <- readRDS( 'StopData_merged.rds' )
+
+## Location info:
+latmax <- max(mergedf$lat, na.rm = TRUE) 
+latmin <- min(mergedf$lat, na.rm = TRUE)
+lonmax <- max(mergedf$long, na.rm = TRUE) 
+lonmin <- min(mergedf$long, na.rm = TRUE)
+latvals <- c(latmin, latmax)
+lonvals <- c(lonmin, lonmax)
+
+## Make the variables factors:
+mergedf$CarSearch <- factor(mergedf$CarSearch)
+levels(mergedf$CarSearch) <- c("No Search", "Search")
+
+mergedf$Race <- factor(mergedf$Race)
+levels(mergedf$Race) <- c("Asian", "Black", "Hispanic", "Other", "White")
+
+mergedf$Enforcement <- factor(mergedf$Enforcement)
+levels(mergedf$Enforcement) <- c("Arrest", "Citation", "Other", "Warning")
+
+mergedf$Reason <- factor(mergedf$Reason)
+levels(mergedf$Reason) <- c("Other", "Investigation", "Probation/Parole", "Reasonable Suscipcion", "Traffic", "Wanted")
+
+mergedf <- mergedf %>%
+  mutate(Arrest = ifelse( 
+    as.character(Enforcement) == "Arrest", 
+    "Arrested",
+    "Not Arrested") ) %>%
+  mutate(Arrest = factor(Arrest))
+
+v <- factor(mergedf$Other)
+levels(v)
+
+mergedf <- mergedf %>%
+  mutate(Emergency.Psych.Eval = ifelse( str_detect(as.character(Other), "MH"), 
+                                        yes = "Yes",
+                                        no = "No") )
+
+mergedf$Emergency.Psych.Eval <- factor(mergedf$Emergency.Psych.Eval)
+
+### BerkeleyMap 
+berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14)
+
 ```
 
+# Get Shapefiles Again
 ```{r, echo=FALSE}
-# Read data
-stop_by <- read_csv("cleaned_stop_data.csv")
-
-stop_by$Call.Date.Time <- mdy_hm(stop_by$Call.Date.Time)
-stop_by$AgeRange <- as.factor(stop_by$AgeRange)
-
-stop_by <- stop_by %>%
-  mutate(Hour = as.integer(hour(Call.Date.Time))) %>%
-  mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))
-```
-### About the Berkeley PD Stop Data
-
-The Berkeley PD stop data reflects accidents that have been visited by the Berkeley Police Department.
-Raw data set is downloaded from City of Berkeley Open Data website and provided by Berkeley Police Department.
-
-Analysis in this part focuses on the information provided by the dispositions variable.
-The dispositions variable can be divided into three types.
-The first type only includes more than one six-character person message.
-Each character in each person message represents the race, gender and age range of the person, reason and enforcement of the stop, and the car search information during the stop.
-The second type of the dispositions includes only additional dispositions, such as P for primary cases.
-The third type of dispositions includes both person information and additional dispositions.
-
-For the analysis in this part, only person information for all the people in each incident and the Call Time/Date of each incident are analyzed.
-
-### Introduction
-
-items                                | count    | percentage
----------------------------------    | -------- |-------------------
-total incidents reported             | 16255    |     
-dispositions of type one             | 13016    | 80.07%
-dispositions of type two             | 2346     | 14.43%
-dispositions of type three           | 845      | 5.20%
-blank dispositions                   | 48       | 0.30%
-incidents with person information    | 13861    | 85.27%
-  incidents with 1 person            | 13494    | 97.352%
-  incidents with 2 people            | 322      | 2.323%              
-  incidents with 3 people            | 33       | 0.238%
-  incidents with 4 people            | 6        | 0.043%
-  incidents with 5 people            | 6        | 0.043%
-total personnel reported             | 14291    |
-
-(1) From 01/26/2015 to 04/30/2016, the Berkeley Police Department(or BPD) stopped for 16225 incidents.
-80.07% of all stops report only personnel information, 14.43% report only additional information, 5.20% report both and 0.30% report none.
-
-(2) Among all the stops, 13851 incidents recorded 14291 personnel dispositions in total.
-Of the incidents with the personal dispositions, 97.352% involves only one person, 2.648% involves more than two people.
-There are at most 5 people involved in one incident.
-
-### Analysis of Race
-#### Picture: count of person recorded in each race
-```{r, echo=FALSE}
-stop_race <- stop_by %>%
-  ggvis(x = ~Race , fill = ~AgeRange) %>%
-  layer_bars() %>%
-  add_tooltip(function(stop_by) stop_by$AgeRange)
-
-stop_race
+blocks <- readOGR("Census_Tract_Polygons2010","Census_tracts_2010", verbose = TRUE)
+locationCensusFiles <- "Census_Tract_Polygons2010"
+blocks <- readOGR(locationCensusFiles,"Census_tracts_2010", verbose = TRUE)
+b2 <- spTransform(blocks, CRS("+proj=longlat +datum=WGS84"))
+b3 <- fortify(b2)
 ```
 
-```{r, echo=FALSE, eval=FALSE}
-stop_race_total_table <- stop_by %>%
-    group_by(Race) %>%
-    summarise(totaLCount = n())
-stop_race_agerange_table <- stop_by %>%
-    group_by(Race, AgeRange) %>%
-    summarise(count = n())
-stop_race_table <- left_join(stop_race_agerange_table, stop_race_total_table, by = "Race")
-stop_race_table <- stop_race_table %>%
-    mutate(percentage = count/totaLCount) %>%
-    select(-count)
-stop_race_table <- stop_race_table %>%
-    spread(key = AgeRange, value = percentage)
-stop_race_table
+## Population by census block group:
+```{r,echo=FALSE}
+shpfile <- "Census_tracts_2010.shp"
+sh <- readShapePoly("Census_Tract_Polygons2010/Census_tracts_2010.shp")
+sh@data$ID <- as.numeric(sh@data$ID)
+
+# Read in the demographic data and merge on Neighbourhood Id
+demo <- read.csv(file="Census_tracts_2010.csv", header = TRUE)
+sh2 <- merge(sh, demo, by='ID')
+
+# Set the palette
+p <- colorRampPalette(c("white", "red"))(128)
+palette(p)
+
+# Scale the total population to the palette
+pop <- sh2@data$TotalPop.x
+
+cols <- (pop - min(pop))/diff(range(pop))*127+1
+plot(sh, col=cols)
+
 ```
-
-####Table: count of person recorded in each race
-
-Race  | Count | 0-18 | 18-29 | 30-39 | 40+
-----  |------ |------|-------|-------|-------
-A     | 1141  | 8.50%| 43.47%| 21.03%| 26.99%
-B     | 4636  | 2.29%| 34.97%| 25.09%| 37.66%
-H     | 1676  | 3.22%| 46.12%| 29.18%| 21.47%
-O     | 1384  | 2.67%| 43.28%| 29.17%| 26.87%
-W     | 5454  | 1.72%| 29.45%| 25.39%| 43.44%
-
-#### Picture: count of person recorded of each race in each hour
-```{r, echo=FALSE}
-stop_time_percentage <- ggplot(stop_by, aes(x = Hour, fill = AgeRange)) + 
-  geom_bar(aes(y = ..count..)) +
-  labs(y = "Count") +
-  facet_wrap(~Race)
-stop_time_percentage
-```
-
-(1) Among the 14291 personnel information recorded, white people contributes the largest percentage.
-38.16% of all records, namely 5454 people are white.
-black people contributes the second largest percentage.
-32.44% of all records, namely 4636 people are black.
-
-(2) The percentage of people aged from 0 to 18 of each race is less than 9%. People aged from 0 to 18 contributes the smallest percentage in each race, which makes sense in daily life.
-
-(3) Among white people and black people recorded by the BPD, people aged greater than 40  contribute the largest percentage, which are 43.44% and 37.66% respectively.
-
-(4) Among Asian, Hispanic and other people recorded by the BPD, people aged from 18 to 29 contribute the largest percentage, which are 43.37%, 46.12% and 43.28%, respectively.
-
-(5) Black people contribute a especially high percentage of the incidents at night, while the white people contribute a especially high percentage of the incidents at the noon.
-
-### Analysis of Day
-####Picture: count of person recorded in each day in a week
-```{r, echo=FALSE}
-stop_day <- stop_by %>%
-  ggvis(x = ~Day, fill = ~Race) %>%
-  layer_bars() %>%
-  add_tooltip(function(stop_by) stop_by$Race)
-stop_day
-```
-
-```{r, eval=FALSE, echo=FALSE}
-stop_day_total_table <- stop_by %>%
-    group_by(Day) %>%
-    summarise(totaLCount = n())
-stop_day_race_table <- stop_by %>%
-    group_by(Day, Race) %>%
-    summarise(count = n())
-stop_day_table <- left_join(stop_day_race_table, stop_day_total_table, by = "Day")
-stop_day_table <- stop_day_table %>%
-    spread(key = Race, value = count)
-stop_day_table
-```
-####Picture: count of person recorded in each day in a week
-Day   | Count | A    | B    | H    | O    | W
-----  |------ |------|------|------|------|------
-1     | 1692  | 116  | 653  | 221  | 159  | 543
-2     | 1665  | 149  | 543  | 182  | 166  | 625
-3     | 2344  | 189  | 645  | 231  | 236  | 1043
-4     | 2287  | 206  | 616  | 251  | 244  | 970
-5     | 1941  | 152  | 592  | 223  | 168  | 806
-6     | 2231  | 186  | 764  | 268  | 203  | 810
-7     | 2131  | 143  | 823  | 300  | 208  | 657
-
-(1) Among all the day in a week, the number of incidents happened on Sunday and Monday are much less than those happened through Tuesday and Saturday.
-
-(2) The count of records of Asian people is much less than that of any other race of people in each day and the whole week.
-
-(3) Asian people are less liable to commit an incident on Sunday, and more liable to commit an incident on Wednesday.
-
-(4) Black and Hispanic people are less liable to commit an incident on Monday, and more liable to commit an incident on Saturday.
-
-(5) White people are less liable to commit an incident on Sunday, and more liable to commit an incident on Tuesday and Wednesday, which is similar to Asian people.
-
-### Analysis of incidents of each hour in a day
-#### Picture: count of person recorded in each hour in a day
-```{r, echo=FALSE}
-stop_time <- stop_by %>%
-  ggvis(x = ~Hour, fill := "lightpink") %>%
-  layer_histograms(width = input_slider(1, 12, value = 1, step = 2, label = "width"))
-stop_time
-```
-Note: you can change the value of the input slider to see the number of incidents happened in specific time span of the day.
-```{r, eval = FALSE, echo=FALSE}
-stop_hour <- stop_by %>%
-  group_by(Hour) %>%
-  summarise(count = n())
-stop_hour
-```
-
-#### Table: count of person recorded in each hour in a day
-Hour  | Count | Hour | Count| Hour | Count| Hour | Count
-----  |------ |------|------|------|------|------|------
-0     | 823   | 6    | 157  | 12   | 802  | 18   | 544
-1     | 618   | 7    | 341  | 13   | 695  | 19   | 685
-2     | 489   | 8    | 489  | 14   | 515  | 20   | 834
-3     | 289   | 9    | 582  | 15   | 542  | 21   | 971
-4     | 135   | 10   | 682  | 16   | 612  | 22   | 1098
-5     | 61    | 11   | 678  | 17   | 578  | 23   | 1071
-
-(1) time span of 22:00 to 22:59 in a day has the highest incidents count, and 23:00 to 23:59 has the second highest, which means that Berkeley area is most dangerous from 10pm to 12am in a day, which corresponds to daily life experience.
-(2) From 12am to 5am, incidents count decreases gradually.
-A reasonable conjecture of this fact is that more and more people choose to sleep as time goes by in this time span.
-For the similar reason, incidents count increases gradually from 5am to 11am.
-(3) After the evening until the midnight, the incidents number increases gradually again in a day. The tally of the statistics and observation in daily life is in a good agreement.
-
-#### Picture: Probability of BPD stop for a specific age range in a specific hour
-
-```{r, echo=FALSE}
-stop_by_hourCount <- stop_by %>%
-  group_by(Hour) %>%
-  summarise(hourcount = n())
-stop_by_ageRangeCount <- stop_by %>%
-  group_by(Hour, AgeRange) %>%
-  summarise(ageCount = n())
-stop_by_percentage <- left_join(stop_by_ageRangeCount, stop_by_hourCount, by = "Hour")
-stop_by_percentage <- stop_by_percentage %>%
-  mutate(percentage = ageCount/hourcount)
-```
-
-```{r, echo=FALSE}
-stop_time_percentage2 <- ggplot(stop_by_percentage, aes(x = Hour, fill = AgeRange)) + 
-  geom_bar(aes(y = percentage), stat="identity") +
-  labs(y = "Density") +
-  facet_wrap(~AgeRange)
-stop_time_percentage2
-```
-
-(1) People aged from 18 to 29 are obviously liable to commit incidents at night.
-The average ratio of people aged from 18 to 29 to all people stoped by the Berkeley Police Department at night is greater than 40%.
-
-(2) People aged greater than 40 are obviously liable to commit incidents in the daytime.
-The average ratio of people aged greater than 40 to all people stoped by the Berkeley Police Deparment during the day is greater than 40%.
-
-(3) Ratio of incidents commited by people aged between 0 and 18 and people aged from 30 to 39 fluctuates during the daytime and the night, with an average ratio of 2.5% and 25% respectively.
-
-#### Probability of BPD stop for a specific race in a specific hour
-
-```{r, echo=FALSE}
-stop_by_hourCount <- stop_by %>%
-  group_by(Hour) %>%
-  summarise(hourcount = n())
-stop_by_raceRangeCount <- stop_by %>%
-  group_by(Hour, Race) %>%
-  summarise(raceCount = n())
-stop_by_percentage1 <- left_join(stop_by_raceRangeCount, stop_by_hourCount, by = "Hour")
-stop_by_percentage1 <- stop_by_percentage1 %>%
-  mutate(percentage = raceCount/hourcount)
-```
-```{r, echo=FALSE}
-stop_time_percentage3 <- ggplot(stop_by_percentage1, aes(x = Hour, fill = Race)) + 
-  geom_bar(aes(y = percentage), stat="identity") +
-  labs(y = "Density") +
-  facet_wrap(~Race)
-stop_time_percentage3
-```
-
-(1) Black people are obviously liable to commit incidents at night. The average ratio of black people to all people stoped by the Berkeley Police Department at night is about 40%.
-
-(2) White people are obviously liable to commit incidents in the daytime. The average ratio of white people to all people stoped by the Berkeley Police Deparment during the day is about 45%.
-
-(3) Ratio of incidents commited by Asian people, Hispanic people and other people fluctuate during the daytime and the night, with an average ratio of 8%, 11% and 9% respectively.
-
-### Analysis of prefence of Berkeley Police Department arresting people
-#### Picture: Probability of arrested by BPD of a specific race for a specific reason
-```{r,echo=FALSE}
-stop_by_reason_race <- stop_by %>%
-  group_by(Race, Reason) %>%
-  summarise(totalCount = n())
-stop_by_arrest <- stop_by %>%
-  filter(Enforcement == "A") %>%
-  group_by(Race, Reason) %>%
-  summarise(arrestCount = n())
-stop_by_percentage_race <- left_join(stop_by_reason_race, stop_by_arrest, by = c("Race", "Reason"))
-stop_by_percentage_race <- stop_by_percentage_race %>%
-  mutate(percentage = arrestCount / totalCount)
-stop_arrest <- ggplot(stop_by_percentage_race, aes(x = Race, fill = Race)) + 
-  geom_bar(aes(y = percentage), stat="identity") +
-  labs(y = "Density") +
-  facet_wrap(~Reason)
-stop_arrest
-```
-
-#### Table: Probability of arrested by BPD of a specific race for a specific reason
-P   | I      | K     | R     | T     | W     |
-----|------- |-------|-------|-------|-------|
-A   | 10.34% | 75%   | 10.34%| 1.24% | 17.60%|
-B   | 8.31%  | 15.79%| 6.44% | 2.28% | 50.00%|  
-H   | 6.98%  | 16.67%| 9.09% | 2.90% | 25.00%| 
-O   | 10.14% | 18.18%| 5.80% | 0.41% | 41.67%| 
-W   | 8.14%  | 26.79%| 3.40% | 1.04% | 32.56%|
-
-Note: in the picture and table above, I for Investigation, T for Traffic, R for Reasonable Suspicion, K for Probation/Parole, W for wanted.
-
-(1) The probability of arrested in a stop with the reason Traffic is much lower than other reasons. The average conditional probability of arrested given reason is traffic is 1.58%.
-
-(2) The probability of arrested in a stop with the reason Wanted is much higher than other reason. The average conditional probability of arrested given reason is wanted is 33.33%.
-
-(3) An interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and race Asian is 75%, which is much higher than that of any other race.
-Asian people is much more liable to be arrested by Berkeley Police Department during Probation or Parole if stopped by the BPD.
-
-#### Picture: Probability of arrested by BPD of a specific age range for a specific reason
-```{r, echo=FALSE}
-stop_by_reason_age <- stop_by %>%
-  group_by(AgeRange, Reason) %>%
-  summarise(totalCount = n())
-stop_by_arrest_age <- stop_by %>%
-  filter(Enforcement == "A") %>%
-  group_by(AgeRange, Reason) %>%
-  summarise(arrestCount = n())
-stop_by_percentage_age <- left_join(stop_by_reason_age, stop_by_arrest_age, by = c("AgeRange", "Reason"))
-stop_by_percentage_age <- stop_by_percentage_age %>%
-  mutate(percentage = arrestCount / totalCount)
-
-stop_arrest_age <- ggplot(stop_by_percentage_age, aes(x = AgeRange, fill = AgeRange)) + 
-  geom_bar(aes(y = percentage), stat="identity") +
-  labs(y = "Density") +
-  facet_wrap(~Reason)
-stop_arrest_age
-```
-
-#### Table: Probability of arrested by BPD of a specific age range for a specific reason
-P     | I     | K     | R     | T     | W     
-------|-------|-------|-------|-------|-------
-0-18  | 17.95%| 67.57%| 13.95%| 3.24% | 50.00%  
-18-29 | 8.45% | 12.50%| 3.87% | 1.82% | 32.35%  
-30-39 | 6.32% | 25.00%| 5.04% | 1.49% | 26.92%  
-40+   | 7.87% | 23.86%| 6.13% | 1.41% | 31.82%
-
-(1) Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.
-
-(2) Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and age range 0-18 is 67.56%, which is much higher than that of any other age range.
-
-(3) The conditional probability of arrested given person involved aged from 0 to 18 in a stop given any reason is more than that of given person with any other age range.
-Therefore, in a stop, teenagers is more liable to be arrested by Berkeley Police Department.
-
-#### Picture: Probability of arrested by BPD of a specific gender for a specific reason
-```{r, echo=FALSE}
-stop_by_reason_gender <- stop_by %>%
-  group_by(Gender, Reason) %>%
-  summarise(totalCount = n())
-stop_by_arrest_gender <- stop_by %>%
-  filter(Enforcement == "A") %>%
-  group_by(Gender, Reason) %>%
-  summarise(arrestCount = n())
-stop_by_percentage_gender <- left_join(stop_by_reason_gender, stop_by_arrest_gender, by = c("Gender", "Reason"))
-stop_by_percentage_gender <- stop_by_percentage_gender %>%
-  mutate(percentage = arrestCount / totalCount)
-
-stop_arrest_gender <- ggplot(stop_by_percentage_gender, aes(x = Gender, fill = Gender)) + 
-  geom_bar(aes(y = percentage), stat="identity") +
-  labs(y = "Density") +
-  facet_wrap(~Reason)
-stop_arrest_gender
-```
-
-#### Table: Probability of arrested by BPD of a specific gender range for a specific reason
-P     | I      | K     | R     | T     | W    
-----  |------- |-------|-------|-------|-------
-F     | 9.05%  | 50.00%| 5.00% | 1.10% | 28.57%  
-M     | 8.26%  | 20.11%| 5.42% | 1.85% | 37.04%
-
-(1) Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.
-
-(2) Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole of female is 50.00%, which is much higher than that of male, which is 20.11%.

### 2016-08-08 eda/PaperDrafts/pdf?.Rmd
still organizing

--- 2016-08-08 eda/PaperDrafts/BDP_Paperdraft.Rmd
@@
 ---
-title: "BPD_draftpaper"
-author: "Amy Zhu; Mengyu Li; Rebecca Reus; Shangjun Jiang"
+title: "testing"
+author: "Shangjun Jiang"
 date: "2 August 2016"
-output: html_document
+output: pdf_document
 ---
@@
 #### Picture: count of person recorded in each race
-```{r, echo=FALSE}
+```{r, echo=FALSE, eval = FALSE}
 stop_race <- stop_by %>%
@@
 ####Picture: count of person recorded in each day in a week
-```{r, echo=FALSE}
+```{r, echo=FALSE, eval = FALSE}
 stop_day <- stop_by %>%
@@
 #### Picture: count of person recorded in each hour in a day
-```{r, echo=FALSE}
+```{r, echo=FALSE, eval= FALSE}
 stop_time <- stop_by %>%

### 2016-08-08 eda/PaperDrafts/testing.Rmd
still organizing

--- 2016-08-08 eda/PaperDrafts/BDPpaper.Rmd
@@
 ---
-title: "paper1"
-author: "Amy Zhu; Mengyu Li; Rebecca Reus; Shangjun Jiang"
+title: "testing"
+author: "Shangjun Jiang"
 date: "2 August 2016"
 output: html_document
-runtime: shiny
 ---
@@
 ```
+
+## Population by census block group:
 
@@
 ## Population by census block group:
+![](population_census.png)
 ```{r,echo=FALSE}
@@
 
+png(file="population_census.png",width=700,height=500)
+plot(sh, col=cols)
+dev.off()
+
 ```
+
+```{r, fig.height = 5, fig.width = 10, echo=FALSE}
+df <- subset(mergedf, as.character(Reason) == "Traffic")
+
+ggmap(berkMap) +
+  geom_point(aes(x = long, y = lat, colour = Race, size= Arrest), data = df, alpha = 0.5) + 
+  scale_size_discrete( range = c(5, 2) ) +
+  theme ( 
+		panel.grid.major = element_blank (), # remove major grid
+		panel.grid.minor = element_blank (),  # remove minor grid
+		axis.text = element_blank (), 
+		axis.title = element_blank (),
+		axis.ticks = element_blank ()
+		) + 
+  ggtitle("BPD Traffic Stops Only, by Race 2015-2016") +
+  facet_grid(~ Race)
+```
+
+```

### 2016-08-08 eda/StopDataExploration/StopData_Maps.Rmd
still organizing

--- 2016-08-03 Rebeccas_location_Code2/StopData_Maps.Rmd
@@
 bMap12
+
+ggmap(berkMap) +
+  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
+    size = 2, bins = 5, data = b3, geom = "polygon") +
+  scale_fill_gradient(low = "black", high = "red") +
+    theme (panel.grid.major = element_blank (), # remove major grid
+		       panel.grid.minor = element_blank (),  # remove minor grid
+		       axis.text = element_blank (), 
+		       axis.title = element_blank (),
+	         axis.ticks = element_blank ()) + 
+  ggtitle("Population by census block group") +
+  labs(alpha = element_blank())+
+  guides(alpha = FALSE)
+
 ```

### 2016-08-08 pdf1.Rmd
update

--- 2016-08-08 eda/PaperDrafts/draft2.Rmd
@@
 ---
-title: "STAT 133 Final Project: Visualizing Berkeley Police Data"
-author: "Amy Zhu, Mengyu Li, Rebecca Reus and Shangjun Jiang"
-date: "August 7, 2016"
-output: html_document
+title: "Draft Paper"
+author: "Shangjun Jiang"
+date: "8 August 2016"
+output: 
+    pdf_document:
+    latex_engine: xelatex
+sansfont: Times New Roman
+
 ---
@@
 library(formattable)
+library(pander)
+library(extrafont)
+font_import()
+loadfonts()
+setwd("~/Desktop/Stats R/PROJECT/finaldraftfiles")
 ```
@@
 ```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
+
 # read in the data and define parameters:
 
-# data file names:
+############################ read in the data ###########################
 
-file_stops <- "../../clean_data/StopData_clean.csv" 
-file_berkcensus2010 <- "../../clean_data/berk_census2010_clean.csv" 
-file_berkcensus2010map <-"../../clean_data/berk_census2010_mapdata.rds" 
-file_tables <-  "../../clean_data/Tables_AgeSexRacePopStop.rds"
-file_calls <-  "../../clean_data/CallsForService.rds"
-file_arrests <- "../../raw_data/Berkeley_PD_Log_-_Arrests.csv"
-file_jail <- "../../raw_data/Berkeley_PD_Log_-_Jail_Bookings.csv"
+############################ Police Stop Data ###########################
+stops <- read_csv("clean_data/StopData_clean.csv")
 
-# read in the data:
+stops$Gender<- stops$Gender %>% 
+  str_replace("F", "Female") %>% 
+  str_replace("M", "Male")
 
-# Police Stop Data:
-stops <- read_csv(file = file_stops)
+stops$AgeRange<- stops$AgeRange %>% 
+  str_replace("1", "0-18") %>% 
+  str_replace("2", "18-29") %>% 
+  str_replace("3", "30-39") %>%  
+  str_replace("4", "40+")
 
-# Census 2010 Data by Polygonal Tract Number:
-berkcensus2010 <- read_csv(file = file_berkcensus2010)
+########################### arrest data ###########################
+arrest <- read_csv("raw_data/Berkeley_PD_Log_-_Arrests.csv")
 
-# Census 2010 Data by Polygonal Tract Number, ready for mapping:
-berkcensus2010map <- readRDS(file_berkcensus2010map)
+############################ jail data ###########################
+jail <- read.csv("raw_data/Berkeley_PD_Log_-_Jail_Bookings.csv")
 
-# Police Calls for service:
-callservice <- read_csv(file = file_calls)
+############################ Census 2010 Data by Polygonal Tract Number ###########################
+berkcensus2010 <- read_csv("clean_data/berk_census2010_clean.csv")
 
-# Tables to use in paper:
-tables <- readRDS(file = file_tables)
+census <- read.csv("raw_data/Census_Data_2000_And_2010.csv")
+census2010 <- census[census$Year==2010,]
+totalpop <- census2010[1,]$Amount
+
+racecensus2010 <- census2010[census2010$Heading=="Not Hispanic or Latino"|census2010$Heading=="HISPANIC OR LATINO AND RACE",]
+sexcensus2010 <- census2010[census2010$Heading=="Sex",]
+agecensus2010 <- census2010[census2010$Heading=="Age",]
+
+race <- select(racecensus2010, Description, Amount) 
+race$Description <- race$Description %>% 
+  str_replace_all("Hispanic or Latino.*", "Hispanic") %>% 
+  str_replace("Black or African American", "Black") %>% 
+  str_replace("American Indian and Alaska Native", "Other") %>% 
+  str_replace("Native Hawaiian and Other Pacific Islander", "Other") %>% 
+  str_replace("Some other race", "Other") %>% 
+  str_replace("Two or more races", "Other")
+
+race <- race[race$Description !="Not Hispanic",]
+
+race <- race %>% 
+  group_by(Description) %>% 
+  tally(Amount)
+
+names(race) <- c("Description", "Counts")
+
+sex <- select(sexcensus2010, Description, Amount)
+
+age <- select(agecensus2010, Description, Amount) 
+age <- age[age$Description != "Median age",]
+age$Description <- c("0-18", "0-18", "18-64", "65+")
+
+age<- age %>% 
+  group_by(Description) %>% 
+  tally(Amount)
+names(age) <- c("Description", "Counts")
+
+################## Census 2010 Data by Polygonal Tract Number, ready for mapping #################
+berkcensus2010map <- readRDS("clean_data/berk_census2010_mapdata.rds")
+
+############################ Police Calls for service ############################ 
+callservice <- readRDS("clean_data/CallsForService.rds")
+
+callservice<- callservice %>% 
+  mutate(latlong= 
+           str_replace_all(callservice$Block_Location,"[0-9]* [A-Za-z]*", "") %>%
+           str_replace_all("\nBerkeley,\n", "") %>% 
+           str_replace("[A-Za-z]*", "") %>% 
+           str_replace(";","") %>% 
+           str_replace("&","") %>% 
+           str_replace("[A-Za-z]*", ""))
+
+callservice$latlong<- str_replace(callservice$latlong, "\\(", "") %>% 
+                    str_replace("\\)", "")
+
+callservice<- callservice %>% 
+  separate(latlong, c("lat", "long"), sep=",")
+
+callservice$lat <- as.numeric(callservice$lat)
+callservice$long <- as.numeric(callservice$long)
+
+#############################  Tables to use in paper ############################ 
+tables <- readRDS("clean_data/Tables_AgeSexRacePopStop.rds")
 ss <- as.data.frame(tables[5]) # stop data by race proportional to population
@@
 
-# arrest data:
-arrest <- read_csv(file_arrests)
+```
 
-# jail data:
-jail <- read.csv(file_jail)
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
 
@@
 # zoomed out berkeley map:
-berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 12) 
+berkMap = map = get_map(location = c(lon = mean(lonvals), 
+                                     lat = mean(latvals)), 
+                        zoom = 12) 
 berkgg_zoom1 <-ggmap(berkMap) +
-  xlim(-122.335, lonmax) + ylim(latvals) +
+  xlim(-122.335, lonmax) + 
+  ylim(latvals) +
   ditch_the_axes
@@
 # zoomed in berkeley map:
-berkMap2 = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14) 
+berkMap2 = map = get_map(location = c(lon = mean(lonvals), 
+                                      lat = mean(latvals)), 
+                         zoom = 14) 
 berkgg_zoom2 <-ggmap(berkMap2) +
@@
 
-### CENSUS DATA MAPS:
+########### CENSUS DATA MAPS: ########
 
-# create color palette:
-
+###############################  create color palette ############################## 
 p <- colorRampPalette(c("white", "red"))(128)
@@
 
-# create the maps:
+############################## multiplot function ############################## 
+multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
+  library(grid)
+
+  # Make a list from the ... arguments and plotlist
+  plots <- c(list(...), plotlist)
+
+  numPlots = length(plots)
+
+  # If layout is NULL, then use 'cols' to determine layout
+  if (is.null(layout)) {
+    # Make the panel
+    # ncol: Number of columns of plots
+    # nrow: Number of rows needed, calculated from # of cols
+    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
+                    ncol = cols, nrow = ceiling(numPlots/cols))
+  }
+
+ if (numPlots==1) {
+    print(plots[[1]])
+
+  } else {
+    # Set up the page
+    grid.newpage()
+    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
+
+    # Make each plot, in the correct location
+    for (i in 1:numPlots) {
+      # Get the i,j matrix positions of the regions that contain this subplot
+      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
+
+      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
+                                      layout.pos.col = matchidx$col))
+    }
+  }
+}
+
+```
+
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
+
+############################ create the maps ###########################
+
+m <- stops %>% 
+  group_by(Race) %>% 
+  tally()
+names(m) <- c("Race", "stop")
+names(race) <- c("Race", "census")
+
+pop1 <- left_join(race,m,by = "Race")
+pop1 <- pop1 %>% 
+  mutate(percentage=(stop/census))
+pop1$percentage<- round(pop1$percentage, digits = 2)
+
+pop1 <- pop1 %>% 
+  mutate(totalp = (census/totalpop))
+pop1$totalp <- round(pop1$totalp, digits =2)
+
+p1 <- ggplot(pop1)+
+  geom_bar(aes(x=Race,
+               y=percentage), 
+           stat = "identity",
+           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+  geom_text(aes(x=Race,
+               y=percentage,
+               label = paste0(percentage*100,"%")), 
+            size=5,
+            colour= "#000000")+
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  labs(title = "The Percentage of Being \nStopped by Police (%)",
+       x = "Race",
+       y = "Percentage") +
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) + 
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
+p2 <- ggplot(pop1)+
+  geom_bar(aes(x=Race,
+               y=totalp), 
+           stat = "identity",
+           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+  geom_text(aes(x=Race,
+               y=totalp,
+               label = paste0(totalp*100,"%")), 
+            size=5,
+            colour= "#000000")+
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  labs(title = "The Percentage of Berkeley \nCensus Data 2010 (%)",
+       x = "Race",
+       y = "Percentage") +
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) + 
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
 
@@
 map_stop1 <- berkgg_zoom2 +
-  geom_point(data = stops, aes(x=long,y=lat), alpha = .1) +
-  ggtitle("Berkeley Police Stops, 2015-2016")
+  geom_point(data = stops, aes(x=long,y=lat), alpha = .2) +
+  ggtitle("Berkeley Police Stops, 2015-2016")+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
 
-map_stop1
+# calls for service
+map_callservice1 <- berkgg_zoom2 +
+  geom_point(data = na.omit(callservice), aes(x=long,y=lat), alpha = .2) +
+  ggtitle("Calls for Service (not criminal reports) \nWithin 180 days (Feb-July 2016)")+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
 
-# populatrion density map
+```
+
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
+
+############################ populatrion density map ###########################
 popdenmap <- berkgg_zoom1 + 
@@
   labs(fill = "Percent") +
-  scale_fill_gradientn(colors = p)
+  scale_fill_gradientn(colors = p) +
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
 
@@
 black2 <- berkgg_zoom1 + 
-  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Black), color = "white", alpha = .7) +
-  scale_fill_gradientn(colors = p, limits = c(0,1)) +
-  guides( fill = "none" ) +
-  ggtitle("% Black")
-
-# white population:
+  geom_polygon(data = berkcensus2010map, 
+               aes(x = long, 
+                   y = lat, 
+                   group = group, 
+                   fill = Percent.Black), 
+               color = "white", 
+               alpha = .7) +
+  scale_fill_gradientn(colors = p, 
+                       limits = c(0,1)) +
+  guides(fill = "none") +
+  ggtitle("% Black")+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+# White population:
 
 w2 <- berkgg_zoom1 + 
-  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.White), color = "white", alpha = .7) +
-  scale_fill_gradientn(colors = p, limits = c(0,1)) +
-  guides( fill = "none" ) +
-  ggtitle("% White")
+  geom_polygon(data = berkcensus2010map, 
+               aes(x = long, 
+                   y = lat, 
+                   group = group, 
+                   fill = Percent.White), 
+               color = "white", 
+               alpha = .7) +
+  scale_fill_gradientn(colors = p, 
+                       limits = c(0,1)) +
+  guides(fill = "none") +
+  ggtitle("% White")+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
 
-# asian population:
+# Asian population:
 
 a2 <- berkgg_zoom1 + 
-  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Asian), color = "white", alpha = .7) +
-  scale_fill_gradientn(colors = p, limits = c(0,1)) +
-  guides( fill = "none" ) +
-  ggtitle("% Asian")
+  geom_polygon(data = berkcensus2010map, 
+               aes(x = long, 
+                   y = lat, 
+                   group = group, 
+                   fill = Percent.Asian), 
+               color = "white", 
+               alpha = .7) +
+  scale_fill_gradientn(colors = p, 
+                       limits = c(0,1)) +
+  guides(fill = "none") +
+  ggtitle("% Asian")+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
 
-# hispanic population:
+# Hispanic population:
 
 h2 <- berkgg_zoom1 + 
-  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Hispanic), color = "white", alpha = .7) +
-  scale_fill_gradientn(colors = p, limits = c(0,1)) +
-  guides( fill = "none" ) +
-  ggtitle( "% Hispanic")
+  geom_polygon(data = berkcensus2010map, 
+               aes(x = long, 
+                   y = lat, 
+                   group = group, 
+                   fill = Percent.Hispanic), 
+               color = "white", 
+               alpha = .7) +
+  scale_fill_gradientn(colors = p, 
+                       limits = c(0,1)) +
+  guides(fill = "none") +
+  ggtitle("% Hispanic")+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
 
-# other population:
+# Other population:
 
 o2 <- berkgg_zoom1 + 
-  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Other), color = "white", alpha = .7) +
-  scale_fill_gradientn(colors = p, limits = c(0,1)) +
-  guides( fill = "none" ) +
-  ggtitle("% Other")
+  geom_polygon(data = berkcensus2010map, 
+               aes(x = long, 
+                   y = lat, 
+                   group = group, 
+                   fill = Percent.Other), 
+               color = "white", 
+               alpha = .7) +
+  scale_fill_gradientn(colors = p, 
+                       limits = c(0,1)) +
+  guides(fill = "none") +
+  ggtitle("% Other")+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
 
+```
+
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
 # bar graphs based on stop data:
 names(ss)[2:3] <- c("Percent Stopped", "Percent Population Stopped")
-bargraph1 <- ggplot(ss, aes(Race, `Percent Stopped`, fill = Race))+
-  geom_bar(stat = "identity") +
-  guides( fill = FALSE)
+
+bargraph1 <- ggplot(ss, 
+                    aes(x=Race, 
+                        y=`Percent Stopped`, 
+                        fill = Race))+
+  geom_bar(stat = "identity",
+           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0")) +
+  geom_text(aes(x=Race,
+               y=`Percent Stopped`,
+               label = paste0(`Percent Stopped`)), 
+            size=5,
+            colour= "#000000")+
+  scale_y_discrete(labels=percent, limits = c(0,1))+
+  labs(title = "The percentage of berkeley being stopped (%)",
+       x = "Race",
+       y = "Percent stopped")+
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank())+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
 bargraph2 <- ggplot(ss, aes(Race, `Percent Population Stopped`, fill = Race))+
-  geom_bar(stat = "identity") +
-  guides( fill = FALSE)
+  geom_bar(stat = "identity")+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
+```
+
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
 
@@
   ggtitle("All BPD Stops Density, 2015-2016") +
-  labs(alpha = element_blank())+
+  labs(alpha = element_blank()) +
   guides(alpha = FALSE)
@@
 stopdensitymap4 <- berkgg_zoom2 +
-  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
+  stat_density2d(aes(x = long, 
+                     y = lat, 
+                     fill = ..level.., 
+                     alpha = ..level..),
     bins = I(5), geom = "polygon", data = stops ) +
@@
   scale_alpha(range = c(.2, .55), guide = FALSE) +
-  ggtitle("Berkeley Police Stop Density Map, 2015-2016") +
-  guides(fill = "none")
-stopdensitymap4
+  ggtitle("Berkeley Police Stop \nDensity Map, 2015-2016")+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
 
@@
 stopdensitymap3 <- berkgg_zoom2 +
-  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
-    bins = I(5), geom = "polygon", data = stops ) +
-  scale_fill_gradient2( "StopDensity",
+  stat_density2d(aes(x = long, 
+                      y = lat, 
+                      fill = ..level.., 
+                      alpha = ..level..),
+    bins = I(5), geom = "polygon", data = stops) +
+  scale_fill_gradient2("StopDensity",
     low = "white", mid = "orange", high = "red", midpoint = 100) +
-  labs(x = "Longitude", y = "Latitude") + facet_grid( ~ Race ) +
+  labs(x = "Longitude", y = "Latitude") + facet_grid( ~ Race) +
   scale_alpha(range = c(.2, .55), guide = FALSE) +
@@
   guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10)) +
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10)) +
   guides(fill = FALSE)
-
-stopdensitymap3
 
@@
 df <- subset(stops, as.character(Enforcement) == "Arrest")
+
 map_arrested1 <- berkgg_zoom1 +
-  geom_point(aes(x = long, y = lat, colour = Race), data = df, alpha = 0.5) + 
+  geom_point(aes(x = long, 
+                 y = lat, 
+                 colour = Race), 
+             data = df, 
+             alpha = 0.7) + 
   theme ( 
@@
 		axis.title = element_blank (),
-		axis.ticks = element_blank ()
-		) + 
+		axis.ticks = element_blank ()) + 
   ggtitle("Police Stops Leading to Arrests") +
   facet_grid(~ Race) +
-  guides( colour = FALSE)
+  guides(colour = FALSE) +
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
 
+```
+
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
 # stops by reason:
+stop_reason <- stops[stops$Reason != "Other",]
+
 map_reasons1 <- berkgg_zoom2 +
-  geom_point(aes(x = long, y = lat, colour = Reason), data = stops, alpha = .7) +
+  geom_point(aes(x = long, 
+                 y = lat, 
+                 colour = Reason), 
+             data = stop_reason, 
+             alpha = .7) +
   theme ( 
@@
 		axis.title = element_blank (),
-		axis.ticks = element_blank ()
-		) + 
-  ggtitle("Police Stops by Reason")
+		axis.ticks = element_blank ()) + 
+  ggtitle("Police Stops by Reason")+
+  facet_grid(~Reason)+
+  guides(colour = FALSE)+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
+stop_reason <- stop_reason %>% 
+  group_by(Race, Reason) %>% 
+  tally()
+
+```
+
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
+
+############################ create the maps ###########################
 
 # arrest data histogram by race
+arrest1 <- na.omit(arrest) %>% 
+  group_by(Race) %>% 
+  tally()
+total <- arrest1 %>% 
+  tally()
+arresttotal <- 192
+arrest1 <- arrest1 %>% 
+  mutate(percentage = n/192)
+arrest1$percentage <- round(arrest1$percentage, digits = 2)
+
+jail1 <-na.omit(jail) %>% 
+  group_by(Race) %>% 
+  tally()
+total <- jail1 %>% 
+  tally()
+jailtotal <- 242
+jail1<- jail1 %>% 
+  mutate(percentage = n/242)
+jail1$percentage <- round(jail1$percentage, digits = 2)
+
+bararrest<- ggplot(arrest1)+
+  geom_bar(aes(x=Race,
+               y= percentage),
+           stat = "identity",
+           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+  geom_text(aes(x=Race,
+               y=percentage,
+               label = paste0(percentage*100,"%")), 
+            size=5,
+            colour= "#000000")+
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  labs(title = "The Percentage of Being Arrested \nAfter being Stopped (%)",
+       x = "Race",
+       y = "Percentage") +
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) + 
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 14),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
+barjail <- ggplot(jail1)+
+  geom_bar(aes(x=Race,
+               y=percentage),
+           stat = "identity",
+           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+  geom_text(aes(x=Race,
+               y=percentage,
+               label = paste0(percentage*100,"%")), 
+            size=5,
+            colour= "#000000")+
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  labs(title = "The Percentage of Being jailed \nAfter being Stopped (%)",
+       x = "Race",
+       y = "Percentage") +
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) + 
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 14),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
 histogram_arrest1 <- ggplot(na.omit(arrest))+
@@
        title = "Arrest Data") +
-  guides( fill = FALSE)
+  guides(fill = FALSE)+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 14),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
 
@@
        title = "Jail Data") + 
-  guides(fill = FALSE)
+  guides(fill = FALSE)+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 14),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
 
@@
   labs(title = "Number of people each race in every hour") +
-  guides( fill = FALSE )
+  guides(fill = FALSE)+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 14),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
 ```
@@
 
-```{r, warning = FALSE, echo = F, fig.width = 5, fig.height = 5, message = FALSE, fig.align="center"}
-map_stop1
+```{r, warning = FALSE, echo = F, fig.width = 10, fig.height = 5, message = FALSE, fig.align="center"}
+#multiplot(map_stop1,map_callservice1, cols =2)
+#map_stop1
+#map_callservice1
+grid.arrange(map_stop1,map_callservice1, ncol = 2)
 ```
@@
 ```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.align = "center"}
-#### !!!!!!!!!!!!!!!!!!!!!INCLUDE BAR GRAPH HERE of % race stopped vs. relative percent, as described above!!!!!!!!!!!!!!!!!!!!!!!!
+
+#!!!!!!!!!!!!!!!!!!!!!INCLUDE BAR GRAPH HERE of % race stopped vs. percent race population, as described above!!!!!!!!!!!!!!!!!!!!!!!!
+multiplot(p1, p2,cols=2)
 ```
@@
 
-!!!!!!!!!!!!!!!!!!!!!INCLUDE BAR GRAPH & TABLE OF ENFORCEMENT BY RACE HERE (Use percentage %, not count) !!!!!!!!!!!!!!!!!!!!!!!!
-
-```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.align = "center"}
-#### INCLUDE BAR GRAPH & TABLE OF ENFORCEMENT BY RACE HERE(Use percentage, not count)!!!!!!!!!!!!!!!!!!!!!!!!
+```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.align = "center", fig.width = 8, fig.height = 3}
+#map_enforcement
 ```
@@
 #### INCLUDE BAR GRAPH & TABLE OF POLICE STOPS BY REASON HERE(Use percentage, not count)!!!!!!!!!!!!!!!!!!!!!!!!
+multiplot(bararrest, barjail, cols = 2)
 ```

### 2016-08-08 clean_data/CleanDataScripts/Geocode_GetLocation.R
organizing

@@ -1,0 +1,321 @@
+###################################################################################################################################
+## Title: STOP DATA LOCATION SCRIPT FOR BERKELEY, CA
+## Author: Rebecca Reus, STAT 133 Summer 2016, UC Berkeley
+## Information:
+## Uses part of Geocoding script for large list of addresses created by Shane Lynn, 10/10/2013 at http://www.r-bloggers.com/batch-geocoding-with-r-and-google-maps/
+## Police Stop Data source: https://data.cityofberkeley.info/
+###################################################################################################################################
+
+
+# Jenny WD:
+# setwd("~/Desktop/Stats R/PROJECT/finalproject_angry_ladies/Rebeccas_location_Code2")
+
+###################################################################################################################################
+## LIBRARIES:
+## Before running, please make sure you have installed ALL of these packages:
+###################################################################################################################################
+library(sp)
+library(ggmap)
+library(tidyr)
+library(plyr)
+library(dplyr)
+library(stringr)
+library(lubridate)
+library(ggplot2)
+library(readr)
+library(data.table)
+library(rworldmap)
+library(maps)
+library(mapdata)
+library(maptools)
+library(scales)
+library(RgoogleMaps)
+library(tmap)
+library(sp)
+library(rgdal)
+library(rgeos)
+
+###################################################################################################################################
+## DATA TIDYING:
+###################################################################################################################################
+
+# Get the data from the file in your working directory:
+infile <- "StopData"
+data <- read.csv(paste0('./', infile, '.csv'))
+
+# Filter out the data that is not already in coordinate form (doesn't have longitude and latitude values):
+stop_df <- data %>%
+  subset( !str_detect( Location, "~-"))
+
+# Since we removed some rows, we need to rename the row names to match the row number (used for indexing):
+rownames(stop_df) <- 1:nrow(stop_df)
+
+# We need to adjust the Location data to account for popular abbreviations and mispellings, in addition to some other data tidying:
+stop_df <- stop_df %>%
+  mutate( Location = as.character( Location ) ) %>%                  # location = character information
[…]

### 2016-08-08 eda/GeolocationExploration/StopDataGetLocation.R

@@ -1,321 +1,0 @@
-###################################################################################################################################
-## Title: STOP DATA LOCATION SCRIPT FOR BERKELEY, CA
-## Author: Rebecca Reus, STAT 133 Summer 2016, UC Berkeley
-## Information:
-## Uses part of Geocoding script for large list of addresses created by Shane Lynn, 10/10/2013 at http://www.r-bloggers.com/batch-geocoding-with-r-and-google-maps/
-## Police Stop Data source: https://data.cityofberkeley.info/
-###################################################################################################################################
-
-
-# Jenny WD:
-# setwd("~/Desktop/Stats R/PROJECT/finalproject_angry_ladies/Rebeccas_location_Code2")
-
-###################################################################################################################################
-## LIBRARIES:
-## Before running, please make sure you have installed ALL of these packages:
-###################################################################################################################################
-library(sp)
-library(ggmap)
-library(tidyr)
-library(plyr)
-library(dplyr)
-library(stringr)
-library(lubridate)
-library(ggplot2)
-library(readr)
-library(data.table)
-library(rworldmap)
-library(maps)
-library(mapdata)
-library(maptools)
-library(scales)
-library(RgoogleMaps)
-library(tmap)
-library(sp)
-library(rgdal)
-library(rgeos)
-
-###################################################################################################################################
-## DATA TIDYING:
-###################################################################################################################################
-
-# Get the data from the file in your working directory:
-infile <- "StopData"
-data <- read.csv(paste0('./', infile, '.csv'))
-
-# Filter out the data that is not already in coordinate form (doesn't have longitude and latitude values):
-stop_df <- data %>%
-  subset( !str_detect( Location, "~-"))
-
-# Since we removed some rows, we need to rename the row names to match the row number (used for indexing):
-rownames(stop_df) <- 1:nrow(stop_df)
-
-# We need to adjust the Location data to account for popular abbreviations and mispellings, in addition to some other data tidying:
-stop_df <- stop_df %>%
-  mutate( Location = as.character( Location ) ) %>%                  # location = character information
[…]

### 2016-08-08 eda/CensusDataExploration/HeatmapExploration2.Rmd

--- 2016-08-06 eda/Heatmap/HeatmapExploration3.Rmd
@@
 
+### All BPD Stops Density, 2015-2016
+```{r}
+# a contour plot
+berkgg +
+  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
+                 size = 2, bins = 5, data = df, geom = "polygon") +
+  scale_fill_gradient(low = "black", high = "red") +
+  theme (panel.grid.major = element_blank (), # remove major grid
+         panel.grid.minor = element_blank ()  # remove minor grid
+  )+ 
+  ggtitle("All BPD Stops Density, 2015-2016") +
+  labs(alpha = element_blank())+
+  guides(alpha = FALSE)
+
+```
+
+### Works, but no google maps:
+```{r, eval = F, include = F}
+### Read in .shp file only for Census:
+shpfile <- str_c(census_tract_folder, "/", census_tract_shp, ".shp")
+sh <- readShapePoly(shpfile)
+```
+
+```{r, eval = F, include = F}
+### Plot the census shape file (polygon outline only)
+plot(sh)
+```
+
+```{r, eval = F, include = F}
+### Modify the shape file to plot the population info:
+sh@data$ID <- as.numeric(sh@data$ID)
+
+# Read in the demographic data and merge on Neighbourhood Id
+census5_2 <- read.csv(file=census_tract_csv, header = TRUE)
+sh2 <- merge(sh, census5_2, by='ID')
+
+# Set the palette
+p <- colorRampPalette(c("white", "red"))(128)
+palette(p)
+
+# Scale the total population to the palette
+popamts <- sh2@data$TotalPop.x
+
+# create the colors:
+cols <- (popamts - min(popamts))/diff(range(popamts))*127+1
+```
+
+```{r, eval = F, include = F}
+### Plot the census shape data with colors for population:
+plot(sh, col=cols)
+```
+
+### Cleaning the Berkeley population info (do not run):
+```{r, eval = F, include = F}
+#write.csv(alameda, file = "alameda_population_by_tract.csv")
+# alameda = alameda[-1,]
+# rownames(alameda) <- 1:nrow(alameda)
+# names(alameda)[1] <- "GEO"
+# 
+# alameda2 <- alameda %>%
+#   separate(GEO, c("Block","Group", "Tract", "County", "State"), ", ")
+# 
+# names(alameda2)[3] <- "NAME10"
+# 
+# alameda2 <- alameda2 %>%
+#   mutate(NAME10 = str_replace(NAME10, "Census Tract ", ""))
+# 
+# alameda2 <- alameda2 %>%
+#   select(-Block,-Group,-County,-State) 
+# 
+# a3 <- alameda2 %>%
+#         group_by(NAME10)%>% 
+#         summarise(Total.population = sum(Total.population),
+#                   Hispanic.or.Latino = sum(Hispanic.or.Latino),
+#                   Total.population..not.Hispanic.or.Latino = sum(Total.population..not.Hispanic.or.Latino),
+#                   One.race.total = sum(One.race.total),
+#                   White = sum(White),
+#                   Black.or.African.American = sum(Black.or.African.American),
+#                   American.Indian.and.Alaska.Native = sum(American.Indian.and.Alaska.Native),
+#                   Asian = sum(Asian),
+#                   Native.Hawaiian.and.Other.Pacific.Islander = sum(Native.Hawaiian.and.Other.Pacific.Islander),
+#                   Some.Other.Race = sum(Some.Other.Race),
+#                   Two.or.More.Races = sum(Two.or.More.Races))
+
+#census_alameda <- left_join(censusCSV, a3, by = "NAMELSAD10")
+```
+
+```{r, eval = F, include = F}
+# cleaning the berkeley population data, do not run me:
+
+# census.df <- fortify(censusT, region = 'NAME10')
+# pop <- censusCSV
+# pop <- pop %>%
+#   select(NAME10, TotalPop) %>%
+#   mutate(NAME10 = as.character(NAME10))
+# 
+# names(a3)[1] <- "id"
+# names(pop)[1] <- "id"
+# 
+# pop <- left_join(pop, a3, by = "id")
+# census.df <- left_join(census.df, pop, by = "id")
+#write.csv2(a3,"alameda_clean_census.csv")
+#write.csv2(pop, "berkeley_census_clean.csv")
+```
+
 ```{r, include = F, echo = F}
@@
   select(-TotalPop) %>%
+  mutate(Relative.Per.Black = Black.or.African.American/sum(Black.or.African.American)) %>%
+  mutate(Relative.Per.Other = Other/sum(Other)) %>%
+  mutate(Relative.Per.White = White/sum(White)) %>%
+  mutate(Realtive.Per.Asian = Asian/sum(Asian)) %>%
+  mutate(Relative.Per.Hispanic = Hispanic.or.Latino/sum(Hispanic.or.Latino)) %>%
   select(-American.Indian.and.Alaska.Native,-Native.Hawaiian.and.Other.Pacific.Islander, -Some.Other.Race)
-
-# %>%
-#   mutate(Relative.Per.Black = Black.or.African.American/sum(Black.or.African.American)) %>%
-#   mutate(Relative.Per.Other = Other/sum(Other)) %>%
-#   mutate(Relative.Per.White = White/sum(White)) %>%
-#   mutate(Realtive.Per.Asian = Asian/sum(Asian)) %>%
-#   mutate(Relative.Per.Hispanic = Hispanic.or.Latino/sum(Hispanic.or.Latino)) %>%
-
-names(bpop2)[2] <- "Population"
-names(bpop2)[3] <- "Hispanic"
-names(bpop2)[5] <- "Black"
-names(bpop2)
 
@@
 ```
-
-### BPD Stops:
-```{r}
-# edit bpd stops:
-# nblacks <- sum(df$Race == "Black")
-# nasians <- sum(df$Race == "Asian")
-# nwhite <- sum(df$Race == "White")
-# nother <- sum(df$Race == "Other")
-# nhisp <- sum(df$Race == "Hispanic")
-# totalpop <- sum(berkpop$TotalPop)
-# stop.summary <- data.frame(Race = unique(df$Race))
-
-stop.summary <- df %>%  
-   group_by(Race) %>%
-   summarise(Stopped=n() ) %>%
-  mutate(Percent.Stopped = Stopped/nrow(df))
-
-racepops <- c(sum(bpop2$Asian), sum(bpop2$Black), sum(bpop2$Hispanic), sum(bpop2$Other), sum(bpop2$White) )
-
-stop.summary$Race.Population = racepops
-stop.summary$Percent.Race.Stopped = stop.summary$Stopped/stop.summary$Race.Population
-
-```
-
-```{r}
-# all stop density:
-allbpd_density <- berkgg +
-  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = 2* ..level..),
-                 size = 2, bins = 5, data = df, geom = "polygon") +
-  scale_fill_gradientn(trans = "log10", colors = p) +
-  theme (panel.grid.major = element_blank (), # remove major grid
-         panel.grid.minor = element_blank ()  # remove minor grid
-  ) + 
-  ggtitle("All BPD Stops Density, 2015-2016") +
-  labs(alpha = element_blank()) +
-  guides(alpha = FALSE)
-allbpd_density
-
-# by race:
-bpd_race <- berkgg +
-  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha =  ..level..),
-                 size = 2, bins = 5, data = df, geom = "polygon") +
-  scale_fill_gradientn(trans = "log10", colors = p) +
-  theme (panel.grid.major = element_blank (), # remove major grid
-         panel.grid.minor = element_blank ()  # remove minor grid
-  ) + 
-  guides(alpha = FALSE) +
-  facet_grid( ~ Race)
-bpd_race
-
-```
-
-### Bar Charts:
-
-```{r}
-p <- ggplot(df, aes(Enforcement, fill = Enforcement)) + 
-    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  facet_grid(~Race) +
-  scale_y_continuous(labels=percent) +
-  labs(y = "Percent Stopped") +
-  ggtitle("Results of Stops by Race")
-p
-```
-
-```{r}
-p1 <- ggplot(df, aes(Enforcement, fill = Enforcement)) + 
-    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
-  scale_y_continuous(labels=percent, limits=c(0,1)) +
-  labs(y = "Percent Stopped") + ggtitle("Results of Stops") +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank()
-  )   + guides(fill = FALSE)
-
-p1race <- ggplot(df, aes(Enforcement, fill = Enforcement)) + 
-    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
-  scale_y_continuous(labels=percent, limits=c(0,1)) +
-  labs(y = "Percent Stopped") +
-  ggtitle("Results of Stops by Race") +
-  facet_grid( ~ Race) +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-
-```
-
-```{r}
-grid.arrange(p1,p1race,ncol=2)
-```
-
-```{r}
-
-p2 <- ggplot(df, aes(Reason, fill = Reason)) + 
-    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
-+  
-  scale_y_continuous(labels=percent, limits=c(0,1)) +
-  labs(y = "Percent") + ggtitle("Reason for Stops") +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank()
-  )   + guides(fill = FALSE)
-
-p2race <- ggplot(df, aes(Reason, fill = Reason)) + 
-    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
-  scale_y_continuous(labels=percent, limits=c(0,1)) +
-  labs(y = "Percent Stopped") +
-  ggtitle("Reason for Stops") +
-  facet_grid( ~ Race) +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-
-p3 <- ggplot(df, aes(CarSearch, fill = CarSearch)) + 
-    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
-  scale_y_continuous(labels=percent, limits=c(0,1)) +
-  labs(y = "Percent Stopped") +
-  ggtitle("Car Search by Race") +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank()
-  )   + guides(fill = FALSE)
-
-p3race <- ggplot(df, aes(Reason, fill = Reason)) + 
-    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
-  scale_y_continuous(labels=percent, limits=c(0,1)) +
-  labs(y = "Percent Stopped") +
-  ggtitle("Car Search by Race") +
-  facet_grid( ~ Race) +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-
-df$AgeRange = factor(df$AgeRange)
-levels(df$AgeRange) <- c("10-19","20-29","30-39","40-49")
-
-p4 <- ggplot(df, aes(AgeRange, fill = AgeRange )) + 
-    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
-  scale_y_continuous(labels=percent, limits=c(0,1)) +
-  labs(y = "Percent Stopped") +
-  ggtitle("Age Range of Car Stops by Race") +
-  labs(x = "Age Range") +
-  labs(fill = "Age Range")
-
-```
-
-```{r}
-
-```

### 2016-08-08 eda/PaperDrafts/Paper_draft_BPD.Rmd

---
title: "Berkeley Police Data"
author: "Shangjun Jiang"
date: "2 August 2016"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Shiny Presentation

This R Markdown presentation is made interactive using Shiny. The viewers of the presentation can change the assumptions underlying what's presented and see the results immediately.

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Interactive Plot

```{r eruptions}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),

  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")

  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## R Output

```{r cars}
summary(cars)
```

### 2016-08-08 eda/PaperDrafts/Some thoughts.Rmd

--- 2016-08-04 final_paper/Some thoughts.Rmd
@@
 # Read data
-
-############################## stop data ############################## 
-stop_by <- read_csv("cleaned_stop_data.csv")
-
-stop_by$Call.Date.Time <- mdy_hm(stop_by$Call.Date.Time)
-stop_by$AgeRange <- as.factor(stop_by$AgeRange)
-
-stop_by <- stop_by %>%
-  mutate(Hour = as.integer(hour(Call.Date.Time))) %>%
-  mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))
-
-stop_by$Race<- stop_by$Race %>% 
-  str_replace("A", "Asian") %>% 
-  str_replace("B", "Black") %>% 
-  str_replace("H", "Hispanic") %>% 
-  str_replace("O", "Other") %>% 
-  str_replace("W", "White")
-
-stop_by$Gender<- stop_by$Gender %>% 
-  str_replace("F", "Female") %>% 
-  str_replace("M", "Male")
-
-stop_by$Reason <- stop_by$Reason %>% 
-  str_replace("I", "Investigation") %>% 
-  str_replace("T", "Traffic") %>% 
-  str_replace("R", "Reasonable Suspicion") %>% 
-  str_replace("R", "Probation/ Parole") %>% 
-  str_replace("W", "Wanted")
-
-stop_by$Enforcement <- stop_by$Enforcement %>% 
-  str_replace("A", "Arrest") %>% 
-  str_replace("C", "Citation") %>% 
-  str_replace("O", "Other") %>% 
-  str_replace("W", "Warning")
-
-stop_by$CarSearch <- stop_by$CarSearch %>% 
-  str_replace("S", "Search") %>% 
-  str_replace("N", "No Search")
 
@@
 mergedf$Emergency.Psych.Eval <- factor(mergedf$Emergency.Psych.Eval)
+
+mergedf$Call.Date.Time <- mdy_hm(mergedf$Call.Date.Time)
+mergedf$AgeRange <- as.factor(mergedf$AgeRange)
+
+mergedf <- mergedf %>%
+  mutate(Hour = as.integer(hour(Call.Date.Time))) %>%
+  mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))
+
+mergedf$Gender<- mergedf$Gender %>% 
+  str_replace("F", "Female") %>% 
+  str_replace("M", "Male")
+
+mergedf$AgeRange<- mergedf$AgeRange %>% 
+  str_replace("1", "0-18") %>% 
+  str_replace("2", "18-29") %>% 
+  str_replace("3", "30-39") %>%  
+  str_replace("4", "40+")
 
@@
 
+```{r, fig.height = 5, fig.width = 10, echo = FALSE}
+
+m <- stop_by %>% 
+  group_by(Race) %>% 
+  tally()
+names(m) <- c("Race", "stop")
+names(race) <- c("Race", "census")
+
+pop <- left_join(race,m,by = "Race")
+pop <- pop %>% 
+  mutate(percentage=(stop/census)) %>% 
+pop$percentage<- round(pop$percentage, digits = 2)
+
+ggplot(pop)+
+  geom_bar(aes(x=Race,
+               y=percentage), 
+           stat = "identity",
+           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40b8d0"))+
+  geom_text(aes(x = Race, 
+                y = percentage, 
+                label = paste0(percentage,"%")), 
+            size=5,
+            colour= "#000000")+
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  labs(title = "The percentage of being stopped by police (%)",
+       x = "Race",
+       y = "Percentage") +
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) + 
+  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"), 
+        text=element_text(family="Tahoma"),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
+```
+
 ```{r}
-p1 <- ggplot(stop_by)+
-  geom_bar(aes(x=Race),
-           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))+
-  labs(title = "Stop data by race",
-       x = "Race",
-       y = "Counts")
-
-p2 <- ggplot(race)+
-  geom_bar(aes(x=Description, 
-               y=Counts),
-           stat = "identity", 
-           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))+
-  labs(title = "Demographic race data",
-       x = "Race",
-       y = "Counts")
-
-multiplot(p1, p2,cols=2)
+race %>%
+  ggvis(x = ~Description, y = ~Counts) %>%
+  layer_bars(fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF")) %>%
+  add_tooltip(function(race) race$Counts) 
+  labs(title = "Counts of people recorded in every day in a week")
 

### 2016-08-08 eda/PaperDrafts/Some_thoughts2.Rmd

---
title: "Some ideas"
author: "Shangjun Jiang"
date: "4 August 2016"
output: html_document
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sp)
library(ggmap)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(data.table)
library(rworldmap)
library(maps)
library(mapdata)
library(maptools)
library(rgeos)
library(scales)
library(RgoogleMaps)
library(tmap)
library(sp)
library(rgdal)
library(rgeos)
library(classInt)
library(gdata)
library(XLConnect)
library(RColorBrewer)

```

```{r, echo=FALSE, warning = FALSE, message = FALSE, include = FALSE}
# Read data

# data file names:

file_stops <- "../../clean_data/StopData_clean.csv" 
file_berkcensus2010 <- "../../clean_data/berk_census2010_clean.csv" 
file_berkcensus2010map <-"../../clean_data/berk_census2010_mapdata.rds" 
file_stopsummary <- "../../clean_data/StopData_summary.csv" 
file_tables <-  "../../clean_data/Tables_AgeSexRacePop.rds"
file_calls <-  "../../clean_data/CallsForService.rds"

# read in the data:

mergedf <- read_csv(file = file_stops)
berkcensus2010 <- read_csv(file = file_berkcensus2010)
berkcensus2010map <- readRDS(file_berkcensus2010map)
stopsummary <- read.csv( file_stopsummary )
stopsummary <- select(stopsummary, -X)
callservice <- read_csv(file = file_calls)
tables <- readRDS(file = file_tables)
pop1 <- as.data.frame(tables[4])
race <- as.data.frame(tables[3])
sex <- as.data.frame(tables[2])
age <- as.data.frame(tables[1])

## Location info:

latmax <- max(mergedf$lat, na.rm = TRUE) 
latmin <- min(mergedf$lat, na.rm = TRUE)
lonmax <- max(mergedf$long, na.rm = TRUE) 
lonmin <- min(mergedf$long, na.rm = TRUE)
latvals <- c(latmin, latmax)
lonvals <- c(lonmin, lonmax)

# get rid of the axes theme:

ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )

# get the plain Berkeley map from Google:

berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 12)

berkgg <-ggmap(berkMap) +
  xlim(-122.335, lonmax) + ylim(latvals) +
  ditch_the_axes

############################## multiplot function ############################## 
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```

## PART I: Racial Discrimination??

### Figure 1: Percentage of being stopped by different races
```{r, fig.height = 5, fig.width = 10, echo = FALSE, warning= FALSE, include = FALSE}

# create some figures, create some new data frames:

p1 <- ggplot(pop1)+
  geom_bar(aes(x=Race,
               y=percentage), 
           stat = "identity",
           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
  geom_text(aes(x=Race,
               y=percentage,
               label = paste0(percentage,"%")), 
            size=5,
            colour= "#000000")+
  scale_y_continuous(labels=percent, limits = c(0,1)) +
  labs(title = "The percentage of being stopped by police (%)",
       x = "Race",
       y = "Percentage") +
  theme(axis.line = element_line(size=1, colour = "black"), 
        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
        panel.border = element_blank(), panel.background = element_blank()) + 
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"), 
        text=element_text(family="Tahoma", size = 15),
        axis.text.x=element_text(colour="black", size = 10), 
        axis.text.y=element_text(colour="black", size = 10))

p2 <- ggplot(pop1)+
  geom_bar(aes(x=Race,
               y=totalp), 
           stat = "identity",
           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
  geom_text(aes(x=Race,
               y=totalp,
               label = paste0(totalp,"%")), 
            size=5,
            colour= "#000000")+
  scale_y_continuous(labels=percent, limits = c(0,1)) +
  labs(title = "The percentage of berkeley census data 2010 (%)",
       x = "Race",
       y = "Percentage") +
  theme(axis.line = element_line(size=1, colour = "black"), 
        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
        panel.border = element_blank(), panel.background = element_blank()) + 
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"), 
        text=element_text(family="Tahoma",size = 15),
        axis.text.x=element_text(colour="black", size = 10), 
        axis.text.y=element_text(colour="black", size = 10))

```

```{r, fig.height = 5, fig.width = 10, include = FALSE}
multiplot(p1, p2,cols=2)
```

Though there are more white people living in Berkeley, if we look at the race data from the percentage, we can clearly see huge differences between the white population and the black population. White people are composed of 55% of the Berkeley population but only 10% of them are being stopped. However, only 10% of the total population are black people but almost half of them are being stopped.

## a): places of being stopped 
We are interested in investigating 'where' they are stopped.

### Figure 2: All BPD Stops Density, 2015-2016
```{r, include = FALSE, warning=FALSE, message = FALSE}
g1 <- ggmap(berkMap) +
  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
    size = 2, bins = 5, data = mergedf, geom = "polygon") +
  scale_fill_gradient(low = "black", high = "red") +
    theme (panel.grid.major = element_blank (), # remove major grid
		       panel.grid.minor = element_blank ()  # remove minor grid
		       )+ 
  ggtitle("All BPD Stops Density, 2015-2016") +
  labs(alpha = element_blank())+
  guides(alpha = FALSE)
```

```{r, fig.width=10, fig.asp=0.8, echo=FALSE}
# a contour plot
g1

```

```{r, include=FALSE, warning=FALSE, message = FALSE }

### MAP: BPD All Stops w/ Car Searches by Race 2015-2016

g2 <- ggmap(berkMap) +
  geom_point(aes(x = long, y = lat, size = CarSearch, color = Race), data = mergedf, alpha = 0.2) + 
  scale_size_discrete( "Car Search", labels = c("No Search", "Search") ) +
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + 
  ggtitle("BPD All Stops w/ Car Searches by Race 2015-2016") +
  facet_wrap(~ Race, ncol = 3) 
```

```{r, fig.width=10, fig.asp=0.8, echo=FALSE}
g2
```

## b): reasons of being stopped 
Then we are interested in investigating **why** they are stopped.

#### arrest + jail data

## c): time of being stopped 
Then we are interested in investigating **when** they are stopped.

## d) outcome of being stopped
Then we are interested in investigating the result **after** they are stopped. Whether they are being car searched, arrested, or even put into jail.

## PART II: Age differences?

## PART IV: Exploration
#### pick location + time
#### time you top 3 stop reasons + top 2 race + top 2 age groups

### 2016-08-08 eda/PaperDrafts/draft.Rmd

--- 2016-08-08 final_paper/draft.Rmd
@@
 ---
-title: "BDP draft paper"
+title: "Berkeley Police Data"
 author: "Amy Zhu; Mengyu Li; Rebecca Reus; Shangjun Jiang"
-date: "2 August 2016"
+date: "August 7, 2016"
 output: html_document
@@
 
-```{r setup, include=FALSE}
+```{r setup, include=FALSE, warning = FALSE, message = FALSE}
 knitr::opts_chunk$set(echo = TRUE)
@@
 * [Berkeley Census Data](https://data.cityofberkeley.info/Demographics/Census-Data-2000-And-2010/bnkq-f2ge)
+*[Census 2010 Population and race data by county by census tract](http://www.bayareacensus.ca.gov/small/small.htm)
 
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
 # Read data
@@
 
-p1 <- ggplot(stop_by)+
+p1 <- ggplot(mergedf)+
   geom_bar(aes(x=Race),
-           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))+
+           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
   labs(title = "Stop data by race",
@@
 p2 <- ggplot(race)+
-  geom_bar(aes(x=Description, 
-               y=Counts),
+  geom_bar(aes(x=Race, 
+               y=census),
            stat = "identity", 
-           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))+
+           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
   labs(title = "Demographic race data",
@@
 
+c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF")
+
 ggplot(mtcars, aes(factor(cyl))) + 
@@
           scale_y_continuous(labels=percent)
+
+c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0")
 
@@
 ```
+
+```{r}
+
+```{r}
+race %>%
+  ggvis(x = ~Description, y = ~Counts) %>%
+  layer_bars(fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF")) %>%
+  add_tooltip(function(race) race$Counts) 
+  labs(title = "Counts of people recorded in every day in a week")
+
+```
+
+```

### 2016-08-08 eda/StopDataExploration/AnalysizeStop.Rmd

--- 2016-08-02 AnalysizeStop.Rmd
@@
 ```
+
+```{r}
+stop_reason <- ggplot(stop_by, aes(x = Reason, fill = Enforcement)) + 
+   geom_bar(aes(y = ..count../sum(..count..))) + 
+   labs(y = "Density") +
+   scale_fill_brewer(palette = "Set1") +
+   facet_grid(AgeRange~Gender)
+stop_reason
+```
+
+```{r}
+stop_enforcement <- ggplot(stop_by, aes(x = Reason, fill = Enforcement)) + 
+  geom_bar(aes(y = ..count../sum(..count..))) + 
+  labs(y = "Density") +
+  facet_wrap(~Race)
+stop_enforcement
+```

### 2016-08-09 clean_data/CleanDataScripts/CleanData_MakePercentageTables1.R
organizing, cleaning script code

@@ -1,5 +1,7 @@
 # CleanData_MakePercentageTables1.R
+
 # Author: Rebecca Reus
+
 # Purpose: 
 # 1. Clean the merged location stop data and create and save summary tables to display in paper.
 # 2. Clean the jail and arrest data and create and save tables to display in paper.
@@ -31,7 +33,7 @@ library(XLConnect)
 library(RColorBrewer)
 
 # data file names:
-file_stops <- "../../clean_data/StopData_clean.csv" 
+file_stops <- "../StopData_clean.csv" 
 file_stopsummary <- "../../clean_data/StopData_summary.csv" 
 mergedf <- read_csv(file = file_stops)
 census <- read.csv("../../raw_data/Census_Data_2000_And_2010.csv")
@@ -122,6 +124,7 @@ pop1$totalp <- round(pop1$totalp, digits =2)
 
 
 # stop summary table:
+
 ss <- stopsummary %>%
   select(Race, Percent.Stopped, Percent.Race.Stopped)
 ss$Percent.Stopped <- percent(ss$Percent.Stopped)
@@ -141,10 +144,10 @@ mergedf <- mergedf %>%
 
 # save the clean data:
 
-#saveRDS(callservice, "../../clean_data/CallsForService.rds")
+# saveRDS(callservice, "../CallsForService.rds")
 # listOfTables <- list(age, sex, race, pop1, ss)
-# saveRDS(listOfTables, "../../clean_data/Tables_AgeSexRacePopStop.rds")
-# saveRDS(mergedf, "../../clean_data/StopData_clean.rds")
-# write.csv(mergedf, "../../clean_data/StopData_clean.csv" )
+# saveRDS(listOfTables, "../Tables_AgeSexRacePopStop.rds")
+# saveRDS(mergedf, "../StopData_clean.rds")
+# write.csv(mergedf, "../clean_data/StopData_clean.csv" )
 
 

### 2016-08-09 Stop_Data_Analysis.Rmd

--- 2016-08-02 Stop_Data_Analysis.Rmd
@@
 
+story: 
+Comparing to any other race, Asian people contributes a lowest percentage of the incidents. This corresponds to the "model-minority" stereotype of the Asian poeple -
+
 ### Analysis of Day

### 2016-08-09 pdf1.Rmd

@@
 library(formattable)
-library(pander)
 library(extrafont)
@@
 map_stop1 <- berkgg_zoom2 +
-  geom_point(data = stops, aes(x=long,y=lat), alpha = .2) +
+  geom_point(data = stops, aes(x=long,y=lat), alpha = .2, color = "red") +
   ggtitle("Berkeley Police Stops, 2015-2016")+
@@
 map_callservice1 <- berkgg_zoom2 +
-  geom_point(data = na.omit(callservice), aes(x=long,y=lat), alpha = .2) +
+  geom_point(data = na.omit(callservice), aes(x=long,y=lat), alpha = .2, color = "red") +
   ggtitle("Calls for Service (not criminal reports) \nWithin 180 days (Feb-July 2016)")+

### 2016-08-09 eda/PaperDrafts/pdf1 (1).Rmd
adding abstract to finalpaper

--- 2016-08-08 pdf1.Rmd
@@
 ---
-title: "Draft Paper"
-author: "Shangjun Jiang"
-date: "8 August 2016"
+title: "STAT 133 Final Project: Visualizing Berkeley Police Data"
+author: "Amy Zhu, Mengyu Li, Rebecca Reus, and Shangjun Jiang"
+date: "9 August 2016"
 output: 
@@
 library(extrafont)
-font_import()
-loadfonts()
 setwd("~/Desktop/Stats R/PROJECT/finaldraftfiles")
@@
 
+arrest$Age <- as.integer(arrest$Age)
+arrest$Age[arrest$Age<18] <- "0-18"
+arrest$Age[arrest$Age>=18 & arrest$Age<30] <- "18-29"
+arrest$Age[arrest$Age>=30 & arrest$Age<40] <- "30-39"
+arrest$Age[arrest$Age>=40] <- "40+"
+arrest$Age <- as.factor(arrest$Age)
+
 ############################ jail data ###########################
 jail <- read.csv("raw_data/Berkeley_PD_Log_-_Jail_Bookings.csv")
+
+jail$Age <- as.integer(jail$Age)
+jail$Age[jail$Age<20] <- "0-19"
+jail$Age[jail$Age>=20 & jail$Age<30] <- "20-29"
+jail$Age[jail$Age>=30 & jail$Age<40] <- "30-39"
+jail$Age[jail$Age>=40] <- "40+"
+jail$Age <- as.factor(jail$Age)
 
@@
 age <- as.data.frame(tables[1]) # age population count
+
+dataframelist <- readRDS("clean_Data/StopData_tbls.rds")
+table1 <- data.frame(dataframelist[1])
+table2 <- data.frame(dataframelist[2])   #Enforcement
+table3 <- data.frame(dataframelist[3])   #Enforcement
 
@@
         panel.border = element_blank(), panel.background = element_blank()) + 
-  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
-        text=element_text(family="Times New Roman", size = 10),
-        axis.text.x=element_text(colour="black", size = 10), 
-        axis.text.y=element_text(colour="black", size = 10))
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))
 
@@
   scale_fill_gradientn(colors = p) +
-  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
-        text=element_text(family="Times New Roman", size = 10),
-        axis.text.x=element_text(colour="black", size = 10), 
-        axis.text.y=element_text(colour="black", size = 10))
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))
 
@@
 # bar graphs based on stop data:
-names(ss)[2:3] <- c("Percent Stopped", "Percent Population Stopped")
-
-bargraph1 <- ggplot(ss, 
+names(ss)[2:3] <- c("Percent_Stopped", "Percent_Population_Stopped")
+
+per <- ss %>% 
+  mutate(Percent_Stopped= str_replace(Percent_Stopped, "%", "")) %>% 
+  mutate(Percent_Population_Stopped=str_replace(Percent_Population_Stopped, "%", ""))
+
+per$Percent_Stopped <- as.numeric(per$Percent_Stopped)
+per$Percent_Stopped <- per$Percent_Stopped/100
+per$Percent_Population_Stopped <- as.numeric(per$Percent_Population_Stopped)
+per$Percent_Population_Stopped <- per$Percent_Population_Stopped/100
+
+bargraph1 <- ggplot(per, 
                     aes(x=Race, 
-                        y=`Percent Stopped`, 
+                        y=Percent_Stopped, 
                         fill = Race))+
@@
   geom_text(aes(x=Race,
-               y=`Percent Stopped`,
-               label = paste0(`Percent Stopped`)), 
+               y=Percent_Stopped,
+               label = paste0(Percent_Stopped*100, "%")), 
             size=5,
-            colour= "#000000")+
-  scale_y_discrete(labels=percent, limits = c(0,1))+
-  labs(title = "The percentage of berkeley being stopped (%)",
+            colour= "#000000") +
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  labs(title = "The percentage of berkeley \nbeing stopped (%)",
        x = "Race",
@@
         panel.border = element_blank(), panel.background = element_blank())+
-  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
-        text=element_text(family="Times New Roman", size = 10),
-        axis.text.x=element_text(colour="black", size = 10), 
-        axis.text.y=element_text(colour="black", size = 10))
-
-bargraph2 <- ggplot(ss, aes(Race, `Percent Population Stopped`, fill = Race))+
-  geom_bar(stat = "identity")+
-  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
-        text=element_text(family="Times New Roman", size = 10),
-        axis.text.x=element_text(colour="black", size = 10), 
-        axis.text.y=element_text(colour="black", size = 10))
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))
+
+bargraph2 <- ggplot(per, 
+                    aes(x=Race, 
+                        y=Percent_Population_Stopped, 
+                        fill = Race))+
+  geom_bar(stat = "identity",
+           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+  geom_text(aes(x=Race,
+               y=Percent_Population_Stopped,
+               label = paste0(Percent_Population_Stopped*100, "%")), 
+            size=5,
+            colour= "#000000") +
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  labs(title = "The percentage of population \nbeing stopped (%)",
+       x = "Race",
+       y = "Percent stopped")+
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank())+
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))
 
@@
   ggtitle("Berkeley Police Stop \nDensity Map, 2015-2016")+
-  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
-        text=element_text(family="Times New Roman", size = 10),
-        axis.text.x=element_text(colour="black", size = 10), 
-        axis.text.y=element_text(colour="black", size = 10))
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))
+
+############################ populatrion density map ###########################
+popdenmap <- berkgg_zoom1 + 
+  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Berkeley), color = "white", alpha = .7) +
+  ggtitle("2010 Population Density") +
+  labs(fill = "Percent") +
+  scale_fill_gradientn(colors = p) +
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))
 
@@
   theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
-        text=element_text(family="Times New Roman", size = 10),
+        text=element_text(family="Times New Roman", size = 14),
         axis.text.x=element_text(colour="black", size = 10), 
@@
              alpha = 0.7) + 
+  scale_color_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
   theme ( 
@@
 stop_reason <- stops[stops$Reason != "Other",]
+
+stop_by_reason_race <- stop_reason %>%
+  group_by(Race, Reason) %>%
+  tally()
+names(stop_by_reason_race) <- c("Race", "Reason", "totalCount")
+
+stop_by_arrest <- stop_reason[stop_reason$Enforcement == "Arrest",] %>% 
+  group_by(Race, Reason) %>%
+  tally()
+names(stop_by_arrest) <- c("Race", "Reason", "arrestCount")
+
+stop_by_percentage_race <- left_join(stop_by_reason_race, stop_by_arrest, by = c("Race", "Reason"))
+stop_by_percentage_race <- stop_by_percentage_race %>%
+  mutate(percentage = arrestCount / totalCount)
+stop_by_percentage_race$percentage <- round(stop_by_percentage_race$percentage, digits = 2)
+
+stop_arrest <- ggplot(stop_by_percentage_race, aes(x = Race, 
+                                                   fill = Race)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+  facet_wrap(~Reason)+
+  geom_text(aes(x=Race,
+               y=percentage,
+               label = paste0(percentage*100, "%")), 
+            size=4,
+            colour= "#000000") +
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  labs(title = "The percentage of population \nbeing stopped (%)",
+       x = "Race",
+       y = "Density")+
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank())+
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))+
+  guides(fill = FALSE)
 
@@
         panel.border = element_blank(), panel.background = element_blank()) + 
-  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
-        text=element_text(family="Times New Roman", size = 14),
-        axis.text.x=element_text(colour="black", size = 10), 
-        axis.text.y=element_text(colour="black", size = 10))
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))
 
@@
   facet_grid(~Race)+
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
  labs(x= "Age",
@@
        title = "Jail Data") + 
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
   guides(fill = FALSE)+
@@
 # histogram of stop data by race and hour:
-histogram_stoptime1 <- ggplot(stops, aes(x = Hour, fill = Race)) + 
+histogram_stoptime1 <- ggplot(stops, 
+                              aes(x = Hour, 
+                                  fill = Race)) + 
   geom_bar(aes(y = ..count..)) +
-  labs(y = "Count") +
   facet_grid(~Race)+
-  labs(title = "Number of people each race in every hour") +
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+  labs(title = "Number of people each race in every hour",
+       y = "Count") +
   guides(fill = FALSE)+
@@
 ```
+
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
+
+test <-read_csv("clean_data/enforcement_race.csv")
+
+map_enforcement <- ggplot(test)+ 
+  geom_bar(aes(x=Race,
+               y=percentage,
+               fill = Race),
+           stat = "identity") +
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+  facet_grid(~Enforcement)+
+  geom_text(aes(x=Race,
+               y=percentage,
+               label = paste0(percentage*100,"%")), 
+            size=4,
+            colour= "#000000")+
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  labs(title = "Percent Stopped by Enforcement Type and Race",
+       x = "Race",
+       y = "Percentage") +
+  scale_x_discrete(labels = function(x) str_wrap(x, width = 5))+
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) + 
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 13),
+        axis.text.x=element_text(colour="black", size = 8), 
+        axis.text.y=element_text(colour="black", size = 8))+
+  guides(fill = FALSE)
+
+```
+
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
+
+race_hour <- read.csv("clean_data/race_hour.csv")
+stop_percentage3 <- ggplot(race_hour, aes(x = Hour, fill = Race)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  labs(title = "Probability of BPD stop for \na specific race in a specific hour",
+       x= "Hour",
+       y = "Density") +
+  facet_wrap(~Race) +
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0")) +
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) +
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7)) +
+  guides(fill = FALSE)
+
+```
+
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
+
+age_hour <- read.csv("clean_data/age_hour.csv")
+
+stop_percentage2 <- ggplot(age_hour , aes(x = Hour, fill = AgeRange)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  facet_wrap(~AgeRange)+
+labs(title = "Probability of BPD stop for a \nspecific age range in a specific hour",
+       x= "Hour",
+       y = "Density") +
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0")) +
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) +
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7)) +
+  guides(fill = FALSE)
+
+```
+
+# Abstract  
+    _In this paper, we analyzed the stop data_
 
@@
 
-## Berkeley Stop Data Cleaning Process
+## Cleaning Data
+
+-  __Berkeley Stop Data Cleaning Process__
 
@@
 
-## Berkeley Arrest and Jail Bookings Cleaning Process
+![](Stop_data_header.png)
+
+- __Berkeley Arrest and Jail Bookings Cleaning Process__
 
@@
 
-## Berkeley Census 2010 Data Cleaning Process
+![](arrest_data_header.png)
+
+![](jail_data_header.png)
+
+- __Berkeley Census 2010 Data Cleaning Process__
 
@@
 
+![](census_data_header.png)
+
 ## Berkeley Police Stop Data Analysis
@@
 ```{r, warning = FALSE, echo = F, fig.width = 10, fig.height = 5, message = FALSE, fig.align="center"}
-#multiplot(map_stop1,map_callservice1, cols =2)
-#map_stop1
-#map_callservice1
 grid.arrange(map_stop1,map_callservice1, ncol = 2)
@@
 
-```{r, warning = FALSE, echo = F, fig.width = 5, fig.height = 5, message = FALSE, fig.align="center"}
-stopdensitymap4
+```{r, warning = FALSE, echo = F, fig.width = 10, fig.height = 5, message = FALSE, fig.align="center"}
+grid.arrange(stopdensitymap4, popdenmap, ncol=2)
 ```
@@
 
-```{r, warning = FALSE, echo = F, fig.width = 5, fig.height = 5, message = FALSE, fig.align = "center"}
-popdenmap # print population density map
-```
-
 In order to break down the stop data further, we mapped the data by each of the disposition variables.
@@
 pander(ss) # table for stop data percentages by race relative to population
-#grid.arrange(bargraph1, bargraph2, ncol = 2)
 ```
@@
 A visualization of the groups of people stopped in proportion to their population is provided by the bar chart below.
+```{r, message = F, echo = FALSE, fig.width = 10, fig.height = 5, warning = FALSE, fig.align="center"}
+grid.arrange(bargraph1, bargraph2, ncol = 2)
+```
 
@@
 
-#!!!!!!!!!!!!!!!!!!!!!INCLUDE BAR GRAPH HERE of % race stopped vs. percent race population, as described above!!!!!!!!!!!!!!!!!!!!!!!!
-multiplot(p1, p2,cols=2)
+grid.arrange(p1, p2,ncol=2)
 ```
@@
 
-```{r, echo=FALSE, warning=F,message=F, fig.align="center", fig.width = 12, fig.height=3}
+```{r, echo=FALSE, warning=F,message=F, fig.align="center", fig.width = 20, fig.height=15}
 map_arrested1 
@@
 ```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.align = "center", fig.width = 8, fig.height = 3}
-#map_enforcement
+map_enforcement
 ```
@@
 
-```{r, fig.height = 5, fig.width = 5, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
 map_reasons1
@@
 
-```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.align="center"}
+```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+stop_arrest
+```
+### Jenny Findings 
+1. The probability of arrested in a stop with the reason Traffic is much lower than other reasons. The average conditional probability of arrested given reason is traffic is 1.58%.
+
+2. The probability of arrested in a stop with the reason Wanted is much higher than other reason. The average conditional probability of arrested given reason is wanted is 33.33%.
+
+3.
+An interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and race Asian is 75%, which is much higher than that of any other race.
+Asian people is much more liable to be arrested by Berkeley Police Department during Probation or Parole if stopped by the BPD.
+
+```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+stop_percentage3
+```
+### Jenny Findings 
+1. Black people are obviously liable to commit incidents at night. The average ratio of black people to all people stoped by the Berkeley Police Department at night is about 40%.
+
+2. White people are obviously liable to commit incidents in the daytime. The average ratio of white people to all people stoped by the Berkeley Police Deparment during the day is about 45%.
+
+3. Ratio of incidents commited by Asian people, Hispanic people and other people fluctuate during the daytime and the night, with an average ratio of 8%, 11% and 9% respectively.
+
+```{r}
+stop_percentage2 
+```
+### Jenny Findings 
+1.
+People aged from 18 to 29 are obviously liable to commit incidents at night.
+The average ratio of people aged from 18 to 29 to all people stoped by the Berkeley Police Department at night is greater than 40%.
+
+2.
+People aged greater than 40 are obviously liable to commit incidents in the daytime.
+The average ratio of people aged greater than 40 to all people stoped by the Berkeley Police Deparment during the day is greater than 40%.
+
+3. Ratio of incidents commited by people aged between 0 and 18 and people aged from 30 to 39 fluctuates during the daytime and the night, with an average ratio of 2.5% and 25% respectively.
+
+```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 5, fig.width = 5, fig.align="center"}
 #### INCLUDE BAR GRAPH & TABLE OF POLICE STOPS BY REASON HERE(Use percentage, not count)!!!!!!!!!!!!!!!!!!!!!!!!
-multiplot(bararrest, barjail, cols = 2)
+grid.arrange(bararrest, barjail, ncol = 2)
 ```

### 2016-08-09 final_paper/finalpaper.Rmd
adding abstract to finalpaper

--- 2016-08-09 eda/PaperDrafts/pdf1 (1).Rmd
@@
 
-# Abstract  
-    _In this paper, we analyzed the stop data_
+## Abstract
+
+For the STAT 133 final project, our group was interested in finding some ways of visualizing the relationship between the Berkeley Police Department and different groups of people, especially within the context of different populations and neighborhoods within Berkeley.
+To obtain our results, we accessed, cleaned and explored the four Berkeley Police Department data sets available to the public on the Berkeley Open Data website using R.
+The police data included police stops, calls for service, arrests, and jail bookings made by the the police department over the past 3-6 month period.
+Further, we accessed 2010 census data in order to obtain demographic information about Berkeley as a whole and about neighborhoods inside of Berkeley.
+The three census data sets that we used are also available through Berkeley's open data site, including the Berkeley 2000-2010 census data, which gave us the total population counts and percentages of each group of people in Berkeley for the 2010 census, the Alameda County 2010 census tract population data, which gave us the demographic information for 33 different neighborhoods (or "tracts") inside Berkeley, and the Berkeley 2010 census tract polygons shapefile (.shp), which provided the information we needed to visually display the 33 neighborhoods.
+Rather than limiting the project to one set of data, incorporating multiple data sets gave us a more holistic understanding of subject and expanded our ability to understand and visualize the data.
+Our findings suggest the hypothesis that (1) black people in Berkeley are more likely than others to be stopped by Berkeley police, and that (2) black people in Berkeley are arrested and jailed at disproportionate rates.
+We recommend future statistical analysis be done to test our hypothesis.
 
@@
 
-For the STAT 133 final project, our group cleaned and explored the four police data sets available on the Berkeley Open Data website.
-We examined police stops, calls for service, arrests, and jail bookings made by the the police department over the past 3-6 month period.
-Rather than limiting the project to one set of data, these four sets were chosen in order to gain a more holistic and comprehensive understanding of the data.
-The additional resources of information each data set provided expanded our ability to understand and visualize the data.
-
-One of our main goals was to find some ways of visualizing the relationship between the Berkeley Police Department and different groups of people.
-In order to compare the proportion of each group represented in the Berkeley police data relative to the population of those groups in Berkeley, we also considered three census data sets that are also available through Berkeley's open data site: first, we included the Berkeley 2000-2010 census data, which gave us the total population counts and percentages of each group of people in Berkeley for the 2010 census; next, we considered the Alameda County 2010 census tract population data, which allowed us to obtain the demographic information for 33 different neighborhoods (or "tracts") inside Berkeley; and finally, the Berkeley 2010 census tract polygons shapefile (.shp), which provided the information we needed to visually display the 33 neighborhoods.
-
-After detailing our process of cleaning the police and census data, we will provided some visualizations of the data and some preliminary findings about the police data.
-Afterwards, we will discuss problems with the data and some ideas for future statistical analysis.
+After detailing our process of cleaning the police and census data, we will provide our results: visualizations of the data and some preliminary findings about the police data.
+Afterwards, we will discuss problems with the data and recommendation for future work on this subject.
 
@@
 
--  __Berkeley Stop Data Cleaning Process__
+###__Berkeley Stop Data Cleaning Process__
 
@@
 The disposition column was a character string typically including 6 characters per individual as follows: the first character was race, the second gender, the third age range, the fourth reason (for the stop), the fifth enforcement (the result of the stop), and the sixth car search (yes or no).
-Each individual that was involved with the stop was recorded using comma separation in the same column followed by an additional 6 characters.
-Additional optional dispositions were sometimes included, again using comma-separation from the 6-character strings (though not necessarily in any order), ranging from one to three characters.
+Each individual that was involved with the stop was recorded using comma separation in the same row and column followed by an additional 6 characters.
+Additional optional dispositions were sometimes included, again using comma-separation from the 6-character strings (though not necessarily in an order), ranging from one to three characters.
 These additional disposition abbreviations represented the following options: Primary Case Report, MDT Narrative Only, Arrest Reporty Only (No Case Report Submitted), Incident Report, Field Card, Collision Investigation Report, Emergency Psychiatric Evaluation, Impounded Vehicle, and Officer Made a Stop of More Than 5 Persons.
@@
 
-Turning the stop data's character locations into location coordinates using a Google Maps geolocation library proved difficult because of common data entry mistakes.
+We used the stop data's character string locations into location coordinates using a Google Maps geolocation service provided by the library `ggmap` with the function `geocode`.
+This task proved difficult because of common data entry mistakes.
 For example, there were common mispellings of street names and the use of several different abbreviations to mean the same thing.
-Additionally, we found that the geolocation provided by Google does not accept forward slash characters the data used in the colum to indicate street intersections, so these forward slash characters were changed to `and` for the location processing.
-Google also does not accept the term "block" to indicate a range of addresses, and so the word "block" was removed from the data set for location processing.
+Additionally, we found that the geolocation provided by Google does not accept forward slash characters the data used in the column to indicate street intersections, so these forward slash characters were changed to `and` for the location processing, which was successful.
+Google also does not accept the term "block" to indicate a range of addresses, and so the word "block" was removed from the location column for location processing.
 Most of these types of errors were delt with on an automated basis before geolocation, using `str_detect` and `str_replace` from the `stringr` package.
@@
 
-- __Berkeley Arrest and Jail Bookings Cleaning Process__
+###__Berkeley Arrest and Jail Bookings Cleaning Process__
 
@@
 
-- __Berkeley Census 2010 Data Cleaning Process__
+###__Berkeley Census 2010 Data Cleaning Process__
 
@@
 
-## Berkeley Police Stop Data Analysis
+### Berkeley Police Stop Data Analysis
 
@@
 ```
+
 ### Jenny Findings 

### 2016-08-09 eda/PaperDrafts/Stop_Data_Analysis.Rmd

--- 2016-08-09 Stop_Data_Analysis.Rmd
@@
 
-```{r, echo=FALSE, eval=FALSE}
+```{r, echo=FALSE}
 stop_race_total_table <- stop_by %>%
@@
 
-```{r, eval=FALSE, echo=FALSE}
+```{r, echo=FALSE}
 stop_day_total_table <- stop_by %>%
@@
 ```{r, echo=FALSE}
-stop_time <- stop_by %>%
-  ggvis(x = ~Hour, fill := "lightpink") %>%
-  layer_histograms(width = input_slider(1, 12, value = 1, step = 2, label = "width"))
-stop_time
+stop_race <- ggplot(stop_by, aes(x = Hour, fill = CarSearch)) + 
+  geom_bar(aes(y = ..count../sum(..count..))) + 
+  scale_fill_brewer(palette = "Pastel2") +
+  labs(y = "Density")
+stop_race
 ```
 Note: you can change the value of the input slider to see the number of incidents happened in specific time span of the day.
-```{r, eval = FALSE, echo=FALSE}
+```{r, echo=FALSE}
 stop_hour <- stop_by %>%
@@
 #### Picture: Probability of arrested by BPD of a specific race for a specific reason
-```{r,echo=FALSE}
+```{r,echo=FALSE, fig.asp=0.5, fig.height=3, fig.width=15}
 stop_by_reason_race <- stop_by %>%
@@
   labs(y = "Density") +
-  facet_wrap(~Reason)
+  facet_grid(~Reason)
 stop_arrest
@@
   labs(y = "Density") +
-  facet_wrap(~Reason)
+  facet_grid(~Reason)
 stop_arrest_age
@@
   labs(y = "Density") +
-  facet_wrap(~Reason)
+  facet_grid(~Reason)
 stop_arrest_gender

### 2016-08-09 final_paper/finalpaper.Rmd
updating final paper

@@
 library(extrafont)
-setwd("~/Desktop/Stats R/PROJECT/finaldraftfiles")
 ```
@@
 ############################ Police Stop Data ###########################
-stops <- read_csv("clean_data/StopData_clean.csv")
+stops <- read_csv("../clean_data/StopData_clean.csv")
 
@@
 ########################### arrest data ###########################
-arrest <- read_csv("raw_data/Berkeley_PD_Log_-_Arrests.csv")
+arrest <- read_csv("../raw_data/Berkeley_PD_Log_-_Arrests.csv")
 
@@
 ############################ jail data ###########################
-jail <- read.csv("raw_data/Berkeley_PD_Log_-_Jail_Bookings.csv")
+jail <- read.csv("../raw_data/Berkeley_PD_Log_-_Jail_Bookings.csv")
 
@@
 ############################ Census 2010 Data by Polygonal Tract Number ###########################
-berkcensus2010 <- read_csv("clean_data/berk_census2010_clean.csv")
-
-census <- read.csv("raw_data/Census_Data_2000_And_2010.csv")
+berkcensus2010 <- read_csv("../clean_data/census2010tractpop_clean.csv")
+
+census <- read.csv("../raw_data/Census_Data_2000_And_2010.csv")
 census2010 <- census[census$Year==2010,]
@@
 ################## Census 2010 Data by Polygonal Tract Number, ready for mapping #################
-berkcensus2010map <- readRDS("clean_data/berk_census2010_mapdata.rds")
+berkcensus2010map <- readRDS("../clean_data/berk_census2010_mapdata.rds")
 
 ############################ Police Calls for service ############################ 
-callservice <- readRDS("clean_data/CallsForService.rds")
+callservice <- readRDS("../clean_data/CallsForService.rds")
 
@@
 #############################  Tables to use in paper ############################ 
-tables <- readRDS("clean_data/Tables_AgeSexRacePopStop.rds")
+tables <- readRDS("../clean_data/Tables_AgeSexRacePopStop.rds")
 ss <- as.data.frame(tables[5]) # stop data by race proportional to population
@@
 
-dataframelist <- readRDS("clean_Data/StopData_tbls.rds")
+dataframelist <- readRDS("../clean_Data/StopData_tbls.rds")
 table1 <- data.frame(dataframelist[1])
-table2 <- data.frame(dataframelist[2])   #Enforcement
-table3 <- data.frame(dataframelist[3])   #Enforcement
+table2 <- data.frame(dataframelist[2])  
+table3 <- data.frame(dataframelist[3])
 
@@
   scale_y_continuous(labels=percent, limits = c(0,1)) +
-  labs(title = "The percentage of berkeley \nbeing stopped (%)",
+  labs(title = "Percent Stopped",
        x = "Race",
@@
   scale_y_continuous(labels=percent, limits = c(0,1)) +
-  labs(title = "The percentage of population \nbeing stopped (%)",
+  labs(title = "Percent Population Stopped",
        x = "Race",
@@
 
-test <-read_csv("clean_data/enforcement_race.csv")
-
-map_enforcement <- ggplot(test)+ 
+table3$Percent.Stopped <- percent(table3$Percent.Stopped)
+
+map_enforcement <- ggplot(table3)+ 
   geom_bar(aes(x=Race,
-               y=percentage,
+               y=Percent.Stopped,
                fill = Race),
@@
   scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
-  facet_grid(~Enforcement)+
+  scale_y_continuous(limits = c(0,1), labels = c("0%", "25%", "50%", "75%", "100%")) +
+  labs(title = "Arrested Individuals, by Race",
+       x = "Race",
+       y = "Percent of Arrested") +
     geom_text(aes(x=Race,
-               y=percentage,
-               label = paste0(percentage*100,"%")), 
-            size=4,
+               y=Percent.Stopped,
+               label = paste0(Percent.Stopped)), 
+            size=5,
             colour= "#000000")+
-  scale_y_continuous(labels=percent, limits = c(0,1)) +
-  labs(title = "Percent Stopped by Enforcement Type and Race",
-       x = "Race",
-       y = "Percentage") +
   scale_x_discrete(labels = function(x) str_wrap(x, width = 5))+
@@
   guides(fill = FALSE)
-
-```
-
-```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
-
-race_hour <- read.csv("clean_data/race_hour.csv")
+#map_enforcement
+
+```
+
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE, eval = F}
+
+race_hour <- read.csv("../clean_data/race_hour.csv")
 stop_percentage3 <- ggplot(race_hour, aes(x = Hour, fill = Race)) + 
@@
 
-```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
-
-age_hour <- read.csv("clean_data/age_hour.csv")
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE, eval = F}
+
+age_hour <- read.csv("../clean_data/age_hour.csv")
 
@@
 The arrest data had 205 observations and the jail bookings data had 223.
-These observations were obtained by the Berkeley Police department over a 3-month period, rather than a 6-month period.
+These observations were obtained by the Berkeley Police department over a 3-month period, rather than a 6-month period (as was for the stop data).
 They both contained similar variables, including a case/arrest/booking number, date and time, type, and subject information (name, race, sex, D.O.B., age, height, weight, hair, eyes, and occupation) and statute information (type, description, agency, and disposition).
-Cleaning required the dates and times to be put into `lubridate` format, and we also created an hour column to consider the number of occurences during each hour of the day.
+Cleaning this data set also required the dates and times to be put into `lubridate` format, and we also created an hour column to consider the number of occurences during each hour of the day.
 
@@
 Finally, to get the map information of these 33 tracts, we also needed the census 2010 Berkeley tract polygons shapefile.
-After converting the shapefile to a dataframe and changing the coordinate system to the one used by Google, we were able to `left_join` this new polygonal mapping data frame to the data frame containing the population information for the Berkeley tracts by the unique census 2010 tract number.
+After converting the shapefile to a dataframe and changing the coordinate system to the one used by Google (this was done using the `readOGR` function from the library `rgdal`), we were able to `left_join` this new polygonal mapping data frame to the data frame containing the population information for the Berkeley tracts by the unique census 2010 tract number.
 Combining the tract map information with the tract population information gave us a data frame containing population information for 33 different areas that we could represent spatially.
-The goal was to visualize the population density of different groups of people within the city by neighborhood, which provided us with a context for the stop data.
+The goal was to visualize the population density of different groups of people within the city by neighborhood, which provided us with a context for the sptatially visualized stop data.
 
@@
 
-```{r, warning = FALSE, echo = F, fig.width = 10, fig.height = 5, message = FALSE, fig.align="center"}
+```{r, warning = FALSE, echo = F, fig.width = 10, fig.height = 5, message = FALSE, fig.align="center", eval = F}
 grid.arrange(map_stop1,map_callservice1, ncol = 2)
@@
 
-A more interesting visualization of this data is the following two-dimensional density map of the police stop data, displayed below.
+Another visualization of the stop data mapped above is the following two-dimensional density map of the police stop data, displayed below.
 One of the most interesting aspects of the stop data in this image is that most police stops seem to be clustered west of UC Berkeley Campus in the Downtown area.
-This was contrary to what some of us might have thought before, which was that the area with the higher police stops would be corrolated to the area with the higher population.
+This was contrary to what some of us might have thought before, which was that the area with the higher police stops would be the area with the higher population.
 As you will see, however, this is not necessarily true.
@@
 Note that the population density visualized by the following map is conveyed by the range between white (0%) and red (100%).
-As you can see, White people make up a striking majority in all census 2010 tract areas, which seems to imply that the population of Black people in Berkeley is being disproportionally affected by police stops.
+As you can see, White people make up a striking majority in all census 2010 tract areas, which seems to suggest the hypothesis that the population of Black people in Berkeley is being disproportionally affected by police stops.
 
@@
 grid.arrange(bargraph1, bargraph2, ncol = 2)
-```
-
-!!!!!!!!!!!!!!!!!!!!!INCLUDE BAR GRAPH HERE of % race stopped vs. percent race population, as described above!!!!!!!!!!!!!!!!!!!!!!!!
-
-```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.align = "center"}
-
-grid.arrange(p1, p2,ncol=2)
 ```

### 2016-08-09 clean_data/CleanDataScripts/CleanData_MakePercentageTables2.R

@@ -58,6 +58,22 @@ tbl_summary <- tbl_summary %>%
   select(-Stopped, -Race.Population)
 tbl_summary
 
+# enforcement = arrest by race
+tbl_arrest <- mergedf %>%  
+  subset(Enforcement == "Arrest") %>%
+  group_by(Race) %>%
+  summarise(Stopped=n() ) %>%
+  mutate(Percent.Stopped = Stopped/nrow(subset(mergedf, Enforcement == "Arrest"))) 
+tbl_arrest
+temp <- c(sum(bpop2$Asian), sum(bpop2$Black), sum(bpop2$Hispanic), sum(bpop2$Other), sum(bpop2$White) )
+tbl_arrest$Race.Population = temp
+tbl_arrest$Percent.Race.Stopped = tbl_arrest$Stopped/tbl_arrest$Race.Population
+tbl_arrest
+
+tbl_arrest <- tbl_arrest %>%
+  select(-Stopped, -Race.Population)
+tbl_arrest
+
 # percent reason:
 
 tbl_reason <- mergedf %>%  
@@ -113,15 +129,15 @@ tbl_enforcement$Percent.White <- tbl_enforcement_W$Percent.White
 tbl_enforcement
 rm(tbl_enforcement_W)
 
-# # save the resulting 3 data frames:
-# saveRDS(list(tbl_summary, tbl_reason, tbl_enforcement), file = file_saveAs)
-# 
-# # how to read in the data: 
-# 
-# dataframelist <- readRDS(file = file_saveAs)
-# 
-# # how to exract the data frames from the list:
-# table1 <- data.frame(dataframelist[1])
-# table2 <- data.frame(dataframelist[2])
-# table3 <- data.frame(dataframelist[3])
+# save the resulting 3 data frames:
+saveRDS(list(tbl_summary, tbl_reason, tbl_arrest), file = file_saveAs)
+
+# how to read in the data:
+
+dataframelist <- readRDS(file = file_saveAs)
+
+# how to exract the data frames from the list:
+table1 <- data.frame(dataframelist[1])
+table2 <- data.frame(dataframelist[2])
+table3 <- data.frame(dataframelist[3])
 

### 2016-08-09 final_paper/finalpaper.Rmd
updatesJenny

@@
 library(extrafont)
+setwd("~/Desktop/Stats R/PROJECT/FINALPAPER/finalproject_angry_ladies/final_paper")
 ```
@@
   scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
-  facet_wrap(~Reason)+
+  facet_wrap(~Reason, nrow = 1)+
   geom_text(aes(x=Race,
@@
   scale_y_continuous(labels=percent, limits = c(0,1)) +
-  labs(title = "The percentage of population \nbeing stopped (%)",
+  labs(title = "Probability of arrested by BPD of \na specific race for a specific reason (%)",
        x = "Race",
@@
 
-```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE, eval = F}
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
 
 race_hour <- read.csv("../clean_data/race_hour.csv")
+
 stop_percentage3 <- ggplot(race_hour, aes(x = Hour, fill = Race)) + 
@@
        y = "Density") +
-  facet_wrap(~Race) +
+  facet_wrap(~Race, nrow = 1) +
   scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0")) +
@@
 
-```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE, eval = F}
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
 
@@
 
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
+# Probability of arrested by BPD of a specific gender for a specific reason
+stop_by_reason_gender <-stops %>% 
+  group_by(Gender, Reason) %>% 
+  summarise(totalCount=n())
+stop_by_arrest_gender <- stops[stops$Enforcement == "Arrest",] %>%
+  group_by(Gender, Reason) %>%
+  summarise(arrestCount = n())
+stop_by_percentage_gender <- left_join(stop_by_reason_gender, stop_by_arrest_gender, by = c("Gender", "Reason"))
+stop_by_percentage_gender <- stop_by_percentage_gender %>%
+  mutate(percentage = arrestCount / totalCount)
+stop_by_percentage_gender$percentage <- round(stop_by_percentage_gender$percentage , digits=2)
+stop_by_percentage_gender<- stop_by_percentage_gender[stop_by_percentage_gender$Reason != "Other",]
+
+stop_arrest_gender <- ggplot(stop_by_percentage_gender , aes(x = Gender, fill = Gender)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  facet_wrap(~Reason, nrow=1)+
+labs(title = "Probability of BPD stop for a \nspecific gender range in a specific reason",
+       x= "Hour",
+       y = "Density") +
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0")) +
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  geom_text(aes(x=Gender,
+               y=percentage,
+               label = paste0(percentage*100, "%")), 
+            size=3,
+            colour= "#000000")+
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) +
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7)) +
+  guides(fill = FALSE)
+
+```
+
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
+stop_by_reason_age <- stops %>%
+  group_by(AgeRange, Reason) %>%
+  summarise(totalCount = n())
+stop_by_arrest_age <- stops[stops$Enforcement == "Arrest",] %>%
+  group_by(AgeRange, Reason) %>%
+  summarise(arrestCount = n())
+stop_by_percentage_age <- left_join(stop_by_reason_age, stop_by_arrest_age, by = c("AgeRange", "Reason"))
+stop_by_percentage_age <- stop_by_percentage_age %>%
+  mutate(percentage = arrestCount / totalCount)
+stop_by_percentage_age$percentage<-round(stop_by_percentage_age$percentage,digits = 2)
+stop_by_percentage_age <- stop_by_percentage_age[stop_by_percentage_age$Reason != "Other",]
+
+stop_arrest_age <- ggplot(stop_by_percentage_age, aes(x = AgeRange, fill = AgeRange)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  facet_wrap(~Reason)+
+  labs(title = "Probability of BPD stop for a \nspecific age range in a specific reason",
+       x= "Hour",
+       y = "Density") +
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0")) +
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  geom_text(aes(x=AgeRange,
+               y=percentage,
+               label = paste0(percentage*100, "%")), 
+            size=3,
+            colour= "#000000")+
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) +
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7)) +
+  guides(fill = FALSE)
+
+```
+
 ## Abstract
@@
 
-```{r, warning = FALSE, echo = F, fig.width = 10, fig.height = 5, message = FALSE, fig.align="center", eval = F}
+```{r, warning = FALSE, echo = F, fig.width = 10, fig.height = 5, message = FALSE, fig.align="center"}
 grid.arrange(map_stop1,map_callservice1, ncol = 2)
@@
 
-```{r}
+```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
 stop_percentage2 
@@
 
-```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 5, fig.width = 5, fig.align="center"}
+```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+stop_arrest_age
+```
+### Jenny Findings 
+1. Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.
+
+2.
+Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and age range 0-18 is 67.56%, which is much higher than that of any other age range.
+
+3.
+The conditional probability of arrested given person involved aged from 0 to 18 in a stop given any reason is more than that of given person with any other age range.
+Therefore, in a stop, teenagers is more liable to be arrested by Berkeley Police Department.
+
+```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+stop_arrest_gender 
+```
+### Jenny Findings 
+1. Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.
+
+2.
+Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole of female is 50.00%, which is much higher than that of male, which is 20.11%.
+
+```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 5, fig.width = 10, fig.align="center"}
 #### INCLUDE BAR GRAPH & TABLE OF POLICE STOPS BY REASON HERE(Use percentage, not count)!!!!!!!!!!!!!!!!!!!!!!!!

### 2016-08-09 final_paper/finalpaper.Rmd
updates

@@
   scale_y_continuous(limits = c(0,1), labels = c("0%", "25%", "50%", "75%", "100%")) +
-  labs(title = "Arrested Individuals, by Race",
+  labs(title = "Police Stops Leading to Arrests",
        x = "Race",
@@
 
+## PLEASE COMPARE THIS TO THE CALLS FOR SERVICE DATA  
+### LOVE JENNY
+
 Another visualization of the stop data mapped above is the following two-dimensional density map of the police stop data, displayed below.
@@
 
-```{r, message = F, echo = FALSE, warning = FALSE, fig.align="center"}
-panderOptions("digits", 3)
-pander(ss) # table for stop data percentages by race relative to population
-```
+## PLEASE CHANGE THE TABLE 
+#### LOVE JENNY
 
@@
 
-### Jenny Findings 
+###  Findings 
 1. The probability of arrested in a stop with the reason Traffic is much lower than other reasons. The average conditional probability of arrested given reason is traffic is 1.58%.
@@
 ```
-### Jenny Findings 
+###  Findings 
 1. Black people are obviously liable to commit incidents at night. The average ratio of black people to all people stoped by the Berkeley Police Department at night is about 40%.
@@
 ```
-### Jenny Findings 
+### Findings 
 1.
@@
 Even though most of the population in Berkeley is white, most of the people recently arrested are black.
-
-Another idea we had was to visualize the arrest data for each community by hour. 
-!!!!!!!!!!!PLEASE FIX LABEL SO THAT IT IS PERCENT ARRESTED, NOT COUNT!!!! ALSO, PLEASE DISPLAY A SMALL TABLE OF THIS DATA FOLLOWING THE HISTOGRAM.!!!!!!!!!!!!!
-
-```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.width = 12, fig.height=3, fig.align = "center"}
-histogram_stoptime1
-#!!!!!!!!!!!PLEASE FIX LABEL SO THAT IT IS PERCENT ARRESTED, NOT COUNT!!!! ALSO, PLEASE DISPLAY A SMALL TABLE OF THIS DATA FOLLOWING THE HISTOGRAM.!!!!!!!!!!!!!
-```
-
-!!!!!!!!!!!!!! 1-2 sentence: ANALYZE FINDINGS FROM chart & TABLE ABOVE HERE!!!!!!!!!!
 
@@
 * [Berkeley Census Data](https://data.cityofberkeley.info/Demographics/Census-Data-2000-And-2010/bnkq-f2ge)
-*[Census 2010 Population and race data by county tract polygons](http://www.bayareacensus.ca.gov/small/small.htm)
-*[Berkeley Census 2010 Tract Polygons](https://data.cityofberkeley.info/Demographics/Census-Tract-Polygons-2010/peq3-2arw)
+* [Census 2010 Population and race data by county tract polygons](http://www.bayareacensus.ca.gov/small/small.htm)
+* [Berkeley Census 2010 Tract Polygons](https://data.cityofberkeley.info/Demographics/Census-Tract-Polygons-2010/peq3-2arw)

### 2016-08-09 final_paper/finalpaper.Rmd
updates stop data

@@
 # histogram of stop data by race and hour:
+
 histogram_stoptime1 <- ggplot(stops, 
@@
   geom_bar(aes(y = percentage), stat="identity") +
-  facet_wrap(~AgeRange)+
+  facet_wrap(~AgeRange, nrow = 1)+
 labs(title = "Probability of BPD stop for a \nspecific age range in a specific hour",
@@
   geom_bar(aes(y = percentage), stat="identity") +
-  facet_wrap(~Reason)+
+  facet_wrap(~Reason, nrow=1)+
   labs(title = "Probability of BPD stop for a \nspecific age range in a specific reason",
@@
 
-## PLEASE CHANGE THE TABLE 
+## PLEASE mention that we delete tHE TABLE 
 #### LOVE JENNY
@@
 A visualization of the groups of people stopped in proportion to their population is provided by the bar chart below.
-```{r, message = F, echo = FALSE, fig.width = 10, fig.height = 5, warning = FALSE, fig.align="center"}
+```{r, message = F, echo = FALSE, fig.width = 10, fig.height = 3, warning = FALSE, fig.align="center"}
 grid.arrange(bargraph1, bargraph2, ncol = 2)
@@
 
-```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+```{r, fig.height = 5, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
 map_reasons1
@@
 
-!!!!!!!!!!!!!!!!!!!!!INCLUDE BAR GRAPH & TABLE OF POLICE STOPS BY REASON HERE (Use percentage, not count) !!!!!!!!!!!!!!!!!!!!!!!!
-
-```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+## PLEASE EDIT THE ANALYSIS 
+###  LOVE JENNY
+
+We also analyzed the preference of Berkeley PD when arresting people.
+We calculated the conditional probability of arrested given a specific reasion and given a specific race/ age range/ gender.
+That is to say, we calculated the probability of a person will be arrested by Berkeley PD if he was stopped by a specific reason and was of a specific race/ age range/ gender.
+
+```{r, fig.height = 3, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
 stop_arrest
 ```
-
-###  Findings 
-1. The probability of arrested in a stop with the reason Traffic is much lower than other reasons. The average conditional probability of arrested given reason is traffic is 1.58%.
-
-2. The probability of arrested in a stop with the reason Wanted is much higher than other reason. The average conditional probability of arrested given reason is wanted is 33.33%.
-
-3.
-An interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and race Asian is 75%, which is much higher than that of any other race.
+## PLEASE EDIT THE ANALYSIS 
+###  LOVE JENNY 
+From the above plots, we can tell that the probability of arrested in a stop with the reason Traffic is much lower than other reasons.
+The average conditional probability of arrested given reason is traffic is 1.58%.
+Additionally, the probability of arrested in a stop with the reason Wanted is much higher than other reason.
+The average conditional probability of arrested given reason is wanted is 33.33%.
+Lastly, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and race Asian is 75%, which is much higher than that of any other race.
 Asian people is much more liable to be arrested by Berkeley Police Department during Probation or Parole if stopped by the BPD.
 
-```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+```{r, fig.height = 3, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+stop_arrest_age
+```
+
+## FINDINGS
+###  LOVE JENNY 
+Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.
+Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and age range 0-18 is 67.56%, which is much higher than that of any other age range.
+Lastly, the conditional probability of arrested given person involved aged from 0 to 18 in a stop given any reason is more than that of given person with any other age range.
+Therefore, in a stop, teenagers is more liable to be arrested by Berkeley Police Department.
+
+```{r, fig.height = 3, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+stop_arrest_gender
+```
+
+## FINDINGS
+####  LOVE JENNY 
+Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.
+Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole of female is 50.00%, which is much higher than that of male, which is 20.11%.
+
+```{r, fig.height = 3, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
 stop_percentage3
 ```
-###  Findings 
-1. Black people are obviously liable to commit incidents at night. The average ratio of black people to all people stoped by the Berkeley Police Department at night is about 40%.
-
-2. White people are obviously liable to commit incidents in the daytime. The average ratio of white people to all people stoped by the Berkeley Police Deparment during the day is about 45%.
-
-3. Ratio of incidents commited by Asian people, Hispanic people and other people fluctuate during the daytime and the night, with an average ratio of 8%, 11% and 9% respectively.
-
-```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+##  Findings 
+#### LOVE JENNY
+Black people are obviously liable to commit incidents at night.
+The average ratio of black people to all people stoped by the Berkeley Police Department at night is about 40%.
+On the other hand, white people are obviously liable to commit incidents in the daytime.
+The average ratio of white people to all people stoped by the Berkeley Police Deparment during the day is about 45%.
+While the Ratio of incidents commited by Asian people, Hispanic people and other people fluctuate during the daytime and the night, with an average ratio of 8%, 11% and 9% respectively.
+
+```{r, fig.height = 3, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
 stop_percentage2
 ```
-### Findings 
-1.
+
+## Findings 
+#### LOVE JENNY  
 People aged from 18 to 29 are obviously liable to commit incidents at night.
 The average ratio of people aged from 18 to 29 to all people stoped by the Berkeley Police Department at night is greater than 40%.
-
-2.
 People aged greater than 40 are obviously liable to commit incidents in the daytime.
 The average ratio of people aged greater than 40 to all people stoped by the Berkeley Police Deparment during the day is greater than 40%.
-
-3. Ratio of incidents commited by people aged between 0 and 18 and people aged from 30 to 39 fluctuates during the daytime and the night, with an average ratio of 2.5% and 25% respectively.
-
-```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
-stop_arrest_age
-```
-### Jenny Findings 
-1. Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.
-
-2.
-Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and age range 0-18 is 67.56%, which is much higher than that of any other age range.
-
-3.
-The conditional probability of arrested given person involved aged from 0 to 18 in a stop given any reason is more than that of given person with any other age range.
-Therefore, in a stop, teenagers is more liable to be arrested by Berkeley Police Department.
-
-```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
-stop_arrest_gender 
-```
-### Jenny Findings 
-1. Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.
-
-2.
-Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole of female is 50.00%, which is much higher than that of male, which is 20.11%.
-
-```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 5, fig.width = 10, fig.align="center"}
-#### INCLUDE BAR GRAPH & TABLE OF POLICE STOPS BY REASON HERE(Use percentage, not count)!!!!!!!!!!!!!!!!!!!!!!!!
-grid.arrange(bararrest, barjail, ncol = 2)
-```
+While Ratio of incidents commited by people aged between 0 and 18 and people aged from 30 to 39 fluctuates during the daytime and the night, with an average ratio of 2.5% and 25% respectively.
 
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 12, fig.height=3, fig.align="center"}
-histogram_jail1
-histogram_arrest1
-# I decided to take these out of the multi.plot function because I was trying to adjust the format a certain way, I'm sorry!
+```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 3, fig.width = 10, fig.align="center"}
+grid.arrange(bararrest, barjail, ncol = 2)
 ```
@@
 ## Problems
+
+## PLEASE MENTION THE HOMELESS PEOPLE. 
+#### LOVE JENNY
 

### 2016-08-09 final_paper/finalpaper.Rmd
git figure

@@
 
-```{r, warning = FALSE, echo = F, fig.width = 10, fig.height = 5, message = FALSE, fig.align="center"}
+```{r, warning = FALSE, echo = F, fig.width = 10, fig.height = 5, message = FALSE, fig.align="center", fig.cap= "BP stop and Calls for service"}
 grid.arrange(map_stop1,map_callservice1, ncol = 2)
@@
 
-```{r, warning = FALSE, echo = F, fig.width = 10, fig.height = 5, message = FALSE, fig.align="center"}
+```{r, warning = FALSE, echo = F, fig.width = 10, fig.height = 5, message = FALSE, fig.align="center", fig.cap= "Berkeley Police Stop Density Map by Race"}
 grid.arrange(stopdensitymap4, popdenmap, ncol=2)
@@
 
-```{r, fig.height = 3, fig.width = 12, echo=FALSE, warning = FALSE, message = FALSE, fig.align="center"}
+```{r, fig.height = 3, fig.width = 12, echo=FALSE, warning = FALSE, message = FALSE, fig.align="center", fig.cap= "Police Stop density map VS. Population Density"}
 stopdensitymap3 
@@
 
-```{r, warning = FALSE, echo = FALSE,  fig.width = 15, fig.height = 3, message = FALSE, fig.align = "center"}
+```{r, warning = FALSE, echo = FALSE,  fig.width = 15, fig.height = 3, message = FALSE, fig.align = "center", fig.cap= "Berkeley Population Density Map by Race"}
 grid.arrange(a2,black2,h2,o2,w2, ncol = 5) # print population density by race.
@@
 A visualization of the groups of people stopped in proportion to their population is provided by the bar chart below.
-```{r, message = F, echo = FALSE, fig.width = 10, fig.height = 3, warning = FALSE, fig.align="center"}
+```{r, message = F, echo = FALSE, fig.width = 10, fig.height = 3, warning = FALSE, fig.align="center", fig.cap= "Percent stopped and Perecent Population Stopped"}
 grid.arrange(bargraph1, bargraph2, ncol = 2)
@@
 
-```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.align = "center", fig.width = 8, fig.height = 3}
+```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.cap= "Police Stops Leading to Arrests", fig.align = "center", fig.width = 8, fig.height = 3}
 map_enforcement
@@
 
-```{r, fig.height = 5, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+```{r, fig.height = 5, fig.width = 10, fig.cap= "Police Stops by Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
 map_reasons1
@@
 
-```{r, fig.height = 3, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Probability of Arrested by BDP Race VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
 stop_arrest
@@
 
-```{r, fig.height = 3, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Probability of Arrested by BDP Age VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
 stop_arrest_age
@@
 
-```{r, fig.height = 3, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Probability of Arrested by BDP Gender VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
 stop_arrest_gender
@@
 
-```{r, fig.height = 3, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Probability of Arrested by BDP Race VS Hour", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
 stop_percentage3
@@
 
-```{r, fig.height = 3, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Probability of Arrested by BDP Age VS Hour", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
 stop_percentage2
@@
 
-```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 3, fig.width = 10, fig.align="center"}
+```{r, echo=FALSE, fig.cap= "Percentage of being arrested and being jailed", warning = FALSE, message=FALSE, fig.height = 3, fig.width = 10, fig.align="center"}
 grid.arrange(bararrest, barjail, ncol = 2)

### 2016-08-09 final_paper/finalpaper.Rmd
figures

@@
 
-```{r, echo=FALSE, warning=F,message=F, fig.align="center", fig.width = 20, fig.height=15}
+```{r, echo=FALSE, warning=F,message=F, fig.cap= "Police Stops Leading to Arrests", fig.align="center", fig.width = 10, fig.height=5}
 map_arrested1 
@@
 
-```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.cap= "Police Stops Leading to Arrests", fig.align = "center", fig.width = 8, fig.height = 3}
+```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.align = "center", fig.width = 5, fig.height = 3}
 map_enforcement

### 2016-08-09 final_paper/finalpaper.Rmd
Rebecca: updating paper

@@
 For the STAT 133 final project, our group was interested in finding some ways of visualizing the relationship between the Berkeley Police Department and different groups of people, especially within the context of different populations and neighborhoods within Berkeley.
-To obtain our results, we accessed, cleaned and explored the four Berkeley Police Department data sets available to the public on the Berkeley Open Data website using R.
-The police data included police stops, calls for service, arrests, and jail bookings made by the the police department over the past 3-6 month period.
-Further, we accessed 2010 census data in order to obtain demographic information about Berkeley as a whole and about neighborhoods inside of Berkeley.
+To obtain our results, we accessed, cleaned and explored the four Berkeley Police Department data sets available to the public on the Berkeley Open Data website using R, the statistical programming language.
+These data sets were of police stops, calls for service, arrests, and jail bookings.
+There were 16,255 stops (including stops for traffic, suscpicious vehicle, bicycle, and pedestrian) from a 16-month period that began on January 26, 2015 (in compliance with General Order B-4, which ordered this data set to be collected for the public) and ending on April 30th, 2016.
+The calls of service included 4,911 calls, which were collected over the most recent a 6-month period, from January 20, 2016 to July 13, 2016.
+The arrest data included 232 arrests from the previous 1-month period, from June 20, 2016 to July 19, 2016.
+The jail bookings data set included 248 jail bookings over the same 1-month period.
+In addition to the police data, we accessed the Berkeley city and the Alameda County 2010 census data in order to obtain demographic information about Berkeley as a whole and about the demographics of the neighborhoods inside of Berkeley.
 The three census data sets that we used are also available through Berkeley's open data site, including the Berkeley 2000-2010 census data, which gave us the total population counts and percentages of each group of people in Berkeley for the 2010 census, the Alameda County 2010 census tract population data, which gave us the demographic information for 33 different neighborhoods (or "tracts") inside Berkeley, and the Berkeley 2010 census tract polygons shapefile (.shp), which provided the information we needed to visually display the 33 neighborhoods.
 Rather than limiting the project to one set of data, incorporating multiple data sets gave us a more holistic understanding of subject and expanded our ability to understand and visualize the data.
-Our findings suggest the hypothesis that (1) black people in Berkeley are more likely than others to be stopped by Berkeley police, and that (2) black people in Berkeley are arrested and jailed at disproportionate rates.
+Our findings suggest the hypothesis that (1) black people in Berkeley are more likely than members of another group of people to be stopped by Berkeley police, and that (2) black people in Berkeley are arrested and jailed at disproportionate rates.
 We recommend future statistical analysis be done to test our hypothesis.
@@
 
-After detailing our process of cleaning the police and census data, we will provide our results: visualizations of the data and some preliminary findings about the police data.
-Afterwards, we will discuss problems with the data and recommendation for future work on this subject.
+First we will detail our process of cleaning the police and census data and obtaining spatial information for the police data.
+Next, we will provide our analysis, which includes visualizations of the police and census data and some preliminary findings related to these visualizations.
+Afterwards, we will discuss problems with the data and recommendations for future study on this subject.
 
@@
 
-The original stop data set included 16,255 stops recorded by the Berkeley police over a 6 month period.
+The original stop data set included 16,255 stops recorded by the Berkeley police over a 16-month period.
 For each stop, the police recorded the date and time, location (as a character string, usually in the form of a street block or a street intersection), incident type (referring to whether the stop was traffic, pedestrian, bike, or suspicious vehicle), and a column called disposition.
@@
 These additional disposition abbreviations represented the following options: Primary Case Report, MDT Narrative Only, Arrest Reporty Only (No Case Report Submitted), Incident Report, Field Card, Collision Investigation Report, Emergency Psychiatric Evaluation, Impounded Vehicle, and Officer Made a Stop of More Than 5 Persons.
-We found it reasonable to assume that most of these additional dispositions were related to police paperwork procedures, and since they were used so sparingly, we decided to not consider most of them.
+We found it reasonable to assume that these additional dispositions were related to police paperwork procedures, and since they were used so sparingly, we decided to not consider most of them.
 However, in spite of only being used less than 20 times, we found the Emergency Psychiatric Evaluation disposition to be of interest, and so we chose to investigate this one in addition to the main 6 character set of dispostions, though we chose not to focus on it during the limited scope of this paper.
@@
 Cleaning the stop data was relatively time-consuming compared to the other data sets.
-First, the dispositions column in the original data set contained many columns and sometimes rows worth of information, though the comma-separated pieces were without order (for example, sometimes the optional disposition letters came before the 6-character string instead).
+First, the dispositions column in the original data set contained many columns and sometimes rows worth of information, though the comma-separated pieces were without order (for example, sometimes the optional disposition abbreviations came before the 6-character string instead).
 We handled this problem using data tidying techniques, such as separating disposition information into separate row entries for each individual assessed, in the case of multiple persons stopped, and splitting the column by isolating the six character dispositions into different columns.
@@
 Additional cleaning for the stop data included changing the stop date and time variable to the `lubridate` date format.
-We also created an hour column in order to examine the stop occurrences by hour over a 24-hour period.
+We also created an hour column and day column in order to examine the stop occurrences by hour over a 24-hour period and accross a 7-day week.
 
@@
 The number of problems with the location data that were left over afterwards were small enough to handle on a case-by-case basis.
-Further, because Google limits geolocation to 2500 queries per 24-hour period, it took about a week to process all the locations into coordinates from the stop data (this includes the time it took to fix spelling errors).
-
-At the end of the cleaning process, the stop data was whittled down from 16,255 observations to 14,291 observations (due to missing disposition values, etc.) that we will use in the following analysis.
+Further, because Google limits geolocation to 2500 queries per 24-hour period, it took approximately a week to process all the locations into coordinates from the stop data (this includes the time it took to fix spelling errors).
+
+The reason why the Incident Number column was not removed was because two teams from our group worked on cleaning different parts of the stop data at the same time; one team worked on separating the disposition column into variables, while the other team worked on creating the columns for longitude and latitude.
+After both teams had finished their cleaning task, the resulting data frames were then merged by Incident Number.
+
+At the end of the cleaning process, the stop data had been whittled down from 16,255 observations to the 14,291 police stops (due to missing disposition values) that we will use in the following analysis.
+Please see Figure 1 for a glimpse at the clean police stops data frame.
 
@@
 The number of observations available for the arrest data set and the jail data set were significantly smaller than the number of observations for the stop data set.
-The arrest data had 205 observations and the jail bookings data had 223.
-These observations were obtained by the Berkeley Police department over a 3-month period, rather than a 6-month period (as was for the stop data).
+The arrest data had 232 observations and the jail bookings data had 248.
+These observations were obtained by the Berkeley Police department over the same month-long time period between June and July of 2016.
 They both contained similar variables, including a case/arrest/booking number, date and time, type, and subject information (name, race, sex, D.O.B., age, height, weight, hair, eyes, and occupation) and statute information (type, description, agency, and disposition).
-Cleaning this data set also required the dates and times to be put into `lubridate` format, and we also created an hour column to consider the number of occurences during each hour of the day.
+Cleaning this data set also required the dates and times to be put into `lubridate` format, and we also created an hour and day column.
+See Figure 2 for the cleaned arrest data frame and Figure 3 for the cleaned jailings data frame.
 
@@
 
-The Berkeley census 2010 data was utilized from 3 different data sets.
-We first considered the summary information provided by the city website in an excel document, which was easy to clean and use in our analysis, since the information was already summarized.
-However, in order to understand the demographics of populations within Berkeley in a way that could provide a spatial visualization, we needed to use the census 2010 Alameda County population tract data, which provided the population and race data by tract number.
-For example, there are 33 census 2010 tracts within the city of Berkeley.
-Finally, to get the map information of these 33 tracts, we also needed the census 2010 Berkeley tract polygons shapefile.
-After converting the shapefile to a dataframe and changing the coordinate system to the one used by Google (this was done using the `readOGR` function from the library `rgdal`), we were able to `left_join` this new polygonal mapping data frame to the data frame containing the population information for the Berkeley tracts by the unique census 2010 tract number.
-Combining the tract map information with the tract population information gave us a data frame containing population information for 33 different areas that we could represent spatially.
-The goal was to visualize the population density of different groups of people within the city by neighborhood, which provided us with a context for the sptatially visualized stop data.
+The Berkeley census 2010 data was obtained from three separate data sets, as previously mentioned.
+We first considered the summary information provided by the city website, which was easy to clean and use in our analysis, since it provided a summary of the demographic information for each group in Berkeley as a whole.
+However, in order to understand the demographics within Berkeley in a way that could provide a spatial visualization, we needed to use the census 2010 Alameda County population tract data, which provided the population and race data by tract number for the entire county.
+We manipulated this data into a data frame containing the 33 census tracts inside Berkeley, each row representing a single tract and the columns providing demographic information for that tract.
+To be able to map these 33 tracts, we also needed the census 2010 Berkeley tract polygons shapefile.
+After converting the shapefile to a mapping dataframe and changing the coordinate system to the one used by Google (this was done using the `readOGR` function from the library `rgdal`), we were able to `left_join` this new polygonal mapping data frame to the data frame of the 33 census tracts by the unique census 2010 tract number.
+This new data frame allowed us to map population information for 33 different areas.
+The goal was to visualize the population density of different groups of people within the city by neighborhood, which provided us with a context for the sptatially visualized police data.
 
@@
 
-### Berkeley Police Stop Data Analysis
-
-Below is the stop data points mapped by location.
+## Data Analysis
+
+### Stop Data Analysis
+
+Figure 5 below provides a map of the location information of the Berkeley police stop data mapped by location.
 We have excluded some observations on the outskirts of Berkeley in order to zoom in to the street level.
@@
 
-```{r, warning = FALSE, echo = F, fig.width = 10, fig.height = 5, message = FALSE, fig.align="center", fig.cap= "BP stop and Calls for service"}
-grid.arrange(map_stop1,map_callservice1, ncol = 2)
-```
-
-## PLEASE COMPARE THIS TO THE CALLS FOR SERVICE DATA  
-### LOVE JENNY
-
-Another visualization of the stop data mapped above is the following two-dimensional density map of the police stop data, displayed below.
+```{r, warning = FALSE, echo = F, fig.width = 10, fig.height = 5, message = FALSE, fig.align="center", fig.cap= "Berkeley Police Stops"}
+map_stop1
+```
+
+Another visualization of the stop data mapped in Figure 5 is the two-dimensional density map of the police stop data in the left pane of Figure 6.
 One of the most interesting aspects of the stop data in this image is that most police stops seem to be clustered west of UC Berkeley Campus in the Downtown area.
 This was contrary to what some of us might have thought before, which was that the area with the higher police stops would be the area with the higher population.
-As you will see, however, this is not necessarily true.
-
-```{r, warning = FALSE, echo = F, fig.width = 10, fig.height = 5, message = FALSE, fig.align="center", fig.cap= "Berkeley Police Stop Density Map by Race"}
-grid.arrange(stopdensitymap4, popdenmap, ncol=2)
-```
-
-To visualize the relationship between police stops and population density, consider the following map, which visualizes the percent population of Berkeley for each of the 33 census 2010 tracts.
+To visualize the relationship between police stops and population density, consider the map in the right pane of Figure 6, which visualizes the percent population of Berkeley for each of the 33 census 2010 tracts.
 As you can see, the least densely populated area is the census 2010 tract containing the University of California.
@@
 
+```{r, warning = FALSE, echo = F, fig.width = 10, fig.height = 5, message = FALSE, fig.align="center", fig.cap= "2015-2016 Police Stop Density (left) and 2010 Census Tract Population Density (right)"}
+grid.arrange(stopdensitymap4, popdenmap, ncol=2)
+```
+
 In order to break down the stop data further, we mapped the data by each of the disposition variables.
 We found race to be one of the most interesting variables.
-Consider the Berkeley stop data density map faceted by race:
+Consider the stop data density map faceted by race in Figure 7.
 
@@
 
-The image above seems to imply that the group of people stopped the most are African Americans in Berkeley.
+Figure 7 seems to suggest that the group of people stopped the most in Berkeley are African Americans.
 For instance, in Centeral Berkeley, the stops for the races defined by the Berkeley Police Department as Asian, Hispanic, White, and Other seem comparatively lower than for Blacks in the same area.
-In order to consider the context of these stop incidents, we considered the population distribution for these areas, as shown below.
-Note that the population density visualized by the following map is conveyed by the range between white (0%) and red (100%).
+In order to consider the context of these stops, we considered the population distribution for these areas, shown in Figure 8.
+Note that the population density visualized by Figure 8 is conveyed by the range between white (0%) and red (100%).
 As you can see, White people make up a striking majority in all census 2010 tract areas, which seems to suggest the hypothesis that the population of Black people in Berkeley is being disproportionally affected by police stops.
@@
 Now, with the population percentages of each group in hand, we were able to compare the results of our stop data analysis to the percent population of each group.
-The `Percent Stopped` column in the table below represents the percent of each group compared to the total number of people stopped, while the `Percent Population Stopped` column represents the proportion of people in each group stopped relative to the population of that group living in Berkeley.
-Even though the people stopped by police are not necessarily Berkeley residents, the `Percent Population Stopped` gives us some perspective of which groups of people in Berkeley might be more likely to be affected by police stops than others.
-
-## PLEASE mention that we delete tHE TABLE 
-#### LOVE JENNY
-
-A visualization of the groups of people stopped in proportion to their population is provided by the bar chart below.
-```{r, message = F, echo = FALSE, fig.width = 10, fig.height = 3, warning = FALSE, fig.align="center", fig.cap= "Percent stopped and Perecent Population Stopped"}
+In Figure 9 (left), the `Percent Stopped` chart conveys the percent of each group stopped compared to the total number of people stopped.
+In Figure 9 (right), the `Percent Population Stopped` chart conveys the proportion of people in each group stopped relative to the population of that group living in Berkeley.
+Even though the people stopped by police are not necessarily Berkeley residents, the `Percent Population Stopped` chart suggests that some groups of people in Berkeley might be more likely to be affected by police stops, especially Black people.
+
+```{r, message = F, echo = FALSE, fig.width = 10, fig.height = 3, warning = FALSE, fig.align="center", fig.cap= "Police Stops in Berkeley by Race"}
 grid.arrange(bargraph1, bargraph2, ncol = 2)
@@
 
-Another disposition variable we found to be of interest was the enforcement variable, which has the options Arrest, Citation, Warning, and Other.
+Another disposition variable we found to be of interest was the Enforcement variable, which has the options Arrest, Citation, Warning, and Other.
 We were interested to see how these enforcements were applied to different groups of people.
-Consider the map below, which appears to suggest that the enforcement of arrest is applied differently to different groups of people.
-
-```{r, echo=FALSE, warning=F,message=F, fig.cap= "Police Stops Leading to Arrests", fig.align="center", fig.width = 10, fig.height=5}
-map_arrested1 
-```
-
-```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.align = "center", fig.width = 5, fig.height = 3}
-map_enforcement
-```
-
-A third interesting disposition variable was the reason disposition, which indicates the reason for the police stop in each case.
+Consider Figure 10 (top), which maps the stops that led to the enforcement of arrest.
+The number of incidents of police stops mapped for Blacks seems higher than for the other groups.
+Upon examination of the chart in Figure 10 (bottom), we see that of the 441 individuals stopped and then arrested as a result of the stop during this 18-month period, Blacks were the majority.
+Though the Berkeley police department is prohibited from the use of biased policing practices, Figure 10 does suggest that the enforcement of arrest may be applied differently to different groups of people.
+
+```{r, echo=FALSE, warning=F,message=F, fig.align="center", fig.width = 10, fig.height=7, fig.cap = "Police Stops Leading to Arrests"}
+grid.arrange(map_arrested1, map_enforcement, ncol=1)
+```
+
+A third interesting disposition variable was the Reason disposition, which indicates the reason for the police stop in each case.
 The reasons available for police stops are Investigation, Other, Probation/Parole, Reasonable Suspicion, Traffic, and Wanted.
@@
 Due to the length and scope of this paper however, we decided to end our discussion of the stop data by showing that most of the Berkeley police stops are traffic stops.
-Mapping the coordinates seems to suggest this fact, and it is confirmed when analyzing the data.
+Mapping the coordinates by reason for the stop, as in Figure 11, suggests that this is the most common reason for a stop in Berkeley.
 
@@
 
-## PLEASE EDIT THE ANALYSIS 
-###  LOVE JENNY
-
-We also analyzed the preference of Berkeley PD when arresting people.
-We calculated the conditional probability of arrested given a specific reasion and given a specific race/ age range/ gender.
-That is to say, we calculated the probability of a person will be arrested by Berkeley PD if he was stopped by a specific reason and was of a specific race/ age range/ gender.
-
-```{r, fig.height = 3, fig.width = 10, fig.cap= "Probability of Arrested by BDP Race VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
-stop_arrest
-```
-## PLEASE EDIT THE ANALYSIS 
-###  LOVE JENNY 
-From the above plots, we can tell that the probability of arrested in a stop with the reason Traffic is much lower than other reasons.
-The average conditional probability of arrested given reason is traffic is 1.58%.
-Additionally, the probability of arrested in a stop with the reason Wanted is much higher than other reason.
-The average conditional probability of arrested given reason is wanted is 33.33%.
-Lastly, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and race Asian is 75%, which is much higher than that of any other race.
-Asian people is much more liable to be arrested by Berkeley Police Department during Probation or Parole if stopped by the BPD.
-
-```{r, fig.height = 3, fig.width = 10, fig.cap= "Probability of Arrested by BDP Age VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
-stop_arrest_age
-```
-
-## FINDINGS
-###  LOVE JENNY 
-Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.
-Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and age range 0-18 is 67.56%, which is much higher than that of any other age range.
-Lastly, the conditional probability of arrested given person involved aged from 0 to 18 in a stop given any reason is more than that of given person with any other age range.
-Therefore, in a stop, teenagers is more liable to be arrested by Berkeley Police Department.
-
-```{r, fig.height = 3, fig.width = 10, fig.cap= "Probability of Arrested by BDP Gender VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
-stop_arrest_gender
-```
-
-## FINDINGS
-####  LOVE JENNY 
-Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.
-Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole of female is 50.00%, which is much higher than that of male, which is 20.11%.
-
-```{r, fig.height = 3, fig.width = 10, fig.cap= "Probability of Arrested by BDP Race VS Hour", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
-stop_percentage3
-```
-##  Findings 
-#### LOVE JENNY
-Black people are obviously liable to commit incidents at night.
-The average ratio of black people to all people stoped by the Berkeley Police Department at night is about 40%.
-On the other hand, white people are obviously liable to commit incidents in the daytime.
-The average ratio of white people to all people stoped by the Berkeley Police Deparment during the day is about 45%.
-While the Ratio of incidents commited by Asian people, Hispanic people and other people fluctuate during the daytime and the night, with an average ratio of 8%, 11% and 9% respectively.
+We also analyzed possible preferences of Berkeley police officers when arresting individuals.
+We calculated the conditional probability of an individual being arrested given a specific reasion for the stop and given a specific race / age range / gender.
+That is to say, we calculated the probability a person will be arrested if he was stopped for a specific reason and was of a specific race / age range / gender.
+
+For Figure 12, we can tell that the probability of a stopped individual being arrested in a stop with the reason being traffic is much lower than other stop reasons.
+The average conditional probability of being arrested for a traffic stop is 1.58%.
+Additionally, the probability of being arrested in a stop with the reason wanted is much higher than other reason.
+The average conditional probability of being arrested given being stopped for the wanted reason is 33.33%.
+Lastly, an interesting fact is that the conditional probability of being arrested in a stop for probation or parole and race Asian is 75%, which is much higher than that of any other race.
+Our conclusion is that Asian people are much more liable to be arrested in Berkeley during their probation or parole if stopped by police.
+
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Police Stop: Race VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+stop_arrest # figure 12
+```
+
+In Figure 13, the probability of being arrested as a result of a police stop in Berkeley for a traffic reason is again the lowest and wanted is the highest.
+Again similar to the conclusion in Figure 12, an interesting fact is that the conditional probability of being arrested in a stop with the reason probation or parole and age range 0-18 is 67.56%, which is much higher than that of any other age range for that reason.
+Lastly, the conditional probability of being arrested in a police stop given an individual in the age range 0-18 given any reason is more than that of any other age range.
+This suggests that in a police stop, teenagers are more liable to be arrested in Berkeley.
+
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Police Stop: Age VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+stop_arrest_age # figure 13
+```
+
+Again for Figure 14, the conditional probability of being arrested during a police stop for the stop reason of traffic is the lowest and wanted is the highest.
+Again similar to the conclusion above, an interesting fact is that the conditional probability of being arrested in a police stop with the stop reason probation or parole given a gender of female is 50.00%, which is much higher than that of male, which is 20.11%.
+
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Police Stop: Gender VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+stop_arrest_gender # figure 14
+```
+
+Figure 15 is interesting because it suggests that black people are more likely to be stopped by the police in Berkeley during the hours of night.
+The average ratio of black people to all people stopped by the Berkeley Police Department at night is about 40%.
+Figure 15 also suggests that white people are more liable to be stopped by the police in Berkeley during the hours of daytime.
+The average ratio of white people to all people stopped by the Berkeley Police Deparment during the day is about 45%.
+While the ratio of Asian people, Hispanic people and other people stopped by police fluctuate during the daytime and the night, with an average ratio of 8%, 11% and 9% respectively.
+
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Poby BDP Race VS Hour", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+stop_percentage3 # figure 15
+```
+Figure 16 implies that people aged from 18 to 29 are more liable to be stopped by police in Berkeley at night.
+The average ratio of people aged from 18 to 29 to all people stopped by the Berkeley Police Department at night is greater than 40%.
+The figure also implies that people aged 40+ are more liable to be stopped by the police during the daytime.
+The average ratio of people aged 40+ to all people stopped by police during the day is greater than 40%.
+Further, the ratio of stops of people aged 0-18 and people aged 30-39 fluctuates during the daytime and the night, with an average ratio of 2.5% and 25% respectively.
 
 ```{r, fig.height = 3, fig.width = 10, fig.cap= "Probability of Arrested by BDP Age VS Hour", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
-stop_percentage2
-```
-
-## Findings 
-#### LOVE JENNY  
-People aged from 18 to 29 are obviously liable to commit incidents at night.
-The average ratio of people aged from 18 to 29 to all people stoped by the Berkeley Police Department at night is greater than 40%.
-People aged greater than 40 are obviously liable to commit incidents in the daytime.
-The average ratio of people aged greater than 40 to all people stoped by the Berkeley Police Deparment during the day is greater than 40%.
-While Ratio of incidents commited by people aged between 0 and 18 and people aged from 30 to 39 fluctuates during the daytime and the night, with an average ratio of 2.5% and 25% respectively.
-
-## Berkeley Police Jail and Arrest Data Set Analysis
-
-Moving on from the stop data, we considered the available jail data and arrest data for the city of Berkeley.
-Two of the most interesting variables to consider in these two data sets were race and age.
+stop_percentage2 # figure 16
+```
+
+### Berkeley Police Jail and Arrest Data Set Analysis
+
+Moving on from the stop data, we considered the available jail data and arrest data for the city of Berkeley, which is not necessarily dependent on the stop data.
+Two of the most interesting variables to consider in these two data sets were also race and age.
 We decided the examine the percent of people arrested and the percent of people jailed in Berkeley by both age and race.
-Below is a visualization.
+Figure 17 is a visualization.
 Note that the following percentages are related to the number of total people jailed or the number of people arrested, respectively, and we have not adjusted the data relative to the population of Berkeley.
 
-```{r, echo=FALSE, fig.cap= "Percentage of being arrested and being jailed", warning = FALSE, message=FALSE, fig.height = 3, fig.width = 10, fig.align="center"}
-grid.arrange(bararrest, barjail, ncol = 2)
-```
-
-One of the most interesting things about the above histograms is the number of black people arrested in Berkeley over the past 3 month period is higher than the numbers of people arrested in Berkeley over the past 3 months for other groups of people.
-Note that this statement is even stronger for young black people.
-Even though most of the population in Berkeley is white, most of the people recently arrested are black.
+```{r, echo=FALSE, fig.cap= "Arrested and Jailed in Berkeley", warning = FALSE, message=FALSE, fig.height = 3, fig.width = 10, fig.align="center"}
+grid.arrange(bararrest, barjail, ncol = 2) # figure 17
+```
 
@@
 
-## PLEASE MENTION THE HOMELESS PEOPLE. 
-#### LOVE JENNY
-
 The data sets provided by the police (stops, arrests, and jail bookings) have several problems.
-First, concerning the methods for recording race, especially for police stops that do not result in an arrest, it seems unreasonable to assume that racial information would have been provided by the individual stopped.
-It seems more likely that police officers are judging the race of the people they stop during stops for themselves in order to record it in the system that ultimately provided us with that data set.
-The problem with that method of data collection is that it opens up an opportunity for error; for example, if the data is recorded by the officers through their judgment of the individuals physical appearance, they may unknowingly classify that individual as a race that the individual does not identify themselves as.
-Even though there doesn't seem to be a means to get around this problem with the actual method of collecting racial data for police stops, it is still important to address it as something that could induce error into a statistical analysis.
-
-In further regards to the racial data available, it is important to note that the census data we used to compare the stop data to included many more categories for race than just the five used by police (Asian, Black, Hispanic, Other, and White).
-To deal with this problem, other races included by the census were moved to the category of other.
+First, the open data site updates the data frequently and we are concerned that the data available to be used for this paper won't be available after the data is updated, since the early data is removed.
+Second, concerning the methods for recording disposition data like race and gender, the Berkeley Open Data website indicates that the police officers conducting the stops are deciding these dispositions on their own, meaning that they are judging whether or not the individual apprehended is male or female, black or hispanic, etc., meaning that the data could be misrecorded.
+Even though there doesn't seem to be a means to get around this problem with the actual method of collecting the disposition data for stops that don't lead to citation or arrest, it is still important to address it as something that could induce error into a statistical analysis of the data.
+
+Another problem with the stop data is the many data-entry problems we faced in the character string location column (which sometimes included only 1 street), since there is likely some stops that are not mapped correctly due to approximations made by the Google geolocation service (for example, in some cases, only having one street name instead of a corner caused Google to estimate the exact location based on few conditions).
+Though most of the location problems were fixed using string detection, there is likely some error in our visual representation of the stop data maps.
+Also, the website indicates that any specific addresses were changed to "block" address form to preserve anonymity; therefore the spatial data cannot be expected to be of a precision smaller than a city block.
+
+In regards to the racial data from all of the police data, it is important to note that the census data we used to compare the stop data to included many more categories for race than just the five used by police (Asian, Black, Hispanic, Other, and White).
+To deal with this problem, other races included by the census data but not included by the police data were combined with the category of other.
 Additionally, the census data had a category for Two or More Races, which we also included as Other in order to compare the data sets.
@@
 
-Secondly, the police data used in this paper comes from a data set maintained by the City of Berkeley that is updated every 3 months (as is the case for arrests and jail bookings) or every 6 months (as is the case for stops).
-A more thorough analysis of this data may want to include an analysis of the data over a longer time frame.
-
-Next, with regard to the stop data in particular, the many data-entry problems we faced in the character string location column (which sometimes included only 1 street), there is likely some stops that are not mapped correctly due to approximations made by the Google geolocation service (for example, in some cases, only having one street name instead of a corner caused Google to estimate the exact location based on few conditions).
-Though most of the location problems were fixed using string detection, there is likely some error in our visual representation of the stop data maps.
-
-Finally, with regard to the calls for service, we were unfortunately unable to incorporate that particular data set due to constraints on time.
-However, a more full analysis of police interactions with the community might consider incorporating the calls for service data.
+Finally, it should be noted that a more accurate analysis of Berkeley police data would also incorporate demographics for the homeless population as well as for the UC student population, both of which are not included in the census data.
+Including these data sets would provide a more nuanced context for the stop data.
 
@@
 Another `shiny` app idea we had was to filter the calls for service data by proximity to certain locations so that users could use the app to select a location and find a visual displaying the most common types of calls in that area.
-It would also be useful to find a way to create an app can easily be updated when the new data is posted (every 3-6 months), allowing users of the app to get the most up-to-date information.
+It would also be useful to find a way to create an app that can easily be updated when the new data is posted (every month), allowing users of the app to get the most up-to-date information.
 

### 2016-08-09 final_paper/finalpaper.Rmd
Rebecca: editing

@@
 ```{r, warning = FALSE, echo = F, fig.width = 10, fig.height = 5, message = FALSE, fig.align="center", fig.cap= "Berkeley Police Stops"}
-map_stop1
+map_stop1 # figure 5
 ```
@@
 ```{r, warning = FALSE, echo = F, fig.width = 10, fig.height = 5, message = FALSE, fig.align="center", fig.cap= "2015-2016 Police Stop Density (left) and 2010 Census Tract Population Density (right)"}
-grid.arrange(stopdensitymap4, popdenmap, ncol=2)
+grid.arrange(stopdensitymap4, popdenmap, ncol=2) # figure 6
 ```
@@
 
-```{r, fig.height = 3, fig.width = 12, echo=FALSE, warning = FALSE, message = FALSE, fig.align="center", fig.cap= "Police Stop density map VS. Population Density"}
-stopdensitymap3 
+```{r, fig.height = 3, fig.width = 12, echo=FALSE, warning = FALSE, message = FALSE, fig.align="center", fig.cap= "Police Stop Density Map by race"}
+stopdensitymap3 # figure 7
 ```

### 2016-08-09 final_paper/finalpaper.Rmd
Rebecca: editting

@@
   scale_y_continuous(labels=percent, limits = c(0,1)) +
-  labs(title = "The Percentage of Being Arrested \nAfter being Stopped (%)",
+  labs(title = "Races Arrested in 30 Days (%)",
        x = "Race",
@@
   scale_y_continuous(labels=percent, limits = c(0,1)) +
-  labs(title = "The Percentage of Being jailed \nAfter being Stopped (%)",
+  labs(title = "Races Jailed in 30 Days (%)",
        x = "Race",
@@
 
-```{r, fig.height = 3, fig.width = 12, echo=FALSE, warning = FALSE, message = FALSE, fig.align="center", fig.cap= "Police Stop Density Map by race"}
+```{r, fig.height = 3, fig.width = 12, echo=FALSE, warning = FALSE, message = FALSE, fig.align="center", fig.cap= "Police Stop Density Map by Race"}
 stopdensitymap3 # figure 7
@@
 ```{r, warning = FALSE, echo = FALSE,  fig.width = 15, fig.height = 3, message = FALSE, fig.align = "center", fig.cap= "Berkeley Population Density Map by Race"}
-grid.arrange(a2,black2,h2,o2,w2, ncol = 5) # print population density by race.
+grid.arrange(a2,black2,h2,o2,w2, ncol = 5) # print population density by race # figure 8
 ```
@@
 
-```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Poby BDP Race VS Hour", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Police Stop: Race VS Hour", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
 stop_percentage3 # figure 15
@@
 
-```{r, fig.height = 3, fig.width = 10, fig.cap= "Probability of Arrested by BDP Age VS Hour", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Arrested  During a Police Stop: Age VS Hour", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
 stop_percentage2 # figure 16

### 2016-08-09 final_paper/finalpaper.Rmd
rebecca edit

@@
 Figure 17 is a visualization.
-Note that the following percentages are related to the number of total people jailed or the number of people arrested, respectively, and we have not adjusted the data relative to the population of Berkeley.
+Note that the following percentages are related to the number of total people jailed or the number of people arrested during a recent 30 day time frame, respectively, and we have not adjusted the data relative to the population of Berkeley.
+It is interesting to note that the proportions of arrests and jailings during this time are nearly identical, since these arrests and jailings occur during the same time frame.
+The data suggests that being jailed after being arrested in Berkeley is almost certain, and that this is true for all races.
+However, note that the proportion of Blacks being arrested and jailed is more than the other races, in spite of Berkeley being a city of mostly White people.
 

### 2016-08-09 final_paper/draft.Rmd
updates the paper

--- 2016-08-03 final_paper/draft.Rmd
@@
 ---
-title: "BDP draft paper"
-author: "Amy Zhu; Mengyu Li; Rebecca Reus; Shangjun Jiang"
-date: "2 August 2016"
-output: html_document
+title: "STAT 133 Final Project: Visualizing Berkeley Police Data"
+author: "Amy Zhu, Mengyu Li, Rebecca Reus, and Shangjun Jiang"
+date: "9 August 2016"
+output: 
+    pdf_document:
+    latex_engine: xelatex
+sansfont: Times New Roman
+
 ---
 
-```{r setup, include=FALSE}
+```{r setup, include=FALSE, warning = FALSE, message = FALSE}
 knitr::opts_chunk$set(echo = TRUE)
@@
 library(data.table)
-library(rworldmap)
 library(maps)
@@
 library(RgoogleMaps)
-library(tmap)
-library(sp)
 library(rgdal)
@@
 library(gridExtra)
+library(pander)
+library(formattable)
+library(pander)
+library(extrafont)
+setwd("~/Desktop/Stats R/PROJECT/FINALPAPER/finalproject_angry_ladies/final_paper")
 ```
 
-## Introduction
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
 
- For the final project, our group cleaned, explored, and analyzed four different data sets from the City of Berkeley that contained information on stops, calls for service, arrests, and jail bookings made by the the police department in 2016 (and 2015 as well for the stop data).
-Rather than limiting the project to one set of data, three sets were chosen in order to gain a more holistic and comprehensive understanding of the data.
-The additional variables included expanded the project’s capacity for manipulating data, examining relationships, and improving result reliability.
+# read in the data and define parameters:
 
-#### Data Collection
+############################ read in the data ###########################
 
-* [Stop Data (16,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Stop-Data/6e9j-pj9p)
-* [Arrest (200)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Arrests/xi7q-nji6)
-* [Jail Bookings (250)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Jail-Bookings/7ykt-c32j)
-* [Berkeley Census Data](https://data.cityofberkeley.info/)
+############################ Police Stop Data ###########################
+stops <- read_csv("../clean_data/StopData_clean.csv")
 
- With the variety of resources and information made available by collating multiple data sets, the objective for the project was to study differences in police activity (in terms of call requests and patrols) and intensity of assessed offenses based on time, race, gender, age, and mental health.
-As a challenge, another project target was to create a map applet depicting the density of police activity in Berkeley with an interactive component allowing the visitor to input an address and observe their proximity and observe the types of incidences that occurred most commonly in the area.
+stops$Gender<- stops$Gender %>% 
+  str_replace("F", "Female") %>% 
+  str_replace("M", "Male")
 
-	For the stop data, there were 16,255 incidents assessed by the Berkeley police.
-In each case, the call date and time, location, incident type, and disposition(s) were recorded.
-The cases typically had a six character disposition, with each character conveying race, gender, age, reason, enforcement, and car search,  respectively, for each subject involved in the incident, although there were additional dispositions, ranging from one to three characters,  that could be input and conveyed other information.
-In order to prepare the data for exploration, the cleaning process included changing the call date and time to the lubridate format, transforming the addresses into longitude and latitude referencing Google Maps, and separating the information on dispositions into separate row entries for each individual assessed in a case and splitting it further by isolating “other” and six character dispositions into different columns.
+stops$AgeRange<- stops$AgeRange %>% 
+  str_replace("1", "0-18") %>% 
+  str_replace("2", "18-29") %>% 
+  str_replace("3", "30-39") %>%  
+  str_replace("4", "40+")
 
-	The data for arrests (205), and jail bookings (223) contained similar information on case/arrest/booking number, date and time, type, and subject information (name, race, sex, D.O.B., age, height, weight, hair, eyes, and occupation) and statute information (type, description, agency, and disposition).
-Cleaning required the dates and times to be put into lubridate format, for the two data sets to be compiled in a reasonable, and for other needed adjustments.
-The same code created for the stop data would be used again to convert address information to longitudinal and latitudinal coordinates.
+########################### arrest data ###########################
+arrest <- read_csv("../raw_data/Berkeley_PD_Log_-_Arrests.csv")
 
-## Data Analysis
+arrest$Age <- as.integer(arrest$Age)
+arrest$Age[arrest$Age<18] <- "0-18"
+arrest$Age[arrest$Age>=18 & arrest$Age<30] <- "18-29"
+arrest$Age[arrest$Age>=30 & arrest$Age<40] <- "30-39"
+arrest$Age[arrest$Age>=40] <- "40+"
+arrest$Age <- as.factor(arrest$Age)
 
-```{r, echo=FALSE}
-# Read data
+############################ jail data ###########################
+jail <- read.csv("../raw_data/Berkeley_PD_Log_-_Jail_Bookings.csv")
 
-############################## stop data ############################## 
-stop_by <- read_csv("cleaned_stop_data.csv")
+jail$Age <- as.integer(jail$Age)
+jail$Age[jail$Age<20] <- "0-19"
+jail$Age[jail$Age>=20 & jail$Age<30] <- "20-29"
+jail$Age[jail$Age>=30 & jail$Age<40] <- "30-39"
+jail$Age[jail$Age>=40] <- "40+"
+jail$Age <- as.factor(jail$Age)
 
-stop_by$Call.Date.Time <- mdy_hm(stop_by$Call.Date.Time)
-stop_by$AgeRange <- as.factor(stop_by$AgeRange)
+############################ Census 2010 Data by Polygonal Tract Number ###########################
+berkcensus2010 <- read_csv("../clean_data/census2010tractpop_clean.csv")
 
-stop_by <- stop_by %>%
-  mutate(Hour = as.integer(hour(Call.Date.Time))) %>%
-  mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))
+census <- read.csv("../raw_data/Census_Data_2000_And_2010.csv")
+census2010 <- census[census$Year==2010,]
+totalpop <- census2010[1,]$Amount
 
-############################## location data ############################## 
-mergedf <- readRDS( 'StopData_merged.rds' )
+racecensus2010 <- census2010[census2010$Heading=="Not Hispanic or Latino"|census2010$Heading=="HISPANIC OR LATINO AND RACE",]
+sexcensus2010 <- census2010[census2010$Heading=="Sex",]
+agecensus2010 <- census2010[census2010$Heading=="Age",]
+
+race <- select(racecensus2010, Description, Amount) 
+race$Description <- race$Description %>% 
+  str_replace_all("Hispanic or Latino.*", "Hispanic") %>% 
+  str_replace("Black or African American", "Black") %>% 
+  str_replace("American Indian and Alaska Native", "Other") %>% 
+  str_replace("Native Hawaiian and Other Pacific Islander", "Other") %>% 
+  str_replace("Some other race", "Other") %>% 
+  str_replace("Two or more races", "Other")
+
+race <- race[race$Description !="Not Hispanic",]
+
+race <- race %>% 
+  group_by(Description) %>% 
+  tally(Amount)
+
+names(race) <- c("Description", "Counts")
+
+sex <- select(sexcensus2010, Description, Amount)
+
+age <- select(agecensus2010, Description, Amount) 
+age <- age[age$Description != "Median age",]
+age$Description <- c("0-18", "0-18", "18-64", "65+")
+
+age<- age %>% 
+  group_by(Description) %>% 
+  tally(Amount)
+names(age) <- c("Description", "Counts")
+
+################## Census 2010 Data by Polygonal Tract Number, ready for mapping #################
+berkcensus2010map <- readRDS("../clean_data/berk_census2010_mapdata.rds")
+
+############################ Police Calls for service ############################ 
+callservice <- readRDS("../clean_data/CallsForService.rds")
+
+callservice<- callservice %>% 
+  mutate(latlong= 
+           str_replace_all(callservice$Block_Location,"[0-9]* [A-Za-z]*", "") %>%
+           str_replace_all("\nBerkeley,\n", "") %>% 
+           str_replace("[A-Za-z]*", "") %>% 
+           str_replace(";","") %>% 
+           str_replace("&","") %>% 
+           str_replace("[A-Za-z]*", ""))
+
+callservice$latlong<- str_replace(callservice$latlong, "\\(", "") %>% 
+                    str_replace("\\)", "")
+
+callservice<- callservice %>% 
+  separate(latlong, c("lat", "long"), sep=",")
+
+callservice$lat <- as.numeric(callservice$lat)
+callservice$long <- as.numeric(callservice$long)
+
+#############################  Tables to use in paper ############################ 
+tables <- readRDS("../clean_data/Tables_AgeSexRacePopStop.rds")
+ss <- as.data.frame(tables[5]) # stop data by race proportional to population
+pop1 <- as.data.frame(tables[4]) # census data population count table.
+race <- as.data.frame(tables[3]) # race population count
+sex <- as.data.frame(tables[2]) # sex population count
+age <- as.data.frame(tables[1]) # age population count
+
+dataframelist <- readRDS("../clean_Data/StopData_tbls.rds")
+table1 <- data.frame(dataframelist[1])
+table2 <- data.frame(dataframelist[2])  
+table3 <- data.frame(dataframelist[3])
+
+```
+
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
 
 ## Location info:
-latmax <- max(mergedf$lat, na.rm = TRUE) 
-latmin <- min(mergedf$lat, na.rm = TRUE)
-lonmax <- max(mergedf$long, na.rm = TRUE) 
-lonmin <- min(mergedf$long, na.rm = TRUE)
+
+latmax <- max(stops$lat, na.rm = TRUE) 
+latmin <- min(stops$lat, na.rm = TRUE)
+lonmax <- max(stops$long, na.rm = TRUE) 
+lonmin <- min(stops$long, na.rm = TRUE)
 latvals <- c(latmin, latmax)
@@
 
-## Make the variables factors:
-mergedf$CarSearch <- factor(mergedf$CarSearch)
-levels(mergedf$CarSearch) <- c("No Search", "Search")
+# get rid of the axes theme:
 
-mergedf$Race <- factor(mergedf$Race)
-levels(mergedf$Race) <- c("Asian", "Black", "Hispanic", "Other", "White")
+ditch_the_axes <- theme(
+  axis.text = element_blank(),
+  axis.line = element_blank(),
+  axis.ticks = element_blank(),
+  panel.border = element_blank(),
+  panel.grid = element_blank(),
+  axis.title = element_blank()
+  )
 
-mergedf$Enforcement <- factor(mergedf$Enforcement)
-levels(mergedf$Enforcement) <- c("Arrest", "Citation", "Other", "Warning")
+# get the plain Berkeley map from Google:
 
-mergedf$Reason <- factor(mergedf$Reason)
-levels(mergedf$Reason) <- c("Other", "Investigation", "Probation/Parole", "Reasonable Suscipcion", "Traffic", "Wanted")
+# zoomed out berkeley map:
+berkMap = map = get_map(location = c(lon = mean(lonvals), 
+                                     lat = mean(latvals)), 
+                        zoom = 12) 
+berkgg_zoom1 <-ggmap(berkMap) +
+  xlim(-122.335, lonmax) + 
+  ylim(latvals) +
+  ditch_the_axes
 
-mergedf <- mergedf %>%
-  mutate(Arrest = ifelse( 
-    as.character(Enforcement) == "Arrest", 
-    "Arrested",
-    "Not Arrested") ) %>%
-  mutate(Arrest = factor(Arrest))
+# zoomed in berkeley map:
+berkMap2 = map = get_map(location = c(lon = mean(lonvals), 
+                                      lat = mean(latvals)), 
+                         zoom = 14) 
+berkgg_zoom2 <-ggmap(berkMap2) +
+  ditch_the_axes
 
-v <- factor(mergedf$Other)
-levels(v)
+########### CENSUS DATA MAPS: ########
 
-mergedf <- mergedf %>%
-  mutate(Emergency.Psych.Eval = ifelse( str_detect(as.character(Other), "MH"), 
-                                        yes = "Yes",
-                                        no = "No") )
-
-mergedf$Emergency.Psych.Eval <- factor(mergedf$Emergency.Psych.Eval)
-
-### BerkeleyMap 
-berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14)
-
-############################## census data ##############################
-
-blocks <- readOGR("Census_Tract_Polygons2010","Census_tracts_2010", verbose = TRUE)
-
-locationCensusFiles <- "Census_Tract_Polygons2010"
-blocks <- readOGR(locationCensusFiles,"Census_tracts_2010", verbose = TRUE)
-b2 <- spTransform(blocks, CRS("+proj=longlat +datum=WGS84"))
-b3 <- fortify(b2)
-
-############################## arrest data ############################## 
-arrest <- read.csv("Berkeley_PD_Log_-_Arrests.csv")
-
-############################## jail data ############################## 
-jail <- read.csv("Berkeley_PD_Log_-_Jail_Bookings.csv")
+###############################  create color palette ############################## 
+p <- colorRampPalette(c("white", "red"))(128)
+palette(p)
 
@@
 }
+
 ```
 
-We found the census data from the Berkeley open data website and created a headmap of the Berkeley population.
-Similarly, we created a heat map based on the stop data.
-Interestingly, we found that the most dense place (around north Berkeley) is a relatively safe place.
-The place where people are more likely to be stopped (downtown berkeley area) is less dense.
-Since Downtown Berkeley area is a transportation hub, numerous people come and go around this area.
-Though there are more people living north Berkeley, mainly residents will visit the area.
-Therefore, the most likely place to get stopped is not the most dense living place.
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
 
-### Heat Map 1: Berkeley Census data.
-```{r, fig.width=10, fig.asp=0.8, echo=FALSE}
-ggmap(berkMap) +
-  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
-    size = 2, bins = 5, data = b3, geom = "polygon") +
-  scale_fill_gradient(low = "black", high = "red") +
-    theme (panel.grid.major = element_blank (), # remove major grid
-		       panel.grid.minor = element_blank ()  # remove minor grid
-		       ) + 
-  ggtitle("Population Density") +
-  labs(alpha = element_blank())+
-  guides(alpha = FALSE)
+############################ create the maps ###########################
+
+m <- stops %>% 
+  group_by(Race) %>% 
+  tally()
+names(m) <- c("Race", "stop")
+names(race) <- c("Race", "census")
+
+pop1 <- left_join(race,m,by = "Race")
+pop1 <- pop1 %>% 
+  mutate(percentage=(stop/census))
+pop1$percentage<- round(pop1$percentage, digits = 2)
+
+pop1 <- pop1 %>% 
+  mutate(totalp = (census/totalpop))
+pop1$totalp <- round(pop1$totalp, digits =2)
+
+p1 <- ggplot(pop1)+
+  geom_bar(aes(x=Race,
+               y=percentage), 
+           stat = "identity",
+           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+  geom_text(aes(x=Race,
+               y=percentage,
+               label = paste0(percentage*100,"%")), 
+            size=5,
+            colour= "#000000")+
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  labs(title = "The Percentage of Being \nStopped by Police (%)",
+       x = "Race",
+       y = "Percentage") +
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) + 
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))
+
+p2 <- ggplot(pop1)+
+  geom_bar(aes(x=Race,
+               y=totalp), 
+           stat = "identity",
+           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+  geom_text(aes(x=Race,
+               y=totalp,
+               label = paste0(totalp*100,"%")), 
+            size=5,
+            colour= "#000000")+
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  labs(title = "The Percentage of Berkeley \nCensus Data 2010 (%)",
+       x = "Race",
+       y = "Percentage") +
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) + 
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))
+
+# stop data map
+map_stop1 <- berkgg_zoom2 +
+  geom_point(data = stops, aes(x=long,y=lat), alpha = .2) +
+  ggtitle("Berkeley Police Stops, 2015-2016")+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
+# calls for service
+map_callservice1 <- berkgg_zoom2 +
+  geom_point(data = na.omit(callservice), aes(x=long,y=lat), alpha = .2) +
+  ggtitle("Calls for Service (not criminal reports) \nWithin 180 days (Feb-July 2016)")+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
 ```
 
-### Heat Map 2: All BPD Stops Density, 2015-2016
-```{r, fig.width=10, fig.asp=0.8, echo=FALSE}
-# a contour plot
-ggmap(berkMap) +
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
+
+############################ populatrion density map ###########################
+popdenmap <- berkgg_zoom1 + 
+  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Berkeley), color = "white", alpha = .7) +
+  ggtitle("2010 Population Density") +
+  labs(fill = "Percent") +
+  scale_fill_gradientn(colors = p) +
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))
+
+# black population:
+
+black2 <- berkgg_zoom1 + 
+  geom_polygon(data = berkcensus2010map, 
+               aes(x = long, 
+                   y = lat, 
+                   group = group, 
+                   fill = Percent.Black), 
+               color = "white", 
+               alpha = .7) +
+  scale_fill_gradientn(colors = p, 
+                       limits = c(0,1)) +
+  guides(fill = "none") +
+  ggtitle("% Black")+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+# White population:
+
+w2 <- berkgg_zoom1 + 
+  geom_polygon(data = berkcensus2010map, 
+               aes(x = long, 
+                   y = lat, 
+                   group = group, 
+                   fill = Percent.White), 
+               color = "white", 
+               alpha = .7) +
+  scale_fill_gradientn(colors = p, 
+                       limits = c(0,1)) +
+  guides(fill = "none") +
+  ggtitle("% White")+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
+# Asian population:
+
+a2 <- berkgg_zoom1 + 
+  geom_polygon(data = berkcensus2010map, 
+               aes(x = long, 
+                   y = lat, 
+                   group = group, 
+                   fill = Percent.Asian), 
+               color = "white", 
+               alpha = .7) +
+  scale_fill_gradientn(colors = p, 
+                       limits = c(0,1)) +
+  guides(fill = "none") +
+  ggtitle("% Asian")+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
+# Hispanic population:
+
+h2 <- berkgg_zoom1 + 
+  geom_polygon(data = berkcensus2010map, 
+               aes(x = long, 
+                   y = lat, 
+                   group = group, 
+                   fill = Percent.Hispanic), 
+               color = "white", 
+               alpha = .7) +
+  scale_fill_gradientn(colors = p, 
+                       limits = c(0,1)) +
+  guides(fill = "none") +
+  ggtitle("% Hispanic")+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
+# Other population:
+
+o2 <- berkgg_zoom1 + 
+  geom_polygon(data = berkcensus2010map, 
+               aes(x = long, 
+                   y = lat, 
+                   group = group, 
+                   fill = Percent.Other), 
+               color = "white", 
+               alpha = .7) +
+  scale_fill_gradientn(colors = p, 
+                       limits = c(0,1)) +
+  guides(fill = "none") +
+  ggtitle("% Other")+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
+```
+
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
+# bar graphs based on stop data:
+names(ss)[2:3] <- c("Percent_Stopped", "Percent_Population_Stopped")
+
+per <- ss %>% 
+  mutate(Percent_Stopped= str_replace(Percent_Stopped, "%", "")) %>% 
+  mutate(Percent_Population_Stopped=str_replace(Percent_Population_Stopped, "%", ""))
+
+per$Percent_Stopped <- as.numeric(per$Percent_Stopped)
+per$Percent_Stopped <- per$Percent_Stopped/100
+per$Percent_Population_Stopped <- as.numeric(per$Percent_Population_Stopped)
+per$Percent_Population_Stopped <- per$Percent_Population_Stopped/100
+
+bargraph1 <- ggplot(per, 
+                    aes(x=Race, 
+                        y=Percent_Stopped, 
+                        fill = Race))+
+  geom_bar(stat = "identity",
+           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0")) +
+  geom_text(aes(x=Race,
+               y=Percent_Stopped,
+               label = paste0(Percent_Stopped*100, "%")), 
+            size=5,
+            colour= "#000000") +
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  labs(title = "Percent Stopped",
+       x = "Race",
+       y = "Percent stopped")+
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank())+
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))
+
+bargraph2 <- ggplot(per, 
+                    aes(x=Race, 
+                        y=Percent_Population_Stopped, 
+                        fill = Race))+
+  geom_bar(stat = "identity",
+           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+  geom_text(aes(x=Race,
+               y=Percent_Population_Stopped,
+               label = paste0(Percent_Population_Stopped*100, "%")), 
+            size=5,
+            colour= "#000000") +
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  labs(title = "Percent Population Stopped",
+       x = "Race",
+       y = "Percent stopped")+
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank())+
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))
+
+```
+
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
+
+# STOP DENSITY MAP:
+stopdensitymap <- berkgg_zoom2 +
   stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
-    size = 2, bins = 5, data = mergedf, geom = "polygon") +
+    size = 2, bins = 5, data = stops, geom = "polygon") +
   scale_fill_gradient(low = "black", high = "red") +
@@
   ggtitle("All BPD Stops Density, 2015-2016") +
-  labs(alpha = element_blank())+
+  labs(alpha = element_blank()) +
   guides(alpha = FALSE)
 
-```
-
-Then we explored the stop data by age range (0-18; 18-29; 30-39; and 40+) and data.
-Similar to the all stop data density, the area that is most likely to be stopped is the same: Downtown Berkeley.
-The differences in age range and race don't play a crucial role in the possibility of being stopped in Berkeley.
-These two analysis confirmed our explanations for the all stop data heat map.
-
-### Heat Map 3: BPD Stop Contour Map of Berkeley by Age Range
-```{r, fig.height = 8, fig.width = 10, echo=FALSE}
-ggmap(berkMap) +
-  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
-    bins = I(5), geom = "polygon", data = mergedf ) +
-  scale_fill_gradient2( "Stop Density",
-    low = "white", mid = "orange", high = "red", midpoint = 25) +
-  labs(x = "Longitude", y = "Latitude") + facet_wrap(~ AgeRange) +
-  scale_alpha(range = c(.2, .55), guide = FALSE) +
-  ggtitle("BPD Stop Contour Map of Berkeley by Age Range") +
-  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
-```
-
-Note: 1 refers to age 0-18; 2 refers to age 18-29; 3 refers to age 30-39; and 4 refers to age 40+.
-
-## Heat Map 4: BPD Stop Contour Map of Berkeley by Race
-```{r, fig.height = 5, fig.width = 10, echo=FALSE}
-ggmap(berkMap) +
-  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
-    bins = I(5), geom = "polygon", data = mergedf ) +
+# stop density 
+stopdensitymap4 <- berkgg_zoom2 +
+  stat_density2d(aes(x = long, 
+                     y = lat, 
+                     fill = ..level.., 
+                     alpha = ..level..),
+    bins = I(5), geom = "polygon", data = stops ) +
   scale_fill_gradient2( "StopDensity",
     low = "white", mid = "orange", high = "red", midpoint = 100) +
-  labs(x = "Longitude", y = "Latitude") + facet_wrap( ~ Race ) +
+  labs(x = "Longitude", y = "Latitude") +
   scale_alpha(range = c(.2, .55), guide = FALSE) +
-  ggtitle("BPD Stop Contour Map of Berkeley by Race") +
-  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
-```
+  ggtitle("Berkeley Police Stop \nDensity Map, 2015-2016")+
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))
 
-The Berkeley PD stop data reflects accidents that have been visited by the Berkeley Police Department.
-After the initial exploration of the Berkeley stop data, we focus on on the information provided by the dispositions variable.
-The dispositions variable provides information on race, gender and age range, stop reason and enforcement of the stop, the car search information during the stop, and additional dispositions.
+############################ populatrion density map ###########################
+popdenmap <- berkgg_zoom1 + 
+  geom_polygon(data = berkcensus2010map, aes(x = long, y = lat, group = group, fill = Percent.Berkeley), color = "white", alpha = .7) +
+  ggtitle("2010 Population Density") +
+  labs(fill = "Percent") +
+  scale_fill_gradientn(colors = p) +
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))
 
-### Analysis of Race
-#### Picture: count of person recorded in each race
-```{r, echo=FALSE}
-stop_race <- stop_by %>%
-  ggvis(x = ~Race , fill = ~AgeRange) %>%
-  layer_bars() %>%
-  add_tooltip(function(stop_by) stop_by$AgeRange)
+# stop density by race
+stopdensitymap3 <- berkgg_zoom2 +
+  stat_density2d(aes(x = long, 
+                      y = lat, 
+                      fill = ..level.., 
+                      alpha = ..level..),
+    bins = I(5), geom = "polygon", data = stops) +
+  scale_fill_gradient2("StopDensity",
+    low = "white", mid = "orange", high = "red", midpoint = 100) +
+  labs(x = "Longitude", y = "Latitude") + facet_grid( ~ Race) +
+  scale_alpha(range = c(.2, .55), guide = FALSE) +
+  ggtitle("Berkeley Police Stop Density Map by Race") +
+  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10)) +
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 14),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10)) +
+  guides(fill = FALSE)
 
-stop_race
-```
+# stops leading to arrests by race:
+df <- subset(stops, as.character(Enforcement) == "Arrest")
 
-####Table: count of person recorded in each race
-
-   Race     | Count | 0-18 | 18-29 | 30-39 | 40+
-----------  |------ |------|-------|-------|-------
-A (Asian)   | 1141  | 8.50%| 43.47%| 21.03%| 26.99%
-B (Black)   | 4636  | 2.29%| 34.97%| 25.09%| 37.66%
-H (Hispanic)| 1676  | 3.22%| 46.12%| 29.18%| 21.47%
-O (Other)   | 1384  | 2.67%| 43.28%| 29.17%| 26.87%
-W (White)   | 5454  | 1.72%| 29.45%| 25.39%| 43.44%
-
-(1) Among the 14291 personnel information recorded, white people is the number one population group of being stopped by Berkeley Police.
-38.16% of all records, namely 5454 people are white.
-black people contributes the second largest percentage.
-32.44% of all records, namely 4636 people are black.
-
-(2) The percentage of people aged from 0 to 18 of each race is less than 9%.
-People aged from 0 to 18 contributes the smallest percentage in each race.
-These teengers are looked after by their parents most of the time.
-
-(3) Among white people and black people recorded by the BPD, people who are over 40 are the number one group that are being stopped, which are 43.44% and 37.66% respectively.
-
-(4) Among Asian, Hispanic and other people recorded by the BPD, people aged from 18 to 29 contribute the largest percentage, which are 43.37%, 46.12% and 43.28%, respectively.
-
-Then we further analyzed the stop data with and without car searches by race.
-Since white people are the largest group to be stopped, we suspect that their possibility of car search will also be the highest.
-However, black people seem to be more likely to be car searched.
-
-### MAP: BPD All Stops w/ Car Searches by Race 2015-2016
-```{r, echo=FALSE}
-ggmap(berkMap) +
-  geom_point(aes(x = long, y = lat, size = CarSearch, color = Race), data = mergedf, alpha = 0.2) + 
-  scale_size_discrete( "Car Search", labels = c("No Search", "Search") ) +
+map_arrested1 <- berkgg_zoom1 +
+  geom_point(aes(x = long, 
+                 y = lat, 
+                 colour = Race), 
+             data = df, 
+             alpha = 0.7) + 
+  scale_color_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
   theme ( 
@@
 		axis.title = element_blank (),
-		axis.ticks = element_blank ()
-		) + 
-  ggtitle("BPD All Stops w/ Car Searches by Race 2015-2016") +
-  facet_wrap(~ Race, ncol = 3) 
+		axis.ticks = element_blank ()) + 
+  ggtitle("Police Stops Leading to Arrests") +
+  facet_grid(~ Race) +
+  guides(colour = FALSE) +
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 14),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
 ```
 
-Even when we look at the arrest data, it is black people who are more likely to get arrest after being stopped.
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
+# stops by reason:
+stop_reason <- stops[stops$Reason != "Other",]
 
-```{r, echo=FALSE}
-df <- subset(mergedf, as.character(Enforcement) == "Arrest")
+stop_by_reason_race <- stop_reason %>%
+  group_by(Race, Reason) %>%
+  tally()
+names(stop_by_reason_race) <- c("Race", "Reason", "totalCount")
 
-ggmap(berkMap) +
-  geom_point(aes(x = long, y = lat, colour = Race), data = df, alpha = 0.5, size = 5) + 
+stop_by_arrest <- stop_reason[stop_reason$Enforcement == "Arrest",] %>% 
+  group_by(Race, Reason) %>%
+  tally()
+names(stop_by_arrest) <- c("Race", "Reason", "arrestCount")
+
+stop_by_percentage_race <- left_join(stop_by_reason_race, stop_by_arrest, by = c("Race", "Reason"))
+stop_by_percentage_race <- stop_by_percentage_race %>%
+  mutate(percentage = arrestCount / totalCount)
+stop_by_percentage_race$percentage <- round(stop_by_percentage_race$percentage, digits = 2)
+
+stop_arrest <- ggplot(stop_by_percentage_race, aes(x = Race, 
+                                                   fill = Race)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+  facet_wrap(~Reason, nrow = 1)+
+  geom_text(aes(x=Race,
+               y=percentage,
+               label = paste0(percentage*100, "%")), 
+            size=4,
+            colour= "#000000") +
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  labs(title = "Probability of arrested by BPD of \na specific race for a specific reason (%)",
+       x = "Race",
+       y = "Density")+
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank())+
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))+
+  guides(fill = FALSE)
+
+map_reasons1 <- berkgg_zoom2 +
+  geom_point(aes(x = long, 
+                 y = lat, 
+                 colour = Reason), 
+             data = stop_reason, 
+             alpha = .7) +
   theme ( 
@@
 		axis.title = element_blank (),
-		axis.ticks = element_blank ()
-		) + 
-  ggtitle("BPD Stops to Arrests, by Race 2015-2016") +
-  facet_wrap(~ Race) 
-```
+		axis.ticks = element_blank ()) + 
+  ggtitle("Police Stops by Reason")+
+  facet_grid(~Reason)+
+  guides(colour = FALSE)+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
 
-### Analysis of stop reasons 
-#### Map: All stop data by reasons.
-```{r, fig.height = 5, fig.width = 10, echo=FALSE}
-ggmap(berkMap) +
-  geom_point(aes(x = long, y = lat, colour = Reason), data = mergedf, alpha = 0.7, size = 3) +
-  theme ( 
-		panel.grid.major = element_blank (), # remove major grid
-		panel.grid.minor = element_blank (),  # remove minor grid
-		axis.text = element_blank (), 
-		axis.title = element_blank (),
-		axis.ticks = element_blank ()
-		) + ggtitle("All BPD Stops, by Reason 2015-2016") +
-    facet_wrap( ~ Reason , ncol = 2)
-```
-
-From the analysis of the stop reasons, we found that the number one reason that people are stopped is traffic reason.
-Then we explored traffic reasons within race.
-As shown, black people are the highest to be stopped and arrested.
-
-#### Map: stop by only traffic reasons
-```{r, fig.height = 5, fig.width = 10, echo=FALSE}
-df <- subset(mergedf, as.character(Reason) == "Traffic")
-
-ggmap(berkMap) +
-  geom_point(aes(x = long, y = lat, colour = Race, size= Arrest), data = df, alpha = 0.5) + 
-  scale_size_discrete( range = c(5, 2) ) +
-  theme ( 
-		panel.grid.major = element_blank (), # remove major grid
-		panel.grid.minor = element_blank (),  # remove minor grid
-		axis.text = element_blank (), 
-		axis.title = element_blank (),
-		axis.ticks = element_blank ()
-		) + 
-  ggtitle("BPD Traffic Stops Only, by Race 2015-2016") +
-  facet_grid(~ Race)
-```
-
-Such contrast of being stopped, being car searched, and being arrested is further demonstrated by the arrest and jail data.
-
-```{r, echo=FALSE}
-p1 <- ggplot(na.omit(arrest))+
-  geom_bar(aes(x=Age, 
-               fill = Race))+
-  facet_wrap(~Race)+
- labs(x= "Age",
-       y = "Number of people",
-       title = "Arrest Data")
-
-p2 <- ggplot(na.omit(jail))+
-  geom_bar(aes(x=Age,
-               fill = Race))+
-  facet_wrap(~Race)
-
-multiplot(p1, p2,cols=2)
+stop_reason <- stop_reason %>% 
+  group_by(Race, Reason) %>% 
+  tally()
 
@@
 
-As we can clearly see from both graphs, black people are the largest group to be arrested and put into jail.
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
 
-Then we looked at the counts of people recorded of different races in every hour.Black people contribute a especially high percentage of the incidents at night, while the white people contribute a especially high percentage of the incidents at the noon.
+############################ create the maps ###########################
 
-```{r, echo=FALSE}
-stop_time_percentage <- ggplot(stop_by, aes(x = Hour, fill = AgeRange)) + 
+# arrest data histogram by race
+arrest1 <- na.omit(arrest) %>% 
+  group_by(Race) %>% 
+  tally()
+total <- arrest1 %>% 
+  tally()
+arresttotal <- 192
+arrest1 <- arrest1 %>% 
+  mutate(percentage = n/192)
+arrest1$percentage <- round(arrest1$percentage, digits = 2)
+
+jail1 <-na.omit(jail) %>% 
+  group_by(Race) %>% 
+  tally()
+total <- jail1 %>% 
+  tally()
+jailtotal <- 242
+jail1<- jail1 %>% 
+  mutate(percentage = n/242)
+jail1$percentage <- round(jail1$percentage, digits = 2)
+
+bararrest<- ggplot(arrest1)+
+  geom_bar(aes(x=Race,
+               y= percentage),
+           stat = "identity",
+           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+  geom_text(aes(x=Race,
+               y=percentage,
+               label = paste0(percentage*100,"%")), 
+            size=5,
+            colour= "#000000")+
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  labs(title = "Races Arrested in 30 Days (%)",
+       x = "Race",
+       y = "Percentage") +
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) + 
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))
+
+barjail <- ggplot(jail1)+
+  geom_bar(aes(x=Race,
+               y=percentage),
+           stat = "identity",
+           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+  geom_text(aes(x=Race,
+               y=percentage,
+               label = paste0(percentage*100,"%")), 
+            size=5,
+            colour= "#000000")+
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  labs(title = "Races Jailed in 30 Days (%)",
+       x = "Race",
+       y = "Percentage") +
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) + 
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7))
+
+histogram_arrest1 <- ggplot(na.omit(arrest))+
+  geom_bar(aes(x=Age, 
+               fill = Race))+
+  facet_grid(~Race)+
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+ labs(x= "Age",
+       y = "Number of people",
+       title = "Arrest Data") +
+  guides(fill = FALSE)+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 14),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
+# jail data histogram by race
+histogram_jail1 <- ggplot(na.omit(jail))+
+  geom_bar(aes(x=Age,
+               fill = Race))+
+  facet_grid(~Race)+
+ labs(x= "Age",
+       y = "Number of people",
+       title = "Jail Data") + 
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+  guides(fill = FALSE)+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 14),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
+# histogram of stop data by race and hour:
+
+histogram_stoptime1 <- ggplot(stops, 
+                              aes(x = Hour, 
+                                  fill = Race)) + 
   geom_bar(aes(y = ..count..)) +
-  labs(y = "Count") +
-  facet_wrap(~Race)+
-  labs(title = "Counts of people recorded of each race in every hour")
-stop_time_percentage
+  facet_grid(~Race)+
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+  labs(title = "Number of people each race in every hour",
+       y = "Count") +
+  guides(fill = FALSE)+
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 14),
+        axis.text.x=element_text(colour="black", size = 10), 
+        axis.text.y=element_text(colour="black", size = 10))
+
 ```
 
-Such analysis made us further digged into the date information.
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
 
-### Analysis of Day
-```{r, echo=FALSE, eval = FALSE}
-stop_day <- stop_by %>%
-  ggvis(x = ~Day, fill = ~Race) %>%
-  layer_bars() %>%
-  add_tooltip(function(stop_by) stop_by$Race)+
-  labs(title = "Counts of people recorded in every day in a week")
-stop_day
+table3$Percent.Stopped <- percent(table3$Percent.Stopped)
+
+map_enforcement <- ggplot(table3)+ 
+  geom_bar(aes(x=Race,
+               y=Percent.Stopped,
+               fill = Race),
+           stat = "identity") +
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
+  scale_y_continuous(limits = c(0,1), labels = c("0%", "25%", "50%", "75%", "100%")) +
+  labs(title = "Police Stops Leading to Arrests",
+       x = "Race",
+       y = "Percent of Arrested") +
+    geom_text(aes(x=Race,
+               y=Percent.Stopped,
+               label = paste0(Percent.Stopped)), 
+            size=5,
+            colour= "#000000")+
+  scale_x_discrete(labels = function(x) str_wrap(x, width = 5))+
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(),
+        panel.border = element_blank(), panel.background = element_blank()) + 
+  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 13),
+        axis.text.x=element_text(colour="black", size = 8), 
+        axis.text.y=element_text(colour="black", size = 8))+
+  guides(fill = FALSE)
+#map_enforcement
+
 ```
 
-####Picture: counts of people recorded in every day in a week
-Day   | Count | A (Asian)| B (Black)| H (Hispanic) | O (Other)| W (White)
-----  |------ |----------|----------|--------------|----------|------
-1     | 1692  | 116      | 653      | 221          | 159      | 543
-2     | 1665  | 149      | 543      | 182          | 166      | 625
-3     | 2344  | 189      | 645      | 231          | 236      | 1043
-4     | 2287  | 206      | 616      | 251          | 244      | 970
-5     | 1941  | 152      | 592      | 223          | 168      | 806
-6     | 2231  | 186      | 764      | 268          | 203      | 810
-7     | 2131  | 143      | 823      | 300          | 208      | 657
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
 
-(1) Among all the day in a week, the number of incidents happened on Sunday and Monday are much less than those happened through Tuesday and Saturday.
+race_hour <- read.csv("../clean_data/race_hour.csv")
 
-(2) The count of records of Asian people is much less than that of any other race of people in each day and the whole week.
+stop_percentage3 <- ggplot(race_hour, aes(x = Hour, fill = Race)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  labs(title = "Probability of BPD stop for \na specific race in a specific hour",
+       x= "Hour",
+       y = "Density") +
+  facet_wrap(~Race, nrow = 1) +
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0")) +
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) +
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7)) +
+  guides(fill = FALSE)
 
-(3) Asian people are less liable to commit an incident on Sunday, and more liable to commit an incident on Wednesday.
-
-(4) Black and Hispanic people are less liable to commit an incident on Monday, and more liable to commit an incident on Saturday.
-
-(5) White people are less liable to commit an incident on Sunday, and more liable to commit an incident on Tuesday and Wednesday, which is similar to Asian people.
-
-## Analysis of incidents of each hour in a day
-#### Picture: count of person recorded in each hour in a day
-```{r, echo=FALSE}
-stop_time <- stop_by %>%
-  ggvis(x = ~Hour, fill := "lightpink") %>%
-  layer_histograms(width = input_slider(1, 12, value = 1, step = 2, label = "width"))
-stop_time
 ```
 
-Note: you can change the value of the input slider to see the number of incidents happened in specific time span of the day.
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
 
-#### Table: count of person recorded in each hour in a day
-Hour  | Count | Hour | Count| Hour | Count| Hour | Count
-----  |------ |------|------|------|------|------|------
-0     | 823   | 6    | 157  | 12   | 802  | 18   | 544
-1     | 618   | 7    | 341  | 13   | 695  | 19   | 685
-2     | 489   | 8    | 489  | 14   | 515  | 20   | 834
-3     | 289   | 9    | 582  | 15   | 542  | 21   | 971
-4     | 135   | 10   | 682  | 16   | 612  | 22   | 1098
-5     | 61    | 11   | 678  | 17   | 578  | 23   | 1071
+age_hour <- read.csv("../clean_data/age_hour.csv")
 
-(1) time span of 22:00 to 22:59 in a day has the highest incidents count, and 23:00 to 23:59 has the second highest, which means that Berkeley area is most dangerous from 10pm to 12am in a day, which corresponds to daily life experience.
-(2) From 12am to 5am, incidents count decreases gradually.
-A reasonable conjecture of this fact is that more and more people choose to sleep as time goes by in this time span.
-For the similar reason, incidents count increases gradually from 5am to 11am.
-(3) After the evening until the midnight, the incidents number increases gradually again in a day. The tally of the statistics and observation in daily life is in a good agreement.
+stop_percentage2 <- ggplot(age_hour , aes(x = Hour, fill = AgeRange)) + 
+  geom_bar(aes(y = percentage), stat="identity") +
+  facet_wrap(~AgeRange, nrow = 1)+
+labs(title = "Probability of BPD stop for a \nspecific age range in a specific hour",
+       x= "Hour",
+       y = "Density") +
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0")) +
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) +
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7)) +
+  guides(fill = FALSE)
 
-#### Picture: Probability of BPD stop for a specific age range in a specific hour
-
-```{r, echo=FALSE}
-stop_by_hourCount <- stop_by %>%
-  group_by(Hour) %>%
-  summarise(hourcount = n())
-stop_by_ageRangeCount <- stop_by %>%
-  group_by(Hour, AgeRange) %>%
-  summarise(ageCount = n())
-stop_by_percentage <- left_join(stop_by_ageRangeCount, stop_by_hourCount, by = "Hour")
-stop_by_percentage <- stop_by_percentage %>%
-  mutate(percentage = ageCount/hourcount)
 ```
 
-```{r, echo=FALSE}
-stop_time_percentage2 <- ggplot(stop_by_percentage, aes(x = Hour, fill = AgeRange)) + 
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
+# Probability of arrested by BPD of a specific gender for a specific reason
+stop_by_reason_gender <-stops %>% 
+  group_by(Gender, Reason) %>% 
+  summarise(totalCount=n())
+stop_by_arrest_gender <- stops[stops$Enforcement == "Arrest",] %>%
+  group_by(Gender, Reason) %>%
+  summarise(arrestCount = n())
+stop_by_percentage_gender <- left_join(stop_by_reason_gender, stop_by_arrest_gender, by = c("Gender", "Reason"))
+stop_by_percentage_gender <- stop_by_percentage_gender %>%
+  mutate(percentage = arrestCount / totalCount)
+stop_by_percentage_gender$percentage <- round(stop_by_percentage_gender$percentage , digits=2)
+stop_by_percentage_gender<- stop_by_percentage_gender[stop_by_percentage_gender$Reason != "Other",]
+
+stop_arrest_gender <- ggplot(stop_by_percentage_gender , aes(x = Gender, fill = Gender)) + 
   geom_bar(aes(y = percentage), stat="identity") +
-  labs(y = "Density") +
-  facet_wrap(~AgeRange)
-stop_time_percentage2
+  facet_wrap(~Reason, nrow=1)+
+labs(title = "Probability of BPD stop for a \nspecific gender range in a specific reason",
+       x= "Hour",
+       y = "Density") +
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0")) +
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  geom_text(aes(x=Gender,
+               y=percentage,
+               label = paste0(percentage*100, "%")), 
+            size=3,
+            colour= "#000000")+
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) +
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7)) +
+  guides(fill = FALSE)
+
 ```
 
-(1) People aged from 18 to 29 are obviously liable to commit incidents at night.
-The average ratio of people aged from 18 to 29 to all people stoped by the Berkeley Police Department at night is greater than 40%.
-
-(2) People aged greater than 40 are obviously liable to commit incidents in the daytime.
-The average ratio of people aged greater than 40 to all people stoped by the Berkeley Police Deparment during the day is greater than 40%.
-
-(3) Ratio of incidents commited by people aged between 0 and 18 and people aged from 30 to 39 fluctuates during the daytime and the night, with an average ratio of 2.5% and 25% respectively.
-
-#### Probability of BPD stop for a specific race in a specific hour
-
-```{r, echo=FALSE}
-stop_by_hourCount <- stop_by %>%
-  group_by(Hour) %>%
-  summarise(hourcount = n())
-stop_by_raceRangeCount <- stop_by %>%
-  group_by(Hour, Race) %>%
-  summarise(raceCount = n())
-stop_by_percentage1 <- left_join(stop_by_raceRangeCount, stop_by_hourCount, by = "Hour")
-stop_by_percentage1 <- stop_by_percentage1 %>%
-  mutate(percentage = raceCount/hourcount)
-```
-```{r, echo=FALSE}
-stop_time_percentage3 <- ggplot(stop_by_percentage1, aes(x = Hour, fill = Race)) + 
-  geom_bar(aes(y = percentage), stat="identity") +
-  labs(y = "Density") +
-  facet_wrap(~Race)
-stop_time_percentage3
-```
-
-(1) Black people are obviously liable to commit incidents at night. The average ratio of black people to all people stoped by the Berkeley Police Department at night is about 40%.
-
-(2) White people are obviously liable to commit incidents in the daytime. The average ratio of white people to all people stoped by the Berkeley Police Deparment during the day is about 45%.
-
-(3) Ratio of incidents commited by Asian people, Hispanic people and other people fluctuate during the daytime and the night, with an average ratio of 8%, 11% and 9% respectively.
-
-### Analysis of preference of Berkeley Police Department arresting people
-#### Picture: Probability of arrested by BPD of a specific race for a specific reason
-```{r,echo=FALSE}
-stop_by_reason_race <- stop_by %>%
-  group_by(Race, Reason) %>%
-  summarise(totalCount = n())
-stop_by_arrest <- stop_by %>%
-  filter(Enforcement == "A") %>%
-  group_by(Race, Reason) %>%
-  summarise(arrestCount = n())
-stop_by_percentage_race <- left_join(stop_by_reason_race, stop_by_arrest, by = c("Race", "Reason"))
-stop_by_percentage_race <- stop_by_percentage_race %>%
-  mutate(percentage = arrestCount / totalCount)
-stop_arrest <- ggplot(stop_by_percentage_race, aes(x = Race, fill = Race)) + 
-  geom_bar(aes(y = percentage), stat="identity") +
-  labs(y = "Density") +
-  facet_wrap(~Reason)
-stop_arrest
-```
-
-#### Table: Probability of arrested by BPD of a specific race for a specific reason
-P   | I      | K     | R     | T     | W     |
-----|------- |-------|-------|-------|-------|
-A   | 10.34% | 75%   | 10.34%| 1.24% | 17.60%|
-B   | 8.31%  | 15.79%| 6.44% | 2.28% | 50.00%|  
-H   | 6.98%  | 16.67%| 9.09% | 2.90% | 25.00%| 
-O   | 10.14% | 18.18%| 5.80% | 0.41% | 41.67%| 
-W   | 8.14%  | 26.79%| 3.40% | 1.04% | 32.56%|
-
-Note: in the picture and table above, I for Investigation, T for Traffic, R for Reasonable Suspicion, K for Probation/Parole, W for wanted.
-
-(1) The probability of arrested in a stop with the reason Traffic is much lower than other reasons. The average conditional probability of arrested given reason is traffic is 1.58%.
-
-(2) The probability of arrested in a stop with the reason Wanted is much higher than other reason. The average conditional probability of arrested given reason is wanted is 33.33%.
-
-(3) An interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and race Asian is 75%, which is much higher than that of any other race.
-Asian people is much more liable to be arrested by Berkeley Police Department during Probation or Parole if stopped by the BPD.
-
-#### Picture: Probability of arrested by BPD of a specific age range for a specific reason
-```{r, echo=FALSE}
-stop_by_reason_age <- stop_by %>%
+```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
+stop_by_reason_age <- stops %>%
   group_by(AgeRange, Reason) %>%
   summarise(totalCount = n())
-stop_by_arrest_age <- stop_by %>%
-  filter(Enforcement == "A") %>%
+stop_by_arrest_age <- stops[stops$Enforcement == "Arrest",] %>%
   group_by(AgeRange, Reason) %>%
@@
   mutate(percentage = arrestCount / totalCount)
+stop_by_percentage_age$percentage<-round(stop_by_percentage_age$percentage,digits = 2)
+stop_by_percentage_age <- stop_by_percentage_age[stop_by_percentage_age$Reason != "Other",]
 
@@
   geom_bar(aes(y = percentage), stat="identity") +
-  labs(y = "Density") +
-  facet_wrap(~Reason)
-stop_arrest_age
+  facet_wrap(~Reason, nrow=1)+
+  labs(title = "Probability of BPD stop for a \nspecific age range in a specific reason",
+       x= "Hour",
+       y = "Density") +
+  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0")) +
+  scale_y_continuous(labels=percent, limits = c(0,1)) +
+  geom_text(aes(x=AgeRange,
+               y=percentage,
+               label = paste0(percentage*100, "%")), 
+            size=3,
+            colour= "#000000")+
+  theme(axis.line = element_line(size=1, colour = "black"), 
+        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
+        panel.border = element_blank(), panel.background = element_blank()) +
+  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
+        text=element_text(family="Times New Roman", size = 10),
+        axis.text.x=element_text(colour="black", size = 7), 
+        axis.text.y=element_text(colour="black", size = 7)) +
+  guides(fill = FALSE)
+
 ```
 
-#### Table: Probability of arrested by BPD of a specific age range for a specific reason
-P     | I (Investigation) |K (Probation/ Parole)| R (Reasonable Suspicion)|T (Traffic)| W (Wanted) 
-----  |------------------ |---------------------|-------------------------|-----------|-------
-0-18  | 17.95%            | 67.57%              | 13.95%                  | 3.24%     | 50.00%  
-18-29 | 8.45%             | 12.50%              | 3.87%                   | 1.82%     | 32.35%  
-30-39 | 6.32%             | 25.00%              | 5.04%                   | 1.49%     | 26.92%  
-40+   | 7.87%             | 23.86%              | 6.13%                   | 1.41%     | 31.82%
+## Abstract
 
-(1) Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.
+_For the STAT 133 final project, our group was interested in finding some ways of visualizing the relationship between the Berkeley Police Department and different groups of people, especially within the context of different populations and neighborhoods within Berkeley.
+To obtain our results, we accessed, cleaned and explored the four Berkeley Police Department data sets available to the public on the Berkeley Open Data website using R, the statistical programming language.
+These data sets were of police stops, calls for service, arrests, and jail bookings.
+There were 16,255 stops (including stops for traffic, suscpicious vehicle, bicycle, and pedestrian) from a 16-month period that began on January 26, 2015 (in compliance with General Order B-4, which ordered this data set to be collected for the public) and ending on April 30th, 2016.
+The calls of service included 4,911 calls, which were collected over the most recent a 6-month period, from January 20, 2016 to July 13, 2016.
+The arrest data included 232 arrests from the previous 1-month period, from June 20, 2016 to July 19, 2016.
+The jail bookings data set included 248 jail bookings over the same 1-month period.
+In addition to the police data, we accessed the Berkeley city and the Alameda County 2010 census data in order to obtain demographic information about Berkeley as a whole and about the demographics of the neighborhoods inside of Berkeley.
+The three census data sets that we used are also available through Berkeley's open data site, including the Berkeley 2000-2010 census data, which gave us the total population counts and percentages of each group of people in Berkeley for the 2010 census, the Alameda County 2010 census tract population data, which gave us the demographic information for 33 different neighborhoods (or "tracts") inside Berkeley, and the Berkeley 2010 census tract polygons shapefile (.shp), which provided the information we needed to visually display the 33 neighborhoods.
+Rather than limiting the project to one set of data, incorporating multiple data sets gave us a more holistic understanding of subject and expanded our ability to understand and visualize the data.
+Our findings suggest the hypothesis that (1) black people in Berkeley are more likely than members of another group of people to be stopped by Berkeley police, and that (2) black people in Berkeley are arrested and jailed at disproportionate rates.
+We recommend future statistical analysis be done to test our hypothesis._
 
-(2) Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and age range 0-18 is 67.56%, which is much higher than that of any other age range.
+## Introduction
 
-(3) The conditional probability of arrested given person involved aged from 0 to 18 in a stop given any reason is more than that of given person with any other age range.
-Therefore, in a stop, teenagers is more liable to be arrested by Berkeley Police Department.
+First we will detail our process of cleaning the police and census data and obtaining spatial information for the police data.
+Next, we will provide our analysis, which includes visualizations of the police and census data and some preliminary findings related to these visualizations.
+Afterwards, we will discuss problems with the data and recommendations for future study on this subject.
 
-#### Picture: Probability of arrested by BPD of a specific gender for a specific reason
-```{r, echo=FALSE}
-stop_by_reason_gender <- stop_by %>%
-  group_by(Gender, Reason) %>%
-  summarise(totalCount = n())
-stop_by_arrest_gender <- stop_by %>%
-  filter(Enforcement == "A") %>%
-  group_by(Gender, Reason) %>%
-  summarise(arrestCount = n())
-stop_by_percentage_gender <- left_join(stop_by_reason_gender, stop_by_arrest_gender, by = c("Gender", "Reason"))
-stop_by_percentage_gender <- stop_by_percentage_gender %>%
-  mutate(percentage = arrestCount / totalCount)
+## Cleaning Data
 
-stop_arrest_gender <- ggplot(stop_by_percentage_gender, aes(x = Gender, fill = Gender)) + 
-  geom_bar(aes(y = percentage), stat="identity") +
-  labs(y = "Density") +
-  facet_wrap(~Reason)
-stop_arrest_gender
+###__Berkeley Stop Data Cleaning Process__
+
+The original stop data set included 16,255 stops recorded by the Berkeley police over a 16-month period.
+For each stop, the police recorded the date and time, location (as a character string, usually in the form of a street block or a street intersection), incident type (referring to whether the stop was traffic, pedestrian, bike, or suspicious vehicle), and a column called disposition.
+The disposition column was a character string typically including 6 characters per individual as follows: the first character was race, the second gender, the third age range, the fourth reason (for the stop), the fifth enforcement (the result of the stop), and the sixth car search (yes or no).
+Each individual that was involved with the stop was recorded using comma separation in the same row and column followed by an additional 6 characters.
+Additional optional dispositions were sometimes included, again using comma-separation from the 6-character strings (though not necessarily in an order), ranging from one to three characters.
+These additional disposition abbreviations represented the following options: Primary Case Report, MDT Narrative Only, Arrest Reporty Only (No Case Report Submitted), Incident Report, Field Card, Collision Investigation Report, Emergency Psychiatric Evaluation, Impounded Vehicle, and Officer Made a Stop of More Than 5 Persons.
+We found it reasonable to assume that these additional dispositions were related to police paperwork procedures, and since they were used so sparingly, we decided to not consider most of them.
+However, in spite of only being used less than 20 times, we found the Emergency Psychiatric Evaluation disposition to be of interest, and so we chose to investigate this one in addition to the main 6 character set of dispostions, though we chose not to focus on it during the limited scope of this paper.
+
+Cleaning the stop data was relatively time-consuming compared to the other data sets.
+First, the dispositions column in the original data set contained many columns and sometimes rows worth of information, though the comma-separated pieces were without order (for example, sometimes the optional disposition abbreviations came before the 6-character string instead).
+We handled this problem using data tidying techniques, such as separating disposition information into separate row entries for each individual assessed, in the case of multiple persons stopped, and splitting the column by isolating the six character dispositions into different columns.
+The optional disposition character strings were moved and separated into more columns.
+
+Additional cleaning for the stop data included changing the stop date and time variable to the `lubridate` date format.
+We also created an hour column and day column in order to examine the stop occurrences by hour over a 24-hour period and accross a 7-day week.
+
+We used the stop data's character string locations into location coordinates using a Google Maps geolocation service provided by the library `ggmap` with the function `geocode`.
+This task proved difficult because of common data entry mistakes.
+For example, there were common mispellings of street names and the use of several different abbreviations to mean the same thing.
+Additionally, we found that the geolocation provided by Google does not accept forward slash characters the data used in the column to indicate street intersections, so these forward slash characters were changed to `and` for the location processing, which was successful.
+Google also does not accept the term "block" to indicate a range of addresses, and so the word "block" was removed from the location column for location processing.
+Most of these types of errors were delt with on an automated basis before geolocation, using `str_detect` and `str_replace` from the `stringr` package.
+The number of problems with the location data that were left over afterwards were small enough to handle on a case-by-case basis.
+Further, because Google limits geolocation to 2500 queries per 24-hour period, it took approximately a week to process all the locations into coordinates from the stop data (this includes the time it took to fix spelling errors).
+
+The reason why the Incident Number column was not removed was because two teams from our group worked on cleaning different parts of the stop data at the same time; one team worked on separating the disposition column into variables, while the other team worked on creating the columns for longitude and latitude.
+After both teams had finished their cleaning task, the resulting data frames were then merged by Incident Number.
+
+At the end of the cleaning process, the stop data had been whittled down from 16,255 observations to the 14,291 police stops (due to missing disposition values) that we will use in the following analysis.
+Please see Figure 1 for a glimpse at the clean police stops data frame.
+
+![](Stop_data_header.png)
+
+###__Berkeley Arrest and Jail Bookings Cleaning Process__
+
+The number of observations available for the arrest data set and the jail data set were significantly smaller than the number of observations for the stop data set.
+The arrest data had 232 observations and the jail bookings data had 248.
+These observations were obtained by the Berkeley Police department over the same month-long time period between June and July of 2016.
+They both contained similar variables, including a case/arrest/booking number, date and time, type, and subject information (name, race, sex, D.O.B., age, height, weight, hair, eyes, and occupation) and statute information (type, description, agency, and disposition).
+Cleaning this data set also required the dates and times to be put into `lubridate` format, and we also created an hour and day column.
+See Figure 2 for the cleaned arrest data frame and Figure 3 for the cleaned jailings data frame.
+
+![](arrest_data_header.png)
+
+![](jail_data_header.png)
+
+###__Berkeley Census 2010 Data Cleaning Process__
+
+The Berkeley census 2010 data was obtained from three separate data sets, as previously mentioned.
+We first considered the summary information provided by the city website, which was easy to clean and use in our analysis, since it provided a summary of the demographic information for each group in Berkeley as a whole.
+However, in order to understand the demographics within Berkeley in a way that could provide a spatial visualization, we needed to use the census 2010 Alameda County population tract data, which provided the population and race data by tract number for the entire county.
+We manipulated this data into a data frame containing the 33 census tracts inside Berkeley, each row representing a single tract and the columns providing demographic information for that tract.
+To be able to map these 33 tracts, we also needed the census 2010 Berkeley tract polygons shapefile.
+After converting the shapefile to a mapping dataframe and changing the coordinate system to the one used by Google (this was done using the `readOGR` function from the library `rgdal`), we were able to `left_join` this new polygonal mapping data frame to the data frame of the 33 census tracts by the unique census 2010 tract number.
+This new data frame allowed us to map population information for 33 different areas.
+The goal was to visualize the population density of different groups of people within the city by neighborhood, which provided us with a context for the sptatially visualized police data.
+
+![](census_data_header.png)
+
+## Data Analysis
+
+### Stop Data Analysis
+
+Figure 5 below provides a map of the location information of the Berkeley police stop data mapped by location.
+We have excluded some observations on the outskirts of Berkeley in order to zoom in to the street level.
+Note that the `alpha` variable in `ggplot` is set to a small number in order to show the incidents that occur multiple times in the same place.
+In other words, the darker the point, the more police stops have occured at that location.
+
+```{r, warning = FALSE, echo = F, fig.width = 8, fig.height = 3, message = FALSE, fig.align="center", fig.cap= "Berkeley Police Stops"}
+map_stop1 # figure 5
 ```
 
-#### Table: Probability of arrested by BPD of a specific gender range for a specific reason
-P     | I (Investigation) |K (Probation/ Parole)| R (Reasonable Suspicion)|T (Traffic)| W (Wanted) 
-----  |------------------ |---------------------|-------------------------|-----------|-------
-F     | 9.05%             | 50.00%              | 5.00%                   | 1.10%     | 28.57%  
-M     | 8.26%             | 20.11%              | 5.42%                   | 1.85%     | 37.04%
+Another visualization of the stop data mapped in Figure 5 is the two-dimensional density map of the police stop data in the left pane of Figure 6.
+One of the most interesting aspects of the stop data in this image is that most police stops seem to be clustered west of UC Berkeley Campus in the Downtown area.
+This was contrary to what some of us might have thought before, which was that the area with the higher police stops would be the area with the higher population.
+To visualize the relationship between police stops and population density, consider the map in the right pane of Figure 6, which visualizes the percent population of Berkeley for each of the 33 census 2010 tracts.
+As you can see, the least densely populated area is the census 2010 tract containing the University of California.
+This makes sense because the area of this tract includes a campus and some student housing areas, where permanent residents are less likeley to live.
+One of the most population-dense areas is Southside, the portion of Berkeley directly south of the UC campus.
+Note that even though the tract including the Southside area is the most densely populated of the 33 census 2010 tracts in Berkeley, it is not the area where most police stops occur.
 
-(1) Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.
+```{r, warning = FALSE, echo = F, fig.width = 8, fig.height = 3, message = FALSE, fig.align="center", fig.cap= "2015-2016 Police Stop Density (left) and 2010 Census Tract Population Density (right)"}
+grid.arrange(stopdensitymap4, popdenmap, ncol=2) # figure 6
+```
 
-(2) Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole of female is 50.00%, which is much higher than that of male, which is 20.11%.
+In order to break down the stop data further, we mapped the data by each of the disposition variables.
+We found race to be one of the most interesting variables.
+Consider the stop data density map faceted by race in Figure 7.
+
+```{r, fig.height = 3, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align="center", fig.cap= "Police Stop Density Map by Race"}
+stopdensitymap3 # figure 7
+```
+
+Figure 7 seems to suggest that the group of people stopped the most in Berkeley are African Americans.
+For instance, in Centeral Berkeley, the stops for the races defined by the Berkeley Police Department as Asian, Hispanic, White, and Other seem comparatively lower than for Blacks in the same area.
+In order to consider the context of these stops, we considered the population distribution for these areas, shown in Figure 8.
+Note that the population density visualized by Figure 8 is conveyed by the range between white (0%) and red (100%).
+As you can see, White people make up a striking majority in all census 2010 tract areas, which seems to suggest the hypothesis that the population of Black people in Berkeley is being disproportionally affected by police stops.
+
+```{r, warning = FALSE, echo = FALSE,  fig.width = 15, fig.height = 3, message = FALSE, fig.align = "center", fig.cap= "Berkeley Population Density Map by Race"}
+grid.arrange(a2,black2,h2,o2,w2, ncol = 5) # print population density by race # figure 8
+```
+
+Now, with the population percentages of each group in hand, we were able to compare the results of our stop data analysis to the percent population of each group.
+In Figure 9 (left), the `Percent Stopped` chart conveys the percent of each group stopped compared to the total number of people stopped.
+In Figure 9 (right), the `Percent Population Stopped` chart conveys the proportion of people in each group stopped relative to the population of that group living in Berkeley.
+Even though the people stopped by police are not necessarily Berkeley residents, the `Percent Population Stopped` chart suggests that some groups of people in Berkeley might be more likely to be affected by police stops, especially Black people.
+
+```{r, message = F, echo = FALSE, fig.width = 10, fig.height = 3, warning = FALSE, fig.align="center", fig.cap= "Police Stops in Berkeley by Race"}
+grid.arrange(bargraph1, bargraph2, ncol = 2)
+```
+
+Another disposition variable we found to be of interest was the Enforcement variable, which has the options Arrest, Citation, Warning, and Other.
+We were interested to see how these enforcements were applied to different groups of people.
+Consider Figure 10 (top), which maps the stops that led to the enforcement of arrest.
+The number of incidents of police stops mapped for Blacks seems higher than for the other groups.
+Upon examination of the chart in Figure 10 (bottom), we see that of the 441 individuals stopped and then arrested as a result of the stop during this 18-month period, Blacks were the majority.
+Though the Berkeley police department is prohibited from the use of biased policing practices, Figure 10 does suggest that the enforcement of arrest may be applied differently to different groups of people.
+
+```{r, echo=FALSE, warning=F,message=F, fig.align="center", fig.width = 10, fig.height=7, fig.cap = "Police Stops Leading to Arrests"}
+grid.arrange(map_arrested1, map_enforcement, ncol=1)
+```
+
+A third interesting disposition variable was the Reason disposition, which indicates the reason for the police stop in each case.
+The reasons available for police stops are Investigation, Other, Probation/Parole, Reasonable Suspicion, Traffic, and Wanted.
+One way we explored the data initially was to see how these reasons affected different groups of people, and we would recommend exploring this idea in depth at a later time.
+Due to the length and scope of this paper however, we decided to end our discussion of the stop data by showing that most of the Berkeley police stops are traffic stops.
+Mapping the coordinates by reason for the stop, as in Figure 11, suggests that this is the most common reason for a stop in Berkeley.
+
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Police Stops by Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+map_reasons1
+```
+
+We also analyzed possible preferences of Berkeley police officers when arresting individuals.
+We calculated the conditional probability of an individual being arrested given a specific reasion for the stop and given a specific race / age range / gender.
+That is to say, we calculated the probability a person will be arrested if he was stopped for a specific reason and was of a specific race / age range / gender.
+
+For Figure 12, we can tell that the probability of a stopped individual being arrested in a stop with the reason being traffic is much lower than other stop reasons.
+The average conditional probability of being arrested for a traffic stop is 1.58%.
+Additionally, the probability of being arrested in a stop with the reason wanted is much higher than other reason.
+The average conditional probability of being arrested given being stopped for the wanted reason is 33.33%.
+Lastly, an interesting fact is that the conditional probability of being arrested in a stop for probation or parole and race Asian is 75%, which is much higher than that of any other race.
+Our conclusion is that Asian people are much more liable to be arrested in Berkeley during their probation or parole if stopped by police.
+
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Police Stop: Race VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+stop_arrest # figure 12
+```
+
+In Figure 13, the probability of being arrested as a result of a police stop in Berkeley for a traffic reason is again the lowest and wanted is the highest.
+Again similar to the conclusion in Figure 12, an interesting fact is that the conditional probability of being arrested in a stop with the reason probation or parole and age range 0-18 is 67.56%, which is much higher than that of any other age range for that reason.
+Lastly, the conditional probability of being arrested in a police stop given an individual in the age range 0-18 given any reason is more than that of any other age range.
+This suggests that in a police stop, teenagers are more liable to be arrested in Berkeley.
+
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Police Stop: Age VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+stop_arrest_age # figure 13
+```
+
+Again for Figure 14, the conditional probability of being arrested during a police stop for the stop reason of traffic is the lowest and wanted is the highest.
+Again similar to the conclusion above, an interesting fact is that the conditional probability of being arrested in a police stop with the stop reason probation or parole given a gender of female is 50.00%, which is much higher than that of male, which is 20.11%.
+
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Police Stop: Gender VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+stop_arrest_gender # figure 14
+```
+
+Figure 15 is interesting because it suggests that black people are more likely to be stopped by the police in Berkeley during the hours of night.
+The average ratio of black people to all people stopped by the Berkeley Police Department at night is about 40%.
+Figure 15 also suggests that white people are more liable to be stopped by the police in Berkeley during the hours of daytime.
+The average ratio of white people to all people stopped by the Berkeley Police Deparment during the day is about 45%.
+While the ratio of Asian people, Hispanic people and other people stopped by police fluctuate during the daytime and the night, with an average ratio of 8%, 11% and 9% respectively.
+
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Police Stop: Race VS Hour", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+stop_percentage3 # figure 15
+```
+Figure 16 implies that people aged from 18 to 29 are more liable to be stopped by police in Berkeley at night.
+The average ratio of people aged from 18 to 29 to all people stopped by the Berkeley Police Department at night is greater than 40%.
+The figure also implies that people aged 40+ are more liable to be stopped by the police during the daytime.
+The average ratio of people aged 40+ to all people stopped by police during the day is greater than 40%.
+Further, the ratio of stops of people aged 0-18 and people aged 30-39 fluctuates during the daytime and the night, with an average ratio of 2.5% and 25% respectively.
+
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Arrested  During a Police Stop: Age VS Hour", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+stop_percentage2 # figure 16
+```
+
+### Berkeley Police Jail and Arrest Data Set Analysis
+
+Moving on from the stop data, we considered the available jail data and arrest data for the city of Berkeley, which is not necessarily dependent on the stop data.
+Two of the most interesting variables to consider in these two data sets were also race and age.
+We decided the examine the percent of people arrested and the percent of people jailed in Berkeley by both age and race.
+Figure 17 is a visualization.
+Note that the following percentages are related to the number of total people jailed or the number of people arrested during a recent 30 day time frame, respectively, and we have not adjusted the data relative to the population of Berkeley.
+It is interesting to note that the proportions of arrests and jailings during this time are nearly identical, since these arrests and jailings occur during the same time frame.
+The data suggests that being jailed after being arrested in Berkeley is almost certain, and that this is true for all races.
+However, note that the proportion of Blacks being arrested and jailed is more than the other races, in spite of Berkeley being a city of mostly White people.
+
+```{r, echo=FALSE, fig.cap= "Arrested and Jailed in Berkeley", warning = FALSE, message=FALSE, fig.height = 3, fig.width = 10, fig.align="center"}
+grid.arrange(bararrest, barjail, ncol = 2) # figure 17
+```
+
+## Problems
+
+The data sets provided by the police (stops, arrests, and jail bookings) have several problems.
+First, the open data site updates the data frequently and we are concerned that the data available to be used for this paper won't be available after the data is updated, since the early data is removed.
+Second, concerning the methods for recording disposition data like race and gender, the Berkeley Open Data website indicates that the police officers conducting the stops are deciding these dispositions on their own, meaning that they are judging whether or not the individual apprehended is male or female, black or hispanic, etc., meaning that the data could be misrecorded.
+Even though there doesn't seem to be a means to get around this problem with the actual method of collecting the disposition data for stops that don't lead to citation or arrest, it is still important to address it as something that could induce error into a statistical analysis of the data.
+
+Another problem with the stop data is the many data-entry problems we faced in the character string location column (which sometimes included only 1 street), since there is likely some stops that are not mapped correctly due to approximations made by the Google geolocation service (for example, in some cases, only having one street name instead of a corner caused Google to estimate the exact location based on few conditions).
+Though most of the location problems were fixed using string detection, there is likely some error in our visual representation of the stop data maps.
+Also, the website indicates that any specific addresses were changed to "block" address form to preserve anonymity; therefore the spatial data cannot be expected to be of a precision smaller than a city block.
+
+In regards to the racial data from all of the police data, it is important to note that the census data we used to compare the stop data to included many more categories for race than just the five used by police (Asian, Black, Hispanic, Other, and White).
+To deal with this problem, other races included by the census data but not included by the police data were combined with the category of other.
+Additionally, the census data had a category for Two or More Races, which we also included as Other in order to compare the data sets.
+While the census data allowed for overlap between racial categories, there was no overlap for the racial categories in the police data.
+This creates a problem when comparing the two data sets, and so it should be acknowledged.
+
+Finally, it should be noted that a more accurate analysis of Berkeley police data would also incorporate demographics for the homeless population as well as for the UC student population, both of which are not included in the census data.
+Including these data sets would provide a more nuanced context for the stop data.
 
@@
 
-1.
-From analyzing the relationship between population density and stop data density, we found that he place that are the most likely to be stopped is not the place where most people live, but the transportation hub.
-2.
-From various analysis of the data by race, such as the stop time by race, car search by race, we found that black people are the number one to be arrested and even put into jail.
-Surprisingly, they are not the number one to be stopped.
-White people are the largest group to be stopped.
-3. During daytime, young people and white people are more likely to be stopped. During nightime, people over 40 and black people are more likely to be stopped by police. 
-4. Asian and female are more likely to be arrested by the BDP for probation/parole.
+The goal of this paper was to investigate and visualize Berkeley Police interactions with the community through the available police data.
+Although our preliminary findings unfortunately seem to suggest that certain groups of people, especially black people, are disproportionately likely to be stopped, arrested, and jailed in the city of Berkeley, we recommend a more rigorous statistical analysis of the data be done in order to assess this claim.
+We also recommend an idea we had for this project to future researchers that could make the Berkeley police data more accessible.
+For example, one could make the data more accessible by creating a `shiny` app that allows the user to toggle between the different dispositions of the data (race, age range, enforcement, reason, etc.) and between modes of view (map or bar chart, for example) or order to play with the data in a visual way.
+Another `shiny` app idea we had was to filter the calls for service data by proximity to certain locations so that users could use the app to select a location and find a visual displaying the most common types of calls in that area.
+It would also be useful to find a way to create an app that can easily be updated when the new data is posted (every month), allowing users of the app to get the most up-to-date information.
+
+## Sources
+
+* [Stop Data (16,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Stop-Data/6e9j-pj9p)
+* [Arrest (200)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Arrests/xi7q-nji6)
+* [Jail Bookings (250)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Jail-Bookings/7ykt-c32j)
+* [Calls for Service (4,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Calls-for-Service/k2nh-s5h5)
+* [Berkeley Census Data](https://data.cityofberkeley.info/Demographics/Census-Data-2000-And-2010/bnkq-f2ge)
+* [Census 2010 Population and race data by county tract polygons](http://www.bayareacensus.ca.gov/small/small.htm)
+* [Berkeley Census 2010 Tract Polygons](https://data.cityofberkeley.info/Demographics/Census-Tract-Polygons-2010/peq3-2arw)

### 2016-08-10 final_paper/STAT 133 Final Project_ Visualizing Berkeley Police Data.Rmd
final paper

--- 2016-08-09 final_paper/draft.Rmd
@@
     latex_engine: xelatex
+    fig_caption: yes
+    keep_tex: yes
+header-includes: \usepackage{float}
 sansfont: Times New Roman
@@
 ```{r setup, include=FALSE, warning = FALSE, message = FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, cache = TRUE, fig.pos = 'H')
 library(sp)
@@
 setwd("~/Desktop/Stats R/PROJECT/FINALPAPER/finalproject_angry_ladies/final_paper")
+
 ```
@@
 
-```{r, warning = FALSE, echo = F, fig.width = 8, fig.height = 3, message = FALSE, fig.align="center", fig.cap= "Berkeley Police Stops"}
+```{r, warning = FALSE, echo = F, fig.width = 8, fig.height = 3, message = FALSE, fig.align="center", fig.cap= "Berkeley Police Stops", fig.show="asis"}
 map_stop1 # figure 5
@@
 
-```{r, warning = FALSE, echo = F, fig.width = 8, fig.height = 3, message = FALSE, fig.align="center", fig.cap= "2015-2016 Police Stop Density (left) and 2010 Census Tract Population Density (right)"}
+```{r, warning = FALSE, echo = F, fig.width = 8, fig.height = 2, message = FALSE, fig.align="center", fig.cap= "2015-2016 Police Stop Density (left) and 2010 Census Tract Population Density (right)", fig.show="asis"}
 grid.arrange(stopdensitymap4, popdenmap, ncol=2) # figure 6
@@
 
-```{r, fig.height = 3, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align="center", fig.cap= "Police Stop Density Map by Race"}
+```{r, fig.height = 2, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align="center", fig.cap= "Police Stop Density Map by Race", fig.show="asis"}
 stopdensitymap3 # figure 7
@@
 
-```{r, warning = FALSE, echo = FALSE,  fig.width = 15, fig.height = 3, message = FALSE, fig.align = "center", fig.cap= "Berkeley Population Density Map by Race"}
+```{r, warning = FALSE, echo = FALSE,  fig.width = 15, fig.height = 3, message = FALSE, fig.align = "center", fig.cap= "Berkeley Population Density Map by Race", fig.show="asis"}
 grid.arrange(a2,black2,h2,o2,w2, ncol = 5) # print population density by race # figure 8
@@
 
-```{r, message = F, echo = FALSE, fig.width = 10, fig.height = 3, warning = FALSE, fig.align="center", fig.cap= "Police Stops in Berkeley by Race"}
+```{r, message = F, echo = FALSE, fig.width = 10, fig.height = 3, warning = FALSE, fig.align="center", fig.cap= "Police Stops in Berkeley by Race", fig.show="asis"}
 grid.arrange(bargraph1, bargraph2, ncol = 2)
@@
 
-```{r, echo=FALSE, warning=F,message=F, fig.align="center", fig.width = 10, fig.height=7, fig.cap = "Police Stops Leading to Arrests"}
+```{r, echo=FALSE, warning=F,message=F, fig.align="center", fig.width = 10, fig.height=5, fig.cap = "Police Stops Leading to Arrests", fig.show="asis"}
 grid.arrange(map_arrested1, map_enforcement, ncol=1)
@@
 
-```{r, fig.height = 3, fig.width = 10, fig.cap= "Police Stops by Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Police Stops by Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center", fig.show="asis"}
 map_reasons1
@@
 
-```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Police Stop: Race VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Police Stop: Race VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center", fig.show="asis"}
 stop_arrest # figure 12
@@
 
-```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Police Stop: Age VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Police Stop: Age VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center", fig.show="asis"}
 stop_arrest_age # figure 13
@@
 
-```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Police Stop: Gender VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Police Stop: Gender VS Reason", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center", fig.show="asis"}
 stop_arrest_gender # figure 14
@@
 
-```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Police Stop: Race VS Hour", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Being Arrested During a Police Stop: Race VS Hour", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center", fig.show="asis"}
 stop_percentage3 # figure 15
@@
 
-```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Arrested  During a Police Stop: Age VS Hour", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
+```{r, fig.height = 3, fig.width = 10, fig.cap= "Conditional Probability of Arrested  During a Police Stop: Age VS Hour", echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center", fig.show="asis"}
 stop_percentage2 # figure 16
@@
 
-```{r, echo=FALSE, fig.cap= "Arrested and Jailed in Berkeley", warning = FALSE, message=FALSE, fig.height = 3, fig.width = 10, fig.align="center"}
+```{r, echo=FALSE, fig.cap= "Arrested and Jailed in Berkeley", warning = FALSE, message=FALSE, fig.height = 3, fig.width = 10, fig.align="center", fig.show="asis"}
 grid.arrange(bararrest, barjail, ncol = 2) # figure 17

### 2016-08-10 final_paper/STAT 133 Final Project_ Visualizing Berkeley Police Data.Rmd
html~

@@
 ---
-title: "STAT 133 Final Project: Visualizing Berkeley Police Data"
+title: 'STAT 133 Final Project: Visualizing Berkeley Police Data'
 author: "Amy Zhu, Mengyu Li, Rebecca Reus, and Shangjun Jiang"
@@
 output:
-    pdf_document:
-    latex_engine: xelatex
+  html_document: default
   fig_caption: yes
   keep_tex: yes
+  latex_engine: xelatex
+  pdf_document: null
+sansfont: Times New Roman
 header-includes: \usepackage{float}
-sansfont: Times New Roman
-
 ---
